
# Study Participants - Disposition and Characteristics

## Enrolment

Participants were enrolled at 4 sites in Australia. There were `r nrow(dat_base)` participants randomised, `r nrow(dat_base)/2` per treatment group. The first participant was randomised on `r as_date(min(dat_base$randdattim))` and the final participant was randomised on `r as_date(max(dat_base$randdattim))` [@fig-enrolment-overall].

```{r}
#| label: fig-enrolment-overall
#| fig-cap: Cumulative trial enrolment by calendar time.
#| fig-height: 3
#| fig-width: 6
dat_enrol <- dat_base |>
  select(record_id, randdattim) |>
  mutate(randdate = as_date(randdattim)) |>
  arrange(randdate) |>
  summarise(n = n(), .by = randdate) |>
  mutate(cn = cumsum(n))

ggplot(
  dat_enrol,
  aes(randdate, cn)
) +
  geom_step() +
  scale_x_date("Calendar time", breaks = "3 months", date_labels = "%d %b %y") +
  scale_y_continuous("Number enrolled", breaks = seq(0, 1000, 100)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Study completion

The study completion status as reported in the database is summarised in @tbl-discharge-outcome and by site in @tbl-discharge-outcome-site. Most of the *early termination - consent withdrawal* and *early termination - protocol violation/deviation* were due to the first 300 PCH participants who were scheduled for the additional blood collections, a number of whom withdrew consent for blood collection at the 18-month and 19-month visits (@tbl-discharge-outcome-site). Data was generally still collected for these participants with the exception of those blood collections, and so these participants may not have been "early termination" in the sense of having withdrawn from participating in the trial. There were 17 participants where contact was lost at or before 15 months of age (based on `discdat`) (@tbl-listing-early-withdrawal).

```{r}
#| label: tbl-discharge-outcome
#| code-summary: Study completion
#| tbl-cap: Number of early study withdrawals and associated reasons.
#| tbl-pos: "H"
#| output: "asis"
tbl_study_termination <- function(dat) {
  st_by_trt <- dat |>
    count(streas, stetrreas, trt) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = trt) |>
    pivot_wider(names_from = trt, values_from = c(n, p))

  st_ovr <- dat |>
    count(streas, stetrreas) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n)) |>
    rename(n_ovr = n, p_ovr = p)

  left_join(st_by_trt, st_ovr, join_by(streas, stetrreas)) |>
    gt() |>
    cols_align(
      align = "left",
      columns = c(streas, stetrreas)
    ) |>
    fmt_percent(starts_with("p_"), decimals = 1) |>
    cols_merge(c(n_aP, p_aP), pattern = "{1} ({2})") |>
    cols_merge(c(n_wP, p_wP), pattern = "{1} ({2})") |>
    cols_merge(c(n_ovr, p_ovr), pattern = "{1} ({2})") |>
    cols_label(
      streas ~ md("**Participant status**"),
      stetrreas ~ md("**Reason**"),
      n_aP ~ "\\makecell{\\textbf{aP} \\\\ N = 486}",
      n_wP ~ "\\makecell{\\textbf{wP} \\\\ N = 486}",
      n_ovr ~ "\\makecell{\\textbf{Overall} \\\\ N = 972}"
    ) |>
    tab_options(table.font.size = "10px") |>
    sub_missing() |>
    tab_style(
      style = cell_text(align = "left"),
      locations = cells_footnotes()
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(n_aP:n_ovr)
    )
}

tbl_study_termination(dat_base) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-discharge-outcome-site
#| code-summary: Study completion by site
#| tbl-cap: Number of early study withdrawals and associated reasons by study site.
#| tbl-pos: "H"
#| output: "asis"
tbl_study_termination_by <- function(dat, by = rand_site) {
  st_by_trt <- dat |>
    count({{ by }}, streas, stetrreas, trt) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = c({{ by }}, trt)) |>
    pivot_wider(names_from = trt, values_from = c(n, p))

  st_ovr <- dat |>
    count({{ by }}, streas, stetrreas) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = {{ by }}) |>
    rename(n_ovr = n, p_ovr = p)

  left_join(st_by_trt, st_ovr, join_by({{ by }}, streas, stetrreas)) |>
    group_by({{ by }}) |>
    gt() |>
    cols_align(
      align = "left",
      columns = c(streas, stetrreas)
    ) |>
    fmt_percent(starts_with("p_"), decimals = 1) |>
    cols_merge(c(n_aP, p_aP), pattern = "{1} ({2})") |>
    cols_merge(c(n_wP, p_wP), pattern = "{1} ({2})") |>
    cols_merge(c(n_ovr, p_ovr), pattern = "{1} ({2})") |>
    cols_label(
      streas ~ md("**Participant status**"),
      stetrreas ~ md("**Reason**"),
      n_aP ~ "\\makecell{\\textbf{aP} \\\\ N = 486}",
      n_wP ~ "\\makecell{\\textbf{wP} \\\\ N = 486}",
      n_ovr ~ "\\makecell{\\textbf{Overall} \\\\ N = 972}"
    ) |>
    tab_options(table.font.size = "10px") |>
    sub_missing() |>
    tab_style(
      style = cell_text(align = "left"),
      locations = cells_footnotes()
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(n_aP:n_ovr)
    )
}

tbl_study_termination_by(dat_base) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-listing-early-withdrawal
#| tbl-cap: |
#|   Listing of infants with last contact occurring at or prior to 15 months of age.
#| tbl-pos: "H"
dat_base |>
  select(record_id, trt, stetrreas, birthdat, disc_age_mth) |>
  left_join(
    filter(dat_grid_vs, visage == "12-month") |>
      select(record_id, visit_attended, windreas, windothspec) |>
      rename(
        visit_attend_12m = visit_attended,
        reas_12m = windreas,
        oth_12m = windothspec
      ),
    join_by(record_id)
  ) |>
  left_join(
    filter(dat_grid_vs, visage == "18-month") |>
      select(record_id, visit_attended, windreas, windothspec) |>
      rename(
        visit_attend_18m = visit_attended,
        reas_18m = windreas,
        oth_18m = windothspec
      ),
    join_by(record_id)
  ) |>
  arrange(disc_age_mth) |>
  filter(disc_age_mth <= 15) |>
  select(-starts_with("reas"), -starts_with("oth")) |>
  gt() |>
  cols_label(
    record_id = md("**Record ID**"),
    trt = md("**Treatment**"),
    birthdat = md("**Date of birth**"),
    stetrreas = md("**Withdrawal reason**"),
    disc_age_mth = md("**Age last contact**"),
    visit_attend_12m = md("**Attend 12-month visit**"),
    visit_attend_18m = md("**Attend 18-month visit**")
  ) |>
  fmt_date(
    c(birthdat)
  ) |>
  sub_missing() |>
  tab_options(table.font.size = px(9), table.width = pct(100))
```

\clearpage

## Protocol deviations

Full deviations are documented in the site deviation log.
The following reports only on those of particular relevance to the analyses.

### Treatment administration {#sec-treatment-admin}

All participants received the first scheduled vaccination at the first visit.

For one participant the pharmacy reportedly dispensed the alternative study vaccine to what the participant had been randomly allocated. The IMP was monitored for all participants, and this was the only documented occasion that this was reported to have occurred. This error was documented in the protocol deviations log [subject 02-333, record id 4629-184]. This participant was analysed as assigned rather than as treated. Given there was only one such case, the analyses were not repeated with their treatment group switched to as treated.

### Completed visits

The majority of participants attended each study visit [@tbl-stage1-visit-attendance; @tbl-stage2-visit-attendance].

```{r}
#| label: tbl-stage1-visit-attendance
#| tbl-cap: |
#|   Number and proportion of infants who attended each visit in stage 1
#|   (based on presence of a reported "visit date").
#|    6-month + 72 hour was an optional visit.
#| tbl-pos: "H"
#| output: asis
vs_by_trt <- dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = c(trt, visage)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

vs_ovr <- dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = visage
  )

left_join(vs_by_trt, vs_ovr) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    wP ~ "\\makecell{\\textbf{wP} \\\\ N = 75}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 150}"
  ) |>
  cols_align(align = "left", columns = visage) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-stage2-visit-attendance
#| tbl-cap: |
#|   Number and proportion of infants who attended each visit in stage 2
#|   (based on presence of a reported "visit date").
#| tbl-pos: "H"
#| output: "asis"
vs_by_trt <- dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = c(trt, visage)
  ) |>
  pivot_wider(names_from = trt, values_from = lab) |>
  relocate(wP, .after = aP)

vs_ovr <- dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = visage
  )

left_join(vs_by_trt, vs_ovr) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    wP ~ "\\makecell{\\textbf{wP} \\\\ N = 411}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 822}"
  ) |>
  cols_align(align = "left", columns = visage) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  tab_footnote(
    footnote = "19-month visit was for first 150 at PCH only: 46/75 (61%) aP and 58/75 (77%) wP of PCH attended.",
    locations = cells_body(columns = visage, rows = visage == "19-month"),
    placement = "right"
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-in-window-attended-st1
#| tbl-cap: |
#|   Amongst stage 1 participants who attended the visit, what percentage were reportedly in the protocolised visit window (`windyn`). Cells are y/n (p%) where y is number in window, n number who attended, and p percentage of attendees in window.
dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    .by = c(visage, trt),
    attended = sum(visit_attended),
    window = sum(windyn & visit_attended, na.rm = TRUE),
    p = window / attended,
    lab = sprintf("%i/%i (%.0f%%)", window, attended, 100 * p)
  ) |>
  select(-attended, -window, -p) |>
  pivot_wider(names_from = trt, values_from = lab) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit age**"),
    aP ~ md("**aP**"),
    wP ~ md("**wP**"),
  ) |>
  cols_align(align = "left", columns = visage) |>
  tab_options(table.font.size = "10px", table.width = pct(50))
```

```{r}
#| label: tbl-in-window-attended-st2
#| tbl-cap: |
#|   Amongst stage 2 participants who attended the visit, what percentage were reportedly (`windyn`) in the protocolised visit window. Cells are y/n (p%) where y is number in window, n number who attended, and p percentage of attendees in window.
dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    .by = c(visage, trt),
    attended = sum(visit_attended),
    window = sum(windyn & visit_attended, na.rm = TRUE),
    p = window / attended,
    lab = sprintf("%i/%i (%.0f%%)", window, attended, 100 * p)
  ) |>
  select(-attended, -window, -p) |>
  pivot_wider(names_from = trt, values_from = lab) |>
  relocate(wP, .after = aP) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit age**"),
    aP ~ md("**aP**"),
    wP ~ md("**wP**"),
  ) |>
  cols_align(align = "left", columns = visage) |>
  tab_options(table.font.size = "10px", table.width = pct(50))
```

\clearpage

### Food and household questionnaires

The majority of food and household questionnaires were completed.

```{r}
#| label: tbl-stage1-fhq-completion
#| tbl-cap: |
#|   Number and proportion of infants who completed food and household questionnaire for each occasion in stage 1.
#| tbl-pos: "H"
#| output: "asis"
fhq_by_trt <- dat_grid_fhq |>
  filter(rand_stage == 1) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = c(trt, visit_age)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

fhq_ovr <- dat_grid_fhq |>
  filter(rand_stage == 1) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = visit_age
  )

left_join(fhq_by_trt, fhq_ovr) |>
  gt() |>
  cols_label(
    visit_age ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    wP ~ "\\makecell{\\textbf{wP} \\\\ N = 75}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 150}"
  ) |>
  cols_align(align = "left", columns = visit_age) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-stage2-fhq-completion
#| tbl-cap: |
#|   Number and proportion of infants who completed food and household questionnaire for each occasion in stage 2.
#| tbl-pos: "H"
#| output: "asis"
fhq_by_trt <- dat_grid_fhq |>
  filter(rand_stage == 2) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = c(trt, visit_age)
  ) |>
  pivot_wider(names_from = trt, values_from = lab) |>
  relocate(wP, .after = aP)

fhq_ovr <- dat_grid_fhq |>
  filter(rand_stage == 2) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = visit_age
  )

left_join(fhq_by_trt, fhq_ovr) |>
  gt() |>
  cols_label(
    visit_age ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    wP ~ "\\makecell{\\textbf{wP} \\\\ N = 411}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 822}"
  ) |>
  cols_align(align = "left", columns = visit_age) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

## Characteristics

The baseline characteristics of all participants by their randomly assigned treatment group are reported in @tbl-baseline-characteristics. Characteristics were similar between the groups.

```{r}
#| label: tbl-baseline-characteristics
#| tbl-cap: Baseline characteristics of trial participants.
#| tbl-pos: "H"
#| output: "asis"
tbl <- dat_base |>
  set_variable_labels(
    etomthr___1 = "Indigenous Australia",
    etomthr___2 = "European Caucasian",
    etomthr___3 = "Asian",
    etomthr___4 = "Indian subcontinent",
    etomthr___5 = "Middle Eastern",
    etomthr___6 = "Pacific Islander",
    etomthr___7 = "South American",
    etomthr___8 = "Black African",
    etofthr___1 = "Indigenous Australia",
    etofthr___2 = "European Caucasian",
    etofthr___3 = "Asian",
    etofthr___4 = "Indian subcontinent",
    etofthr___5 = "Middle Eastern",
    etofthr___6 = "Pacific Islander",
    etofthr___7 = "South American",
    etofthr___8 = "Black African",
    etoinf___1 = "Indigenous Australia",
    etoinf___2 = "European Caucasian",
    etoinf___3 = "Asian",
    etoinf___4 = "Indian subcontinent",
    etoinf___5 = "Middle Eastern",
    etoinf___6 = "Pacific Islander",
    etoinf___7 = "South American",
    etoinf___8 = "Black African",
  ) |>
  mutate(
    bcmoth = fct_lump_min(bcmoth, min = 10) |>
      # fct_recode(Other = "OTHER") |>
      fct_infreq() |>
      fct_relevel("Other", after = Inf),
    bcfath = fct_lump_min(bcfath, min = 10) |>
      fct_recode(Other = "OTHER") |>
      fct_infreq() |>
      fct_relevel("Other", after = Inf)
  ) |>
  tbl_summary(
    include = c(
      # Demographics and infant history
      rand_stage,
      rand_site,
      gender,
      v1_age_days,
      parinc,
      bcmoth,
      bcfath,
      etomthr___2,
      etomthr___3,
      etomthr___4,
      etomthr___7,
      etomthr___1,
      etomthr___6,
      etomthr___8,
      etofthr___2,
      etofthr___3,
      etofthr___4,
      etofthr___6,
      etofthr___7,
      etofthr___1,
      etofthr___8,
      etoinf___2,
      etoinf___3,
      etoinf___4,
      etoinf___7,
      etoinf___6,
      etoinf___1,
      etoinf___8,
      fborn,
      fecurr,
      fecat,
      fedog,
      # Birth history
      mgrav_fac,
      mpar_fac,
      delt,
      ipab,
      neoab,
      gestadelw,
      wgt,
      len,
      inhc,
      aps1_fac,
      aps5_fac,
      # Maternal immunisation
      inmpertvp,
      vacad,
      prevpert,
      minfvp,
      # Family history
      fha
    ),
    label = list(
      rand_stage ~ "Protocol stage",
      rand_site ~ "Participating site",
      bcmoth ~ "Birth country of mother",
      bcfath ~ "Birth country of father",
      v1_age_days ~ "Age at enrolment (days)",
      fborn ~ "First born",
      mgrav_fac ~ "Maternal gravidity",
      mpar_fac ~ "Maternal parity",
      aps1_fac ~ "Apgar score (1 minute)",
      aps5_fac ~ "Apgar score (5 minutes)",
      prevpert ~ "Other maternal pertussis vaccination (last 5 years)",
      fha ~ "Family history of atopy",
      fecat ~ "Cat ownership",
      fedog ~ "Dog ownership",
      fecurr ~ "Feeding status"
    ),
    by = trt,
    missing_text = "(Missing)",
    digits = ~0
  ) |>
  add_n() |>
  add_overall(last = TRUE) |>
  add_variable_group_header(header = "Mother's ethnicity", variables = starts_with("etomthr")) |>
  add_variable_group_header(header = "Father's ethnicity", variables = starts_with("etofthr")) |>
  add_variable_group_header(header = "Infant's ethnicity", variables = starts_with("etoinf")) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      mutate(
        n = if_else(grepl("___", variable), NA_character_, n),
        n = if_else(row_type == "variable_group", '972', n)
      )
  )
tbl |>
  as_gt() |>
  cols_width(label ~ pct(30)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = (row_type == "label" | row_type == "variable_group") & !grepl("___", variable)
    )
  ) |>
  tab_options(
    table.width = pct(100),
    table.font.size = px(9),
    latex.use_longtable = TRUE
  ) |>
  tab_footnote(
    footnote = "Multiple may apply to each participant",
    locations = cells_body(rows = row_type == "variable_group", columns = label)
  ) |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```

```{=latex}
\clearpage
```

Reactogenicity and immunogenicity data was only collected for the first 300 participants at PCH.
The baseline characteristics of this cohort of participants by their randomly assigned treatment group are reported in @tbl-baseline-characteristics-pch-300.

```{r}
#| label: tbl-baseline-characteristics-pch-300
#| tbl-cap: Baseline characteristics of trial participants.
#| tbl-pos: "H"
#| output: "asis"
tbl <- dat_base |>
  filter(rand_site == "PCH") |>
  filter(substr(subjid, 4, 6) %in% str_pad(1:300, 3, pad = "0")) |>
  set_variable_labels(
    etomthr___1 = "Indigenous Australia",
    etomthr___2 = "European Caucasian",
    etomthr___3 = "Asian",
    etomthr___4 = "Indian subcontinent",
    etomthr___5 = "Middle Eastern",
    etomthr___6 = "Pacific Islander",
    etomthr___7 = "South American",
    etomthr___8 = "Black African",
    etofthr___1 = "Indigenous Australia",
    etofthr___2 = "European Caucasian",
    etofthr___3 = "Asian",
    etofthr___4 = "Indian subcontinent",
    etofthr___5 = "Middle Eastern",
    etofthr___6 = "Pacific Islander",
    etofthr___7 = "South American",
    etofthr___8 = "Black African",
    etoinf___1 = "Indigenous Australia",
    etoinf___2 = "European Caucasian",
    etoinf___3 = "Asian",
    etoinf___4 = "Indian subcontinent",
    etoinf___5 = "Middle Eastern",
    etoinf___6 = "Pacific Islander",
    etoinf___7 = "South American",
    etoinf___8 = "Black African",
  ) |>
  mutate(
    bcmoth = fct_lump_min(bcmoth, min = 5) |>
      fct_recode(Other = "OTHER") |>
      fct_infreq() |>
      fct_relevel("Other", after = Inf),
    bcfath = fct_lump_min(bcfath, min = 5) |>
      fct_recode(Other = "OTHER") |>
      fct_infreq() |>
      fct_relevel("Other", after = Inf)
  ) |>
  tbl_summary(
    include = c(
      # Demographics and infant history
      rand_stage,
      gender,
      v1_age_days,
      parinc,
      bcmoth,
      bcfath,
      etomthr___2,
      etomthr___3,
      etomthr___4,
      etomthr___7,
      etomthr___6,
      etomthr___8,
      etomthr___1,
      etofthr___2,
      etofthr___3,
      etofthr___4,
      etofthr___1,
      etofthr___8,
      etofthr___6,
      etofthr___7,
      etoinf___2,
      etoinf___3,
      etoinf___4,
      etoinf___7,
      etoinf___1,
      etoinf___8,
      etoinf___6,
      fborn,
      fecurr,
      fecat,
      fedog,
      # Birth history
      mgrav_fac,
      mpar_fac,
      delt,
      ipab,
      neoab,
      gestadelw,
      wgt,
      len,
      inhc,
      aps1_fac,
      aps5_fac,
      # Maternal immunisation
      inmpertvp,
      vacad,
      prevpert,
      minfvp,
      # Family history
      fha
    ),
    label = list(
      rand_stage ~ "Protocol stage",
      bcmoth ~ "Birth country of mother",
      bcfath ~ "Birth country of father",
      v1_age_days ~ "Age at enrolment (days)",
      fborn ~ "First born",
      mgrav_fac ~ "Maternal gravidity",
      mpar_fac ~ "Maternal parity",
      aps1_fac ~ "Apgar score (1 minute)",
      aps5_fac ~ "Apgar score (5 minutes)",
      prevpert ~ "Other maternal pertussis vaccination (last 5 years)",
      fha ~ "Family history of atopy",
      fecat ~ "Cat ownership",
      fedog ~ "Dog ownership",
      fecurr ~ "Feeding status"
    ),
    by = trt,
    missing_text = "(Missing)",
    digits = ~0
  ) |>
  add_n() |>
  add_overall(last = TRUE) |>
  add_variable_group_header(header = "Mother's ethnicity", variables = starts_with("etomthr")) |>
  add_variable_group_header(header = "Father's ethnicity", variables = starts_with("etofthr")) |>
  add_variable_group_header(header = "Infant's ethnicity", variables = starts_with("etoinf")) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      mutate(
        n = if_else(grepl("___", variable), NA_character_, n),
        n = if_else(row_type == "variable_group", '300', n)
      )
  )
tbl |>
  as_gt() |>
  cols_width(label ~ pct(30)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = (row_type == "label" | row_type == "variable_group") & !grepl("___", variable)
    )
  ) |>
  tab_options(
    table.width = pct(100),
    table.font.size = px(9),
    latex.use_longtable = TRUE
  ) |>
  tab_footnote(
    footnote = "Multiple may apply to each participant",
    locations = cells_body(rows = row_type == "variable_group", columns = label)
  ) |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```


\clearpage
