
# Study Participants - Disposition and Characteristics

## Enrolment

Participants were enrolled at 4 sites in Australia. There were `r nrow(dat_base)` participants randomised, `r nrow(dat_base)/2` per treatment group. The first participant was randomised on `r as_date(min(dat_base$randdattim))` and the final participant was randomised on `r as_date(max(dat_base$randdattim))` [@fig-enrolment-overall].

```{r}
#| label: fig-enrolment-overall
#| fig-cap: Cumulative trial enrolment by calendar time.
#| fig-height: 3
#| fig-width: 6
dat_enrol <- dat_base |>
  select(record_id, randdattim) |>
  mutate(randdate = as_date(randdattim)) |>
  arrange(randdate) |>
  summarise(n = n(), .by = randdate) |>
  mutate(cn = cumsum(n))

ggplot(
  dat_enrol,
  aes(randdate, cn)
) +
  geom_step() +
  scale_x_date("Calendar time", breaks = "3 months", date_labels = "%d %b %y") +
  scale_y_continuous("Number enrolled", breaks = seq(0, 1000, 100)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Study completion

The study completion status as reported in the database is summarised in @tbl-discharge-outcome and by site in @tbl-discharge-outcome-site. Most of the *early termination - consent withdrawal* and *early termination - protocol violation/deviation* were due to the first 300 PCH participants who were scheduled for the additional blood collections, a number of whom withdrew consent for blood collection at the 18-month and 19-month visits (@tbl-discharge-outcome-site). Data was generally still collected for these participants with the exception of those blood collections, and so these participants may not have been "early termination" in the sense of having withdrawn from participating in the trial. There were 17 participants where contact was lost at or before 15 months of age (based on `discdat`) (@tbl-listing-early-withdrawal).

```{r}
#| label: tbl-discharge-outcome
#| code-summary: Study completion
#| tbl-cap: Number of early study withdrawals and associated reasons.
#| tbl-pos: "H"
#| output: "asis"
tbl_study_termination <- function(dat) {
  st_by_trt <- dat |>
    count(streas, stetrreas, trt) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = trt) |>
    pivot_wider(names_from = trt, values_from = c(n, p))

  st_ovr <- dat |>
    count(streas, stetrreas) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n)) |>
    rename(n_ovr = n, p_ovr = p)

  left_join(st_by_trt, st_ovr, join_by(streas, stetrreas)) |>
    gt() |>
    cols_align(
      align = "left",
      columns = c(streas, stetrreas)
    ) |>
    fmt_percent(starts_with("p_"), decimals = 1) |>
    cols_merge(c(n_aP, p_aP), pattern = "{1} ({2})") |>
    cols_merge(c(n_wP, p_wP), pattern = "{1} ({2})") |>
    cols_merge(c(n_ovr, p_ovr), pattern = "{1} ({2})") |>
    cols_label(
      streas ~ md("**Participant status**"),
      stetrreas ~ md("**Reason**"),
      n_aP ~ "\\makecell{\\textbf{aP} \\\\ N = 486}",
      n_wP ~ "\\makecell{\\textbf{wP} \\\\ N = 486}",
      n_ovr ~ "\\makecell{\\textbf{Overall} \\\\ N = 972}"
    ) |>
    tab_options(table.font.size = "10px") |>
    sub_missing() |>
    tab_style(
      style = cell_text(align = "left"),
      locations = cells_footnotes()
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(n_aP:n_ovr)
    )
}

tbl_study_termination(dat_base) |>
  tab_footnote(
    footnote = "In most cases, this was due to no blood collection at visit 3 resulting in no visit 4. Only applicable at PCH.",
    location = cells_body(columns = stetrreas, rows = stetrreas == "Protocol violation/deviation")
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-discharge-outcome-site
#| code-summary: Study completion by site
#| tbl-cap: Number of early study withdrawals and associated reasons by study site.
#| tbl-pos: "H"
#| output: "asis"
tbl_study_termination_by <- function(dat, by = rand_site) {
  st_by_trt <- dat |>
    count({{ by }}, streas, stetrreas, trt) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = c({{ by }}, trt)) |>
    pivot_wider(names_from = trt, values_from = c(n, p))

  st_ovr <- dat |>
    count({{ by }}, streas, stetrreas) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = {{ by }}) |>
    rename(n_ovr = n, p_ovr = p)

  left_join(st_by_trt, st_ovr, join_by({{ by }}, streas, stetrreas)) |>
    group_by({{ by }}) |>
    gt() |>
    cols_align(
      align = "left",
      columns = c(streas, stetrreas)
    ) |>
    fmt_percent(starts_with("p_"), decimals = 1) |>
    cols_merge(c(n_aP, p_aP), pattern = "{1} ({2})") |>
    cols_merge(c(n_wP, p_wP), pattern = "{1} ({2})") |>
    cols_merge(c(n_ovr, p_ovr), pattern = "{1} ({2})") |>
    cols_label(
      streas ~ md("**Participant status**"),
      stetrreas ~ md("**Reason**"),
      n_aP ~ "\\makecell{\\textbf{aP} \\\\ N = 486}",
      n_wP ~ "\\makecell{\\textbf{wP} \\\\ N = 486}",
      n_ovr ~ "\\makecell{\\textbf{Overall} \\\\ N = 972}"
    ) |>
    tab_options(table.font.size = "10px") |>
    sub_missing() |>
    tab_style(
      style = cell_text(align = "left"),
      locations = cells_footnotes()
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(n_aP:n_ovr)
    )
}

tbl_study_termination_by(dat_base) |>
  tab_footnote(
    footnote = "In most cases, this was due to no blood collection at visit 3 resulting in no visit 4",
    location = cells_body(columns = stetrreas, rows = stetrreas == "Protocol violation/deviation")
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-listing-early-withdrawal
#| tbl-cap: |
#|   Listing of infants with last contact occurring at or prior to 15 months of age.
#| tbl-pos: "H"
dat_base |>
  select(record_id, trt, stetrreas, birthdat, disc_age_mth) |>
  left_join(
    filter(dat_grid_vs, visage == "12-month") |>
      select(record_id, visit_attended, windreas, windothspec) |>
      rename(
        visit_attend_12m = visit_attended,
        reas_12m = windreas,
        oth_12m = windothspec
      ),
    join_by(record_id)
  ) |>
  left_join(
    filter(dat_grid_vs, visage == "18-month") |>
      select(record_id, visit_attended, windreas, windothspec) |>
      rename(
        visit_attend_18m = visit_attended,
        reas_18m = windreas,
        oth_18m = windothspec
      ),
    join_by(record_id)
  ) |>
  arrange(disc_age_mth) |>
  filter(disc_age_mth <= 15) |>
  select(-starts_with("reas"), -starts_with("oth")) |>
  gt() |>
  cols_label(
    record_id = md("**Record ID**"),
    trt = md("**Treatment**"),
    birthdat = md("**Date of birth**"),
    stetrreas = md("**Withdrawal reason**"),
    disc_age_mth = md("**Age last contact**"),
    visit_attend_12m = md("**Attend 12-month visit**"),
    visit_attend_18m = md("**Attend 18-month visit**")
  ) |>
  fmt_date(
    c(birthdat)
  ) |>
  sub_missing() |>
  tab_options(table.font.size = px(9), table.width = pct(100))
```

\clearpage

## Protocol deviations

Full deviations are documented in the site deviation log.
The following reports only on those of particular relevance to the analyses.

### Treatment administration {#sec-treatment-admin}

All participants received the first scheduled vaccination at the first visit.

For one participant the pharmacy reportedly dispensed the alternative study vaccine to what the participant had been randomly allocated. The IMP was monitored for all participants, and this was the only documented occasion that this was reported to have occurred. This error was documented in the protocol deviations log [subject 02-333, record id 4629-184]. This participant was analysed as assigned rather than as treated. Given there was only one such case, the analyses were not repeated with their treatment group switched to as treated.

### Completed visits

The majority of participants attended each study visit [@tbl-stage1-visit-attendance; @tbl-stage2-visit-attendance].

```{r}
#| label: tbl-stage1-visit-attendance
#| tbl-cap: |
#|   Number and proportion of infants who attended each visit in stage 1
#|   (based on presence of a reported "visit date").
#|    6-month + 72 hour was an optional visit.
#| tbl-pos: "H"
#| output: asis
vs_by_trt <- dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = c(trt, visage)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

vs_ovr <- dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = visage
  )

left_join(vs_by_trt, vs_ovr) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    wP ~ "\\makecell{\\textbf{wP} \\\\ N = 75}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 150}"
  ) |>
  cols_align(align = "left", columns = visage) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-stage2-visit-attendance
#| tbl-cap: |
#|   Number and proportion of infants who attended each visit in stage 2
#|   (based on presence of a reported "visit date").
#| tbl-pos: "H"
#| output: "asis"
vs_by_trt <- dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = c(trt, visage)
  ) |>
  pivot_wider(names_from = trt, values_from = lab) |>
  relocate(wP, .after = aP)

vs_ovr <- dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = visage
  )

left_join(vs_by_trt, vs_ovr) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    wP ~ "\\makecell{\\textbf{wP} \\\\ N = 411}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 822}"
  ) |>
  cols_align(align = "left", columns = visage) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  tab_footnote(
    footnote = "19-month visit was for first 150 at PCH only: 46/75 (61%) aP and 58/75 (77%) wP of PCH attended.",
    locations = cells_body(columns = visage, rows = visage == "19-month"),
    placement = "right"
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-in-window-attended-st1
#| tbl-cap: |
#|   Amongst stage 1 participants who attended the visit, what percentage were reportedly in the protocolised visit window (`windyn`). Cells are y/n (p%) where y is number in window, n number who attended, and p percentage of attendees in window.
dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    .by = c(visage, trt),
    attended = sum(visit_attended),
    window = sum(windyn & visit_attended, na.rm = TRUE),
    p = window / attended,
    lab = sprintf("%i/%i (%.0f%%)", window, attended, 100 * p)
  ) |>
  select(-attended, -window, -p) |>
  pivot_wider(names_from = trt, values_from = lab) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit age**"),
    aP ~ md("**aP**"),
    wP ~ md("**wP**"),
  ) |>
  cols_align(align = "left", columns = visage) |>
  tab_options(table.font.size = "10px", table.width = pct(50))
```

```{r}
#| label: tbl-in-window-attended-st2
#| tbl-cap: |
#|   Amongst stage 2 participants who attended the visit, what percentage were reportedly (`windyn`) in the protocolised visit window. Cells are y/n (p%) where y is number in window, n number who attended, and p percentage of attendees in window.
dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    .by = c(visage, trt),
    attended = sum(visit_attended),
    window = sum(windyn & visit_attended, na.rm = TRUE),
    p = window / attended,
    lab = sprintf("%i/%i (%.0f%%)", window, attended, 100 * p)
  ) |>
  select(-attended, -window, -p) |>
  pivot_wider(names_from = trt, values_from = lab) |>
  relocate(wP, .after = aP) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit age**"),
    aP ~ md("**aP**"),
    wP ~ md("**wP**"),
  ) |>
  cols_align(align = "left", columns = visage) |>
  tab_options(table.font.size = "10px", table.width = pct(50))
```

\clearpage

### Food and household questionnaires

The majority of food and household questionnaires were completed.

```{r}
#| label: tbl-stage1-fhq-completion
#| tbl-cap: |
#|   Number and proportion of infants who completed food and household questionnaire for each occasion in stage 1.
#| tbl-pos: "H"
#| output: "asis"
fhq_by_trt <- dat_grid_fhq |>
  filter(rand_stage == 1) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = c(trt, visit_age)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

fhq_ovr <- dat_grid_fhq |>
  filter(rand_stage == 1) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = visit_age
  )

left_join(fhq_by_trt, fhq_ovr) |>
  gt() |>
  cols_label(
    visit_age ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    wP ~ "\\makecell{\\textbf{wP} \\\\ N = 75}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 150}"
  ) |>
  cols_align(align = "left", columns = visit_age) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-stage2-fhq-completion
#| tbl-cap: |
#|   Number and proportion of infants who completed food and household questionnaire for each occasion in stage 2.
#| tbl-pos: "H"
#| output: "asis"
fhq_by_trt <- dat_grid_fhq |>
  filter(rand_stage == 2) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = c(trt, visit_age)
  ) |>
  pivot_wider(names_from = trt, values_from = lab) |>
  relocate(wP, .after = aP)

fhq_ovr <- dat_grid_fhq |>
  filter(rand_stage == 2) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = visit_age
  )

left_join(fhq_by_trt, fhq_ovr) |>
  gt() |>
  cols_label(
    visit_age ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    wP ~ "\\makecell{\\textbf{wP} \\\\ N = 411}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 822}"
  ) |>
  cols_align(align = "left", columns = visit_age) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

## Baseline characteristics

The baseline characteristics of all participants by their randomly assigned treatment group are reported in @tbl-baseline-characteristics. Characteristics were similar between the groups.

```{r}
#| label: tbl-baseline-characteristics
#| tbl-cap: Baseline characteristics of trial participants.
#| tbl-pos: "H"
#| output: "asis"
tbl <- dat_base |>
  set_variable_labels(
    etomthr___1 = "Indigenous Australia",
    etomthr___2 = "European Caucasian",
    etomthr___3 = "Asian",
    etomthr___4 = "Indian subcontinent",
    etomthr___5 = "Middle Eastern",
    etomthr___6 = "Pacific Islander",
    etomthr___7 = "South American",
    etomthr___8 = "Black African",
    etofthr___1 = "Indigenous Australia",
    etofthr___2 = "European Caucasian",
    etofthr___3 = "Asian",
    etofthr___4 = "Indian subcontinent",
    etofthr___5 = "Middle Eastern",
    etofthr___6 = "Pacific Islander",
    etofthr___7 = "South American",
    etofthr___8 = "Black African",
    etoinf___1 = "Indigenous Australia",
    etoinf___2 = "European Caucasian",
    etoinf___3 = "Asian",
    etoinf___4 = "Indian subcontinent",
    etoinf___5 = "Middle Eastern",
    etoinf___6 = "Pacific Islander",
    etoinf___7 = "South American",
    etoinf___8 = "Black African",
  ) |>
  mutate(
    bcmoth = fct_lump_min(bcmoth, min = 10) |>
      # fct_recode(Other = "OTHER") |>
      fct_infreq() |>
      fct_relevel("Other", after = Inf),
    bcfath = fct_lump_min(bcfath, min = 10) |>
      fct_recode(Other = "OTHER") |>
      fct_infreq() |>
      fct_relevel("Other", after = Inf)
  ) |>
  tbl_summary(
    include = c(
      # Demographics and infant history
      rand_stage,
      rand_site,
      gender,
      v1_age_days,
      parinc,
      bcmoth,
      bcfath,
      etomthr___2,
      etomthr___3,
      etomthr___4,
      etomthr___7,
      etomthr___1,
      etomthr___6,
      etomthr___8,
      etofthr___2,
      etofthr___3,
      etofthr___4,
      etofthr___6,
      etofthr___7,
      etofthr___1,
      etofthr___8,
      etoinf___2,
      etoinf___3,
      etoinf___4,
      etoinf___7,
      etoinf___6,
      etoinf___1,
      etoinf___8,
      fborn,
      fecurr,
      fecat,
      fedog,
      # Birth history
      mgrav_fac,
      mpar_fac,
      delt,
      ipab,
      neoab,
      gestadelw,
      wgt,
      len,
      inhc,
      aps1_fac,
      aps5_fac,
      # Maternal immunisation
      inmpertvp,
      vacad,
      prevpert,
      minfvp,
      # Family history
      fha
    ),
    label = list(
      rand_stage ~ "Protocol stage",
      rand_site ~ "Participating site",
      bcmoth ~ "Birth country of mother",
      bcfath ~ "Birth country of father",
      v1_age_days ~ "Age at enrolment (days)",
      fborn ~ "First born",
      mgrav_fac ~ "Maternal gravidity",
      mpar_fac ~ "Maternal parity",
      aps1_fac ~ "Apgar score (1 minute)",
      aps5_fac ~ "Apgar score (5 minutes)",
      prevpert ~ "Other maternal pertussis vaccination (last 5 years)",
      fha ~ "Family history of atopy",
      fecat ~ "Cat ownership",
      fedog ~ "Dog ownership",
      fecurr ~ "Feeding status"
    ),
    by = trt,
    missing_text = "(Missing)",
    digits = ~0
  ) |>
  add_n() |>
  add_overall(last = TRUE) |>
  add_variable_group_header(header = "Mother's ethnicity", variables = starts_with("etomthr")) |>
  add_variable_group_header(header = "Father's ethnicity", variables = starts_with("etofthr")) |>
  add_variable_group_header(header = "Infant's ethnicity", variables = starts_with("etoinf")) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      mutate(
        n = if_else(grepl("___", variable), NA_character_, n),
        n = if_else(row_type == "variable_group", '972', n)
      )
  )
tbl |>
  as_gt() |>
  cols_width(label ~ pct(30)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = (row_type == "label" | row_type == "variable_group") & !grepl("___", variable)
    )
  ) |>
  tab_options(
    table.width = pct(100),
    table.font.size = px(9),
    latex.use_longtable = TRUE
  ) |>
  tab_footnote(
    footnote = "Multiple may apply to each participant",
    locations = cells_body(rows = row_type == "variable_group", columns = label)
  ) |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```

```{=latex}
\clearpage
```

Reactogenicity and immunogenicity data was only collected for the first 300 participants at PCH.
The baseline characteristics of this cohort of participants by their randomly assigned treatment group are reported in @tbl-baseline-characteristics-pch-300.

```{r}
#| label: tbl-baseline-characteristics-pch-300
#| tbl-cap: Baseline characteristics of trial participants.
#| tbl-pos: "H"
#| output: "asis"
tbl <- dat_base |>
  filter(rand_site == "PCH") |>
  filter(substr(subjid, 4, 6) %in% str_pad(1:300, 3, pad = "0")) |>
  set_variable_labels(
    etomthr___1 = "Indigenous Australia",
    etomthr___2 = "European Caucasian",
    etomthr___3 = "Asian",
    etomthr___4 = "Indian subcontinent",
    etomthr___5 = "Middle Eastern",
    etomthr___6 = "Pacific Islander",
    etomthr___7 = "South American",
    etomthr___8 = "Black African",
    etofthr___1 = "Indigenous Australia",
    etofthr___2 = "European Caucasian",
    etofthr___3 = "Asian",
    etofthr___4 = "Indian subcontinent",
    etofthr___5 = "Middle Eastern",
    etofthr___6 = "Pacific Islander",
    etofthr___7 = "South American",
    etofthr___8 = "Black African",
    etoinf___1 = "Indigenous Australia",
    etoinf___2 = "European Caucasian",
    etoinf___3 = "Asian",
    etoinf___4 = "Indian subcontinent",
    etoinf___5 = "Middle Eastern",
    etoinf___6 = "Pacific Islander",
    etoinf___7 = "South American",
    etoinf___8 = "Black African",
  ) |>
  mutate(
    bcmoth = fct_lump_min(bcmoth, min = 5) |>
      fct_recode(Other = "OTHER") |>
      fct_infreq() |>
      fct_relevel("Other", after = Inf),
    bcfath = fct_lump_min(bcfath, min = 5) |>
      fct_recode(Other = "OTHER") |>
      fct_infreq() |>
      fct_relevel("Other", after = Inf)
  ) |>
  tbl_summary(
    include = c(
      # Demographics and infant history
      rand_stage,
      gender,
      v1_age_days,
      parinc,
      bcmoth,
      bcfath,
      etomthr___2,
      etomthr___3,
      etomthr___4,
      etomthr___7,
      etomthr___6,
      etomthr___8,
      etomthr___1,
      etofthr___2,
      etofthr___3,
      etofthr___4,
      etofthr___1,
      etofthr___8,
      etofthr___6,
      etofthr___7,
      etoinf___2,
      etoinf___3,
      etoinf___4,
      etoinf___7,
      etoinf___1,
      etoinf___8,
      etoinf___6,
      fborn,
      fecurr,
      fecat,
      fedog,
      # Birth history
      mgrav_fac,
      mpar_fac,
      delt,
      ipab,
      neoab,
      gestadelw,
      wgt,
      len,
      inhc,
      aps1_fac,
      aps5_fac,
      # Maternal immunisation
      inmpertvp,
      vacad,
      prevpert,
      minfvp,
      # Family history
      fha
    ),
    label = list(
      rand_stage ~ "Protocol stage",
      bcmoth ~ "Birth country of mother",
      bcfath ~ "Birth country of father",
      v1_age_days ~ "Age at enrolment (days)",
      fborn ~ "First born",
      mgrav_fac ~ "Maternal gravidity",
      mpar_fac ~ "Maternal parity",
      aps1_fac ~ "Apgar score (1 minute)",
      aps5_fac ~ "Apgar score (5 minutes)",
      prevpert ~ "Other maternal pertussis vaccination (last 5 years)",
      fha ~ "Family history of atopy",
      fecat ~ "Cat ownership",
      fedog ~ "Dog ownership",
      fecurr ~ "Feeding status"
    ),
    by = trt,
    missing_text = "(Missing)",
    digits = ~0
  ) |>
  add_n() |>
  add_overall(last = TRUE) |>
  add_variable_group_header(header = "Mother's ethnicity", variables = starts_with("etomthr")) |>
  add_variable_group_header(header = "Father's ethnicity", variables = starts_with("etofthr")) |>
  add_variable_group_header(header = "Infant's ethnicity", variables = starts_with("etoinf")) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      mutate(
        n = if_else(grepl("___", variable), NA_character_, n),
        n = if_else(row_type == "variable_group", '300', n)
      )
  )
tbl |>
  as_gt() |>
  cols_width(label ~ pct(30)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = (row_type == "label" | row_type == "variable_group") & !grepl("___", variable)
    )
  ) |>
  tab_options(
    table.width = pct(100),
    table.font.size = px(9),
    latex.use_longtable = TRUE
  ) |>
  tab_footnote(
    footnote = "Multiple may apply to each participant",
    locations = cells_body(rows = row_type == "variable_group", columns = label)
  ) |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```

```{=latex}
\clearpage
```

## Introduction of solid foods

For some foods, there are multiple sources of information from the food and household questionnaire. The survey first asked:"does you child currently eat egg?", "does your child currently eat/drink cow's milk?", and "does your child currently eat peanuts/other nuts?".

The survey subsequently asked for each of: rice cereal, cow's milk, fruit, vegetables, meat, fish, egg, and peanut, whether that food had been given for the first time since the last visit, and if so, the date it was first given and the quantity that was given.

### Egg, cow's milk, and nuts

On the food and household questionnaire, parents were asked whether their child *currently* eats egg, cow's milk, or peanuts/other nuts. @tbl-food-status-st1 and @tbl-food-status-st2 report on the responses to these questions by treatment group and the visit occasion (scheduled age) for stage 1 and stage 2 participants respectively. Ages are for the scheduled age of assessment, not necessarily the age at which that survey response was obtained. Proportions and denominators only include those participants who responded to the survey and so vary with the scheduled age of assessment.

```{r}
#| label: tbl-food-status-st1
#| tbl-cap: |
#|   Number and percentage of infants who consume egg, cow's milk, and peanut by treatment group and visit age, stage 1 participants.
#| tbl-pos: "H"
#| results: asis
dat_grid_fhq |>
  filter(rand_stage == 1, !(visit_age %in% c("6-week", "4-month"))) |>
  droplevels() |>
  tbl_strata2(
    strata = "trt",
    .tbl_fun = ~ .x |>
      tbl_summary(
        include = c(feeggyn, femilyn, fenutyn),
        label = list(
          feeggyn ~ "Does your child currently eat egg?",
          femilyn ~ "Does your child currently eat/drink cow's milk?",
          fenutyn ~ "Does your child currently eat peanuts/other nuts?"
        ),
        by = visit_age,
        type = list(feeggyn ~ "dichotomous"),
        value = list(feeggyn ~ "Yes"),
        statistic = all_dichotomous() ~ "{n}/{N} ({p})",
        missing = "no"
      ),
    .combine_with = "tbl_stack"
  ) |>
  modify_header(label = "**Treatment group and item**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(8)) |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```

```{r}
#| label: tbl-food-status-st2
#| tbl-cap: |
#|   Number and percentage of infants who consume egg, cow's milk, and peanut by treatment group and visit age, stage 2 participants.
#| tbl-pos: "H"
#| results: asis
dat_grid_fhq |>
  filter(rand_stage == 2, visit_age != "6-week") |>
  droplevels() |>
  tbl_strata2(
    strata = "trt",
    .tbl_fun = ~ .x |>
      tbl_summary(
        include = c(feeggyn, femilyn, fenutyn),
        label = list(
          feeggyn ~ "Does your child currently eat egg?",
          femilyn ~ "Does your child currently eat/drink cow's milk?",
          fenutyn ~ "Does your child currently eat peanuts/other nuts?"
        ),
        by = visit_age,
        type = list(feeggyn ~ "dichotomous"),
        value = list(feeggyn ~ "Yes"),
        statistic = all_dichotomous() ~ "{n}/{N} ({p})",
        missing = "no"
      ),
    .combine_with = "tbl_stack"
  ) |>
  modify_header(label = "**Treatment group and item**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(8)) |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```

### First time foods

Parent's were subsequently asked whether any of the specified foods had been introduced since the previous visit, and if so, when they were introduced and what quantity was given. In stage 1, rather than a date of introduction, the survey collected the *age* of introduction in days. In stage 2, only the *date* of introduction was elicited. For stage 1 responses, the reported age in days was used unless no record of introduction existed, in which case the response was treated as right-censored at the date of the last completed food and household questionnaire.

In some cases, the stage 2 date at which the food was introduced (`fefooddat`) contained clear errors (e.g. date was before date of birth or after the survey completion date). To calculate age (in days) at which the food was first introduce we first corrected any obvious errors in dates and then took the difference between the infants date of birth and the minimum of `fefooddat` (date food was first given) and `fedat` (date survey was completed). Again, not all infants had a record of food introduction. For infants with no responses of yes for the food in question the age of introduction was treated as right censored at the date of the last completed food and household questionnaire.

Some inconsistencies were noted between the responses to these items and those from the previous section. For example, where `femilyn` is "Yes" to all surveys from a certain date but `fefoodng1` (has cow's milk been introduced since last visit) is "No" or "N/A" on all questionnaires and `fefooddat2` (date cow's milk was introduced) is missing. The `fefooddat` data are reported as documented without cross-checking against other available fields given that it is only possible to cross-check for cow's milk, egg, and nuts. Missing dates were treated as right-censored which may not be accurate (e.g. food could have been introduced by xx-months of age but date of introduction was not documented on any questionnaire).

Kaplan-Meier curves for age at which a food was introduced for stage 1 participants stratified by the assigned treatment group are shown in @fig-kaplan-meier-food-intro-stage1. Curves for stage 2 participants are shown in @fig-kaplan-meier-food-intro-stage2.

```{r}
#| label: food-intro-timing-helpers
make_km_plot <- function(dat, time_var, event_var, trt_var, lab = "") {
  params <- list(
    time_var = substitute(time_var),
    event_var = substitute(event_var),
    trt_var = substitute(trt_var),
    dat = substitute(dat)
  )
  fit <- eval.parent(substitute(
    survfit2(
      Surv(
        time = time_var,
        event = event_var,
      ) ~ trt_var,
      data = dat
    ),
    params
  ))
  fit |>
    ggsurvfit() +
    add_confidence_interval() +
    add_censor_mark() +
    scale_colour_manual("Treatment group", values = c("black", "orange")) +
    scale_fill_manual("Treatment group", values = c("black", "orange")) +
    labs(title = lab, y = "Percentage yet to be introduced to food") +
    scale_ggsurvfit(
      x_scales = list(name = "Age (days)", breaks = seq(0, 630, 90), limits = c(0, 650))
    ) +
    guides(colour = guide_legend(override.aes = list(shape = NA))) +
    theme_get() +
    theme(legend.direction = "horizontal", title = element_text(size = 8))
}

plot_food_introduction_stage1 <- function(
  food_var = fefoodng2,
  age_var = fefoodag2,
  trt_var = trt
) {
  # If they have "Yes", we want that record
  # If they don't we want the last response
  food_obs <- dat_grid_fhq |>
    filter(rand_stage == 1) |>
    select(record_id, birthdat, {{ trt_var }}, fedat, visit_age, {{ food_var }}, {{ age_var }}) |>
    filter({{ food_var }} == "Yes") |>
    # Some have multiple, take the first
    filter(row_number() == 1, .by = record_id) |>
    mutate(fedat_age = time_length(interval(birthdat, fedat), "days")) |>
    mutate(y = pmin(fedat_age, {{ age_var }}), event = 1)
  food_cens <- dat_grid_fhq |>
    filter(rand_stage == 1, !(record_id %in% food_obs$record_id)) |>
    select(record_id, birthdat, {{ trt_var }}, visit_age, fedat, fecurr, {{ food_var }}, {{ age_var }}) |>
    filter(!is.na(fedat)) |>
    filter(row_number() == n(), .by = record_id) |>
    mutate(y = time_length(interval(birthdat, fedat), "days"), event = 0)

  food <- bind_rows(food_obs, food_cens)
  food
}

plot_food_introduction_stage2 <- function(
  food_var = fefoodng2,
  food_date = fefooddat2,
  trt_var = trt,
  food_lab = ""
) {
  food_obs <- dat_grid_fhq |>
    filter(rand_stage == 2) |>
    select(record_id, birthdat, {{ trt_var }}, fedat, visit_age, {{ food_var }}, {{ food_date }}) |>
    filter({{ food_var }} == "Yes") |>
    filter(row_number() == 1, .by = record_id) |>
    mutate(
      feage1 = time_length(interval(birthdat, {{ food_date }}), "days"),
      feage2 = time_length(interval(birthdat, fedat), "days"),
      y = pmin(feage1, feage2),
      event = 1
    )
  food_cens <- dat_grid_fhq |>
    filter(rand_stage == 2, !(record_id %in% food_obs$record_id)) |>
    select(record_id, birthdat, {{ trt_var }}, fedat, visit_age, {{ food_var }}, {{ food_date }}) |>
    filter(!is.na(fedat)) |>
    filter(row_number() == n(), .by = record_id) |>
    mutate(
      feage1 = time_length(interval(birthdat, {{ food_date }}), "days"),
      feage2 = time_length(interval(birthdat, fedat), "days"),
      y = feage2,
      event = 0
    )
  food <- bind_rows(food_obs, food_cens)
  food
}

# Corrections to some dates identified manually
dat_grid_fhq <- dat_grid_fhq |>
  mutate(
    fefooddat1 = case_when(
      record_id == "4629-223" & fefooddat1 == date("2023-01-23") ~ date("2024-01-23"),
      record_id == "4629-263" & fefooddat1 == date("2023-03-30") ~ date("2024-03-30"),
      record_id == "5785-133" & fefooddat1 == date("2022-02-01") ~ date("2023-02-01"),
      .default = fefooddat1
    ),
    fefooddat2 = case_when(
      record_id == "4629-230" & fefooddat2 == date("2023-01-04") ~ date("2024-01-04"),
      record_id == "5785-101" & fefooddat2 == date("2021-10-23") ~ date("2022-10-23"),
      .default = as_date(fefooddat2, format = "%Y-%m-%d")
    ),
    fefooddat2 = replace_values(fefooddat2, date("2010-08-01") ~ date("2020-08-01")),
    fefooddat4 = case_when(
      record_id == "5785-174" & fefooddat4 == date("2022-01-29") ~ date("2023-01-29"),
      .default = fefooddat4
    ),
    fefooddat6 = case_when(
      record_id == "5785-88" & fefooddat6 == ymd("2021-03-10") ~ ymd("2023-03-10"),
      record_id == "5785-126" & fefooddat6 == ymd("2022-02-10") ~ ymd("2023-02-10"),
      record_id == "5785-270" & fefooddat6 == ymd("2023-01-22") ~ ymd("2024-01-22"),
      .default = fefooddat6
    ),
    fefooddat8 = case_when(
      record_id == "4629-27" & fefooddat8 == ymd("2020-01-11") ~ ymd("2021-01-11"),
      record_id == "4629-141" & fefooddat8 == ymd("423-05-02") ~ ymd("2023-05-02"),
      record_id == "5785-124" & fefooddat8 == ymd("2022-03-30") ~ ymd("2023-03-30"),
      record_id == "5785-394" & fefooddat8 == ymd("2023-02-10") ~ ymd("2024-02-10"),
      .default = fefooddat8
    )
  )
```

```{r}
#| label: fig-kaplan-meier-food-intro-stage1
#| fig-cap: |
#|   Age of food introduction for stage 1 participants.
#|   Censored observations are marked by a cross.
#| fig-width: 5
#| fig-height: 7
# Stage 1
# dat_grid_fhq |>
#   filter(any(femilyn == "Yes"), !any(fefoodng2 == "Yes", na.rm = TRUE), .by = record_id) |>
#   select(record_id, visit_age, fedat, femilyn, feeggyn, fenutyn, starts_with("fefood")) |>
#   filter(record_id == "4629-7") |>
#   select(-record_id) |>
#   print(width = Inf)
# dat_grid_fhq |>
#   filter(any(femilyn == "Yes"), !any(fefoodng2 == "Yes", na.rm = TRUE), .by = record_id) |>
#   select(record_id, visit_age, fedat, femilyn, feeggyn, fenutyn, starts_with("fefood")) |>
#   filter(record_id == "4630-34") |>
#   select(-record_id) |>
#   print(width = Inf)
# dat_grid_fhq |>
#   filter(any(femilyn == "Yes"), !any(fefoodng2 == "Yes", na.rm = TRUE), .by = record_id) |>
#   select(record_id, visit_age, fedat, femilyn, feeggyn, fenutyn, starts_with("fefood")) |>
#   filter(record_id == "4629-42") |>
#   select(-record_id)

# dat_grid_fhq |>
#   filter(!any(femilyn == "Yes"), any(fefoodng2 == "Yes", na.rm = TRUE), .by = record_id) |>
#   select(record_id, visit_age, fedat, femilyn, feeggyn, fenutyn, starts_with("fefooddat")) |>
#   filter(record_id == "8644-17") |>
#   select(-record_id)

# dat_grid_fhq |>
#   filter(any(femilyn == "Yes"), !any(fefoodng2 == "Yes", na.rm = TRUE), .by = record_id) |>
#   select(record_id, visit_age, fedat, femilyn, feeggyn, fenutyn, starts_with("fefooddat")) |>
#   pull(record_id) |>
#   unique()

# dat_grid_fhq |>
#   filter(any(feeggyn == "Yes"), !any(fefoodng7 == "Yes", na.rm = TRUE), .by = record_id) |>
#   select(record_id, visit_age, fedat, femilyn, feeggyn, fenutyn, starts_with("fefooddat")) |>
#   filter(record_id == "8644-12") |>
#   select(-record_id)

p1 <- plot_food_introduction_stage1(fefoodng1, fefoodag1, trt)
p2 <- plot_food_introduction_stage1(fefoodng2, fefoodag2, trt)
p3 <- plot_food_introduction_stage1(fefoodng3, fefoodag3, trt)
p4 <- plot_food_introduction_stage1(fefoodng4, fefoodag4, trt)
p5 <- plot_food_introduction_stage1(fefoodng5, fefoodag5, trt)
p6 <- plot_food_introduction_stage1(fefoodng6, fefoodag6, trt)
p7 <- plot_food_introduction_stage1(fefoodng7, fefoodag7, trt)
p8 <- plot_food_introduction_stage1(fefoodng8, fefoodag8, trt)

des <- "
12
34
56
78
99
"
plot_stage1 <- (make_km_plot(p1, y, event, trt, "Rice cereal") +
  make_km_plot(p2, y, event, trt, "Cow's milk") +
  (make_km_plot(p3, y, event, trt, "Fruit") +
    theme(legend.position = "none")) +
  (make_km_plot(p4, y, event, trt, "Vegetables") +
    theme(legend.position = "none")) +
  (make_km_plot(p5, y, event, trt, "Meat") +
    theme(legend.position = "none")) +
  (make_km_plot(p6, y, event, trt, "Fish") +
    theme(legend.position = "none")) +
  (make_km_plot(p7, y, event, trt, "Egg") +
    theme(legend.position = "none")) +
  (make_km_plot(p8, y, event, trt, "Peanut") +
    theme(legend.position = "none")) +
  guide_area()) +
  plot_layout(
    design = des,
    heights = c(1, 1, 1, 1, 0.25),
    guides = "collect",
    axes = "collect_y"
  )

plot_stage1
```


```{r}
#| label: fig-kaplan-meier-food-intro-stage2
#| fig-cap: |
#|   Age of food introduction for stage 2 participants.
#|   Censored observations are marked by a cross.
#| fig-width: 5
#| fig-height: 7
# Stage 2
p1 <- plot_food_introduction_stage2(fefoodng1, fefooddat1, trt)
p2 <- plot_food_introduction_stage2(fefoodng2, fefooddat2, trt)
p3 <- plot_food_introduction_stage2(fefoodng3, fefooddat3, trt)
p4 <- plot_food_introduction_stage2(fefoodng4, fefooddat4, trt)
p5 <- plot_food_introduction_stage2(fefoodng5, fefooddat5, trt)
p6 <- plot_food_introduction_stage2(fefoodng6, fefooddat6, trt)
p7 <- plot_food_introduction_stage2(fefoodng7, fefooddat7, trt)
p8 <- plot_food_introduction_stage2(fefoodng8, fefooddat8, trt)

des <- "
12
34
56
78
99
"
plot_stage2 <- (make_km_plot(p1, y, event, trt, "Rice cereal") +
  make_km_plot(p2, y, event, trt, "Cow's milk") +
  (make_km_plot(p3, y, event, trt, "Fruit") +
    theme(legend.position = "none")) +
  (make_km_plot(p4, y, event, trt, "Vegetables") +
    theme(legend.position = "none")) +
  (make_km_plot(p5, y, event, trt, "Meat") +
    theme(legend.position = "none")) +
  (make_km_plot(p6, y, event, trt, "Fish") +
    theme(legend.position = "none")) +
  (make_km_plot(p7, y, event, trt, "Egg") +
    theme(legend.position = "none")) +
  (make_km_plot(p8, y, event, trt, "Peanut") +
    theme(legend.position = "none")) +
  guide_area()) +
  plot_layout(
    design = des,
    heights = c(1, 1, 1, 1, 0.25),
    guides = "collect",
    axes = "collect_y"
  )

plot_stage2
```



```{=latex}
\clearpage
```
