
# Clinical and mechanistic outcomes {#sec-clinical-mechanistic-outcomes}

Analyses of outcomes grouped participants according to their *assigned* treatment (@sec-treatment-admin).

All of the trial outcomes depend upon a positive SPT to either a specific allergen or to any allergen. Therefore, this section first reports on the SPT data, and outcomes 2c and 2d, followed by the primary (IgE-mediated food allergy) and other secondary clinical (eczema) outcomes.

## Skin prick tests

### Overview

```{r}
#| label: spt-no-records
std_allergens <- c(
  "D.pteronyssinus",
  "cat dander",
  "perennial ryegrass",
  "whole egg",
  "cashew",
  "cow's milk",
  "peanut",
  "seasme"
)

no_spt_record <- dat_spt |>
  filter(row_number() == 1, .by = record_id) |>
  filter(!priyn)
```

There were `r nrow(no_spt_record)` participants with no documented SPTs (@tbl-spt-listing-no-records). In many cases this was due to early withdrawal from the study, with the remaining due to withdrawn consent from the parent. All other participants had a record for at least the scheduled SPT, and in a few cases additional unscheduled SPT (@tbl-spt-number).

```{r}
#| label: tbl-spt-listing-no-records
#| tbl-cap: |
#|   Listing of infants with no SPT records.
#| tbl-pos: "H"
no_spt_record |>
  select(record_id, rand_site, streas, dis_age) |>
  gt() |>
  cols_label(
    rand_site = "Study site",
    dis_age = "Last contact age (months)"
  ) |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(90)
  )
```

```{r}
#| label: tbl-spt-number
#| tbl-cap: |
#|   Number and percentage of infants with scheduled and unscheduled SPTs.
#| tbl-pos: "H"
#| output: "asis"
spt_nums <- dat_spt |>
  filter(row_number() == 1, .by = c(record_id, trt, spt_occasion, spt_num)) |>
  summarise(
    spt_num_sch = sum(spt_occasion == "scheduled" & priyn, na.rm = TRUE),
    spt_num_unsch = sum(
      spt_occasion %in% c("unscheduled", "other") & priyn,
      na.rm = TRUE
    ),
    .by = c(record_id, trt)
  )
tbl_summary(
  spt_nums,
  by = trt,
  include = c(spt_num_sch, spt_num_unsch),
  label = c(spt_num_sch ~ "Scheduled SPT", spt_num_unsch ~ "Unscheduled SPTs")
) |>
  add_overall(last = TRUE) |>
  remove_footnote_header(columns = everything()) |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

For the outcomes, a requirement was positive SPT by 18 months of age. Age of SPT in months was derived from `birthdat` (date of birth) and `pridat` (date of skin prick test) or in some cases `univdat` (date of unscheduled visit) when `pridat` was missing. 

Most SPTs were performed by 13 months of age, although some were delayed, with a few occurring after 18-months of age (@fig-spt-age). There were 5 infants who did have an SPT, but none by 18-months of age (record ID: 3, 25, 34, 36, 5785-10) (@tbl-spt-listing-after-18months). 

Participants who were enrolled under early stage 1 protocol versions had their SPT scheduled for 18 months of age rather than 12 months of age. Delays meant some of these SPT's did not occur until 19 or 20 months of age. For the outcomes, "approximately 12 months of age" was taken to mean "by 18 months of age" which then currently excludes some of these later SPTs.

```{r}
#| label: fig-spt-age
#| fig-cap: |
#|   Age at skin prick test by protocol stage and occasion type (scheduled or unscheduled).
#|   Y-axis scale varies by panel.
#| fig-height: 4
dat_spt |>
  filter(row_number() == 1, .by = c(record_id, trt, spt_occasion, spt_num)) |>
  filter(!is.na(spt_age)) |>
  mutate(
    rand_stage = paste0("Stage ", rand_stage),
    spt_occasion = str_to_sentence(if_else(
      spt_occasion == "other",
      "unscheduled",
      spt_occasion
    ))
  ) |>
  ggplot(aes(spt_age)) +
  ggh4x::facet_grid2(
    spt_occasion ~ rand_stage,
    scales = "free_y",
    independent = "y"
  ) +
  geom_bar() +
  scale_x_continuous("Age (months)", breaks = seq(0, 30, 2)) +
  scale_y_continuous(
    "Number of skin prick tests",
    breaks = integer_breaks(10)
  )
```

```{r}
#| label: tbl-spt-listing-after-18months
#| tbl-cap: Listing of SPTs after 18-months of age.
#| tbl-pos: "H"
dat_spt |>
  filter(
    row_number() == 1,
    .by = c(record_id, trt, spt_occasion, spt_num, spt_age)
  ) |>
  filter(spt_age > 18) |>
  select(record_id, spt_occasion, spt_num, spt_age) |>
  mutate(
    spt_occasion = str_to_sentence(if_else(
      spt_occasion == "other",
      "unscheduled",
      spt_occasion
    ))
  ) |>
  distinct() |>
  gt() |>
  cols_label(
    spt_occasion ~ "SPT type",
    spt_num = "SPT occasion",
    spt_age = "Age (months)"
  ) |>
  tab_options(table.font.size = px(10), table.width = pct(50))
```

\clearpage

### Skin prick test sensitisation

@tbl-spt-positive reports on the number and percentage of infants who had a SPT result of \>1mm and ≥3mm relative to negative control for at least one tested allergen (outcomes 2c and 2d) and also the number with any SPT result >0mm. Any infants with no SPT records by 18 months of age are reported as missing.

```{r}
#| label: tbl-spt-positive
#| tbl-cap: |
#|   Number and percentage of infants with a positive SPT by 18 months of age.
#|   Infants with no SPT records counted as missing.
#| tbl-pos: "H"
#| output: asis
dat_spt |>
  summarise(
    across(
      spt_0mm:spt_3mm,
      ~ if_else(
        all(!priyn) | all(spt_age > 18),
        NA,
        any(.x == 1 & spt_age < 19, na.rm = TRUE)
      )
    ),
    .by = c(record_id, trt)
  ) |>
  tbl_summary(
    by = trt,
    include = c(spt_0mm, spt_1mm, spt_3mm),
    label = c(
      spt_0mm ~ "Any SPT >0mm",
      spt_1mm ~ "Any SPT >1mm vs. neg control",
      spt_3mm ~ "Any SPT ≥3mm vs neg. control"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(70)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

@tbl-spt-positive-tested provides a breakdown of the specific allergens from the standard panel which satisfied the sensitisation criteria. @tbl-spt-positive-tested-other reports on any other tested allergens beyond the standard panel.

Whole egg had a higher proportion of missingness than other allergens from the standard panel. This was usually because either egg white or egg yolk specifically was assessed and reported in the additional allergens instead of whole egg.

```{r}
#| label: spt-check
#| include: false
dat_spt |>
  select(
    record_id,
    trt,
    spt_occasion,
    spt_age,
    spt_tested,
    spt_result,
    spt_reaction,
    spt_0mm,
    spt_1mm,
    spt_3mm
  ) |>
  filter(spt_tested %in% std_allergens) |>
  filter(spt_occasion == "scheduled") |>
  summarise(
    N = n(),
    n = sum(!is.na(spt_1mm)),
    y = sum(spt_1mm, na.rm = TRUE),
    p = y / n,
    m = sum(is.na(spt_1mm)),
    .by = c(trt, spt_tested)
  ) |>
  pivot_wider(
    names_from = trt,
    values_from = c(N, n, y, p, m),
    names_vary = "slowest"
  ) |>
  gt(rowname_col = "spt_tested") |>
  tab_spanner(ends_with("aP"), label = "aP") |>
  tab_spanner(ends_with("wP"), label = "wP") |>
  cols_label(
    starts_with("n_", ignore.case = FALSE) ~ "Tests",
    starts_with("y_") ~ ">1mm",
    starts_with("m_") ~ "Missing"
  )
```

```{r}
#| label: tbl-spt-positive-tested
#| tbl-cap: |
#|   Number and percentage of infants with a positive SPT (scheduled only)
#|   by 18 months of age by tested allergen (for those included in the standard panel).
#|   Each cell gives positive/tested (%) [missing result%].
#| tbl-pos: "H"
#| output: asis
dat_spt |>
  filter(spt_tested %in% std_allergens) |>
  filter(spt_occasion == "scheduled") |>
  select(record_id, trt, spt_age, spt_tested, spt_0mm, spt_1mm, spt_3mm) |>
  droplevels() |>
  complete(nesting(select(dat_base, record_id, trt)), spt_tested) |>
  mutate(
    across(spt_0mm:spt_3mm, ~ .x * (spt_age <= 18))
  ) |>
  pivot_longer(
    spt_0mm:spt_3mm,
    names_pattern = "spt_([0-3]+mm)",
    names_to = "size",
    values_to = "diameter"
  ) |>
  mutate(spt_tested = str_to_sentence(spt_tested)) |>
  pivot_wider(names_from = spt_tested, values_from = diameter) |>
  mutate(
    size = factor(
      size,
      levels = c("0mm", "1mm", "3mm"),
      labels = c(
        "Wheal size >0mm",
        "Wheal size >1mm vs. neg control",
        "Wheal size ≥3mm vs neg. control"
      )
    )
  ) |>
  tbl_strata(
    strata = size,
    .tbl_fun = ~ .x |>
      tbl_summary(
        include = `D.pteronyssinus`:Peanut,
        by = trt,
        statistic = all_dichotomous() ~ "{n}/{N_nonmiss} ({p}%) [{p_miss}%]",
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Allergen**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_column_alignment(starts_with("stat"), align = "right"),
    .combine_with = "tbl_stack"
  ) |>
  as_gt() |>
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_body(columns = label, rows = label == "D.pteronyssinus")
  ) |>
  tab_options(table.font.size = px(10), table.width = pct(80)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-spt-positive-tested-other
#| tbl-cap: |
#|   Number and percentage of infants with a positive SPT (scheduled only)
#|   by 18 months of age by tested allergen for additional allergens which were tested.
#|   Each cell gives positive/tested (%). Totals restricted to only those infants with at least one
#|   additional allergen tested
#| tbl-pos: "H"
#| output: asis
dat_spt |>
  unique() |>
  filter(
    !(spt_tested %in% std_allergens),
    !is.na(spt_tested),
    spt_reaction != "Not Done"
  ) |>
  filter(spt_occasion == "scheduled") |>
  select(record_id, trt, spt_age, spt_tested, spt_0mm, spt_1mm, spt_3mm) |>
  droplevels() |>
  # complete(nesting(select(dat_base, record_id, trt)), spt_tested) |>
  mutate(
    across(spt_0mm:spt_3mm, ~ .x * (spt_age <= 18))
  ) |>
  pivot_longer(
    spt_0mm:spt_3mm,
    names_pattern = "spt_([0-3]+mm)",
    names_to = "size",
    values_to = "diameter"
  ) |>
  mutate(
    diameter = as.logical(diameter),
    spt_tested = fct_infreq(fct_relevel(spt_tested, sort))
  ) |>
  arrange(spt_tested) |>
  mutate(spt_tested = str_to_sentence(spt_tested)) |>
  pivot_wider(names_from = spt_tested, values_from = diameter) |>
  mutate(
    size = factor(
      size,
      levels = c("0mm", "1mm", "3mm"),
      labels = c(
        "Wheal size >0mm",
        "Wheal size >1mm vs. neg control",
        "Wheal size ≥3mm vs neg. control"
      )
    )
  ) |>
  tbl_strata(
    strata = size,
    .tbl_fun = ~ .x |>
      tbl_summary(
        include = `Sesame`:`Whole wheat`,
        by = trt,
        type = list(everything() ~ "dichotomous"),
        value = list(everything() ~ TRUE),
        statistic = all_dichotomous() ~ "{n}/{N_nonmiss} ({p}%)",
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Allergen**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_column_alignment(starts_with("stat"), align = "right"),
    .combine_with = "tbl_stack"
  ) |>
  as_gt() |>
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_body(columns = label, rows = label == "D. farinae")
  ) |>
  tab_options(
    table.font.size = px(10),
    table.width = pct(50),
    latex.use_longtable = TRUE
  ) |>
  tab_footnote(
    footnote = "As documented in free-text field",
    location = cells_column_labels(columns = label)
  ) |>
  as_latex() |>
  latex_unescape() |>
  gsub("NA", "-", x = _) |>
  cat()
```


\clearpage

### Skin prick test sensitisation model

```{r}
#| label: spt-analysis
set.seed(54123)

dat_outcome <- dat_spt_analysis |>
  pivot_longer(
    spt_0mm:spt_3mm,
    names_pattern = "spt_([0-3]mm)",
    names_to = "size",
    values_to = "outcome"
  ) |>
  filter(!is.na(outcome)) |>
  summarise(
    n = n(),
    y = sum(outcome),
    p = y / n,
    a = 1 + y,
    b = 1 + n - y,
    post = dist_beta(a, b),
    .by = c(trt, size)
  ) |>
  mutate(p_sup = c(NA, beta_ineq_sim(a, b)), .by = size)
diff_outcome <- dat_outcome |>
  mutate(post_drws = generate(post, 1e6)) |>
  select(size, trt, post_drws) |>
  unnest(post_drws) |>
  group_by(size, trt) |>
  mutate(.draw = 1:n()) |>
  ungroup() |>
  pivot_wider(names_from = "trt", values_from = "post_drws") |>
  mutate(delta = wP - aP)

diff_2c <- filter(diff_outcome, size == "1mm")
diff_2d <- filter(diff_outcome, size == "3mm")

prior <- c(
  prior(logistic(0, 1), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = "b")
)

fit_outcome <- dat_spt_analysis |>
  pivot_longer(
    spt_0mm:spt_3mm,
    names_pattern = "spt_([0-3]mm)",
    names_to = "size",
    values_to = "outcome"
  ) |>
  nest(-size) |>
  mutate(
    fit = map(
      data,
      \(x) {
        brm(
          outcome ~
            trt + rand_site + gender + bfed + fborn + parinc_imp + ces + fha,
          data = x,
          family = bernoulli(),
          prior = prior,
          warmup = 500,
          iter = 5500,
          refresh = 0,
          silent = 2,
          seed = 54123
        )
      }
    ),
    rd = map2(data, fit, \(x, y) {
      wgt <- c(rdo(rdirichlet(1, rep(1, nrow(x))), ndraws = ndraws(y)))
      x |>
        mutate(trt = "aP") |>
        add_epred_rvars(y, value = ".epred_aP") |>
        mutate(trt = "wP") |>
        add_epred_rvars(y, value = ".epred_wP") |>
        mutate(wgt = wgt) |>
        summarise(
          aP = rvar_mean(.epred_aP),
          wP = rvar_mean(.epred_wP),
          .rd_empirical = rvar_mean(.epred_wP - .epred_aP),
          .rd_bootstrap = rvar_sum((.epred_wP - .epred_aP) * wgt)
        )
    })
  )
```

@tbl-posterior-spt summarises the observed counts and estimated model parameters for skin prick test sensitsation under each treatment group and by the size requirement (see @sec-secondary-outcomes numbers 2c and 2d), and their estimated differences (the posterior densities are shown in @fig-posterior-difference-spt-unadjusted-adjusted). The results between the analyses (unadjusted and adjusted, @sec-primary-analysis) were near identical. The most plausible values for the difference in skin prick test sensitisation at >1mm [2.c] between wP and aP were `r do.call(sprintf, c(fmt = "%.2f (%.2f, %.2f)", as.list(c(quantile2(diff_2c$delta, c(0.5, 0.025, 0.975))))))` [posterior median, (95% credible interval)]. The probability of a reduction in sensitisation under wP compared to aP was `r sprintf("%.2f", mean(diff_2c$delta < 0))`. At ≥3mm [2.d] they were `r do.call(sprintf, c(fmt = "%.2f (%.2f, %.2f)", as.list(c(quantile2(diff_2d$delta, c(0.5, 0.025, 0.975))))))` and `r sprintf("%.2f", mean(diff_2d$delta < 0))`.

```{r}
#| label: tbl-posterior-spt
#| tbl-cap: |
#|   Posterior summaries for risk and difference in risk of positive skin prick test by size.
#| tbl-pos: "H"
diff_agg <- diff_outcome |>
  group_by(size) |>
  nest_rvars() |>
  ungroup() |>
  mutate(
    model = "Unadjusted",
    across(
      aP:delta,
      list(
        med = median,
        lo = ~ quantile2(.x, 0.025),
        hi = ~ quantile2(.x, 0.975),
        pr = ~ Pr(.x < 0)
      ),
      .names = "{.fn}_{.col}"
    )
  ) |>
  select(-aP, -wP, -delta) |>
  bind_rows(
    fit_outcome |>
      unnest(rd) |>
      rename(delta = .rd_empirical) |>
      select(size, aP, wP, delta) |>
      mutate(model = "Adjusted") |>
      mutate(across(
        aP:delta,
        list(
          med = median,
          lo = ~ quantile2(.x, 0.025),
          hi = ~ quantile2(.x, 0.975),
          pr = ~ Pr(.x < 0)
        ),
        .names = "{.fn}_{.col}"
      )) |>
      select(-aP, -wP, -delta)
  ) |>
  select(-pr_aP, -pr_wP)
dat_outcome |>
  select(size, trt, n, y, p) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p)
  ) |>
  left_join(diff_agg, join_by(size)) |>
  mutate(
    size = factor(
      size,
      levels = c("0mm", "1mm", "3mm"),
      labels = c(">0mm", ">1mm vs ctrl. [2.c]", "≥3mm vs ctrl. [2.d]")
    )
  ) |>
  gt(groupname_col = "model", rowname_col = "size") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_align(align = "left", columns = 1) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_aP, lo_aP, hi_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_wP, lo_wP, hi_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "aP"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "wP"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_aP, med_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  tab_options(table.font.size = px(9))
```

```{r}
#| label: fig-posterior-density-spt
#| fig-cap: |
#|    Posterior density for difference in risk of skin prick test sensitisation by size requirement,
#|    unadjusted primary model.
#| fig-height: 3
#| fig-width: 5
#| include: false
p1 <- diff_outcome |>
  select(size, delta) |>
  group_by(size) |>
  nest_rvars() |>
  mutate(
    size = factor(
      size,
      levels = c("0mm", "1mm", "3mm"),
      labels = c(">0mm", ">1mm vs ctrl. [2.c]", "≥3mm vs ctrl. [2.d]")
    )
  ) |>
  ggplot(aes(xdist = delta)) +
  facet_wrap(size ~ ., ncol = 1) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    point_size = 1.5,
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = seq(-0.1, 0.1, 0.02), limits = c(-0.08, 0.08)) +
  scale_y_continuous(limits = ~ range(.x, -0.05)) +
  labs(
    fill = "Interval",
    x = "Difference in skin prick test sensitisation, wP - aP",
    y = "Density"
  ) +
  guides(fill = guide_legend(override.aes = list(point_size = NA)))
p1
```

```{r}
#| label: fig-posterior-difference-spt-adjusted
#| fig-cap: |
#|   Posterior density for difference in risk of skin prick test sensitisation by size requirement, adjusted model.
#| fig-height: 3
#| fig-width: 5
#| include: false
p2 <- fit_outcome |>
  unnest(rd) |>
  select(size, .rd_bootstrap) |>
  mutate(
    size = factor(
      size,
      levels = c("0mm", "1mm", "3mm"),
      labels = c(">0mm", ">1mm vs ctrl. [2.c]", "≥3mm vs ctrl. [2.d]")
    )
  ) |>
  ggplot(aes(xdist = .rd_bootstrap)) +
  facet_wrap(~size, ncol = 1) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    point_size = 1.5,
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = seq(-0.1, 0.1, 0.02), limits = c(-0.08, 0.08)) +
  scale_y_continuous(limits = ~ range(.x, -0.05)) +
  labs(
    fill = "Interval",
    x = "Difference in skin prick test sensitisation, wP - aP",
    y = "Density"
  ) +
  guides(fill = guide_legend(override.aes = list(point_size = NA)))
p2
```

```{r}
#| label: fig-posterior-difference-spt-unadjusted-adjusted
#| fig-cap: |
#|   Posterior density for difference in risk of skin prick test sensitisation by size and model.
#| fig-height: 3.25
#| fig-width: 6
(p1 + labs(title = "Unadjusted") | p2 + labs(title = "Adjusted")) +
  plot_layout(axes = "collect", axis_titles = "collect", guides = "collect") &
  theme(
    legend.position = "bottom",
    legend.margin = margin(t = 0, unit = "cm"),
    legend.key.size = unit(0.9, "line"),
    plot.title = element_text(size = rel(1)),
    text = element_text(size = 9)
  )
```

\clearpage

### Skin prick test magnitude

@tbl-spt-magnitude and @fig-spt-magnitude report on the magnitude of the SPT results (in millimetres) from scheduled SPTs, in particular, amongst infants with SPT sensitisation. There was little evidence of a difference in sensitisation and, given the small number of infants with any sensitisation, in difference in SPT magnitudes, whether overall or conditional on sensitisation.

```{r}
#| label: tbl-spt-magnitude
#| tbl-cap: SPT sensitisation and magnitude by assigned treatment for scheduled SPTs.
#| tbl-pos: "H"
dat_1 <- dat_spt |>
  select(subjid, trt, priyn, spt_num, spt_occasion, pridat, spt_age, spt_tested, spt_neg, spt_pos, spt_result) |>
  filter(spt_age < 19, spt_occasion == "scheduled", spt_tested %in% std_allergens) |>
  mutate(y = spt_result) |>
  filter(!is.na(y))

dat_1 |>
  mutate(spt_tested = factor(str_to_sentence(spt_tested), levels = str_to_sentence(std_allergens))) |>
  summarise(
    n = sum(!is.na(y)),
    ngt0 = sum(y > 0, na.rm = TRUE),
    pgt0 = mean(y > 0, na.rm = TRUE),
    m = mean(y, na.rm = TRUE),
    mgt0 = mean(y[y > 0], na.rm = TRUE),
    .by = c(trt, spt_tested)
  ) |>
  pivot_wider(names_from = trt, values_from = n:mgt0) |>
  gt(rowname_col = "spt_tested") |>
  tab_spanner(columns = ends_with("aP"), label = "aP") |>
  tab_spanner(columns = ends_with("wP"), label = "wP") |>
  fmt_percent(columns = contains("pgt0"), decimals = 1) |>
  fmt_number(columns = contains("m_"), decimals = 2) |>
  fmt_number(columns = contains("mgt0"), decimals = 1) |>
  cols_merge(columns = c(n_aP, ngt0_aP, pgt0_aP), pattern = "{2}/{1} ({3})") |>
  cols_merge(columns = c(n_wP, ngt0_wP, pgt0_wP), pattern = "{2}/{1} ({3})") |>
  cols_label(
    n_aP ~ "y/n (p%)",
    m_aP ~ "E[y]",
    mgt0_aP ~ "E[y|y>0]",
    n_wP ~ "y/n (p%)",
    m_wP ~ "E[y]",
    mgt0_wP ~ "E[y|y>0]"
  ) |>
  cols_align(columns = spt_tested, align = "left") |>
  tab_footnote(
    footnote = "y: number with any sensitisation (>0mm), n: number of scheduled SPT results, p: percentage",
    locations = cells_column_labels(starts_with("n_"))
  ) |>
  tab_footnote(
    footnote = "E[y]: mean of SPT results (mm) amongst all infants",
    locations = cells_column_labels(starts_with("m_"))
  ) |>
  tab_footnote(
    footnote = "E[y|y>0]: mean of SPT results (mm) amongst infants with any sensitisation",
    locations = cells_column_labels(starts_with("mgt0"))
  ) |>
  tab_options(table.font.size = px(9))
```

```{r}
#| label: fig-spt-magnitude
#| fig-cap: |
#|   SPT magnitude amongst infants with any sensitisation (>0mm).
#| fig-width: 6
#| fig-height: 4
dat_1 |>
  filter(spt_result > 0) |>
  mutate(spt_tested = factor(str_to_sentence(spt_tested), levels = str_to_sentence(std_allergens))) |>
  arrange(spt_tested) |>
  ggplot(aes(spt_result)) +
  facet_wrap(~spt_tested) +
  stat_ecdf(aes(colour = trt)) +
  labs(colour = "Treatment group", x = "SPT result (mm)", y = "eCDF") +
  scale_colour_manual(values = c("black", "orange")) +
  theme(legend.position = "top")
```

As a post-hoc exploratory analysis of SPT results, hurdle log-normal models were estimated for each allergen on the standard panel for scheduled SPTs. These models conditioned on assigned treatment only. The estimate differences in mean SPT result, difference in probability of sensitisation, and difference in mean SPT result conditional on sensitisation are reported in @fig-spt-magnitude-estimates.

```{r}
#| label: spt-magnitude-models
dat_1_fits <- dat_1 |>
  nest(dat = -spt_tested) |>
  mutate(
    fit = map(
      dat,
      ~ brm(
        bf(y ~ trt, hu ~ trt),
        data = .x,
        family = hurdle_lognormal(),
        backend = "cmdstanr",
        seed = 71235,
        cores = 8,
        chains = 8,
        refresh = 0,
        silent = 2
      )
    ),
    params = map(
      fit,
      ~ do.call(c, as_draws_rvars(.x))
    )
  )
```

```{r}
#| label: fig-spt-magnitude-estimates
#| fig-cap: | 
#|   Posterior summaries for model estimated difference in mean SPT result, sensitisation probability, and mean SPT result conditional on sensitisation, wP vs aP.
#| fig-width: 6
#| fig-height: 3
dat_1_fits |>
  unnest_wider(params) |>
  select(spt_tested, starts_with("b_"), sigma) |>
  mutate(
    diff_E = (1 - rfun(plogis)(b_hu_Intercept + b_hu_trtwP)) *
      exp(b_Intercept + b_trtwP + sigma^2 / 2) -
      (1 - rfun(plogis)(b_hu_Intercept)) * exp(b_Intercept + sigma^2 / 2),
    diff_Pr_gt0 = 100 * ((1 - rfun(plogis)(b_hu_Intercept + b_hu_trtwP)) - (1 - rfun(plogis)(b_hu_Intercept))),
    diff_E_gt0 = exp(b_Intercept + b_trtwP + sigma^2 / 2) - exp(b_Intercept + sigma^2 / 2)
  ) |>
  select(spt_tested, starts_with("diff")) |>
  pivot_longer(starts_with("diff")) |>
  mutate(
    spt_tested = fct_inorder(str_to_sentence(spt_tested)),
    name = factor(
      name,
      levels = c("diff_E", "diff_Pr_gt0", "diff_E_gt0"),
      labels = c("Difference in E[y]", "Difference in Pr(y > 0)", "Difference in E[y|y > 0]")
    )
  ) |>
  ggplot(aes(y = spt_tested, xdist = value)) +
  facet_wrap(~name, scales = "free_x") +
  stat_pointinterval(point_size = 1.5, .width = c(0.8, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_y_discrete("Allergen tested", limits = rev) +
  labs(x = "Effect size")
```

\clearpage

## IgE-mediated food allergy

### Overview

The primary source of information for IgE-mediated food allergy was the `outcome_report` form. Reported outcomes were cross checked against the food and household questionnaire form (`food_and_household_questionnaire`) and the oral food challenge form (`food_challenge`).

Age at first reaction was derived from `outageval`. If this was missing then `birthdat` (date of birth) and `outdiagdat` (clinician-diagnosis date) was used to derive age in months.

In some cases participants may have withdrawn from the study without sufficient follow-up for IgE-mediated food allergy to have been identified. The report makes no allowance for "missing" IgE-mediated food allergy status, with absence of clinician diagnosed allergy (with or without adequate follow-up) counted as not meeting the primary outcome. Only 17 participants had an age of last contact 15 months or younger and may have been those most likely to have insufficient follow-up.

The following sections provide a summary of the number of infants who met each criteria. 
The infants with an outcome reported are line listed in the appendix (@sec-primary-outcome-listing) along with the reported allergens.

### Oral food challenge

There were `r nrow(filter(dat_fa, outalltp == "Oral food challenge"))` outcome records for food allergy diagnoses resulting from oral food challenge amongst `r length(unique(filter(dat_fa, outalltp == "Oral food challenge")$record_id))` infants. One reaction (record ID: `8644-34`) reportedly first occurred after 18 months of age and all others were by 18 months of age. All records had a matching positive skin prick test to the reported allergen and so 22 infants met the primary outcome according to the definition (1.i).

```{r}
#| label: check-positive-spt-ofc
#| include: false
# Total outcomes reported
dat_fa |>
  filter(
    outalltp == "Oral food challenge"
  )
# Invalid age
dat_fa |>
  filter(
    outalltp == "Oral food challenge",
    outageval_months2 > 18
  )

dat_ofc_spt <- dat_fa |>
  filter(
    outalltp == "Oral food challenge",
    outageval_months2 <= 18,
    !is.na(outdiagdat)
  ) |>
  select(
    record_id,
    trt,
    outdiagdat,
    outageval_months2,
    out_age_months,
    outfrafood
  ) |>
  left_join(
    dat_spt_pos |>
      filter(any_spt_pos_1mm) |>
      select(record_id, spt_age, spt_pos_1mm),
    join_by(record_id),
    relationship = "many-to-many"
  ) |>
  rowwise() |>
  mutate(
    splt_foods = gsub("( |, )", "|", spt_pos_1mm),
    match_food = str_match(tolower(outfrafood), splt_foods)[1]
  ) |>
  ungroup()

dat_po_ofc <- dat_ofc_spt |>
  summarise(
    outfrafood = paste(unique(tolower(outfrafood)), collapse = ", "),
    any_food_match = any(!is.na(match_food) & outageval_months2 <= 18),
    matched_foods = paste(
      unique(match_food[!is.na(match_food) & outageval_months2 <= 18]),
      collapse = ", "
    ),
    .by = c(record_id, trt)
  ) |>
  rename(
    po_ofc = any_food_match,
    po_ofc_foods = outfrafood,
    po_ofc_match = matched_foods
  ) |>
  # Manual fixes for reporting purposes
  mutate(
    po_ofc_match = gsub("egg", "whole egg", po_ofc_match),
    po_ofc_match = gsub("milk", "cow's milk", po_ofc_match),
    po_ofc_match = replace_when(
      po_ofc_match,
      po_ofc_match == "whole" ~ "whole egg",
    )
  )

# Number who met this condition
dat_po_ofc |> count(po_ofc)
```

### Anaphylaxis

There were `r nrow(filter(dat_fa, outalltp == "Anaphylaxis"))` outcome records for food allergy diagnoses resulting from anaphylaxis amongst `r length(unique(filter(dat_fa, outalltp == "Anaphylaxis")$record_id))` infants. Two of these infants (record ID: `68, 5785-283`) were diagnosed after 18 months of age with first reported reaction occuring after 18 months of age. One infant (record ID: `8644-34`) had no skin prick test for the reported allergens. So of the `r nrow(filter(dat_fa, outalltp == "Anaphylaxis"))` infants with analphylaxis reported, 3 did not meet the criteria for the primary outcome according to the definition (1.ii) leaving `r nrow(filter(dat_fa, outalltp == "Anaphylaxis")) - 3` who met the condition.

```{r}
#| label: check-positive-spt-ana
#| include: false
dat_ana_spt <- dat_fa |>
  filter(
    outalltp == "Anaphylaxis",
    outageval_months2 <= 18,
    !is.na(outdiagdat)
  ) |>
  select(
    record_id,
    trt,
    outdiagdat,
    outageval_months2,
    out_age_months,
    outfrafood
  ) |>
  left_join(
    dat_spt_pos |>
      filter(any_spt_pos_1mm) |>
      select(record_id, spt_age, spt_pos_1mm),
    join_by(record_id),
    relationship = "many-to-many"
  )

dat_po_ana <- dat_ana_spt |>
  rowwise() |>
  mutate(
    splt_foods = gsub("( |, )", "|", spt_pos_1mm),
    match_food = str_match(tolower(outfrafood), splt_foods)[1]
  ) |>
  ungroup() |>
  summarise(
    outfrafood = paste(unique(tolower(outfrafood)), collapse = ", "),
    any_food_match = any(!is.na(match_food) & outageval_months2 <= 18),
    matched_foods = paste(
      unique(match_food[!is.na(match_food) & outageval_months2 <= 18]),
      collapse = ", "
    ),
    .by = c(record_id, trt)
  ) |>
  rename(
    po_ana = any_food_match,
    po_ana_foods = outfrafood,
    po_ana_match = matched_foods
  ) |>
  # Manual fixes for reporting purposes
  mutate(
    po_ana_match = replace_when(
      po_ana_match,
      po_ana_match == "milk" ~ "cow's milk",
      po_ana_match == "egg" ~ "whole egg"
    )
  )
```

### Highly probable IgE-mediated food allergy

There were `r nrow(filter(dat_fa, outalltp == "Highly probably IgE mediated food allergy"))` outcome records for food allergy diagnoses resulting from highly probable IgE-mediated allergy amongst `r length(unique(filter(dat_fa, outalltp == "Highly probably IgE mediated food allergy")$record_id))` infants. For one infant (record ID: `68`), the age at first reaction and age at diagnosis were after 18 months of age. Four infants (record ID: `75, 5785-239, 5785-243, 5785-331`) had no positive SPT to the reported allergen. One infant (record ID: `5785-7`) had no date of allergy confirmation of clinician and did not meet the outcome due to a negative food challenge. Of the `r length(unique(filter(dat_fa, outalltp == "Highly probably IgE mediated food allergy")$record_id))` infants with IgE-mediated food allergy reported, 6 did not meet the primary outcome according to the definition (2.) leaving 38 infants who met the condition.

```{r}
#| label: check-positive-spt-ige
#| include: false
# Cross check food challenge data
summ_fc <- dat_fc |>
  filter(ofcresult != "Not challenged") |>
  summarise(
    ofcfood = paste(ofcfood, collapse = ", "),
    ofcresult = paste(ofcresult, collapse = ", "),
    ofcage = paste(ofcage, collapse = ", "),
    .by = record_id
  )

# Check reasons we dont count them
dat_fa |>
  filter(
    outalltp == "Highly probably IgE mediated food allergy",
    outageval_months2 > 18
  )
dat_fa |>
  filter(
    outalltp == "Highly probably IgE mediated food allergy",
    is.na(outdiagdat)
  )

dat_ige_spt <- dat_fa |>
  filter(
    outalltp == "Highly probably IgE mediated food allergy",
    outageval_months2 <= 18,
    !is.na(outdiagdat)
  ) |>
  select(
    record_id,
    trt,
    outdiagdat,
    outageval_months2,
    out_age_months,
    outfrafood
  ) |>
  left_join(
    dat_spt_pos |>
      filter(any_spt_pos_1mm) |>
      select(record_id, spt_age, spt_pos_1mm),
    join_by(record_id),
    relationship = "many-to-many"
  )

dat_po_ige_1 <- dat_ige_spt |>
  rowwise() |>
  mutate(
    splt_foods = gsub("( |, )", "|", spt_pos_1mm),
    splt_foods = replace_na(splt_foods, "^$"),
    outfrafood_sub = gsub(
      "Whole Egg",
      "egg",
      gsub("cows ", "", gsub("Dairy", "milk", outfrafood))
    ),
    match_food = gsub(
      "NA",
      "",
      paste(
        unique(str_match_all(tolower(outfrafood_sub), splt_foods)[[1]]),
        collapse = ", "
      )
    )
  ) |>
  ungroup()

dat_po_ige <- dat_po_ige_1 |>
  summarise(
    outfrafood = paste(unique(tolower(outfrafood)), collapse = ", "),
    any_food_match = any(match_food != "" & outageval_months2 <= 18),
    matched_foods = gsub(
      "^, ",
      "",
      gsub(
        "(, )+",
        ", ",
        paste(
          unique(match_food[!is.na(match_food) & outageval_months2 <= 18]),
          collapse = ", "
        )
      )
    ),
    .by = c(record_id, trt)
  ) |>
  rename(
    po_ige = any_food_match,
    po_ige_foods = outfrafood,
    po_ige_match = matched_foods
  ) |>
  # Manual fixes for reporting purposes
  mutate(
    po_ige_match = gsub("egg", "whole egg", po_ige_match),
    po_ige_match = gsub("milk", "cow's milk", po_ige_match),
    po_ige_match = replace_when(
      po_ige_match,
      record_id == "5785-242" ~ "whole egg",
      record_id == "4629-226" ~ "peanut, egg yolk, egg white",
      record_id == "4630-7" ~ "cow's milk, egg yolk, egg white, soy, beef",
      record_id == "4630-17" ~ "egg yolk, egg white"
    )
  )
```

### Primary outcome (IgE-mediated food allergy)

```{r}
#| label: primary-outcome-data
dat_po <- dat_base |>
  left_join(select(dat_po_ofc, -trt), join_by(record_id)) |>
  left_join(select(dat_po_ana, -trt), join_by(record_id)) |>
  left_join(select(dat_po_ige, -trt), join_by(record_id)) |>
  mutate(
    po_ofc = replace_na(po_ofc, 0),
    po_ana = replace_na(po_ana, 0),
    po_ige = replace_na(po_ige, 0),
    po = po_ofc | po_ana | po_ige,
    po_criteria = case_when(
      po_ofc == 1 ~ "Criteria 1.i",
      po_ana == 1 ~ "Criteria 1.ii",
      po_ige == 1 ~ "Criteria 2.",
      .default = "None"
    )
  ) |>
  select(
    record_id,
    rand_site,
    trt,
    birthdat,
    bcmoth,
    bcfath,
    gender,
    parinc_imp,
    delt,
    fecurr,
    fborn,
    bfed,
    ces,
    fhaast,
    fhaecz,
    fhaar,
    fhafa,
    fha,
    starts_with("po")
  )

# No positive SPT
dat_po <- dat_po |>
  # rowwise() |>
  mutate(
    po_ana_match = na_if(po_ana_match, ""),
    po_foods = pmap_chr(across(c(po_ofc_match, po_ana_match, po_ige_match)), function(...) {
      c(...) %>%
        str_split(", ") %>% # split each column into items
        unlist() %>%
        unique() %>% # drop duplicates
        sort() %>% # alphabetical order
        paste(collapse = ", ")
    }),
    po_foods = na_if(po_foods, ""),
    po_foods_split = po_foods
  ) |>
  separate_longer_delim(po_foods_split, delim = ", ") |>
  mutate(value = 1) |>
  pivot_wider(
    names_from = po_foods_split,
    values_from = value,
    values_fill = 0L
  ) |>
  rename(None = `NA`)
```

Overall `r sprintf("%i/%i (%.0f%%)", sum(dat_po$po), nrow(dat_po), 100 * mean(dat_po$po))` infants met the primary outcome, `r with(dat_po, sprintf("%i/%i (%.0f%%)", sum(po[trt == "aP"]), sum(trt=="aP"), 100 * mean(po[trt == "aP"])))` who were assigned to aP and `r with(dat_po, sprintf("%i/%i (%.0f%%)", sum(po[trt == "wP"]), sum(trt=="wP"), 100 * mean(po[trt == "wP"])))` who were assigned to wP.

@tbl-primary-outcome-criteria reports on the number of participants who met each of the primary outcome criteria (1.i, 1.ii, and 2, see @sec-primary-outcome) as outlined in the sections above by their assigned treatment group.

@tbl-primary-outcome-overall reports the total number of infants who met the primary outcome, and the primary reason grouped hierarchically according to the ordering in the definition (i.e. OFC → Anaphylaxis → IgE-mediated) (irrespective of the timing of those criteria).

@tbl-primary-outcome-foods reports the number of infants who met the primary outcome by the food allergy. Each infant may meet the primary outcome for more than one food.

```{r}
#| label: tbl-primary-outcome-criteria
#| tbl-cap: |
#|   Number and percentage of infants who met each primary outcome condition and treatment group.
#| tbl-pos: "H"
#| output: asis
dat_po |>
  tbl_summary(
    by = trt,
    include = c(po_ofc, po_ana, po_ige),
    label = list(
      po_ofc ~ "Met condition 1.i.",
      po_ana ~ "Met condition 1.ii.",
      po_ige ~ "Met condition 2."
    ),
    missing = "no"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(70)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-primary-outcome-overall
#| tbl-cap: |
#|   Number and percentage of infants who met the primary outcome by primary reason and treatment group.
#| tbl-pos: "H"
#| output: asis
dat_po |>
  tbl_summary(
    by = trt,
    include = c(po, po_criteria),
    label = list(
      po ~ "Met primary outcome",
      po_criteria ~ "Reason"
    ),
    missing = "no"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      filter(label != "None")
  ) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(70)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-primary-outcome-foods
#| tbl-cap: |
#|   Number and percentage of infants who met the primary outcome by the reported food.
#|   The same infant may meet the primary outcome for more than one allergen.
#| tbl-pos: "H"
#| output: asis
tbl <- dat_po |>
  rename_with(.cols = None:sesame, .fn = str_to_sentence) |>
  tbl_summary(
    by = trt,
    include = None:Sesame,
    missing = "no",
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**IgE-mediated food allergy**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  modify_table_body(
    ~ .x |>
      mutate(
        variable = fct_reorder(
          variable,
          as.numeric(str_remove(stat_0, "\\s*\\(.*\\)")),
          .fun = mean,
          .desc = TRUE
        )
      ) |>
      arrange(variable)
  )
tbl |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(70)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

### Primary outcome model

```{r}
#| label: primary-outcome-analysis
set.seed(7761)
po_sum <- dat_po |>
  summarise(
    n = n(),
    y = sum(po),
    p = y / n,
    a = 1 + y,
    b = 1 + n - y,
    l = 1000 - n,
    post = dist_beta(a, b),
    .by = trt
  ) |>
  mutate(
    p_sup = c(NA, beta_ineq_sim(a, b)),
    ppos1 = c(NA, calc_ppos(a, b, l, 0.95, post_method = "sim", post_sim = 2e4))
  )
ppos1 <- sprintf("%.3f", po_sum$ppos1[2])
po_diff <- po_sum |>
  mutate(post_drws = generate(post, 1e6)) |>
  select(trt, post_drws) |>
  unnest(post_drws) |>
  group_by(trt) |>
  mutate(.draw = 1:n()) |>
  ungroup() |>
  pivot_wider(names_from = "trt", values_from = "post_drws") |>
  mutate(delta = wP - aP)

prior <- c(
  prior(logistic(0, 1), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = "b")
)
fit <- brm(
  po ~ trt + rand_site + gender + bfed + fborn + parinc_imp + ces + fha,
  data = dat_po,
  family = bernoulli(),
  prior = prior,
  warmup = 500,
  iter = 5500,
  refresh = 0,
  silent = 2,
  seed = 54123
)

wgt <- c(rdo(rdirichlet(1, rep(1, 972)), ndraws = ndraws(fit)))
dat_rd_po <- dat_po |>
  mutate(trt = "aP") |>
  add_epred_rvars(fit, value = ".epred_aP") |>
  mutate(trt = "wP") |>
  add_epred_rvars(fit, value = ".epred_wP") |>
  mutate(wgt = wgt) |>
  summarise(
    aP = rvar_mean(.epred_aP),
    wP = rvar_mean(.epred_wP),
    .rd_empirical = rvar_mean(.epred_wP - .epred_aP),
    .rd_bootstrap = rvar_sum((.epred_wP - .epred_aP) * wgt)
  )
```

@tbl-posterior-primary summarises the estimated model parameters for the probability of IgE-mediated food allergy under each treatment group, and their difference. The posterior densities for the difference in risk of IgE-mediated food allergy (wP vs aP) are shown in @fig-posterior-difference-primary-unadjusted-adjusted.

The results between the primary analysis (unadjusted) and secondary (adjusted) analyses were near identical. Given the data, under the assumed model the most plausible values for the difference in IgE-mediated food allergy between wP and aP were `r do.call(sprintf, c(fmt = "%.2f (%.2f, %.2f)", as.list(c(quantile2(po_diff$delta, c(0.5, 0.025, 0.975))))))` [posterior median, (95% credible interval)]. The probability of a reduction in IgE-medidated food allergy under wP compared to aP was `r sprintf("%.2f", mean(po_diff$delta < 0))`.

A pre-specified stopping rule was to assess the predicted probability of success (PPoS) up to the planned maximum sample size of 2,000 infants (1,000 per arm). Success required that the probability of a reduction in IgE-mediated food allergy exceed 0.95. Calculation of PPoS used the posterior predictive distribution under each arm to impute outcomes for future participantts. If PPoS < 0.05 then the trial was to be stopped for  futility. At the time of stopping, 486 participants had been assigned to each arm with 514 further infants required to reach 1,000 per arm. At this stage, $\text{PPoS} \approx `{r} ppos1` < 0.05$ and would have met the futility criteria. 

```{r}
#| label: tbl-posterior-primary
#| tbl-cap: |
#|   Posterior summaries for risk and difference in risk of IgE-mediated food allergy.
#| tbl-pos: "H"
diff_agg <- po_diff |>
  nest_rvars() |>
  ungroup() |>
  mutate(
    model = "Unadjusted",
    across(
      aP:delta,
      list(
        med = median,
        lo = ~ quantile2(.x, 0.025),
        hi = ~ quantile2(.x, 0.975),
        pr = ~ Pr(.x < 0)
      ),
      .names = "{.fn}_{.col}"
    )
  ) |>
  select(-aP, -wP, -delta) |>
  bind_rows(
    dat_rd_po |>
      rename(delta = .rd_empirical) |>
      select(aP, wP, delta) |>
      mutate(model = "Adjusted") |>
      mutate(across(
        aP:delta,
        list(
          med = median,
          lo = ~ quantile2(.x, 0.025),
          hi = ~ quantile2(.x, 0.975),
          pr = ~ Pr(.x < 0)
        ),
        .names = "{.fn}_{.col}"
      )) |>
      select(-aP, -wP, -delta)
  ) |>
  select(-pr_aP, -pr_wP)

po_sum |>
  select(trt, n, y, p) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p)
  ) |>
  bind_cols(diff_agg) |>
  gt(rowname_col = "model") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_aP, lo_aP, hi_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_wP, lo_wP, hi_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "aP"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "wP"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_aP, med_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  tab_options(table.font.size = px(9))
```

```{r}
#| label: fig-posterior-density
#| fig-cap: |
#|   Posterior density for difference in risk of primary outcome, unadjusted primary model.
#| fig-height: 2
#| fig-width: 5
#| include: false
p1 <- po_diff |>
  nest_rvars() |>
  ggplot(aes(xdist = delta)) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    point_size = 1.5,
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = seq(-0.1, 0.1, 0.02), limits = c(-0.08, 0.08)) +
  scale_y_continuous(limits = ~ range(.x, -0.1)) +
  labs(
    fill = "Interval",
    x = "Difference in IgE-mediated food allergy risk, wP - aP",
    y = "Density"
  ) +
  guides(fill = guide_legend(override.aes = list(point_size = NA)))
p1
```


```{r}
#| label: fig-posterior-difference-adjusted
#| fig-cap: Posterior difference in risk of IgE-mediated food allergy for wP compared to aP, adjusted model.
#| fig-height: 2
#| fig-width: 5
#| include: false
p2 <- ggplot(dat_rd_po, aes(xdist = .rd_bootstrap)) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    point_size = 1.5,
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = seq(-0.1, 0.1, 0.02), limits = c(-0.08, 0.08)) +
  scale_y_continuous(limits = ~ range(.x, -0.1)) +
  labs(
    fill = "Interval",
    x = "Difference in IgE-mediated food allergy risk, wP - aP",
    y = "Density"
  ) +
  guides(fill = guide_legend(override.aes = list(point_size = NA)))
p2
```

```{r}
#| label: fig-posterior-difference-primary-unadjusted-adjusted
#| fig-cap: |
#|   Posterior density for difference in risk of IgE-mediated food allergy by model.
#|   Values in regions with high density are those most plausible given the data and model assumptions.
#| fig-height: 2
#| fig-width: 6
(p1 + labs(title = "Unadjusted") | p2 + labs(title = "Adjusted")) +
  plot_layout(axes = "collect", axis_titles = "collect", guides = "collect") &
  theme(
    legend.position = "bottom",
    legend.margin = margin(t = 0, unit = "cm"),
    legend.key.size = unit(1, "line"),
    plot.title = element_text(size = rel(1)),
    text = element_text(size = 9)
  )
```

\clearpage

::: {.landscape}

### Subgroup analyses

#### Site

```{r}
#| label: tbl-primary-outcome-criteria-site
#| tbl-cap: |
#|   Number and percentage of infants who met each primary outcome condition by treatment group and participating site.
#| tbl-pos: "H"
#| output: asis
tbl <- dat_po |>
  mutate(rand_site = fct_infreq(rand_site)) |>
  tbl_strata2(
    strata = rand_site,
    .tbl_fun = ~ .x |>
      tbl_summary(
        by = trt,
        include = c(po_ofc, po_ana, po_ige),
        label = list(
          po_ofc ~ "Met condition 1.i.",
          po_ana ~ "Met condition 1.ii.",
          po_ige ~ "Met condition 2."
        ),
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Outcome**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_column_alignment(all_stat_cols(), "right")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(9), table.width = pct(100))
tbl |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-primary-outcome-overall-site
#| tbl-cap: |
#|   Number and percentage of infants who met the primary outcome by primary reason, treatment group, and participating site.
#| tbl-pos: "H"
#| output: asis
tbl <- dat_po |>
  mutate(
    rand_site = fct_infreq(rand_site),
    po_criteria = factor(po_criteria, levels = c("Criteria 1.i", "Criteria 1.ii", "Criteria 2.", "None"))
  ) |>
  tbl_strata(
    strata = rand_site,
    .tbl_fun = ~ .x |>
      tbl_summary(
        by = trt,
        include = c(po, po_criteria),
        label = list(
          po ~ "Met primary outcome",
          po_criteria ~ "Reason"
        ),
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Outcome**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_table_body(
        ~ .x |>
          filter(label != "None")
      ) |>
      modify_column_alignment(all_stat_cols(), "right")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(9), table.width = pct(100))
tbl |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: primary-outcome-subgroup-model-site
dat_po <- dat_po |>
  mutate(rand_site = fct_infreq(rand_site))
# So we can specify prior in terms of deviation from overall "mean" effect
contrasts(dat_po$rand_site) <- bayestestR::contr.equalprior(4)

trt_prior <- prior(normal(0, 2.5), coef = "trtwP")
dev_prior <- "normal(0, 2)"

prior <- c(
  trt_prior,
  prior_string(dev_prior, coef = "rand_site1:trtwP"),
  prior_string(dev_prior, coef = "rand_site2:trtwP"),
  prior_string(dev_prior, coef = "rand_site3:trtwP")
)

fit <- brm(
  po ~ 0 + rand_site + trt + trt:rand_site,
  data = dat_po,
  family = bernoulli(),
  prior = prior,
  warmup = 500,
  iter = 5500,
  refresh = 0,
  silent = 2,
  seed = 54123
)

fit_pars_site <- dat_po |>
  distinct(trt, rand_site) |>
  arrange(rand_site) |>
  add_linpred_rvars(fit) |>
  add_epred_rvars(fit) |>
  pivot_wider(names_from = trt, values_from = c(.linpred, .epred)) |>
  mutate(or = exp(.linpred_wP - .linpred_aP), rd = .epred_wP - .epred_aP)

# fit_pars_site |>
#   select(rand_site, or) |>
#   # add_row(rand_site = "Average", or = exp(rvar_mean(log(fit_pars_site$or))), .before = 1) |>
#   # mutate(rand_site = fct_inorder(rand_site)) |>
#   ggplot(aes(y = rand_site, xdist = or)) +
#   stat_pointinterval(
#     .width = c(0.5, 0.8, 0.95),
#     point_size = 1.5
#   ) +
#   geom_vline(xintercept = 1, linetype = 2) +
#   scale_x_log10(breaks = c(0.1, 0.25, 0.5, 1, 2, 4, 10)) +
#   scale_y_discrete(limits = rev) +
#   labs(
#     x = "Odds ratio for IgE-mediated food allergy, wP versus aP",
#     y = "Density"
#   )

# fit_pars_site |>
#   select(rand_site, rd) |>
#   ggplot(aes(y = rand_site, xdist = rd)) +
#   # facet_wrap(~name, scales = "free") +
#   stat_pointinterval(
#     .width = c(0.5, 0.8, 0.95),
#     point_size = 1.5
#   ) +
#   geom_vline(xintercept = 0, linetype = 2) +
#   scale_x_continuous(breaks = seq(-0.2, 0.2, 0.02)) +
#   scale_y_discrete(limits = rev) +
#   labs(
#     x = "Difference in IgE-mediated food allergy risk, wP - aP",
#     y = "Density"
#   )
```

:::

\clearpage

::: {.landscape}

#### Season of birth

```{r}
#| label: tbl-primary-outcome-criteria-season
#| tbl-cap: |
#|   Number and percentage of infants who met each primary outcome condition by treatment group and season of birth.
#| tbl-pos: "H"
#| output: asis
dat_po <- dat_po |>
  mutate(
    birth_season = factor(
      case_when(
        month(birthdat) %in% c(12, 1, 2) ~ "Summer",
        month(birthdat) %in% 3:5 ~ "Autumn",
        month(birthdat) %in% 6:8 ~ "Winter",
        month(birthdat) %in% 9:11 ~ "Spring"
      ),
      levels = c("Autumn", "Winter", "Spring", "Summer")
    )
  )
tbl <- dat_po |>
  tbl_strata2(
    strata = birth_season,
    .tbl_fun = ~ .x |>
      tbl_summary(
        by = trt,
        include = c(po_ofc, po_ana, po_ige),
        label = list(
          po_ofc ~ "Met condition 1.i.",
          po_ana ~ "Met condition 1.ii.",
          po_ige ~ "Met condition 2."
        ),
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Outcome**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_column_alignment(all_stat_cols(), "right")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(9), table.width = pct(100))
tbl |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-primary-outcome-overall-season
#| tbl-cap: |
#|   Number and percentage of infants who met the primary outcome by primary reason, treatment group, and season of birth.
#| tbl-pos: "H"
#| output: asis
tbl <- dat_po |>
  mutate(
    po_criteria = factor(po_criteria, levels = c("Criteria 1.i", "Criteria 1.ii", "Criteria 2.", "None"))
  ) |>
  tbl_strata(
    strata = birth_season,
    .tbl_fun = ~ .x |>
      tbl_summary(
        by = trt,
        include = c(po, po_criteria),
        label = list(
          po ~ "Met primary outcome",
          po_criteria ~ "Reason"
        ),
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Outcome**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_table_body(
        ~ .x |>
          filter(label != "None")
      ) |>
      modify_column_alignment(all_stat_cols(), "right")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(9), table.width = pct(100))
tbl |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: primary-outcome-subgroup-model-season
contrasts(dat_po$birth_season) <- bayestestR::contr.equalprior(4)

prior <- c(
  trt_prior,
  prior_string(dev_prior, coef = "birth_season1:trtwP"),
  prior_string(dev_prior, coef = "birth_season2:trtwP"),
  prior_string(dev_prior, coef = "birth_season3:trtwP")
)

fit <- brm(
  po ~ 0 + birth_season + trt + trt:birth_season,
  data = dat_po,
  family = bernoulli(),
  # prior = prior,
  warmup = 500,
  iter = 5500,
  refresh = 0,
  silent = 2,
  seed = 54123
)

fit_pars_season <- dat_po |>
  distinct(trt, birth_season) |>
  arrange(birth_season) |>
  add_linpred_rvars(fit) |>
  add_epred_rvars(fit) |>
  pivot_wider(names_from = trt, values_from = c(.linpred, .epred)) |>
  mutate(or = exp(.linpred_wP - .linpred_aP), rd = .epred_wP - .epred_aP)
```

:::

\clearpage

::: {.landscape}

#### Family history of atopy

Family history of atopy was derived from responses on the corresponding baseline form. If an infant had any parent or sibling with a reported history of asthma, eczema, allergic rhinitis, or food allergy, then they were considered to have a family history of atopy.

```{r}
#| label: tbl-primary-outcome-criteria-fha
#| tbl-cap: |
#|   Number and percentage of infants who met each primary outcome condition by treatment group and family history of atopy (yes or no).
#| tbl-pos: "H"
#| output: asis
tbl <- dat_po |>
  tbl_strata2(
    strata = fha,
    .tbl_fun = ~ .x |>
      tbl_summary(
        by = trt,
        include = c(po_ofc, po_ana, po_ige),
        label = list(
          po_ofc ~ "Met condition 1.i.",
          po_ana ~ "Met condition 1.ii.",
          po_ige ~ "Met condition 2."
        ),
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Outcome**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_column_alignment(all_stat_cols(), "right")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(9), table.width = pct(70))
tbl |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-primary-outcome-overall-fha
#| tbl-cap: |
#|   Number and percentage of infants who met the primary outcome by primary reason, treatment group, and family history of atopy (yes or no).
#| tbl-pos: "H"
#| output: asis
tbl <- dat_po |>
  mutate(
    po_criteria = factor(po_criteria, levels = c("Criteria 1.i", "Criteria 1.ii", "Criteria 2.", "None"))
  ) |>
  tbl_strata(
    strata = fha,
    .tbl_fun = ~ .x |>
      tbl_summary(
        by = trt,
        include = c(po, po_criteria),
        label = list(
          po ~ "Met primary outcome",
          po_criteria ~ "Reason"
        ),
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Outcome**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_table_body(
        ~ .x |>
          filter(label != "None")
      ) |>
      modify_column_alignment(all_stat_cols(), "right")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(9), table.width = pct(70))
tbl |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: primary-outcome-subgroup-model-fha
contrasts(dat_po$fha) <- bayestestR::contr.equalprior(2)

prior <- c(
  trt_prior,
  prior_string(dev_prior, coef = "fha1:trtwP")
)

fit <- brm(
  po ~ 0 + fha + trt + trt:fha,
  data = dat_po,
  family = bernoulli(),
  # prior = prior,
  warmup = 500,
  iter = 5500,
  refresh = 0,
  silent = 2,
  seed = 54123
)

fit_pars_fha <- dat_po |>
  distinct(trt, fha) |>
  arrange(fha) |>
  add_linpred_rvars(fit) |>
  add_epred_rvars(fit) |>
  pivot_wider(names_from = trt, values_from = c(.linpred, .epred)) |>
  mutate(or = exp(.linpred_wP - .linpred_aP), rd = .epred_wP - .epred_aP)
```

:::

\clearpage

::: {.landscape}

#### Family history of food allergy

Family history of food allergy was derived from responses on the corresponding baseline form. If an infant had any parent or sibling with a reported history of food allergy, then they were considered to have a family history of food allergy.

```{r}
#| label: tbl-primary-outcome-criteria-fhafa
#| tbl-cap: |
#|   Number and percentage of infants who met each primary outcome condition by treatment group and family history of atopy (yes or no).
#| tbl-pos: "H"
#| output: asis
tbl <- dat_po |>
  mutate(fhafa = factor(fhafa, levels = c(0, 1), labels = c("No", "Yes"))) |>
  tbl_strata2(
    strata = fhafa,
    .tbl_fun = ~ .x |>
      tbl_summary(
        by = trt,
        include = c(po_ofc, po_ana, po_ige),
        label = list(
          po_ofc ~ "Met condition 1.i.",
          po_ana ~ "Met condition 1.ii.",
          po_ige ~ "Met condition 2."
        ),
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Outcome**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_column_alignment(all_stat_cols(), "right")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(9), table.width = pct(70))
tbl |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-primary-outcome-overall-fhafa
#| tbl-cap: |
#|   Number and percentage of infants who met the primary outcome by primary reason, treatment group, and family history of atopy (yes or no).
#| tbl-pos: "H"
#| output: asis
tbl <- dat_po |>
  mutate(
    fhafa = factor(fhafa, levels = c(0, 1), labels = c("No", "Yes")),
    po_criteria = factor(po_criteria, levels = c("Criteria 1.i", "Criteria 1.ii", "Criteria 2.", "None"))
  ) |>
  tbl_strata(
    strata = fhafa,
    .tbl_fun = ~ .x |>
      tbl_summary(
        by = trt,
        include = c(po, po_criteria),
        label = list(
          po ~ "Met primary outcome",
          po_criteria ~ "Reason"
        ),
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Outcome**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_table_body(
        ~ .x |>
          filter(label != "None")
      ) |>
      modify_column_alignment(all_stat_cols(), "right")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(9), table.width = pct(70))
tbl |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: primary-outcome-subgroup-model-fhafa
dat_po$fhafa <- factor(dat_po$fhafa, levels = 0:1, labels = c("No", "Yes"))
contrasts(dat_po$fhafa) <- bayestestR::contr.equalprior(2)

prior <- c(
  trt_prior,
  prior_string(dev_prior, coef = "fhafaYes:trtwP")
)

fit <- brm(
  po ~ 0 + fhafa + trt + trt:fhafa,
  data = dat_po,
  family = bernoulli(),
  # prior = prior,
  warmup = 500,
  iter = 5500,
  refresh = 0,
  silent = 2,
  seed = 54123
)

fit_pars_fhafa <- dat_po |>
  distinct(trt, fhafa) |>
  arrange(fhafa) |>
  add_linpred_rvars(fit) |>
  add_epred_rvars(fit) |>
  pivot_wider(names_from = trt, values_from = c(.linpred, .epred)) |>
  mutate(or = exp(.linpred_wP - .linpred_aP), rd = .epred_wP - .epred_aP)
```

:::

\clearpage

::: {.landscape}

#### Parental country of birth

Whether an infant had one or more parent born in East Asia was derived from the mother's country of birth (`bcmoth`) and the father's country of birth (`bcfath`) classified according to the Standard Australian Classification of Countries (SACC). If either of these was unknown (being missing or reported as "other") than the infant's status was considered "unknown".

```{r}
#| label: subgroup-asia-derive
#| include: false
# Which countries (of those appearing) should be mapped to Asia?
# We focus on "East Asia" as per Standard Australian Classification of Countries from ABS
countries_asia <- c(
  "Malaysia",
  "China (excludes SARs and Taiwan)",
  # "India",
  "Indonesia",
  "Philippines",
  "Hong Kong (SAR of China)",
  "Singapore",
  "Vietnam",
  # "Nepal",
  # "Sri Lanka",
  # "Maldives",
  "Taiwan",
  "Thailand",
  "Korea, Republic of (South)",
  "Brunei Darussalam",
  "Japan",
  "Myanmar"
  # "Pakistan"
)

country_to_region <- tribble(
  ~country                           , ~region     ,
  countries_asia                     , "East Asia" ,
  c(NA, "Other", "OTHER", "Unknown") , "Unknown"
)

# Relabel country of birth
dat_po <- dat_po |>
  mutate(
    bcmoth_region = recode_values(
      bcmoth,
      from = country_to_region$country,
      to = country_to_region$region,
      default = "Not East Asia"
    ),
    bcfath_region = recode_values(
      bcfath,
      from = country_to_region$country,
      to = country_to_region$region,
      default = "Not East Asia"
    ),
    any_asia = factor(
      case_when(
        bcmoth_region == "East Asia" | bcfath_region == "East Asia" ~ "Yes",
        bcmoth_region == "Unknown" ~ "Unknown",
        bcfath_region == "Unknown" ~ "Unknown",
        bcmoth_region == "Not East Asia" & bcfath_region == "Not East Asia" ~ "No"
      ),
      levels = c("No", "Yes", "Unknown")
    )
  )

dat_po |>
  count(bcmoth_region, bcmoth) |>
  print(n = Inf)
dat_po |>
  count(bcfath_region, bcfath) |>
  print(n = Inf)
dat_po |>
  count(any_asia, bcmoth_region, bcfath_region)
```


```{r}
#| label: tbl-primary-outcome-criteria-asia
#| tbl-cap: |
#|   Number and percentage of infants who met each primary outcome condition by treatment group and whether at least one parent was born in East Asia. Unknown cases occurred where either the mother or father's country of birth was not known (either missing or reported as "other").
#| tbl-pos: "H"
#| output: asis
tbl <- dat_po |>
  tbl_strata2(
    strata = any_asia,
    .tbl_fun = ~ .x |>
      tbl_summary(
        by = trt,
        include = c(po_ofc, po_ana, po_ige),
        label = list(
          po_ofc ~ "Met condition 1.i.",
          po_ana ~ "Met condition 1.ii.",
          po_ige ~ "Met condition 2."
        ),
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Outcome**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_column_alignment(all_stat_cols(), "right")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(9), table.width = pct(70))
tbl |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-primary-outcome-overall-asia
#| tbl-cap: |
#|   Number and percentage of infants who met the primary outcome by primary reason, treatment group, and whether at least one parent was born in East Asia. Unknown cases occurred where either the mother or father's country of birth was not known (either missing or reported as "other").
#| tbl-pos: "H"
#| output: asis
tbl <- dat_po |>
  mutate(
    po_criteria = factor(po_criteria, levels = c("Criteria 1.i", "Criteria 1.ii", "Criteria 2.", "None"))
  ) |>
  tbl_strata(
    strata = any_asia,
    .tbl_fun = ~ .x |>
      tbl_summary(
        by = trt,
        include = c(po, po_criteria),
        label = list(
          po ~ "Met primary outcome",
          po_criteria ~ "Reason"
        ),
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Outcome**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_table_body(
        ~ .x |>
          filter(label != "None")
      ) |>
      modify_column_alignment(all_stat_cols(), "right")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(9), table.width = pct(70))
tbl |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: primary-outcome-subgroup-model-asia
contrasts(dat_po$any_asia) <- bayestestR::contr.equalprior(3)

prior <- c(
  trt_prior,
  prior_string(dev_prior, coef = "any_asia1:trtwP"),
  prior_string(dev_prior, coef = "any_asia2:trtwP")
)

fit <- brm(
  po ~ 0 + any_asia + trt + trt:any_asia,
  data = dat_po,
  family = bernoulli(),
  # prior = prior,
  warmup = 500,
  iter = 5500,
  refresh = 0,
  silent = 2,
  seed = 54123
)

fit_pars_asia <- dat_po |>
  distinct(trt, any_asia) |>
  arrange(any_asia) |>
  add_linpred_rvars(fit) |>
  add_epred_rvars(fit) |>
  pivot_wider(names_from = trt, values_from = c(.linpred, .epred)) |>
  mutate(or = exp(.linpred_wP - .linpred_aP), rd = .epred_wP - .epred_aP)
```

:::

\clearpage

#### Subgroup forest plot

For each subgroup we estimated the difference in risk of IgE-mediated food allergy for wP compared to aP within each grouping level. The estimated risk differences for each factor are reported in @fig-subgroup-plot.

```{r}
#| label: fig-subgroup-plot
#| fig-cap: Subgroup specific estimates of difference in risk of IgE-mediated food allergy, wP versus aP.
#| fig-width: 4
#| fig-height: 6.5
bind_rows(
  fit_pars_site |>
    select(rand_site, rd) |>
    mutate(var = "Site of enrolment") |>
    rename(level = rand_site),
  fit_pars_season |>
    select(birth_season, rd) |>
    mutate(var = "Season of birth") |>
    rename(level = birth_season),
  fit_pars_fha |>
    select(fha, rd) |>
    mutate(var = "Family history\nof atopy") |>
    rename(level = fha),
  fit_pars_fhafa |>
    select(fhafa, rd) |>
    mutate(var = "Family history\nof food allergy") |>
    rename(level = fhafa),
  fit_pars_asia |>
    select(any_asia, rd) |>
    mutate(var = "One parent born\nin East Asia") |>
    rename(level = any_asia)
) |>
  mutate(var = fct_inorder(var)) |>
  ggplot(aes(y = level, xdist = rd)) +
  facet_grid(var ~ ., scales = "free_y", space = "free_y", switch = "y") +
  stat_pointinterval(.width = c(0.5, 0.8, 0.95), point_size = 1.5) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_y_discrete("", limits = rev) +
  scale_x_continuous(
    "Difference in risk of IgE-mediated food allergy\nwP - aP",
    breaks = seq(-0.2, 0.2, 0.05)
  ) +
  theme(strip.placement = "outside")
```


\clearpage

## Eczema

### Overview

The primary source of information for eczema was the `outcome_report` form. The timing was based on the documented age at which eczema first occurred. Reported outcomes were cross checked against the food and household questionnaire form and medical history forms. Infants whose first reaction date for clinician-diagnosed eczema was prior to enrolment were considered to have "pre-existing" eczema and were excluded from the analysis set. There were `r sum(dat_ecz$out_ecz_preexisting)` such infants.

The number and percentage of infants with any clinician-diagnosed eczema is reported in @tbl-eczema-trt. For the protocol defined outcomes \[2a. and 2b.\], the eczema diagnosis needed to be new-onset and accompanied by a positive SPT to at least one allergen by approximately 12 months of age. These are reported in @tbl-eczema-trt-newonset.

```{r}
#| label: tbl-eczema-trt
#| tbl-cap: |
#|   Number and percentage of infants with any clinician-diagnosed eczema.
#| tbl-pos: "H"
#| output: asis
# dat_ecz |>
#   filter(out_ecz_preexisting) |>
#   summarise(n = sum(out_ecz_preexisting), .by = trt)

tbl0 <- dat_ecz_spt |>
  mutate(
    ecz = ecz,
    ecz_6m = ecz & outageval_months2 <= 6,
    ecz_12m = ecz & outageval_months2 <= 12,
    ecz_18m = ecz & outageval_months2 <= 18
  ) |>
  tbl_summary(
    by = trt,
    include = c(ecz, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      filter(label != "None")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(55)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-eczema-trt-newonset
#| tbl-cap: |
#|   Number and percentage of infants with new-onset clinician-diagnosed eczema.
#|   To be new-onset the first reaction date needed to be after the infants randomisation date.
#|   Those with missing were due to eczema diagnosis being present but no SPT records being available.
#| tbl-pos: "H"
#| output: asis
tbl1 <- dat_ecz_spt |>
  filter(!out_ecz_preexisting) |>
  mutate(
    ecz_any = ecz & !out_ecz_preexisting,
    ecz_6m = ecz & !out_ecz_preexisting & outageval_months2 <= 6,
    ecz_12m = ecz & !out_ecz_preexisting & outageval_months2 <= 12,
    ecz_18m = ecz & !out_ecz_preexisting & outageval_months2 <= 18
  ) |>
  tbl_summary(
    by = trt,
    include = c(ecz_any, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz_any ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**")

# With positive SPT
tbl2 <- dat_ecz_spt |>
  filter(!out_ecz_preexisting) |>
  mutate(
    ecz_any = ecz & !out_ecz_preexisting & any_pos_spt,
    ecz_6m = ecz & !out_ecz_preexisting & outageval_months2 <= 6 & any_pos_spt,
    ecz_12m = ecz &
      !out_ecz_preexisting &
      outageval_months2 <= 12 &
      any_pos_spt,
    ecz_18m = ecz & !out_ecz_preexisting & outageval_months2 <= 18 & any_pos_spt
  ) |>
  tbl_summary(
    by = trt,
    include = c(ecz_any, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz_any ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**")

tbl_stack(
  list(tbl1, tbl2),
  group_header = c(
    "New-onset cases",
    "New-onset cases with positive SPT"
  )
) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      filter(label != "None")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(55)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

### Eczema model

```{r}
#| label: eczema-analysis
set.seed(54123)

dat_outcome <- dat_ecz_analysis |>
  pivot_longer(
    ecz_6m:ecz_18m,
    names_pattern = "ecz_([0-9]+m)",
    names_to = "age",
    values_to = "outcome"
  ) |>
  mutate(
    age = factor(
      age,
      levels = c("6m", "12m", "18m"),
      labels = c("6m [2.a]", "12m [2.b]", "18m")
    )
  ) |>
  filter(!is.na(outcome)) |>
  summarise(
    n = n(),
    y = sum(outcome),
    p = y / n,
    a = 1 + y,
    b = 1 + n - y,
    post = dist_beta(a, b),
    .by = c(trt, age)
  ) |>
  mutate(p_sup = c(NA, beta_ineq_sim(a, b)), .by = age)

diff_outcome <- dat_outcome |>
  mutate(post_drws = generate(post, 1e6)) |>
  select(age, trt, post_drws) |>
  unnest(post_drws) |>
  group_by(age, trt) |>
  mutate(.draw = 1:n()) |>
  ungroup() |>
  pivot_wider(names_from = "trt", values_from = "post_drws") |>
  mutate(delta = wP - aP)

diff_2a <- filter(diff_outcome, age == "6m [2.a]")
diff_2b <- filter(diff_outcome, age == "12m [2.b]")

prior <- c(
  prior(logistic(0, 1), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = "b")
)

fit_outcome <- dat_ecz_analysis |>
  pivot_longer(
    ecz_6m:ecz_18m,
    names_pattern = "ecz_([0-9]+m)",
    names_to = "age",
    values_to = "outcome"
  ) |>
  mutate(
    age = factor(
      age,
      levels = c("6m", "12m", "18m"),
      labels = c("6m [2.a]", "12m [2.b]", "18m")
    )
  ) |>
  nest(-age) |>
  mutate(
    fit = map(
      data,
      \(x) {
        brm(
          outcome ~
            trt + rand_site + gender + bfed + fborn + parinc_imp + ces + fha,
          data = x,
          family = bernoulli(),
          prior = prior,
          warmup = 500,
          iter = 5500,
          refresh = 0,
          silent = 2,
          seed = 54123
        )
      }
    ),
    rd = map2(data, fit, \(x, y) {
      wgt <- c(rdo(rdirichlet(1, rep(1, nrow(x))), ndraws = ndraws(y)))
      x |>
        mutate(trt = "aP") |>
        add_epred_rvars(y, value = ".epred_aP") |>
        mutate(trt = "wP") |>
        add_epred_rvars(y, value = ".epred_wP") |>
        mutate(wgt = wgt) |>
        summarise(
          aP = rvar_mean(.epred_aP),
          wP = rvar_mean(.epred_wP),
          .rd_empirical = rvar_mean(.epred_wP - .epred_aP),
          .rd_bootstrap = rvar_sum((.epred_wP - .epred_aP) * wgt)
        )
    })
  )
```

@tbl-posterior-eczema summarises the estimated model parameters for the probability of eczema by age under each treatment group, and their difference (the posterior densities are shown in @fig-posterior-difference-eczema-unadjusted-adjusted). The results between the analyses (unadjusted and adjusted, @sec-primary-analysis) were near identical. The most plausible values for the difference in eczema at 6 months [2.a] between wP and aP were `r do.call(sprintf, c(fmt = "%.2f (%.2f, %.2f)", as.list(c(quantile2(diff_2a$delta, c(0.5, 0.025, 0.975))))))` [posterior median, (95% credible interval)]. The probability of a reduction in new-onset eczema under wP compared to aP was `r sprintf("%.2f", mean(diff_2a$delta < 0))`. At 12m [2.b] they were `r do.call(sprintf, c(fmt = "%.2f (%.2f, %.2f)", as.list(c(quantile2(diff_2b$delta, c(0.5, 0.025, 0.975))))))` and `r sprintf("%.2f", mean(diff_2b$delta < 0))`.


```{r}
#| label: tbl-posterior-eczema
#| tbl-cap: |
#|   Posterior summaries for risk and difference in risk of eczema by age of onset.
#| tbl-pos: "H"
diff_agg <- diff_outcome |>
  group_by(age) |>
  nest_rvars() |>
  ungroup() |>
  mutate(
    model = "Unadjusted",
    across(
      aP:delta,
      list(
        med = median,
        lo = ~ quantile2(.x, 0.025),
        hi = ~ quantile2(.x, 0.975),
        pr = ~ Pr(.x < 0)
      ),
      .names = "{.fn}_{.col}"
    )
  ) |>
  select(-aP, -wP, -delta) |>
  bind_rows(
    fit_outcome |>
      unnest(rd) |>
      rename(delta = .rd_empirical) |>
      select(age, aP, wP, delta) |>
      mutate(model = "Adjusted") |>
      mutate(across(
        aP:delta,
        list(
          med = median,
          lo = ~ quantile2(.x, 0.025),
          hi = ~ quantile2(.x, 0.975),
          pr = ~ Pr(.x < 0)
        ),
        .names = "{.fn}_{.col}"
      )) |>
      select(-aP, -wP, -delta)
  ) |>
  select(-pr_aP, -pr_wP)

dat_outcome |>
  select(age, trt, n, y, p) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p)
  ) |>
  left_join(diff_agg, join_by(age)) |>
  gt(groupname_col = "model", rowname_col = "age") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_aP, lo_aP, hi_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_wP, lo_wP, hi_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "aP"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "wP"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_aP, med_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  tab_options(table.font.size = px(9))
```

```{r}
#| label: fig-posterior-density-eczema
#| fig-cap: |
#|    Posterior density for difference in risk of eczema diagnosis by age at reaction,
#|    unadjusted primary model.
#| fig-height: 3
#| fig-width: 5
#| include: false
p1 <- diff_outcome |>
  group_by(age) |>
  nest_rvars() |>
  ggplot(aes(xdist = delta)) +
  facet_wrap(~age, ncol = 1) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    point_size = 1.5,
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = seq(-0.1, 0.1, 0.02), limits = c(-0.08, 0.08)) +
  scale_y_continuous(limits = ~ range(.x, -0.1)) +
  labs(
    fill = "Interval",
    x = "Difference in eczema risk, wP - aP",
    y = "Density"
  ) +
  guides(fill = guide_legend(override.aes = list(point_size = NA)))
p1
```


```{r}
#| label: fig-posterior-difference-eczema-adjusted
#| fig-cap: |
#|   Posterior difference in risk of eczema diagnosis for wP compared to aP, adjusted model.
#| fig-height: 3
#| fig-width: 5
#| include: false
p2 <- fit_outcome |>
  unnest(rd) |>
  ggplot(aes(xdist = .rd_bootstrap)) +
  facet_wrap(~age, ncol = 1) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    point_size = 1.5,
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = seq(-0.1, 0.1, 0.02), limits = c(-0.08, 0.08)) +
  scale_y_continuous(limits = ~ range(.x, -0.15)) +
  labs(
    fill = "Interval",
    x = "Difference in eczema risk, wP - aP",
    y = "Density"
  ) +
  guides(fill = guide_legend(override.aes = list(point_size = NA)))
p2
```

```{r}
#| label: fig-posterior-difference-eczema-unadjusted-adjusted
#| fig-cap: |
#|   Posterior density for difference in risk of eczema by age and model.
#|   Values in regions with high density are those most plausible given the data and model assumptions.
#| fig-height: 3.25
#| fig-width: 6
(p1 + labs(title = "Unadjusted") | p2 + labs(title = "Adjusted")) +
  plot_layout(axes = "collect", axis_titles = "collect", guides = "collect") &
  theme(
    legend.position = "bottom",
    legend.margin = margin(t = 0, unit = "cm"),
    legend.key.size = unit(0.9, "line"),
    plot.title = element_text(size = rel(1)),
    text = element_text(size = 9)
  )
```

\clearpage
