
# Safety {#sec-safety}

## Serious adverse events

A listing of serious adverse events are reported in @tbl-listing-sae. These are any events reported either in the SAE reporting log (`sae_reporting_log`, stage 2 only) or on the adverse events form (`adverse_events`) with `aetype = "SAE"`. In some cases, a record was on the SAE log but not the adverse event log, resulting in some missing information below.

```{r}
#| label: tbl-listing-sae
#| tbl-cap: |
#|   Listing of serious adverse events.
#|   Missing values were due to no adverse event record for corresponding SAE log record.
#| tbl-pos: "H"
dat_sae <- dat_raw |>
  select_form("serious_adverse_events") |>
  select(record_id, saestdat, saemeddra) |>
  distinct() |>
  arrange(str_rank(record_id, numeric = TRUE))

dat_ae_sae <- dat_raw |>
  select_form("adverse_events") |>
  filter(aetype == "SAE") |>
  select(record_id, aecode1, aesev, aetymead, aerel, aestdat, aeenddat) |>
  rename(saestdat = aestdat) |>
  arrange(str_rank(record_id, numeric = TRUE))

dat_saes <- full_join(dat_sae, dat_ae_sae, join_by(record_id, saestdat)) |>
  arrange(str_rank(record_id, numeric = TRUE), saestdat) |>
  mutate(saemeddra = if_else(is.na(aecode1), saemeddra, aecode1)) |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  relocate(trt, .after = record_id)

dat_saes |>
  left_join(select(dat_base, record_id, visdat1)) |>
  mutate(aeday = interval(visdat1, saestdat) %/% days(1)) |>
  select(-visdat1, -saestdat, -aeenddat, -aecode1) |>
  gt() |>
  cols_label(
    trt ~ "Treatment",
    saemeddra ~ "MedDRA code",
    aeday ~ "Days after visit 1"
  ) |>
  sub_missing() |>
  tab_options(table.font.size = px(8), table.width = pct(100)) |>
  cols_width(c(aerel, aetymead) ~ px(100))
```

\clearpage

## Adverse events

```{r}
#| label: ae-data-prep
dat_ae <- dat_raw |>
  select_form("adverse_events") |>
  left_join(select(dat_base, record_id, rand_site), join_by(record_id))
dat_base_ae <- dat_base |>
  select(
    record_id,
    rand_site,
    rand_stage,
    randdattim,
    trt,
    birthdat,
    visdat1
  ) |>
  left_join(
    select_form(dat_raw, "vaccine_administration_v3") |>
      select(record_id, vac7admdat),
    join_by(record_id)
  ) |>
  left_join(dat_ae, join_by(record_id, rand_site)) |>
  mutate(
    ae_age = interval(birthdat, aestdat) %/% days(1),
    aeday_v1 = interval(visdat1, aestdat) %/% days(1),
    aeday_v3 = interval(vac7admdat, aestdat) %/% days(1)
  )
```

There were `r nrow(dat_ae)` adverse events documented during the trial.
Most adverse events were for infants randomised at PCH (`r sum(dat_ae$rand_site == "PCH")`).
Most of these events were due to the additional monitoring performed as part of stage 1 (@tbl-events-by-stage-and-site) and for the first 150 participants recruited as part of stage 2 (@tbl-pch-adverse-events). There were `r sum(dat_ae$rand_site != "PCH")` adverse events for infants randomised at a site other than PCH (listing in @tbl-listing-ae-non-pch).

There was a higher number of adverse events in the wP group compared to aP. There was also a higher number of infants who experienced at least one AE (@tbl-summary-ae-pch). Some of the difference in total AEs was due to infants with mulitple events. For both groups only a few AEs were documented as related to the blinded study vaccine . A listing of the related AEs is given in @tbl-listing-ae-related-pch.

A summary of the number and proportion of infants randomised at PCH with any reported adverse event by the MeDRA preferred term is given in @tbl-any-event-ae.  A summary of the total number of events is given in @tbl-events-ae.

```{r}
#| label: tbl-events-by-stage-and-site
#| tbl-cap: |
#|   Total number of adverse events (event rate) by study stage and site.
#| tbl-pos: "H"
#| results: "asis"
dat_base_ae |>
  summarise(
    participants = length(unique(record_id)),
    aes = sum(!is.na(aecode1)),
    .by = c(trt, rand_site, rand_stage)
  ) |>
  pivot_wider(names_from = trt, values_from = c(participants, aes)) |>
  mutate(rand_stage = paste0("Stage ", rand_stage)) |>
  gt(groupname_col = "rand_stage", rowname_col = "rand_site") |>
  cols_merge(
    c(aes_aP, participants_aP),
    pattern = "{1}/{2}"
  ) |>
  cols_merge(
    c(aes_wP, participants_wP),
    pattern = "{1}/{2}"
  ) |>
  cols_label(
    aes_aP ~ "**aP<br>(events/participants)**",
    aes_wP ~ "**wP<br>(events/participants)**",
    .fn = md
  ) |>
  cols_width(
    aes_aP ~ px(130),
    aes_wP ~ px(130)
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  cols_align(align = "center", columns = c(aes_aP, aes_wP)) |>
  tab_options(table.font.size = px(9), table.width = pct(50)) |>
  as_latex() |>
  cat()
```

```{r}
#| label: tbl-pch-adverse-events
#| tbl-cap: |
#|   Number of adverse events amongst stage 2 PCH participants for the first 150 and for the remaining participants.
#| tbl-pos: "H"
#| results: "asis"
dat_base_ae |>
  filter(rand_site == "PCH", rand_stage == 2) |>
  mutate(
    cohort = if_else(
      as.numeric(substr(record_id, 6, 9)) <= 150,
      "First 150 at PCH",
      "Remaining PCH participants"
    )
  ) |>
  summarise(
    participants = length(unique(record_id)),
    aes = sum(!is.na(aecode1)),
    .by = c(trt, cohort)
  ) |>
  arrange(trt) |>
  pivot_wider(names_from = trt, values_from = c(participants, aes)) |>
  gt(rowname_col = "cohort") |>
  cols_merge(
    c(aes_aP, participants_aP),
    pattern = "{1}/{2}"
  ) |>
  cols_merge(
    c(aes_wP, participants_wP),
    pattern = "{1}/{2}"
  ) |>
  cols_label(
    aes_aP ~ "**aP<br>(events/participants)**",
    aes_wP ~ "**wP<br>(events/participants)**",
    .fn = md
  ) |>
  cols_width(
    aes_aP ~ px(130),
    aes_wP ~ px(130)
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  cols_align(align = "center", columns = c(aes_aP, aes_wP)) |>
  tab_options(table.font.size = px(9), table.width = pct(70)) |>
  as_latex() |>
  cat()
```

```{r}
#| label: tbl-listing-ae-non-pch
#| tbl-cap: |
#|   Adverse event listing of infants randomised at a site other than PCH.
#| tbl-pos: "H"
dat_base_ae |>
  filter(rand_site != "PCH", !is.na(aecode1)) |>
  select(record_id, rand_site, trt, aecode1, aesev, aerel, aetymead) |>
  gt() |>
  cols_label(
    rand_site ~ "Site",
    trt ~ "Treatment"
  ) |>
  tab_options(table.font.size = px(9), table.width = pct(100)) |>
  cols_width(aecode1 ~ px(100), aetymead ~ px(100), aerel ~ px(100)) |>
  cols_align(align = "left")
```

```{r}
#| label: tbl-summary-ae-pch
#| tbl-cap: |
#|   Number and proportion of infants with any adverse events and the number of adverse events per infant, at PCH.
#| tbl-pos: "H"
#| results: "asis"
dat_base_ae |>
  filter(rand_site == "PCH") |>
  summarise(
    `Any AE` = any(!is.na(aecode1)),
    `Any SAE` = any(aetype == "SAE", na.rm = TRUE),
    `Any related AE` = any(aerel == "Related", na.rm = TRUE),
    `Any severe AE` = any(aesev == "Severe", na.rm = TRUE),
    `Number of AEs` = sum(!is.na(aecode1)),
    `Number of related AEs` = sum(aerel == "Related", na.rm = TRUE),
    .by = c(record_id, trt)
  ) |>
  tbl_summary(
    by = "trt",
    include = `Any AE`:`Number of related AEs`,
    digits = ~0
  ) |>
  add_overall(last = TRUE) |>
  modify_column_alignment(columns = starts_with("stat"), align = "right") |>
  remove_footnote_header() |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(70),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-listing-ae-related-pch
#| tbl-cap: |
#|   Listing of adverse events designated as "Related".
#| tbl-pos: "H"
dat_base_ae |>
  filter(aerel == "Related") |>
  select(record_id, trt, aecode1, aeday_v1, aetype, aesev, aerel, aetymead) |>
  gt() |>
  cols_label(
    trt ~ "Treatment",
    aeday_v1 ~ "Days after visit 1"
  ) |>
  tab_options(table.font.size = px(9), table.width = pct(100)) |>
  cols_width(
    record_id ~ px(70),
    trt ~ px(70),
    aecode1 ~ px(100),
    aetymead ~ px(100),
    aerel ~ px(80),
    aetype ~ px(80),
    aesev ~ px(80),
    aerel ~ px(80),
    aeday_v1 ~ px(50)
  ) |>
  cols_align(align = "left") |>
  cols_align(aeday_v1, align = "right") |>
  sub_missing()
```

```{r}
#| label: tbl-any-event-ae
#| tbl-cap: |
#|   Proportion of infants with any event by MeDRA preferred terms, PCH only.
#| tbl-pos: "H"
#| results: "asis"
dat_base_ae |>
  filter(rand_site == "PCH") |>
  tbl_hierarchical(
    variables = aecode1,
    by = trt,
    id = record_id,
    overall_row = TRUE,
    denominator = filter(dat_base, rand_site == "PCH")
  ) |>
  add_overall(last = TRUE) |>
  modify_column_alignment(columns = starts_with("stat"), "right") |>
  remove_footnote_header() |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(70),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-events-ae
#| tbl-cap: |
#|   Number of events by MeDRA preferred terms, PCH only.
#| tbl-pos: "H"
#| results: "asis"
dat_base_ae |>
  filter(rand_site == "PCH") |>
  tbl_hierarchical_count(
    variables = aecode1,
    by = trt,
    overall_row = TRUE,
    denominator = filter(dat_base, rand_site == "PCH")
  ) |>
  add_overall(last = TRUE) |>
  modify_column_alignment(columns = starts_with("stat"), "right") |>
  remove_footnote_header() |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(70),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage
