
# Reactogenicity

## Diary Cards

Combined data from stage 1 and stage 2 diary card responses are reported in the following sections.
During stage 1, diary cards were provided following 6-week, 4-month, 6-month and 18-month vaccinations.
During stage 2, at PCH only, for the first 153 participants, diary cards were provided following 6-week and 18-month vaccinations.
Given the different collection schedules for the protocol versions, the denominators for the 6-week and 18-month responses differ to those for the 4-month and 6-month responses.

This report focuses on 6-week and 18-month occasions given the addition of stage 2 data. 
The complete 4-month and 6-month data from stage 1 has been reported previously but is included here for completeness.

The number and percentage of returned diary cards for stage 1 are shown in @tbl-diary-returned-stage1 and stage 2 in @tbl-diary-returned-stage2.

```{r}
#| label: diary-data
st1_pth <- file.path(
  Sys.getenv("RDS_PATH"),
  "data",
  "derived",
  "tidy",
  "stage1"
)

plot_react <- function(
  dat,
  trt = treatment,
  var = intensity,
  slab = "Intensity",
  ref = 0,
  ymax = 1
) {
  pdat <- dat |>
    group_by({{ trt }}, vaxage, day) |>
    count({{ var }}) |>
    mutate(
      p = n / if_else(is.na({{ var }}), sum(n), sum(n[!is.na({{ var }})]))
    ) |>
    ungroup() |>
    complete({{ trt }}, vaxage, day, {{ var }}, fill = list(n = 0, p = 0))
  plab <- dat |>
    summarise(
      n = sum({{ var }} != ref, na.rm = TRUE),
      p = mean({{ var }} != ref, na.rm = TRUE),
      lab = sprintf(
        "%i/%i",
        sum({{ var }} != ref, na.rm = TRUE),
        sum(!is.na({{ var }}))
      ),
      .by = c({{ trt }}, vaxage, day)
    )
  p <- ggplot(
    pdat |> filter({{ var }} != ref),
    aes(interaction(day, to_factor({{ trt }}), sep = "!"), p)
  ) +
    facet_wrap(~ fct_na_value_to_level(to_factor(vaxage)), ncol = 2) +
    geom_col(aes(fill = to_factor({{ var }})), width = 0.8) +
    geom_text(
      data = plab,
      aes(label = lab),
      vjust = -1,
      size = 1.5,
      family = "TeX Gyre Pagella"
    ) +
    scale_fill_viridis_d(slab) +
    scale_x_discrete(
      guide = legendry::guide_axis_nested(key = key_range_auto(sep = "!")),
      name = "Treatment and day following vaccination"
    ) +
    scale_y_continuous("Proportion", limit = c(0, ymax), expand = c(0, 0)) +
    labs(y = "Proportion") +
    theme(legend.position = "bottom", panel.grid.major.x = element_blank())
  return(p)
}

plot_sys_reaction <- function(diary_sol, x, ymax = 1) {
  p <- diary_sol |>
    mutate(intensity = ordered(as.numeric(intensity) - 1)) |>
    filter(term == x) |>
    plot_react(trt, intensity, x, "0", ymax = ymax) +
    geom_vline(xintercept = 7.5, linetype = 1, colour = "grey80") +
    labs(
      caption = paste0(
        "Grade 1: easily tolerated\n",
        "Grade 2: interferes with normal everyday activities\n",
        "Grade 3: prevents normal everyday activities or requires significant medical intervention\n"
      )
    ) +
    theme(
      plot.caption = element_text(hjust = 0, size = rel(0.5), lineheight = 1)
    )
  return(p)
}

plot_isr <- function(
  dat,
  trt = treatment,
  var = size_cat,
  ref = 0,
  ylim = 1,
  var_lab = "Size (mm)"
) {
  library(ggtext)
  make_plot_data <- function(dat, trt = trt, var = var, ref = ref) {
    pdat <- dat |>
      group_by({{ trt }}, vaxage, vaccine, day) |>
      count({{ var }}) |>
      mutate(
        p = n / if_else(is.na({{ var }}), sum(n), sum(n[!is.na({{ var }})]))
      ) |>
      ungroup() |>
      complete(
        {{ trt }},
        nesting(vaxage, vaccine),
        day,
        {{ var }},
        fill = list(n = 0, p = 0)
      ) |>
      to_factor()
    plab <- dat |>
      summarise(
        n = sum({{ var }} != ref, na.rm = TRUE),
        N = sum(!is.na({{ var }})),
        p = mean({{ var }} != ref, na.rm = TRUE),
        # lab = paste0("frac(", n, ",", N, ")"),
        # lab = paste0("<sup>", n, "</sup>&frasl;<sub>", N, "</sub>"),
        lab = sprintf("%i/%i", n, N),
        .by = c({{ trt }}, vaxage, vaccine, day)
      )
    return(list(pdat, plab))
  }
  out <- make_plot_data(dat, {{ trt }}, {{ var }}, ref)
  pdat <- out[[1]]
  plab <- out[[2]]

  design <- "
  AB#
  CD#
  E##
  FGH
  "

  base_plot <- function(pdat, plab) {
    ggplot(
      pdat |> filter({{ var }} != ref),
      aes(interaction(day, {{ trt }}, sep = "!"), p)
    ) +
      # For landscape plot
      facet_wrap2(
        vaxage ~ vaccine,
        strip = strip_nested(
          background_x = elem_list_rect(
            fill = c("grey90", "grey80"),
            colour = c("white", "grey20")
          ),
          by_layer_x = TRUE
        ),
        ncol = 4
      ) +
      # For vertical plot
      # facet_manual(
      #   vaxage ~ vaccine,
      #   strip = strip_nested(
      #     background_x = elem_list_rect(
      #       fill = c("grey90", "grey80"),
      #       colour = c("white", "grey20")
      #     ),
      #     by_layer_x = TRUE
      #   ),
      #   design = design
      # ) +
      geom_rect(
        data = pdat |>
          filter({{ var }} != ref, vaccine == "aP/wP"),
        xmin = -Inf,
        xmax = Inf,
        ymin = -Inf,
        ymax = Inf,
        fill = "red",
        alpha = 0.005
      ) +
      geom_col(aes(fill = {{ var }})) +
      geom_vline(xintercept = 7.5, colour = "grey10") +
      # geom_richtext(
      #   data = plab,
      #   aes(label = lab),
      #   fill = NA,
      #   label.color = NA,
      #   size = 1.4,
      #   vjust = 0,
      #   family = "serif"
      # ) +
      geom_text(
        data = plab,
        aes(label = lab),
        vjust = -0.5,
        size = 1.4,
        parse = TRUE,
        family = "serif"
      ) +
      scale_fill_viridis_d(var_lab) +
      scale_x_discrete(
        guide = legendry::guide_axis_nested(key_range_auto(sep = "!")),
        name = "Day following vaccination"
      ) +
      scale_y_continuous("Proportion", limit = c(0, ylim), expand = c(0, 0)) +
      theme(
        panel.grid.major.x = element_blank(),
        legend.position = "bottom",
        legend.margin = margin(t = 0, unit = "cm"),
        legend.key.size = unit(1, "line"),
        plot.title = element_text(size = rel(1)),
        text = element_text(size = 9)
      )
  }
  g <- base_plot(pdat, plab)
  return(g)
}

st1_diary <- read_parquet(file.path(st1_pth, "diary-temperature.parquet"))
st2_diary <- select_form(dat_raw, "diary")
```

```{r}
tab_dat <- st1_diary |>
  to_factor() |>
  select(subjectid, vaxage, diaretyn) |>
  filter(!is.na(vaxage)) |>
  distinct() |>
  mutate(stage = "Stage 1") |>
  mutate(subjectid = gsub("^01", "01-", subjectid)) |>
  left_join(select(dat_base, subjid, trt), join_by(subjectid == subjid)) |>
  rename(record_id = subjectid) |>
  bind_rows(
    st2_diary |>
      filter(reaction == "shared") |>
      pull(data) |>
      pluck(1) |>
      select(record_id, vaxage, diaretyn) |>
      distinct() |>
      mutate(
        diaretyn = if_else(!diaretyn | is.na(diaretyn), "No", "Yes"),
        stage = "Stage 2"
      ) |>
      left_join(select(dat_base, record_id, trt), join_by(record_id))
  ) |>
  mutate(
    vaxage = factor(
      vaxage,
      levels = c("6-week", "4-month", "6-month", "18-month")
    )
  )
```

```{r}
#| label: tbl-diary-returned-stage1
#| tbl-cap: |
#|   Number and proportion of parents from stage 1 who returned diary cards following each vaccination occasion.
#| tbl-pos: "H"
#| results: "asis"
tab_dat |>
  filter(stage == "Stage 1") |>
  tbl_strata2(
    strata = vaxage,
    .tbl_fun = ~ .x |>
      tbl_summary(
        include = diaretyn,
        by = trt,
        label = list(diaretyn = .y)
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header(contains("stat_")),
    .combine_with = "tbl_stack",
    .combine_args = list(group_header = NULL)
  ) |>
  modify_header(label = "**Vaccination**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(75),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-diary-returned-stage2
#| tbl-cap: |
#|   Number and proportion of parents from stage 2 who returned diary cards following each vaccination occasion.
#| tbl-pos: "H"
#| results: "asis"
tab_dat |>
  filter(stage == "Stage 2") |>
  tbl_strata2(
    strata = vaxage,
    .tbl_fun = ~ .x |>
      tbl_summary(
        include = diaretyn,
        by = trt,
        label = list(diaretyn = .y)
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header(contains("stat_")),
    .combine_with = "tbl_stack",
    .combine_args = list(group_header = NULL)
  ) |>
  modify_header(label = "**Vaccination**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(75),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

## Parental Satisfaction

Following vaccination, parent's were asked whether they would be willing to have another child have the combination of immunisations at the given age. There were 3 occasions were parent disagreed or strongly disagreed.

```{r}
#| label: diary-likert-data
lik <- filter(dat_raw, form == "diary") |>
  pull(data) |>
  pluck(1) |>
  filter(reaction == "parental") |>
  pull(data) |>
  pluck(1) |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  mutate(
    vaxage = factor(
      vaxage,
      levels = c("6-week", "4-month", "6-month", "18-month")
    )
  )

st1_lik <- read_parquet(file.path(st1_pth, "diary-satisfaction.parquet")) |>
  select(subjectid, vaxage, diaimmscale) |>
  mutate(subjid = gsub("^01", "01-", subjectid)) |>
  to_factor() |>
  left_join(select(dat_base, subjid, trt), join_by(subjid)) |>
  mutate(
    diaimmscale = replace_values(diaimmscale, "Not completed" ~ NA_character_)
  )
st2_lik <- filter(st2_diary, reaction == "shared") |>
  pull(data) |>
  pluck(1) |>
  filter(vaxage == "6-week", vaxloc_id == 1) |>
  left_join(select(dat_base, record_id, subjid, trt), join_by(record_id)) |>
  select(subjid, trt, vaxage, diaimmscale) |>
  mutate(
    diaimmscale = factor(
      recode_values(
        diaimmscale,
        "Not completed" ~ NA_character_,
        "Strongly Disagree" ~ "Strongly disagree",
        "Disagree" ~ "Disagree",
        "Neither Agree or Disagree" ~ "Neither agree nor disagree",
        "Agree" ~ "Agree",
        "Strongly Agree" ~ "Strongly agree"
      ),
      levels = c(
        "Strongly disagree",
        "Disagree",
        "Neither agree nor disagree",
        "Agree",
        "Strongly agree"
      )
    )
  )

diary_likert <- bind_rows(st1_lik, st2_lik) |>
  mutate(
    diaimmscale = fct_drop(diaimmscale, only = "Not completed")
  )
```

```{r}
#| label: tbl-diary-likert
#| tbl-cap: |
#|   Number and percentage of parents giving each response to the given statement following 6-week vaccination, by assigned treatment group. Unknown is due to either a failure to return the diary card, or non-response to the item.
#| tbl-pos: "H"
#| output: "asis"
#| include: false
tbl_summary(
  diary_likert,
  by = trt,
  include = diaimmscale,
  label = list(
    diaimmscale ~ "I would be willing to have another child have this combination of immunisations at this age"
  )
) |>
  add_overall(last = TRUE) |>
  remove_footnote_header(contains("stat_")) |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  cols_width(label ~ pct(30)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = row_type == "label"
    )
  ) |>
  tab_options(
    table.width = pct(100),
    table.font.size = px(9),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```

```{r}
#| label: tbl-diary-likert-all
#| tbl-cap: |
#|   Number and percentage of parents giving each response to the given statement following 6-week, 4-month, 6-month, and 18-month vaccination, by assigned treatment group. Unknown is due to either a failure to return the diary card, or non-response to the item.
#| tbl-pos: "H"
#| output: "asis"
tbl_strata2(
  lik,
  strata = "vaxage",
  .tbl_fun = ~ . |>
    tbl_summary(
      by = trt,
      include = diaimmscale,
      label = list(
        diaimmscale ~ "I would be willing to have another child have this combination of immunisations at this age"
      )
    ) |>
    remove_footnote_header(contains("stat_")) |>
    modify_header(
      all_stat_cols() ~
        paste0(
          "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
          "N = {formatC(n, format='d', big.mark=',')}}}"
        )
    )
) |>
  as_gt() |>
  cols_width(label ~ pct(30)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = row_type == "label"
    )
  ) |>
  tab_options(
    table.width = pct(100),
    table.font.size = px(9),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```

\clearpage

## Solicited systemic reactions

```{r}
#| label: diary-sol-data
st1_sol <- read_parquet(file.path(
  st1_pth,
  "diary-solicited-reactions.parquet"
)) |>
  to_factor() |>
  filter(!is.na(vaxage)) |>
  mutate(subjid = gsub("^01", "01-", subjectid)) |>
  left_join(select(dat_base, subjid, trt), join_by(subjid)) |>
  rename(record_id = subjid) |>
  select(record_id, trt, vaxage, term, day, intensity) |>
  mutate(term = str_to_sentence(term), stage = 1)

st2_sol <- filter(st2_diary, reaction == "sol") |>
  pull(data) |>
  pluck(1) |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  select(record_id, trt, vaxage, term, day, intensity) |>
  mutate(day = as.integer(day), stage = 2)

diary_sol <- bind_rows(st1_sol, st2_sol) |>
  mutate(
    vaxage = factor(
      vaxage,
      levels = c("6-week", "4-month", "6-month", "18-month")
    )
  )

diary_sol_max <- diary_sol |>
  mutate(intensity = ordered(intensity)) |>
  summarise(
    .by = c(record_id, trt, vaxage, term),
    intensity_max = max(intensity)
  )

diary_sol_any <- diary_sol |>
  summarise(
    y = case_when(
      all(is.na(intensity)) ~ NA_real_,
      any(intensity == "Severe") ~ 1,
      TRUE ~ 0
    ),
    .by = c(record_id, trt, vaxage)
  ) |>
  summarise(
    n = sum(!is.na(y)),
    f = sum(y, na.rm = TRUE),
    p = mean(y, na.rm = TRUE),
    .by = c(trt, vaxage)
  ) |>
  pivot_wider(
    names_from = trt,
    values_from = c(f, n, p)
  )
```

The number and percentage of participants who experienced any severe reaction over the 7 days following vaccination, and the highest reaction grade by type over those 7 days are reported in @tbl-sol-intensity-any-severe, @tbl-sol-maximum-intensity, and @fig-sol-maximum-intensity. 

The subsequent sections report on the daily reaction grades following each vaccination occasion by the type of reaction.

```{r}
#| label: tbl-sol-intensity-any-severe
#| tbl-cap: |
#|   Experience of any severe intensity systemic reaction following vaccination (day 0 to 6) by assigned treatment.
#|   Excludes participants with missing diary cards or non-response for all days.
#| tbl-pos: "H"
out <- diary_sol_any |>
  # filter(vaxage %in% c("6-week", "18-month")) |>
  gt(rowname_col = "vaxage") |>
  fmt_percent(
    starts_with("p_"),
    decimals = 0
  ) |>
  cols_merge(
    c(f_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(f_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_label(
    "f_aP" = "aP",
    "f_wP" = "wP",
    .fn = md
  ) |>
  cols_align(align = "right", columns = c(f_aP, f_wP)) |>
  tab_options(table.font.size = px(9), table.width = pct(50)) |>
  as_latex()
# out_new <- paste("\\scriptsize", out, "\\normalsize", sep = "\n")
out_new <- gsub("aP", r"(\\makecell{aP\\\\N = 152})", out)
out_new <- gsub("wP", r"(\\makecell{wP\\\\N = 151})", out_new)
attributes(out_new) <- attributes(out)
out_new
```

```{r}
#| label: sol-intensity-max
pdat <- diary_sol_max |>
  select(record_id, vaxage, term, trt, intensity_max) |>
  count(vaxage, term, trt, intensity_max) |>
  group_by(vaxage, term, trt) |>
  mutate(
    p = n / sum(n[!is.na(intensity_max)]),
    lab = case_when(
      is.na(intensity_max) ~ sprintf("%i", n),
      TRUE ~ sprintf("%i (%.0f)", n, 100 * p)
    )
  ) |>
  ungroup() |>
  # filter(vaxage %in% c("6-week", "18-month")) |>
  mutate(
    intensity_max = recode_factor(
      intensity_max,
      "Mild" = "1",
      "Moderate" = "2",
      "Severe" = "3",
      .ordered = TRUE
    )
  )

plab <- diary_sol_max |>
  select(record_id, vaxage, term, trt, intensity_max) |>
  summarise(
    n = sum(intensity_max != "None", na.rm = TRUE),
    p = mean(intensity_max != "None", na.rm = TRUE),
    lab = sprintf(
      "%i/%i",
      sum(intensity_max != "None", na.rm = TRUE),
      sum(!is.na(intensity_max))
    ),
    .by = c(term, trt, vaxage)
  )
```

```{r}
#| label: tbl-sol-maximum-intensity
#| tbl-cap: |
#|   Highest intensity grade for systemic reactions after the 6-week and 18-month vaccine doses.
#| tbl-pos: "H"
tdat <- pdat |>
  select(-n, -p) |>
  pivot_wider(
    names_from = c(trt, intensity_max),
    values_from = lab,
    values_fill = list(lab = "0 (0)")
  )
out <- gt(tdat, rowname_col = "term", groupname_col = "vaxage") |>
  tab_spanner(
    label = "aP",
    columns = starts_with("aP")
  ) |>
  tab_spanner(
    label = "wP",
    columns = starts_with("wP")
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  cols_label(
    ends_with("_None") ~ "None",
    ends_with("_1") ~ "1",
    ends_with("_2") ~ "2",
    ends_with("_3") ~ "3",
    ends_with("_NA") ~ "Missing"
  ) |>
  tab_footnote(
    footnote = "For participants with non-responses for some days, the maximum intensity was taken as the greatest
    intensity reported on days with responses available. Percentages exclude participants with a missing highest grade (diary card not returned, or non-response for all days)."
  ) |>
  tab_options(table.font.size = px(9))
out
```

```{r}
#| label: fig-sol-maximum-intensity
#| fig-cap: |
#|    Highest intensity grade for systemic reactions after the 6-week and 18-month vaccination doses. *aP, acellular pertussis vaccine; wP, whole-cell pertussis vaccine; grade 1: easily tolerated; grade 2, interferes with normal everyday activities; grade 3, prevents normal everyday activities or requires significant medical intervention. Above each bar (n/N), we indicate the number of infants reporting systemic reactions grade 1 and above (numerator; n) versus the total number of infants reporting on systemic reactions (none, grade 1, grade 2, or grade 3; denominator; N).*
#| fig-height: 5
p <- pdat |>
  filter(!is.na(intensity_max), intensity_max != "None") |>
  ggplot(aes(interaction(trt, vaxage, sep = "!"), p)) +
  facet_wrap(~term, ncol = 2) +
  geom_col(aes(fill = intensity_max)) +
  geom_vline(xintercept = c(2.5, 4.5), colour = "grey80") +
  geom_text(
    data = plab,
    aes(label = lab),
    vjust = -0.5,
    size = 2,
    family = "TeX Gyre Pagella"
  ) +
  labs(
    fill = "Maximum grade",
    y = "Cumulative proportion"
  ) +
  scale_x_discrete(
    guide = legendry::guide_axis_nested(key = "!"),
    name = "Vaccination age and treatment"
  ) +
  ylim(0, 1) +
  theme(
    plot.caption = element_text(hjust = 0, size = rel(0.7), lineheight = 1),
    panel.grid.major.x = element_blank(),
    legend.position = "bottom",
    axis.title = element_text(size = rel(0.9))
  )
p
```

\clearpage

### Irritability

```{r}
#| label: fig-irritability-daily
#| fig-cap: |
#|   Irritability in the 7 days following the 6-week, 4-month, 6-month, and 18-month doses. *aP, acellular pertussis vaccine; wP, whole-cell pertussis vaccine; grade 1: easily tolerated; grade 2, interferes with normal everyday activities; grade 3, prevents normal everyday activities or requires significant medical intervention. Above each bar (n/N), we indicate the number of infants reporting systemic reactions grade 1 and above (numerator; n) versus the total number of infants reporting on systemic reactions (none, grade 1, grade 2, or grade 3; denominator; N).*
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Irritability")
```

\clearpage

### Vomiting

@fig-vomiting-daily presents the distribution of daily grades for vomiting following each vaccination occassion by assigned treatment.

```{r}
#| label: fig-vomiting-daily
#| fig-cap: |
#|   Vomiting in the 7 days following the 6-week, 4-month, 6-month, and 18-month doses. *aP, acellular pertussis vaccine; wP, whole-cell pertussis vaccine; grade 1: easily tolerated; grade 2, interferes with normal everyday activities; grade 3, prevents normal everyday activities or requires significant medical intervention. Above each bar (n/N), we indicate the number of infants reporting systemic reactions grade 1 and above (numerator; n) versus the total number of infants reporting on systemic reactions (none, grade 1, grade 2, or grade 3; denominator; N).*
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Vomiting")
```

\clearpage

### Diarrhoea

@fig-diarrhoea-daily presents the distribution of daily grades for diarrhoea following each vaccination occassion by assigned treatment.

```{r}
#| label: fig-diarrhoea-daily
#| fig-cap: |
#|   Diarrhoea in the 7 days following the 6-week, 4-month, 6-month, and 18-month doses. *aP, acellular pertussis vaccine; wP, whole-cell pertussis vaccine; grade 1: easily tolerated; grade 2, interferes with normal everyday activities; grade 3, prevents normal everyday activities or requires significant medical intervention. Above each bar (n/N), we indicate the number of infants reporting systemic reactions grade 1 and above (numerator; n) versus the total number of infants reporting on systemic reactions (none, grade 1, grade 2, or grade 3; denominator; N).*
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Diarrhoea")
```

\clearpage

### Decreased Feeding

@fig-decreasedfeeding-daily presents the distribution of daily intensity grades for decreased feeding following each vaccination occassion by assigned treatment.

```{r}
#| label: fig-decreasedfeeding-daily
#| fig-cap: |
#|   Decreased feeding in the 7 days following the 6-week, 4-month, 6-month, and 18-month doses. *aP, acellular pertussis vaccine; wP, whole-cell pertussis vaccine; grade 1: easily tolerated; grade 2, interferes with normal everyday activities; grade 3, prevents normal everyday activities or requires significant medical intervention. Above each bar (n/N), we indicate the number of infants reporting systemic reactions grade 1 and above (numerator; n) versus the total number of infants reporting on systemic reactions (none, grade 1, grade 2, or grade 3; denominator; N).*
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Decreased feeding")
```

\clearpage

### Drowsiness

@fig-drowsiness-daily presents the distribution of daily intensity grades for drowsiness following each vaccination occassion by assigned treatment.

```{r}
#| label: fig-drowsiness-daily
#| fig-cap: |
#|   Drowsiness in the 7 days following the 6-week, 4-month, 6-month, and 18-month doses. *aP, acellular pertussis vaccine; wP, whole-cell pertussis vaccine; grade 1: easily tolerated; grade 2, interferes with normal everyday activities; grade 3, prevents normal everyday activities or requires significant medical intervention. Above each bar (n/N), we indicate the number of infants reporting systemic reactions grade 1 and above (numerator; n) versus the total number of infants reporting on systemic reactions (none, grade 1, grade 2, or grade 3; denominator; N).*
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Drowsiness")
```

\clearpage

### Restlessness

@fig-restlessness-daily presents the distribution of daily intensity grades for restlessness following each vaccination occassion by assigned treatment.

```{r}
#| label: fig-restlessness-daily
#| fig-cap: |
#|   Restlessness in the 7 days following the 6-week, 4-month, 6-month, and 18-month doses. *aP, acellular pertussis vaccine; wP, whole-cell pertussis vaccine; grade 1: easily tolerated; grade 2, interferes with normal everyday activities; grade 3, prevents normal everyday activities or requires significant medical intervention. Above each bar (n/N), we indicate the number of infants reporting systemic reactions grade 1 and above (numerator; n) versus the total number of infants reporting on systemic reactions (none, grade 1, grade 2, or grade 3; denominator; N).*
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Restlessness")
```

\clearpage

### Fever

```{r}
#| label: diary-temp-data
st1_temp <- read_parquet(file.path(
  st1_pth,
  "diary-temperature.parquet"
)) |>
  to_factor() |>
  filter(!is.na(vaxage)) |>
  rename(temp_fac = temp_cat) |>
  mutate(subjid = gsub("^01", "01-", subjectid)) |>
  left_join(select(dat_base, subjid, trt), join_by(subjid)) |>
  rename(record_id = subjid) |>
  select(record_id, trt, vaxage, day, temp_fac) |>
  mutate(temp_fac = fct_recode(temp_fac, "38.0-38.4" = "38-38.4"))

st2_temp <- filter(st2_diary, reaction == "temp") |>
  pull(data) |>
  pluck(1) |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  select(record_id, trt, vaxage, day, temp_fac) |>
  mutate(day = as.integer(day))

diary_temp <- bind_rows(st1_temp, st2_temp) |>
  mutate(
    vaxage = factor(
      vaxage,
      levels = c("6-week", "4-month", "6-month", "18-month")
    )
  )
```

```{r}
#| label: tbl-fever-max
#| tbl-cap: Highest fever following each vaccination by assigned treatment.
#| tbl-pos: "H"
pdat <- diary_temp |>
  mutate(temp_fac = as.ordered(temp_fac)) |>
  summarise(
    temp_max = ordered(
      max(temp_fac, na.rm = TRUE),
      levels = levels(diary_temp$temp_fac)
    ),
    .by = c(record_id, trt, vaxage)
  ) |>
  mutate(
    temp_max = fct_na_value_to_level(
      fct_collapse(
        temp_max,
        "<38" = "None (<38)",
        "38.0-38.9" = c("38.0-38.4", "38.5-38.9"),
        "39.0-39.9" = c("39.0-39.4", "39.5-39.9"),
        "≥40.0" = c("40.0-40.4", "40.5-40.9", "41.0 or more"),
      ),
      "Missing"
    )
  ) |>
  count(vaxage, trt, temp_max) |>
  group_by(vaxage, trt) |>
  mutate(
    p = n / sum(n[temp_max != "Missing"]),
    lab = case_when(
      temp_max == "Missing" ~ sprintf("%i", n),
      TRUE ~ sprintf("%i (%.0f)", n, 100 * p)
    )
  ) |>
  ungroup()
tdat <- pdat |>
  complete(temp_max, vaxage, trt, fill = list(lab = "0 (0)")) |>
  select(-n, -p) |>
  arrange(temp_max) |>
  pivot_wider(
    names_from = c(trt, temp_max),
    values_from = lab,
    values_fill = list(lab = "0 (0)")
  )
out <- gt(tdat, rowname_col = "vaxage") |>
  tab_spanner(
    label = "aP",
    columns = starts_with("aP")
  ) |>
  tab_spanner(
    label = "wP",
    columns = starts_with("wP")
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  cols_label_with(
    fn = \(x) gsub("aP_", "", x)
  ) |>
  cols_label_with(
    fn = \(x) gsub("wP_", "", x)
  ) |>
  tab_options(table.font.size = px(9)) |>
  as_latex()
out_new <- gsub(
  "aP",
  r"(\\makecell{aP\\\\N = 152 (6-week, 18-month)\\\\N = 75 (4-month, 6-month)})",
  out
)
out_new <- gsub(
  "wP",
  r"(\\makecell{wP\\\\N = 151 (6-week, 18-month)\\\\N = 75 (4-month, 6-month)})",
  out_new
)
attributes(out_new) <- attributes(out)
out_new
```

```{r}
#| label: fig-fever
#| fig-cap: Fever by vaccination age and assigned treatment, stage 1 and 2 combined.
plot_fever <- function() {
  p <- diary_temp |>
    plot_react(trt, temp_fac, "Temperature", "None (<38)") +
    ylim(0, 0.1)
  return(p)
}
plot_fever()
```

\clearpage

## Solicited Injection Site Reactions

The number and percentage of participants who experienced any severe injection site reaction over the 7 days following vaccination, and the highest reaction size by type over those 7 days are reported in @tbl-sol-intensity-any-severe, @tbl-maximum-site-reaction, and @fig-maximum-site-reaction. 

The subsequent sections report on the daily injection site reaction sizes following each vaccination occasion by the type of reaction.

```{r}
#| label: data-isr
st1_isr <- read_parquet(file.path(st1_pth, "diary-site-reactions.parquet")) |>
  to_factor() |>
  mutate(
    vaccine = fct_recode(
      vaccine,
      "aP/wP" = "aP or wP",
      "MMR" = "MMRV",
      "DTPa-hepB-IPV-Hib" = "DTaP-Hib-HepB-IPV",
      "PCV-13" = "13vPCV"
    )
  ) |>
  mutate(subjid = gsub("^01", "01-", subjectid)) |>
  left_join(select(dat_base, subjid, trt), join_by(subjid)) |>
  rename(record_id = subjid, size_fac = size_cat) |>
  select(record_id, trt, term, vaxage, vaccine, day, size, size_fac) |>
  mutate(stage = 1)

st2_isr <- st2_diary |>
  filter(reaction == "shared") |>
  pull(data) |>
  pluck(1) |>
  complete(record_id, nesting(vaxage, vaccine)) |>
  filter(!is.na(vaccine)) |>
  left_join(
    st2_diary |>
      filter(reaction == "isr") |>
      pull(data) |>
      pluck(1),
    join_by(record_id, visit, vaxage, vaxloc_id)
  ) |>
  complete(record_id, nesting(vaxage, vaccine), day, term) |>
  filter(!is.na(term), !is.na(day)) |>
  mutate(
    day = as.integer(day),
    vaccine = fct_recode(vaccine, "PCV-13" = "PCv13")
  ) |>
  select(record_id, vaxage, term, vaccine, day, size, size_fac) |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  mutate(stage = 2)

# st2_isr <- st2_diary |>
#   filter(reaction == "isr") |>
#   pull(data) |>
#   pluck(1) |>
#   left_join(
# ,
#     join_by(record_id, visit, vaxage, vaxloc_id)
#   ) |>
#   mutate(
#     day = as.integer(day),
#     vaccine = fct_recode(vaccine, "PCV-13" = "PCv13")
#   ) |>
#   select(record_id, vaxage, term, vaccine, day, size, size_fac) |>
#   left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
#   mutate(stage = 2) |>
#   filter(!is.na(vaccine))

diary_isr <- bind_rows(st1_isr, st2_isr) |>
  mutate(
    vaxage = factor(
      vaxage,
      levels = c("6-week", "4-month", "6-month", "18-month")
    )
  )
```

```{r}
#| label: tbl-maximum-site-reaction
#| tbl-cap: |
#|    Maximum injection site reaction size in the 7 days following vaccination,
#|    by vaccination age, vaccine, and assigned treatment.
#| tbl-pos: "H"
pdat <- diary_isr |>
  mutate(size_fac = ordered(size_fac)) |>
  summarise(
    size_max = fct_na_value_to_level(
      suppressWarnings(max(size_fac, na.rm = TRUE)),
      "Missing"
    ),
    .by = c(record_id, term, trt, vaxage, vaccine)
  ) |>
  count(term, vaxage, vaccine, trt, size_max) |>
  complete(term, nesting(vaxage, vaccine), trt, size_max, fill = list(n = 0)) |>
  mutate(
    p = n / sum(n[size_max != "Missing"]),
    lab = case_when(
      size_max == "Missing" ~ sprintf("%i", n),
      TRUE ~ sprintf("%i (%.0f)", n, 100 * p)
    ),
    .by = c(term, vaxage, vaccine, trt)
  )
plab <- diary_isr |>
  mutate(size_fac = ordered(size_fac)) |>
  mutate(
    size_max = suppressWarnings(max(size_fac, na.rm = TRUE)),
    .by = c(record_id, term, trt, vaxage, vaccine)
  ) |>
  select(record_id, vaxage, vaccine, term, trt, size_max) |>
  distinct() |>
  summarise(
    n = sum(size_max != "None", na.rm = TRUE),
    p = mean(size_max != "None", na.rm = TRUE),
    lab = sprintf("%i/%i", n, sum(!is.na(size_max))),
    .by = c(term, vaccine, trt, vaxage)
  )
tdat <- pdat |>
  select(-n, -p) |>
  pivot_wider(
    names_from = c(trt, size_max),
    values_from = lab,
    values_fill = list(lab = "0 (0)")
  )
out <- gt(tdat, rowname_col = "term", groupname_col = c("vaxage", "vaccine")) |>
  tab_spanner(
    id = "A",
    label = "aP",
    columns = starts_with("aP")
  ) |>
  tab_spanner(
    id = "B",
    label = "wP",
    columns = starts_with("wP")
  ) |>
  tab_spanner(
    label = "Size (mm)",
    spanners = c("A", "B")
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  cols_label(
    ends_with("_None") ~ "None",
    ends_with("_>0 to 10") ~ ">0--10",
    ends_with("_>10 to 25") ~ ">10--25",
    ends_with("_>25 to 50") ~ ">25--50",
    ends_with("_>50") ~ ">50",
    ends_with("_Missing") ~ "Missing"
  ) |>
  tab_footnote(
    footnote = "For participants with non-responses for some days, the maximum size is taken as the largest size reported on days with responses available. Percentages exclude participants with a missing highest grade (diary card not returned, or non-response for all days)."
  ) |>
  tab_options(
    table.font.size = px(7),
    table.width = pct(80)
  ) |>
  as_latex()
out_new <- gsub("\\{aP\\}", r"({\\makecell{aP\\\\(n = 152)}})", out)
out_new <- gsub("\\{wP\\}", r"({\\makecell{wP\\\\(n = 151)}})", out_new)
attributes(out_new) <- attributes(out)
out_new
```

```{r}
#| label: fig-maximum-site-reaction
#| fig-cap: |
#|   Maximum injection site reaction size in 7 days following each vaccination,
#|   by vaccination age, vaccine, and assigned treatment.
#|   Stage 1: 6-week, 4-month, 6-month, 18-month.
#|   Stage 2: 6-week, 18-month.
#| fig-width: 7
#| fig-height: 4
p <- pdat |>
  filter(size_max != "None", size_max != "Missing") |>
  ggplot(aes(interaction(trt, term, sep = "!"), p)) +
  facet_wrap2(
    vaxage ~ vaccine,
    strip = strip_nested(
      background_x = elem_list_rect(
        fill = c("grey90", "grey80"),
        colour = c("white", "grey20")
      ),
      by_layer_x = TRUE
    ),
    ncol = 4
  ) +
  geom_rect(
    data = pdat |>
      filter(size_max != "None", size_max != "Missing", vaccine == "aP/wP"),
    xmin = -Inf,
    xmax = Inf,
    ymin = -Inf,
    ymax = Inf,
    fill = "red",
    alpha = 0.005
  ) +
  geom_col(aes(fill = size_max)) +
  geom_text(
    data = plab,
    aes(label = lab),
    vjust = -1,
    size = 1.8,
    family = "TeX Gyre Pagella"
  ) +
  scale_fill_viridis_d("Maximum size (mm)") +
  scale_x_discrete(
    guide = legendry::guide_axis_nested("!"),
    name = "Term and treatment"
  ) +
  ylim(0, 1) +
  labs(y = "Cumulative proportion") +
  theme(axis.title = element_text(size = rel(0.9)), legend.position = "top")
p
```

\clearpage

\newgeometry{top=15mm, bottom=25mm}

::: {.landscape}

### Erythema

```{r}
#| label: fig-redness-daily
#| fig-cap: |
#|    Erythema in the 7 days following the 6-week, 4-month, 6-month, and 18-month vaccinations. *aP, acellular pertussis vaccine; wP, whole-cell pertussis vaccine. Above each bar (n/N), we indicate the number of infants reporting any injection site reaction (numerator; n) versus the total number of infants reporting on injection site reactions (denominator; N).*
#| fig-width: 10
#| fig-height: 4
plot_isr(
  diary_isr |> filter(term == "Erythema"),
  trt = trt,
  var = size_fac,
  ref = "None",
  var_lab = "Redness size (mm)"
)
```

:::

\clearpage

::: {.landscape}

### Swelling

```{r}
#| label: fig-swelling-daily
#| fig-cap: |
#|    Swelling in the 7 days following the 6-week, 4-month, 6-month, and 18-month vaccinations. *aP, acellular pertussis vaccine; wP, whole-cell pertussis vaccine. Above each bar (n/N), we indicate the number of infants reporting any injection site reaction (numerator; n) versus the total number of infants reporting on injection site reactions (denominator; N).*
#| fig-width: 10
#| fig-height: 4
plot_isr(
  diary_isr |> filter(term == "Swelling"),
  trt = trt,
  var = size_fac,
  ref = "None",
  var_lab = "Swelling size (mm)"
)
```

:::

\clearpage

::: {.landscape}

### Induration

```{r}
#| label: fig-induration-daily
#| fig-cap:  |
#|    Induration in the 7 days following the 6-week, 4-month, 6-month, and 18-month vaccinations. *aP, acellular pertussis vaccine; wP, whole-cell pertussis vaccine. Above each bar (n/N), we indicate the number of infants reporting any injection site reaction (numerator; n) versus the total number of infants reporting on injection site reactions (denominator; N).*
#| fig-width: 10
#| fig-height: 4
plot_isr(
  diary_isr |> filter(term == "Induration"),
  trt = trt,
  var = size_fac,
  ref = "None",
  var_lab = "Induration size (mm)"
)
```

:::

\clearpage

\restoregeometry

### Pain

```{r}
#| label: diary-pain-data
st1_pain <- read_parquet(file.path(st1_pth, "diary-pain.parquet")) |>
  to_factor() |>
  mutate(
    vaccine = fct_recode(
      vaccine,
      "aP/wP" = "aP or wP",
      "MMR" = "MMRV",
      "DTPa-hepB-IPV-Hib" = "DTaP-Hib-HepB-IPV",
      "PCV-13" = "13vPCV"
    )
  ) |>
  mutate(subjid = gsub("^01", "01-", subjectid)) |>
  left_join(select(dat_base, subjid, trt), join_by(subjid)) |>
  rename(record_id = subjid) |>
  select(record_id, trt, vaxage, vaccine, day, pain) |>
  mutate(stage = 1)
st2_pain <- st2_diary |>
  filter(reaction == "pain") |>
  pull(data) |>
  pluck(1) |>
  left_join(
    st2_diary |>
      filter(reaction == "shared") |>
      pull(data) |>
      pluck(1),
    join_by(record_id, visit, vaxage, vaxloc_id)
  ) |>
  mutate(
    day = as.integer(day),
    vaccine = fct_recode(vaccine, "PCV-13" = "PCv13")
  ) |>
  select(record_id, vaxage, vaccine, day, pain) |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  mutate(stage = 2) |>
  filter(!is.na(vaccine))

# Fill in missing records
# ids <- st2_diary |>
#   filter(reaction == "pain") |>
#   pull(data) |>
#   pluck(1) |> pull(record_id) |> unique()
# st2_pain <- st2_pain |>
#   complete(record_id = ids, nesting(vaxage, vaccine), day)

diary_pain <- bind_rows(st1_pain, st2_pain) |>
  mutate(
    vaxage = factor(
      vaxage,
      levels = c("6-week", "4-month", "6-month", "18-month")
    )
  )
```


```{r}
#| label: tbl-maximum-pain-reaction
#| tbl-cap: |
#|    Maximum injection site pain in the 7 days following vaccination,
#|    by vaccination age, vaccine, and assigned treatment.
#| tbl-pos: "H"
pdat <- diary_pain |>
  mutate(pain = ordered(pain)) |>
  summarise(
    pain_max = fct_na_value_to_level(
      suppressWarnings(max(pain, na.rm = TRUE)),
      "Missing"
    ),
    .by = c(record_id, trt, vaxage, vaccine)
  ) |>
  count(vaxage, vaccine, trt, pain_max) |>
  complete(nesting(vaxage, vaccine), trt, pain_max, fill = list(n = 0)) |>
  mutate(
    p = n / sum(n[pain_max != "Missing"]),
    lab = case_when(
      pain_max == "Missing" ~ sprintf("%i", n),
      TRUE ~ sprintf("%i (%.0f)", n, 100 * p)
    ),
    .by = c(vaxage, vaccine, trt)
  )
plab <- diary_pain |>
  mutate(pain = ordered(pain)) |>
  mutate(
    pain_max = suppressWarnings(max(pain, na.rm = TRUE)),
    .by = c(record_id, trt, vaxage, vaccine)
  ) |>
  select(record_id, vaxage, vaccine, trt, pain_max) |>
  distinct() |>
  summarise(
    n = sum(pain_max != "None", na.rm = TRUE),
    p = mean(pain_max != "None", na.rm = TRUE),
    lab = sprintf("%i/%i", n, sum(!is.na(pain_max))),
    .by = c(vaccine, trt, vaxage)
  )
tdat <- pdat |>
  select(-n, -p) |>
  pivot_wider(
    names_from = c(trt, pain_max),
    values_from = lab,
    values_fill = list(lab = "0 (0)")
  )
out <- gt(tdat, rowname_col = "vaccine", groupname_col = "vaxage") |>
  tab_spanner(
    id = "A",
    label = "aP",
    columns = starts_with("aP")
  ) |>
  tab_spanner(
    id = "B",
    label = "wP",
    columns = starts_with("wP")
  ) |>
  tab_spanner(
    label = "Pain",
    spanners = c("A", "B")
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  cols_label(
    ends_with("_None") ~ "None",
    ends_with("_Mild") ~ "Mild",
    ends_with("_Moderate") ~ "Moderate",
    ends_with("_Severe") ~ "Severe",
    ends_with("_Missing") ~ "Missing"
  ) |>
  tab_footnote(
    footnote = "For participants with non-responses for some days, the maximum pain is taken as the greatest pain grade reported on days with responses available. Percentages exclude participants with a missing highest grade (diary card not returned, or non-response for all days)."
  ) |>
  tab_options(table.font.size = px(8)) |>
  as_latex()
out_new <- gsub("\\{aP\\}", r"({\\makecell{aP\\\\(n = 152)}})", out)
out_new <- gsub("\\{wP\\}", r"({\\makecell{wP\\\\(n = 151)}})", out_new)
attributes(out_new) <- attributes(out)
out_new
```

```{r}
#| label: fig-maximum-site-pain
#| fig-cap: |
#|    Maximum injection site pain in the 7 days following vaccination,
#|    by vaccination age, vaccine, and assigned treatment.
#| fig-width: 7
#| fig-height: 4
p <- pdat |>
  filter(pain_max != "None", pain_max != "Missing") |>
  ggplot(aes(trt, p)) +
  facet_wrap2(
    vaxage ~ vaccine,
    strip = strip_nested(
      background_x = elem_list_rect(
        fill = c("grey90", "grey80"),
        colour = c("white", "grey20")
      ),
      by_layer_x = TRUE
    ),
    ncol = 4
  ) +
  geom_rect(
    data = pdat |>
      filter(pain_max != "None", pain_max != "Missing", vaccine == "aP/wP"),
    xmin = -Inf,
    xmax = Inf,
    ymin = -Inf,
    ymax = Inf,
    fill = "red",
    alpha = 0.01
  ) +
  geom_col(aes(fill = pain_max)) +
  geom_text(
    data = plab,
    aes(label = lab),
    vjust = -1,
    size = 2,
    family = "TeX Gyre Pagella"
  ) +
  scale_fill_viridis_d("Maximum pain") +
  scale_x_discrete(
    guide = legendry::guide_axis_nested("!"),
    name = "Term and treatment"
  ) +
  ylim(0, 1) +
  labs(y = "Cumulative proportion") +
  theme(axis.title = element_text(size = rel(0.9)), legend.position = "top")
p
```

\newgeometry{top=15mm, bottom=25mm}  

::: {.landscape}

```{r}
#| label: fig-pain-daily
#| fig-cap: |
#|   Pain in the 7 days following the 6-week, 4-month, 6-month, and 18-month vaccinations. *aP, acellular pertussis vaccine; wP, whole-cell pertussis vaccine. Above each bar (n/N), we indicate the number of infants reporting any injection site reaction (numerator; n) versus the total number of infants reporting on injection site reactions
#| fig-width: 10
#| fig-height: 4
plot_isr(diary_pain, trt, pain, ref = "None", var_lab = "Pain intensity")
```

:::

\clearpage

\restoregeometry
