# Immunological Outcomes

## IgE Concentrations

The 6 and 7 month IgE responses were previously reported. What follows additionally reports on 19-month IgE responses.

```{r}
source(here("reports", "R", "functions.R"))

dat_ige <- select_form(dat_raw, "ige")
dat_ige <- dat_base |>
  right_join(dat_ige, join_by(record_id, subjid)) |>
  select(
    record_id,
    subjid,
    rand_site,
    rand_stage,
    trt,
    gender,
    bfed,
    fborn,
    parinc_imp,
    ces,
    fha,
    visit_age,
    allergen,
    ige,
    comment
  ) |>
  mutate(
    any_obs_concentration = any(!is.na(ige)),
    all_obs_concentration = all(!is.na(ige)),
    .by = subjid
  )

multinorm <- cmdstan_model(here("stan", "multinorrm-cens.stan"))

make_stan_data <- function(
  dd,
  form = ~trt,
  var = log_ige,
  L = log10(c(0.005, 0.005, 0.0005))
) {
  tmp <- dd |>
    filter(row_number() == 1, .by = subjid)
  X <- model.matrix(form, data = tmp)
  X <- cbind(
    X[, 1:2, drop = FALSE],
    scale(X[, -(1:2), drop = FALSE], scale = F)
  )
  out <- list(
    N = nrow(tmp),
    J = 3,
    X = X,
    P = ncol(X),
    Y = dd |>
      select(subjid, visit_age, {{ var }}) |>
      spread(visit_age, {{ var }}) |>
      select(-subjid) |>
      mutate(across(everything(), ~ replace_na(.x, -99))) |>
      as.matrix(),
    R = dd |>
      mutate(
        R = case_when(
          is.na({{ var }}) ~ 1,
          {{ var }} == -Inf ~ 2,
          !is.na({{ var }}) ~ 0
        )
      ) |>
      select(subjid, visit_age, R) |>
      spread(visit_age, R) |>
      select(-subjid) |>
      as.matrix(),
    L = L
  )
  return(out)
}

gmr_table_dat <- function(gmr) {
  tdat <- gmr |>
    mutate(
      across(
        V1,
        list(
          mean = ~ mean(.x),
          sd = ~ sd(.x),
          med = ~ median(.x),
          lo = ~ quantile(.x, 0.025)[1, ],
          hi = ~ quantile(.x, 0.975)[1, ],
          pr = ~ Pr(.x < 1)
        )
      )
    ) |>
    select(-V1)
  return(tdat)
}

gmr_table <- function(tdat) {
  gt(tdat, rowname_col = "visit_age", groupname_col = "group") |>
    fmt_number(everything(), decimals = 2) |>
    cols_merge(
      c(V1_mean, V1_sd),
      pattern = "<<{1} $\\pm$ {2}>>"
    ) |>
    cols_merge(
      c(V1_lo, V1_hi),
      pattern = "<<({1}, {2})>>"
    ) |>
    cols_label(
      V1_mean = md("Mean &#177; std"),
      V1_med = "Median",
      V1_lo = md("95% CrI"),
      V1_pr = "Pr(< 1)",
    ) |>
    tab_stub_indent(
      rows = everything(),
      indent = 1
    ) |>
    tab_options(table.font.size = px(9))
}

make_ate <- function(dat, fit) {
  nd <- filter(dat, visit_age == "6-month")
  ndat <- expand_grid(
    visit_age = c("6-month", "7-month", "19-month"),
    trt = c("aP", "wP")
  )
  tmp <- brmsmargins(
    fit,
    newdata = nd,
    at = ndat,
    effects = "integrateoutRE",
    CI = 0.95,
    k = 100L,
    seed = 123,
    contrasts = cbind(
      "6-month_wP - aP" = c(-1, 1, 0, 0, 0, 0),
      "7-month_wP - aP" = c(0, 0, -1, 1, 0, 0),
      "19-month_wP - aP" = c(0, 0, 0, 0, -1, 1)
    )
  )
  r <- rvar(cbind(tmp$Posterior, tmp$Contrasts))
  names(r)[seq_len(nrow(ndat))] <- apply(ndat, 1, paste, collapse = "_")
  out <- as_tibble(t(r)) |>
    pivot_longer(
      everything(),
      names_to = c("visit_age", "trt"),
      names_sep = "_",
      values_to = "posterior"
    ) |>
    mutate(measure = "RD")
  return(out)
}


miss_pattern_trt_tab_ige <- function(dat, var = concentration) {
  pic <- c("X", "-")
  pal <- setNames(c("black", "red"), pic)
  tab <- dat |>
    to_factor() |>
    mutate(missing = pic[1 + as.integer(is.na({{ var }}))]) |>
    select(subjid, allergen, trt, visit_age, missing) |>
    pivot_wider(
      names_from = visit_age,
      values_from = missing,
      values_fill = "-"
    ) |>
    count(allergen, trt, `6-month`, `7-month`, `19-month`) |>
    arrange(trt, desc(n)) |>
    mutate(p = n / sum(n)) |>
    pivot_wider(
      names_from = "trt",
      values_from = c("n", "p"),
      values_fill = 0
    ) |>
    arrange(
      desc(`6-month`),
      desc(`7-month`),
      desc(`19-month`)
    )

  gt(tab, groupname_col = "allergen") |>
    fmt_percent(starts_with("p_"), decimals = 0) |>
    cols_merge(ends_with("_aP"), pattern = "<<{1} ({2})>>") |>
    cols_merge(ends_with("_wP"), pattern = "<<{1} ({2})>>") |>
    cols_label(
      `n_aP` = md("aP<br>(n = 75)"),
      `n_wP` = md("wP<br>(n = 75)")
    ) |>
    cols_align(align = "center", columns = 2:4) |>
    tab_spanner(
      label = "Count (%)",
      columns = 2:4
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(columns = everything())
    )
}
```

### Missingness

```{r}
#| label: tbl-overall-missingness-ige
#| tbl-cap: |
#|    Number and proportion of participants with at least one observed IgE response,
#|    and with all planned IgE responses observed.
#| tbl-pos: "H"
#| results: "asis"
dat_ige |>
  select(record_id, trt, any_obs_concentration, all_obs_concentration) |>
  distinct() |>
  tbl_summary(
    include = c(any_obs_concentration, all_obs_concentration),
    label = list(
      any_obs_concentration ~ "At least one IgE response available",
      all_obs_concentration ~ "All planned IgE responses available"
    ),
    by = trt
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = row_type == "label"
    )
  ) |>
  tab_options(
    table.font.size = px(8),
    table.width = pct(60)
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-igg-missingness-ige
#| tbl-cap:
#|    Missingness pattern of IgG concentrations for DTaP and Hib-PRP by study stage.
#| tbl-pos: H
miss_pattern_trt_tab_ige(filter(dat_ige), var = ige) |>
  tab_options(table.font.size = px(8), table.width = pct(50))
```

\clearpage

### Descriptive

Descriptive statistics are reported in @tbl-sample-ige-gmc.

```{r}
#| label: tbl-sample-ige-gmc
#| code-summary: Sample summaries
#| tbl-cap: |
#|    Sample IgE concentration summaries (kU/L) by allergen type, visit age,
#|    and assigned vaccination schedule.
#| tbl-pos: "H"
#| results: "asis"
tab <- dat_ige |>
  mutate(
    ige = if_else(ige == 0, 0.0005, ige)
  ) |>
  summarise(
    obs = sum(!is.na(ige)),
    gmc = 10^mean(log10(ige), na.rm = TRUE),
    min = min(ige, na.rm = TRUE),
    max = max(ige, na.rm = TRUE),
    det = mean(ige >= 0.01, na.rm = TRUE),
    con = mean(ige >= 0.1, na.rm = TRUE),
    .by = c(allergen, visit_age, trt)
  ) |>
  pivot_wider(
    names_from = trt,
    values_from = c(obs, gmc, min, max, det, con),
    names_vary = "slowest"
  ) |>
  to_factor() |>
  arrange(allergen, visit_age)

gt(tab, groupname_col = "allergen", rowname_col = "visit_age") |>
  fmt_integer(starts_with("obs")) |>
  fmt_number(starts_with("gmc"), decimals = 2) |>
  fmt_number(starts_with(c("min", "max")), decimals = 2) |>
  fmt_percent(starts_with("con"), decimals = 0) |>
  fmt_percent(starts_with("det"), decimals = 0) |>
  cols_merge(
    columns = c(min_aP, max_aP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(min_wP, max_wP),
    pattern = "{1}–{2}"
  ) |>
  cols_label(
    obs_aP = "n",
    obs_wP = "n",
    gmc_aP = "GMC",
    gmc_wP = "GMC",
    min_aP = "Min–Max",
    min_wP = "Min–Max",
    det_aP = md("&ge;0.01"),
    det_wP = md("&ge;0.01"),
    con_aP = md("&ge;0.1"),
    con_wP = md("&ge;0.1")
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{aP} \\\\ N = 75}",
    columns = ends_with("_aP")
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{wP} \\\\ N = 75}",
    columns = ends_with("_wP")
  ) |>
  tab_footnote(
    footnote = "Values of 0.00 kU/L and 0.000 kU/L set to 0.005 and 0.0005 respectively for calculation of sample GMC.",
    locations = cells_column_labels(columns = starts_with("gmc"))
  ) |>
  tab_options(table.font.size = px(9)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

### Geometric mean concentration

Model estimated GMRs are reported in @tbl-gmr-gmfir-ige-total.
For completeness, all IgE types are reported, however, in some cases the GMR may not be meaningful due to the small number of participants with measurable quantities (e.g. tetanus toxoid at 6-months).

```{r}
#| label: model-ige-total
#| cache: true
fit_ige_model <- function(ige_ref) {
  age_labs <- c("6-month", "7-month", "19-month")

  mdat <- dat_ige |>
    filter(allergen == ige_ref) |>
    filter(!all(is.na(ige)), .by = subjid) |>
    mutate(log_ige = log10(ige))

  sdat1 <- make_stan_data(mdat, var = log_ige)
  sdat2 <- make_stan_data(
    mdat,
    var = log_ige,
    ~ trt + gender + bfed + fborn + parinc_imp + ces + fha
  )
  snk <- capture.output(
    sfit1 <- multinorm$sample(
      data = sdat1,
      seed = 71356,
      refresh = 0,
      chains = 8,
      iter_sampling = 1250
    )
  )
  snk <- capture.output(
    sfit2 <- multinorm$sample(
      data = sdat2,
      seed = 71356,
      refresh = 0,
      chains = 8,
      iter_warmup = 500,
      iter_sampling = 1250
    )
  )
  sdrw1 <- as_draws_rvars(sfit1)
  sdrw2 <- as_draws_rvars(sfit2)

  rownames(sdrw1$b) <- age_labs
  rownames(sdrw2$b) <- age_labs

  gmr1 <- 10^sdrw1$b[, 2]
  gmr2 <- 10^sdrw2$b[, 2]
  gmfir1 <- 10^(sdrw1$b[-1, 2] - sdrw1$b[-3, 2])
  gmfir2 <- 10^(sdrw2$b[-1, 2] - sdrw2$b[-3, 2])
  rownames(gmfir1) <- rownames(gmfir2) <- c(
    "7/6-month (GMFR)",
    "19/7-month (GMFR)"
  )

  # Fill missing observations with conditional predictions
  # yfill <- t(as_rvar(sdat1$Y))
  # yfill[which(t(sdat1$R == 1))] <- sdrw1$y_missing
  # yfill[which(t(sdat1$R == 2))] <- sdrw1$y_censored
  # yfill <- t(yfill)
  return(as_tibble(rbind(gmr2, gmfir2), rownames = "visit_age"))
}

res_total_ige <- fit_ige_model("Total IgE")
res_tetanus_ige <- fit_ige_model("Tetanus-toxoid")
res_eggwhite_ige <- fit_ige_model("Egg white")
res_wholeegg_ige <- fit_ige_model("Whole egg")

gmr <- bind_rows(
  mutate(res_total_ige, group = "Total IgE"),
  mutate(res_tetanus_ige, group = "Tetanus toxoid"),
  mutate(res_eggwhite_ige, group = "Egg white"),
  mutate(res_wholeegg_ige, group = "Whole egg")
)
gmr_dat <- gmr_table_dat(gmr)
```

```{r}
#| label: tbl-gmr-gmfir-ige-total
#| tbl-cap: Geometric mean ratio (GMR) and geometric mean fold rise ratio (GMFRR) for total IgE.
#| tbl-pos: "H"
#| results: "asis"
gmr_table(gmr_dat) |>
  tab_options(table.width = pct(60)) |>
  as_latex()
```

\clearpage

### Sensitisation

```{r}
#| label: model-ige-sens
#| cache: true
fit_ige_sens_model <- function(ige_ref, ref_val) {
  mdat <- dat_ige |>
    filter(allergen == ige_ref) |>
    filter(!is.na(ige)) |>
    mutate(sens = as.numeric(ige >= ref_val))

  glmm_prior <- c(
    prior(student_t(4, 0, 1.75), class = "b"),
    prior(student_t(4, 0, 1.75), class = "Intercept"),
    prior(exponential(1), class = "sd")
  )

  mfit1 <- brm(
    sens ~ visit_age * trt + (1 | subjid),
    data = mdat,
    family = bernoulli(),
    backend = "cmdstanr",
    prior = glmm_prior,
    seed = 71235,
    refresh = 0,
    chains = 8,
    iter = 1750,
    warmup = 500
  )

  mfit2 <- update(
    mfit1,
    formula = ~ . +
      visit_age *
        (gender +
          bfed +
          fborn +
          ces +
          fha +
          parinc_imp),
    newdata = mdat
  )

  # Marginal risk difference estimate
  dat_ige |>
    mutate(trt = "aP") |>
    add_epred_rvars(mfit2, value = "aP", allow_new_levels = TRUE) |>
    mutate(trt = "wP") |>
    add_epred_rvars(mfit2, value = "wP", allow_new_levels = TRUE) |>
    mutate(dif = wP - aP) |>
    summarise(across(c(aP, wP, dif), rvar_mean), .by = visit_age)

  # mdrws1 <- rvar(as_draws_matrix(mfit1))
  # beta1 <- mdrws1[grepl("(^b_Inter|^b_age|^b_tr)", names(mdrws1))]
  # omega1 <- mdrws1["sd_subjectid__Intercept"]
  # beta1_marg <- do.call(c, as_draws_rvars(t(marginalcoef(mfit1, posterior = TRUE)$Posterior)))
  # names(beta1_marg) <- names(beta1)
  # cor1 <- exp(c(beta1[3], beta1[3] + beta1[4]))
  # mor1 <- exp(c(beta1_marg[3], beta1_marg[3] + beta1_marg[4]))

  mdrws2 <- rvar(as_draws_matrix(mfit2))
  beta2 <- mdrws2[grepl("(^b_Inter|^b_visit_age|^b_tr)", names(mdrws2))]
  omega2 <- mdrws2["sd_subjid__Intercept"]
  beta2_marg <- do.call(
    c,
    as_draws_rvars(t(marginalcoef(mfit2, posterior = TRUE)$Posterior))
  )
  names(beta2_marg) <- variables(mfit2)[1:30]
  beta2_marg <- beta2_marg[grepl(
    "(^b_Inter|^b_visit_age|^b_tr)",
    names(beta2_marg)
  )]
  cor2 <- exp(c(beta2[4], beta2[4] + beta2[5], beta2[4] + beta2[6]))
  mor2 <- exp(c(
    beta2_marg[4],
    beta2_marg[4] + beta2_marg[5],
    beta2_marg[4] + beta2_marg[6]
  ))
  p1 <- make_ate(dat_ige, mfit1) |>
    pivot_wider(names_from = trt, values_from = posterior)
  p2 <- make_ate(dat_ige, mfit2) |>
    pivot_wider(names_from = trt, values_from = posterior)
  tdat2 <- p2 |>
    select(-measure) |>
    bind_cols(cor = cor2, mor = mor2) |>
    mutate(group = ige_ref)
  return(tdat2)
}

snk <- capture.output({
  res_sens_tetanus <- fit_ige_sens_model("Tetanus-toxoid", 0.1)
  res_sens_eggwhite <- fit_ige_sens_model("Egg white", 0.1)
  res_sens_wholeegg <- fit_ige_sens_model("Whole egg", 0.1)
})

res_sens <- bind_rows(res_sens_tetanus, res_sens_eggwhite, res_sens_wholeegg) |>
  mutate(age = fct_rev(fct_inorder(visit_age)), group = fct_inorder(group)) |>
  relocate(group, .before = visit_age)
```

```{r}
#| label: tbl-sensitisation-summary
#| tbl-cap: Summary of sensitisation model by response type.
#| tbl-pos: "H"
out <- res_sens |>
  transmute(
    group,
    age,
    across(
      3:7,
      list(
        med = ~ median(.x),
        lo = ~ quantile(.x, 0.025)[1, ],
        hi = ~ quantile(.x, 0.975)[1, ]
      )
    ),
    across(5, list(pr = ~ Pr(.x < 0)))
  ) |>
  relocate(`wP - aP_pr`, .after = `wP - aP_hi`) |>
  gt(rowname_col = "age", groupname_col = "group") |>
  fmt_number(-(1:2), decimals = 2) |>
  cols_merge(
    starts_with("wP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("aP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    c("wP - aP_med", "wP - aP_lo", "wP - aP_hi"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("cor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("mor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_label(
    "aP_med" = "aP",
    "wP_med" = "wP",
    "wP - aP_med" = "wP \u2212 aP",
    "wP - aP_pr" = "Pr(wP \u2212 aP < 0)",
    "cor_med" = "Conditional",
    "mor_med" = "Marginal"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    columns = c("aP_med", "wP_med", "wP - aP_med"),
    label = md("Standardised Pr(IgE $\\geq$ 0.1), Median (95% CrI)"),
  ) |>
  tab_spanner(
    columns = c("cor_med", "mor_med"),
    label = "Odds ratios"
  ) |>
  tab_options(table.font.size = px(9))
out
```

\clearpage

### Quantitation

```{r}
#| label: model-ige-quant
#| cache: true
snk <- capture.output({
  res_sens_tetanus <- fit_ige_sens_model("Tetanus-toxoid", 0.01)
  res_sens_eggwhite <- fit_ige_sens_model("Egg white", 0.01)
  res_sens_wholeegg <- fit_ige_sens_model("Whole egg", 0.01)
})

res_sens <- bind_rows(res_sens_tetanus, res_sens_eggwhite, res_sens_wholeegg) |>
  mutate(age = fct_rev(fct_inorder(visit_age)), group = fct_inorder(group)) |>
  relocate(group, .before = visit_age)
```

```{r}
#| label: tbl-sensitisation-tetanus-summary
#| tbl-cap: Summary of sensitisation model for tetanus toxoid IgE.
#| tbl-pos: "H"
out <- res_sens |>
  transmute(
    group,
    age,
    across(
      3:7,
      list(
        med = ~ median(.x),
        lo = ~ quantile(.x, 0.025)[1, ],
        hi = ~ quantile(.x, 0.975)[1, ]
      )
    ),
    across(5, list(pr = ~ Pr(.x < 0)))
  ) |>
  relocate(`wP - aP_pr`, .after = `wP - aP_hi`) |>
  gt(rowname_col = "age", groupname_col = "group") |>
  fmt_number(-(1:2), decimals = 2) |>
  cols_merge(
    starts_with("wP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("aP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    c("wP - aP_med", "wP - aP_lo", "wP - aP_hi"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("cor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("mor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_label(
    "aP_med" = "aP",
    "wP_med" = "wP",
    "wP - aP_med" = "wP \u2212 aP",
    "wP - aP_pr" = "Pr(wP \u2212 aP < 0)",
    "cor_med" = "Conditional",
    "mor_med" = "Marginal"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    columns = c("aP_med", "wP_med", "wP - aP_med"),
    label = md("Standardised Pr(IgE $\\geq$ 0.1), Median (95% CrI)"),
  ) |>
  tab_spanner(
    columns = c("cor_med", "mor_med"),
    label = "Odds ratios"
  ) |>
  tab_options(table.font.size = px(9))
out
```

\clearpage

## IgG Concentrations

\clearpage
