# Immunological Outcomes

This section reports on data relating to the immunological outcomes.

```{r}
#| label: immunological-data
#| include: false
source(here("reports", "R", "functions.R"))

# Blood collection data
dat_bc <- select_form(dat_raw, "blood_collection") |>
  left_join(select(dat_base, record_id, subjid), join_by(record_id)) |>
  select(-record_id, -starts_with("lbsite___")) |>
  filter(visage != "6-month + 72hrs") |>
  arrange(subjid, visage)

# Check for date inconsistencies
dat_dates <- dat_grid_vs |>
  select(record_id, subjid, birthdat, randdat, visdat1, visage, visdat, windyn) |>
  filter(visage != "6-month + 72 hrs") |>
  left_join(dat_bc, join_by(subjid, visage)) |>
  mutate(
    age_months = interval(birthdat, visdat) %/% months(1),
    .after = visage
  ) |>
  mutate(
    .by = subjid,
    days_gap_visit = interval(lag(visdat), visdat) %/% days(1),
    .after = visdat
  ) |>
  mutate(
    .by = subjid,
    days_gap_lbipdat = interval(lag(lbipdat), lbipdat) %/% days(1),
    .after = lbipdat
  )
# dat_dates |>
#   filter(visdat != lbipdat)
# dat_dates |>
#   filter(
#     abs(age_months - as.numeric(gsub("-month", "", visage))) > 1
#   )

# Apply obvious corrections then derive gaps
# E.g. 7-month blood collection before 6-month, but 7-month visit date after
dat_bc <- dat_bc |>
  mutate(
    lbipdat = replace_when(
      lbipdat,
      subjid == "01-017" & visage == "7-month" ~ date("2018-12-04"),
      subjid == "01-094" & visage == "7-month" ~ date("2020-01-10")
    )
  ) |>
  mutate(
    .by = subjid,
    days_gap = interval(lag(lbipdat), lbipdat) %/% days(1)
  ) |>
  mutate(
    bc_window = case_when(
      !(visage %in% c("7-month", "19-month")) ~ NA,
      !lbipvst ~ NA,
      visage == "7-month" & between(days_gap, 21, 35) ~ TRUE,
      visage == "19-month" & between(days_gap, 21, 35) ~ TRUE,
      .default = FALSE
    )
  )

# IgG data
dat_igg <- select_form(dat_raw, "igg") |>
  left_join(dat_bc, join_by(subjid, visage)) |>
  left_join(dat_base, join_by(subjid)) |>
  select(
    record_id,
    subjid,
    rand_site,
    rand_stage,
    randdat,
    birthdat,
    trt,
    gender,
    bfed,
    fborn,
    parinc_imp,
    ces,
    fha,
    visage,
    type,
    antigen,
    units,
    ref,
    concentration,
    log_concentration,
    positive,
    lbipvst,
    lbncireas,
    lbvol,
    lbsite,
    lbipdat
  ) |>
  mutate(stage = if_else(substr(subjid, 1, 2) == "01", "1", "2")) |>
  filter(!(antigen %in% c("PnPs 11A"))) |>
  mutate(antigen = fct_inorder(antigen)) |>
  mutate(
    any_obs_concentration = any(!is.na(concentration)),
    all_obs_concentration = all(!is.na(concentration)),
    .by = subjid
  )

# IgE data
dat_ige <- select_form(dat_raw, "ige") |>
  left_join(dat_base, join_by(record_id, subjid)) |>
  select(
    record_id,
    subjid,
    rand_site,
    rand_stage,
    trt,
    gender,
    bfed,
    fborn,
    parinc_imp,
    ces,
    fha,
    visit_age,
    allergen,
    ige,
    comment
  ) |>
  mutate(
    fold_rise = if_else(visit_age == "7-month", ige / lag(ige), NA_real_),
    .by = c(subjid, allergen)
  ) |>
  mutate(
    any_obs_concentration = any(!is.na(ige)),
    all_obs_concentration = all(!is.na(ige)),
    .by = subjid
  )
```

```{r}
#| label: immunological-functions
multinorm <- cmdstan_model(here("stan", "multinorrm-cens.stan"))

make_stan_data <- function(
  dd,
  form = ~trt,
  var = log_ige,
  L = log10(c(0.005, 0.005, 0.0005))
) {
  tmp <- dd |>
    filter(row_number() == 1, .by = subjid)
  X <- model.matrix(form, data = tmp)
  X <- cbind(
    X[, 1:2, drop = FALSE],
    scale(X[, -(1:2), drop = FALSE], scale = F)
  )
  out <- list(
    N = nrow(tmp),
    J = 3,
    X = X,
    P = ncol(X),
    Y = dd |>
      select(subjid, visit_age, {{ var }}) |>
      spread(visit_age, {{ var }}) |>
      select(-subjid) |>
      mutate(across(everything(), ~ replace_na(.x, -99))) |>
      as.matrix(),
    R = dd |>
      mutate(
        R = case_when(
          is.na({{ var }}) ~ 1,
          {{ var }} == -Inf ~ 2,
          !is.na({{ var }}) ~ 0
        )
      ) |>
      select(subjid, visit_age, R) |>
      spread(visit_age, R) |>
      select(-subjid) |>
      as.matrix(),
    L = L
  )
  return(out)
}

gmr_table_dat <- function(gmr) {
  tdat <- gmr |>
    mutate(
      across(
        V1,
        list(
          mean = ~ mean(.x),
          sd = ~ sd(.x),
          med = ~ median(.x),
          lo = ~ quantile(.x, 0.025)[1, ],
          hi = ~ quantile(.x, 0.975)[1, ],
          pr = ~ Pr(.x < 1)
        )
      )
    ) |>
    select(-V1)
  return(tdat)
}

gmr_table <- function(tdat) {
  gt(tdat, rowname_col = "visit_age", groupname_col = "group") |>
    fmt_number(everything(), decimals = 2) |>
    cols_merge(
      c(V1_mean, V1_sd),
      pattern = "<<{1} $\\pm$ {2}>>"
    ) |>
    cols_merge(
      c(V1_lo, V1_hi),
      pattern = "<<({1}, {2})>>"
    ) |>
    cols_label(
      V1_mean = md("Mean &#177; std"),
      V1_med = "Median",
      V1_lo = md("95% CrI"),
      V1_pr = "Pr(< 1)",
    ) |>
    tab_stub_indent(
      rows = everything(),
      indent = 1
    ) |>
    tab_options(table.font.size = px(9))
}

make_ate <- function(dat, fit) {
  nd <- filter(dat, visit_age == "6-month")
  ndat <- expand_grid(
    visit_age = c("6-month", "7-month", "19-month"),
    trt = c("aP", "wP")
  )
  tmp <- brmsmargins(
    fit,
    newdata = nd,
    at = ndat,
    effects = "integrateoutRE",
    CI = 0.95,
    k = 100L,
    seed = 123,
    contrasts = cbind(
      "6-month_wP - aP" = c(-1, 1, 0, 0, 0, 0),
      "7-month_wP - aP" = c(0, 0, -1, 1, 0, 0),
      "19-month_wP - aP" = c(0, 0, 0, 0, -1, 1)
    )
  )
  r <- rvar(cbind(tmp$Posterior, tmp$Contrasts))
  names(r)[seq_len(nrow(ndat))] <- apply(ndat, 1, paste, collapse = "_")
  out <- as_tibble(t(r)) |>
    pivot_longer(
      everything(),
      names_to = c("visit_age", "trt"),
      names_sep = "_",
      values_to = "posterior"
    ) |>
    mutate(measure = "RD")
  return(out)
}


miss_pattern_trt_tab_ige <- function(dat, var = concentration) {
  pic <- c("X", "-")
  pal <- setNames(c("black", "red"), pic)
  tab <- dat |>
    to_factor() |>
    mutate(missing = pic[1 + as.integer(is.na({{ var }}))]) |>
    select(subjid, allergen, trt, visit_age, missing) |>
    pivot_wider(
      names_from = visit_age,
      values_from = missing,
      values_fill = "-"
    ) |>
    count(allergen, trt, `6-month`, `7-month`, `19-month`) |>
    arrange(trt, desc(n)) |>
    mutate(p = n / sum(n)) |>
    pivot_wider(
      names_from = "trt",
      values_from = c("n", "p"),
      values_fill = 0
    ) |>
    arrange(
      desc(`6-month`),
      desc(`7-month`),
      desc(`19-month`)
    )

  gt(tab, groupname_col = "allergen") |>
    fmt_percent(starts_with("p_"), decimals = 0) |>
    cols_merge(ends_with("_aP"), pattern = "<<{1} ({2})>>") |>
    cols_merge(ends_with("_wP"), pattern = "<<{1} ({2})>>") |>
    cols_label(
      `n_aP` = md("aP<br>N = 75"),
      `n_wP` = md("wP<br>N = 75")
    ) |>
    cols_align(align = "center", columns = 2:4) |>
    tab_spanner(
      label = "Count (%)",
      columns = 2:4
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(columns = everything())
    )
}
```

## Blood collection

For participants enrolled under stage 1, blood collection was to occur at visit 3 (6 to <7-months), visit 5 (21 to 35 days after visit 3, i.e. at 7-months), visit 7 (18 to <19-months), and visit 8 (21 to 35 days after visit 7, i.e. at 19-months). For those enrolled under stage 2, blood collection was at 18 to <19-months and 21 to 35 days after that collection.

The number and percentage of infants with blood collected at each visit is reported in @tbl-blood-collection. Most collections occurred within the protocolised window but were late on occasion (@fig-blood-collection-gaps).

```{r}
#| label: tbl-blood-collection
#| tbl-cap: |
#|   Number and proportion of infants who had blood collected at the specified ages.
#| tbl-pos: "H"
#| results: "asis"
dat_bc |>
  mutate(lbipvst = if_else(is.na(lbipvst) | !lbipvst, FALSE, TRUE)) |>
  left_join(select(dat_base, subjid, rand_stage, trt), join_by(subjid)) |>
  summarise(
    .by = c(rand_stage, visage, trt),
    y = sum(lbipvst),
    n = n(),
    p = y / n,
    lab = sprintf("%i/%i (%.0f%%)", y, n, 100 * p)
  ) |>
  arrange(rand_stage, visage) |>
  select(-y, -n, -p) |>
  pivot_wider(names_from = trt, values_from = lab) |>
  mutate(rand_stage = paste("Stage ", rand_stage)) |>
  gt(groupname_col = "rand_stage") |>
  cols_label(
    visage ~ "Visit age"
  ) |>
  cols_label(
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    wP ~ "\\makecell{\\textbf{wP} \\\\ N = 75}"
  ) |>
  tab_options(table.font.size = px(9), table.width = pct(35)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: fig-blood-collection-gaps
#| fig-cap: |
#|   Gap times (days) between blood collection and previous visit for 7-month (21-35 days after 6-month vaccination) and 19-month (21-35 days after 18-month vaccination) based on documented visit dates. The protocolised windows of 21 to 35 days are indicated by dotted lines.
#| fig-height: 2.5
dat_bc |>
  filter(visage %in% c("7-month", "19-month")) |>
  ggplot(aes(days_gap)) +
  facet_wrap(~visage) +
  geom_histogram(boundary = 0.5, binwidth = 1) +
  geom_vline(xintercept = c(20.5, 35.5), linetype = 2) +
  scale_x_continuous(breaks = c(10, 15, 21, 28, 35, 40, 45, 50, 55, 60)) +
  labs(y = "Count", x = "Gap (days) between vaccination and blood collection")
```

\clearpage

## IgE concentrations {#sec-ige}

### Missingness

Even if blood collection occurred, IgE concentrations may have been missing (e.g. insufficient sample collected). The majority of participants had at least one IgE measurement and a small majority had all planned measurements (@tbl-overall-missingness-ige). Missingness varied by the specific allergen being tested (@tbl-igg-missingness-ige).

```{r}
#| label: tbl-overall-missingness-ige
#| tbl-cap: |
#|    Number and proportion of participants with at least one observed IgE response (from any visit or allergen),
#|    and with all planned IgE responses observed (all visits and allergens).
#| tbl-pos: "H"
#| results: "asis"
dat_ige |>
  select(record_id, trt, any_obs_concentration, all_obs_concentration) |>
  distinct() |>
  tbl_summary(
    include = c(any_obs_concentration, all_obs_concentration),
    label = list(
      any_obs_concentration ~ "At least one IgE response available",
      all_obs_concentration ~ "All planned IgE responses available"
    ),
    by = trt
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = row_type == "label"
    )
  ) |>
  tab_options(
    table.font.size = px(8),
    table.width = pct(60)
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-igg-missingness-ige
#| tbl-cap:
#|    Missingness pattern of IgE concentrations by allergen, "X" indicates a measurement was available and "-" means it was missing.
#| tbl-pos: H
miss_pattern_trt_tab_ige(filter(dat_ige), var = ige) |>
  tab_options(table.font.size = px(8), table.width = pct(50))
```

\clearpage

### Descriptive

Descriptive summaries of IgE concentrations (kU/L) are reported in @tbl-sample-ige-gmc including the number of observations (n), the geometric mean concentration (GMC), the range of concentrations (min-max), and the proportion of concentrations exceeding either at least 0.01 or 0.1 kU/L.

```{r}
#| label: tbl-sample-ige-gmc
#| code-summary: Sample summaries
#| tbl-cap: |
#|    Sample IgE concentration summaries (kU/L) by allergen type, visit age,
#|    and assigned vaccination schedule.
#| tbl-pos: "H"
#| results: "asis"
tab <- dat_ige |>
  mutate(
    ige = if_else(ige == 0, 0.0005, ige)
  ) |>
  summarise(
    obs = sum(!is.na(ige)),
    gmc = 10^mean(log10(ige), na.rm = TRUE),
    min = min(ige, na.rm = TRUE),
    max = max(ige, na.rm = TRUE),
    det = mean(ige >= 0.01, na.rm = TRUE),
    con = mean(ige >= 0.1, na.rm = TRUE),
    .by = c(allergen, visit_age, trt)
  ) |>
  pivot_wider(
    names_from = trt,
    values_from = c(obs, gmc, min, max, det, con),
    names_vary = "slowest"
  ) |>
  to_factor() |>
  arrange(allergen, visit_age)

gt(tab, groupname_col = "allergen", rowname_col = "visit_age") |>
  fmt_integer(starts_with("obs")) |>
  fmt_number(starts_with("gmc"), decimals = 2) |>
  fmt_number(starts_with(c("min", "max")), decimals = 2) |>
  fmt_percent(starts_with("con"), decimals = 0) |>
  fmt_percent(starts_with("det"), decimals = 0) |>
  cols_merge(
    columns = c(min_aP, max_aP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(min_wP, max_wP),
    pattern = "{1}–{2}"
  ) |>
  cols_label(
    obs_aP = "n",
    obs_wP = "n",
    gmc_aP = "GMC",
    gmc_wP = "GMC",
    min_aP = "Min–Max",
    min_wP = "Min–Max",
    det_aP = md("&ge;0.01"),
    det_wP = md("&ge;0.01"),
    con_aP = md("&ge;0.1"),
    con_wP = md("&ge;0.1")
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{aP} \\\\ N = 75}",
    columns = ends_with("_aP")
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{wP} \\\\ N = 75}",
    columns = ends_with("_wP")
  ) |>
  tab_footnote(
    footnote = "Values of 0.00 kU/L and 0.000 kU/L set to 0.005 and 0.0005 respectively for calculation of sample GMC.",
    locations = cells_column_labels(columns = starts_with("gmc"))
  ) |>
  tab_options(table.font.size = px(9)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

### Geometric mean concentration

Model estimated geometric mean ratios (GMRs) and geometric mean fold-rise ratios (GMFRRs) are reported in @tbl-gmr-gmfir-ige-total. For completeness, all IgE types are reported, however, in some cases the GMR and GMFRR may not be meaningful due to the small number of participants with measurable quantities (e.g. tetanus toxoid at 6-months and 7-months).

For total IgE, the GMFRR from 6 to 7 months was 0.94 (0.72, 1.22) and the probability of a reduction in the GMFR was 0.69.

```{r}
#| label: model-ige-total
#| cache: true
fit_ige_model <- function(ige_ref) {
  age_labs <- c("6-month", "7-month", "19-month")

  mdat <- dat_ige |>
    filter(allergen == ige_ref) |>
    filter(!all(is.na(ige)), .by = subjid) |>
    mutate(log_ige = log10(ige))

  sdat1 <- make_stan_data(mdat, var = log_ige)
  sdat2 <- make_stan_data(
    mdat,
    var = log_ige,
    ~ trt + gender + bfed + fborn + parinc_imp + ces + fha
  )
  snk <- capture.output(
    sfit1 <- multinorm$sample(
      data = sdat1,
      seed = 71356,
      refresh = 0,
      chains = 8,
      iter_sampling = 1250
    )
  )
  snk <- capture.output(
    sfit2 <- multinorm$sample(
      data = sdat2,
      seed = 71356,
      refresh = 0,
      chains = 8,
      iter_warmup = 500,
      iter_sampling = 1250
    )
  )
  sdrw1 <- as_draws_rvars(sfit1)
  sdrw2 <- as_draws_rvars(sfit2)

  rownames(sdrw1$b) <- age_labs
  rownames(sdrw2$b) <- age_labs

  gmr1 <- 10^sdrw1$b[, 2]
  gmr2 <- 10^sdrw2$b[, 2]
  gmfir1 <- 10^(sdrw1$b[-1, 2] - sdrw1$b[-3, 2])
  gmfir2 <- 10^(sdrw2$b[-1, 2] - sdrw2$b[-3, 2])
  rownames(gmfir1) <- rownames(gmfir2) <- c(
    "7/6-month (GMFR)",
    "19/7-month (GMFR)"
  )

  # Fill missing observations with conditional predictions
  # yfill <- t(as_rvar(sdat1$Y))
  # yfill[which(t(sdat1$R == 1))] <- sdrw1$y_missing
  # yfill[which(t(sdat1$R == 2))] <- sdrw1$y_censored
  # yfill <- t(yfill)
  return(as_tibble(rbind(gmr2, gmfir2), rownames = "visit_age"))
}

res_total_ige <- fit_ige_model("Total IgE")
res_tetanus_ige <- fit_ige_model("Tetanus-toxoid")
res_eggwhite_ige <- fit_ige_model("Egg white")
res_wholeegg_ige <- fit_ige_model("Whole egg")

gmr <- bind_rows(
  mutate(res_total_ige, group = "Total IgE"),
  mutate(res_tetanus_ige, group = "Tetanus toxoid"),
  mutate(res_eggwhite_ige, group = "Egg white"),
  mutate(res_wholeegg_ige, group = "Whole egg")
)
gmr_dat <- gmr_table_dat(gmr)
```

```{r}
#| label: tbl-gmr-gmfir-ige-total
#| tbl-cap: Geometric mean ratio (GMR) and geometric mean fold rise ratio (GMFRR) for wP compared to aP by age.
#| tbl-pos: "H"
#| results: "asis"
gmr_table(gmr_dat) |>
  tab_options(table.width = pct(60)) |>
  tab_footnote(
    "All models adjsuted for sex, birth order, breastfeeding status, delivery method, family history of atopic disease, and parental income at baseline."
  ) |>
  as_latex()
```

\clearpage

### Quantitation

Model estimate difference in probability of concentrations exceeding ≥0·01 kU/L are reported in @tbl-quantitation-summary.

```{r}
#| label: model-ige-quant
#| cache: true
fit_ige_sens_model <- function(ige_ref, ref_val) {
  mdat <- dat_ige |>
    filter(allergen == ige_ref) |>
    filter(!is.na(ige)) |>
    mutate(sens = as.numeric(ige >= ref_val))

  glmm_prior <- c(
    prior(student_t(4, 0, 1.75), class = "b"),
    prior(student_t(4, 0, 1.75), class = "Intercept"),
    prior(exponential(1), class = "sd")
  )

  mfit1 <- brm(
    sens ~ visit_age * trt + (1 | subjid),
    data = mdat,
    family = bernoulli(),
    backend = "cmdstanr",
    prior = glmm_prior,
    seed = 71235,
    refresh = 0,
    chains = 8,
    iter = 1750,
    warmup = 500
  )

  mfit2 <- update(
    mfit1,
    formula = ~ . +
      visit_age *
        (gender +
          bfed +
          fborn +
          ces +
          fha +
          parinc_imp),
    newdata = mdat
  )

  # Marginal risk difference estimate
  dat_ige |>
    mutate(trt = "aP") |>
    add_epred_rvars(mfit2, value = "aP", allow_new_levels = TRUE) |>
    mutate(trt = "wP") |>
    add_epred_rvars(mfit2, value = "wP", allow_new_levels = TRUE) |>
    mutate(dif = wP - aP) |>
    summarise(across(c(aP, wP, dif), rvar_mean), .by = visit_age)

  # mdrws1 <- rvar(as_draws_matrix(mfit1))
  # beta1 <- mdrws1[grepl("(^b_Inter|^b_age|^b_tr)", names(mdrws1))]
  # omega1 <- mdrws1["sd_subjectid__Intercept"]
  # beta1_marg <- do.call(c, as_draws_rvars(t(marginalcoef(mfit1, posterior = TRUE)$Posterior)))
  # names(beta1_marg) <- names(beta1)
  # cor1 <- exp(c(beta1[3], beta1[3] + beta1[4]))
  # mor1 <- exp(c(beta1_marg[3], beta1_marg[3] + beta1_marg[4]))

  mdrws2 <- rvar(as_draws_matrix(mfit2))
  beta2 <- mdrws2[grepl("(^b_Inter|^b_visit_age|^b_tr)", names(mdrws2))]
  omega2 <- mdrws2["sd_subjid__Intercept"]
  beta2_marg <- do.call(
    c,
    as_draws_rvars(t(marginalcoef(mfit2, posterior = TRUE)$Posterior))
  )
  names(beta2_marg) <- variables(mfit2)[1:30]
  beta2_marg <- beta2_marg[grepl(
    "(^b_Inter|^b_visit_age|^b_tr)",
    names(beta2_marg)
  )]
  cor2 <- exp(c(beta2[4], beta2[4] + beta2[5], beta2[4] + beta2[6]))
  mor2 <- exp(c(
    beta2_marg[4],
    beta2_marg[4] + beta2_marg[5],
    beta2_marg[4] + beta2_marg[6]
  ))
  p1 <- make_ate(dat_ige, mfit1) |>
    pivot_wider(names_from = trt, values_from = posterior)
  p2 <- make_ate(dat_ige, mfit2) |>
    pivot_wider(names_from = trt, values_from = posterior)
  tdat2 <- p2 |>
    select(-measure) |>
    bind_cols(cor = cor2, mor = mor2) |>
    mutate(group = ige_ref)
  return(tdat2)
}

snk <- capture.output({
  res_sens_tetanus <- fit_ige_sens_model("Tetanus-toxoid", 0.01)
  res_sens_eggwhite <- fit_ige_sens_model("Egg white", 0.01)
  res_sens_wholeegg <- fit_ige_sens_model("Whole egg", 0.01)
})

res_sens <- bind_rows(res_sens_tetanus, res_sens_eggwhite, res_sens_wholeegg) |>
  mutate(age = fct_rev(fct_inorder(visit_age)), group = fct_inorder(group)) |>
  relocate(group, .before = visit_age)
```

```{r}
#| label: tbl-quantitation-summary
#| tbl-cap: Summary of quantitation model results.
#| tbl-pos: "H"
out <- res_sens |>
  transmute(
    group,
    age,
    across(
      3:7,
      list(
        med = ~ median(.x),
        lo = ~ quantile(.x, 0.025)[1, ],
        hi = ~ quantile(.x, 0.975)[1, ]
      )
    ),
    across(5, list(pr = ~ Pr(.x < 0)))
  ) |>
  relocate(`wP - aP_pr`, .after = `wP - aP_hi`) |>
  gt(rowname_col = "age", groupname_col = "group") |>
  fmt_number(-(1:2), decimals = 2) |>
  cols_merge(
    starts_with("wP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("aP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    c("wP - aP_med", "wP - aP_lo", "wP - aP_hi"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("cor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("mor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_label(
    "aP_med" = "aP",
    "wP_med" = "wP",
    "wP - aP_med" = "wP \u2212 aP",
    "wP - aP_pr" = "Pr(wP \u2212 aP < 0)",
    "cor_med" = "Conditional",
    "mor_med" = "Marginal"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    columns = c("aP_med", "wP_med", "wP - aP_med"),
    label = md("Standardised Pr(IgE $\\geq$ 0.1), Median (95% CrI)"),
  ) |>
  tab_spanner(
    columns = c("cor_med", "mor_med"),
    label = "Odds ratios"
  ) |>
  tab_footnote(
    "All models adjsuted for sex, birth order, breastfeeding status, delivery method, family history of atopic disease, and parental income at baseline."
  ) |>
  tab_options(table.font.size = px(9))
out
```

\clearpage

### Sensitisation

```{r}
#| label: model-ige-sens
#| cache: true
snk <- capture.output({
  res_sens_tetanus <- fit_ige_sens_model("Tetanus-toxoid", 0.1)
  res_sens_eggwhite <- fit_ige_sens_model("Egg white", 0.1)
  res_sens_wholeegg <- fit_ige_sens_model("Whole egg", 0.1)
})

res_sens <- bind_rows(res_sens_tetanus, res_sens_eggwhite, res_sens_wholeegg) |>
  mutate(age = fct_rev(fct_inorder(visit_age)), group = fct_inorder(group)) |>
  relocate(group, .before = visit_age)
```

```{r}
#| label: tbl-sensitisation-summary
#| tbl-cap: Summary of sensitisation model by response type.
#| tbl-pos: "H"
out <- res_sens |>
  transmute(
    group,
    age,
    across(
      3:7,
      list(
        med = ~ median(.x),
        lo = ~ quantile(.x, 0.025)[1, ],
        hi = ~ quantile(.x, 0.975)[1, ]
      )
    ),
    across(5, list(pr = ~ Pr(.x < 0)))
  ) |>
  relocate(`wP - aP_pr`, .after = `wP - aP_hi`) |>
  gt(rowname_col = "age", groupname_col = "group") |>
  fmt_number(-(1:2), decimals = 2) |>
  cols_merge(
    starts_with("wP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("aP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    c("wP - aP_med", "wP - aP_lo", "wP - aP_hi"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("cor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("mor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_label(
    "aP_med" = "aP",
    "wP_med" = "wP",
    "wP - aP_med" = "wP \u2212 aP",
    "wP - aP_pr" = "Pr(wP \u2212 aP < 0)",
    "cor_med" = "Conditional",
    "mor_med" = "Marginal"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    columns = c("aP_med", "wP_med", "wP - aP_med"),
    label = md("Standardised Pr(IgE $\\geq$ 0.1), Median (95% CrI)"),
  ) |>
  tab_spanner(
    columns = c("cor_med", "mor_med"),
    label = "Odds ratios"
  ) |>
  tab_footnote(
    "All models adjsuted for sex, birth order, breastfeeding status, delivery method, family history of atopic disease, and parental income at baseline."
  ) |>
  tab_options(table.font.size = px(9))
out
```

\clearpage

## IgG concentrations {#sec-igg}

### Missingness

```{r}
#| label: tbl-overall-missingness
#| tbl-cap: |
#|    Number and proportion of participants with at least one observed response,
#|    and with all planned responses observed.
#| tbl-pos: "H"
#| results: "asis"
dat_igg |>
  select(record_id, trt, any_obs_concentration, all_obs_concentration) |>
  distinct() |>
  tbl_summary(
    include = c(any_obs_concentration, all_obs_concentration),
    label = list(
      any_obs_concentration ~ "At least one IgG response available",
      all_obs_concentration ~ "All planned IgG responses available"
    ),
    by = trt
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = row_type == "label"
    )
  ) |>
  tab_options(
    table.font.size = px(8),
    table.width = pct(80)
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-igg-missingness-dtap
#| tbl-cap:
#|    Missingness pattern of IgG concentrations by study stage.
#| tbl-pos: H
miss_pattern_trt_tab(filter(dat_igg, antigen == "PT")) |>
  tab_options(
    table.font.size = px(8),
    table.width = pct(60)
  )
```

\clearpage

### Descriptive

```{r}
#| label: tbl-igg-summary-dtap
#| tbl-cap: Descriptive summaries for IgG concentrations by assigned treatment group.
#| results: "asis"
igg_gmc_summary_tab(
  filter(dat_igg, type != "PnPs")
) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(8)
  ) |>
  as_latex() |>
  gsub(pattern = "\\\\textendash", replacement = "\\\\textendash ") |>
  cat()
```

```{r}
#| label: tbl-igg-summary-pnps
#| tbl-cap: Descriptive summaries for IgG concentrations for 13-valent pneumococcal vaccine serotypes by assigned treatment group.
#| results: "asis"
igg_gmc_summary_tab(
  filter(dat_igg, type == "PnPs"),
  rn = "visage"
) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(8)
  ) |>
  as_latex() |>
  gsub(pattern = "\\\\textendash", replacement = "\\\\textendash ") |>
  cat()
```

\clearpage
