# Immunological Outcomes

This section reports on data relating to the immunological outcomes.

```{r}
#| label: immunological-data
#| include: false
source(here("reports", "R", "functions.R"))

# Blood collection data
dat_bc <- select_form(dat_raw, "blood_collection") |>
  left_join(select(dat_base, record_id, subjid), join_by(record_id)) |>
  select(-record_id, -starts_with("lbsite___")) |>
  filter(visage != "6-month + 72hrs") |>
  arrange(subjid, visage)

# Check for date inconsistencies
dat_dates <- dat_grid_vs |>
  select(record_id, subjid, birthdat, randdat, visdat1, visage, visdat, windyn) |>
  filter(visage != "6-month + 72 hrs") |>
  left_join(dat_bc, join_by(subjid, visage)) |>
  mutate(
    age_months = interval(birthdat, visdat) %/% months(1),
    .after = visage
  ) |>
  mutate(
    .by = subjid,
    days_gap_visit = interval(lag(visdat), visdat) %/% days(1),
    .after = visdat
  ) |>
  mutate(
    .by = subjid,
    days_gap_lbipdat = interval(lag(lbipdat), lbipdat) %/% days(1),
    .after = lbipdat
  )
# dat_dates |>
#   filter(visdat != lbipdat)
# dat_dates |>
#   filter(
#     abs(age_months - as.numeric(gsub("-month", "", visage))) > 1
#   )

# Apply obvious corrections then derive gaps
# E.g. 7-month blood collection before 6-month, but 7-month visit date after
dat_bc <- dat_bc |>
  mutate(
    lbipdat = replace_when(
      lbipdat,
      subjid == "01-017" & visage == "7-month" ~ date("2018-12-04"),
      subjid == "01-094" & visage == "7-month" ~ date("2020-01-10")
    )
  ) |>
  mutate(
    .by = subjid,
    days_gap = interval(lag(lbipdat), lbipdat) %/% days(1)
  ) |>
  mutate(
    bc_window = case_when(
      !(visage %in% c("7-month", "19-month")) ~ NA,
      !lbipvst ~ NA,
      visage == "7-month" & between(days_gap, 21, 35) ~ TRUE,
      visage == "19-month" & between(days_gap, 21, 35) ~ TRUE,
      .default = FALSE
    )
  )

# IgG data
dat_igg <- select_form(dat_raw, "igg") |>
  left_join(dat_bc, join_by(subjid, visage)) |>
  left_join(dat_base, join_by(subjid)) |>
  select(
    record_id,
    subjid,
    rand_site,
    rand_stage,
    randdat,
    birthdat,
    trt,
    gender,
    bfed,
    fborn,
    parinc_imp,
    ces,
    fha,
    visage,
    type,
    antigen,
    units,
    ref,
    concentration,
    log_concentration,
    positive,
    lbipvst,
    lbncireas,
    lbvol,
    lbsite,
    lbipdat
  ) |>
  mutate(stage = if_else(substr(subjid, 1, 2) == "01", "1", "2")) |>
  filter(!(antigen %in% c("PnPs 11A"))) |>
  mutate(antigen = fct_inorder(antigen)) |>
  mutate(
    fold_rise = concentration / lag(concentration),
    .by = c(subjid, antigen),
    .after = positive
  ) |>
  mutate(
    any_obs_concentration = any(!is.na(concentration)),
    all_obs_concentration = all(!is.na(concentration)),
    .by = subjid
  )

# IgE data
dat_ige <- select_form(dat_raw, "ige") |>
  left_join(dat_base, join_by(record_id, subjid)) |>
  select(
    record_id,
    subjid,
    rand_site,
    rand_stage,
    trt,
    gender,
    bfed,
    fborn,
    parinc_imp,
    ces,
    fha,
    visit_age,
    allergen,
    ige,
    comment
  ) |>
  mutate(
    fold_rise = if_else(visit_age == "7-month", ige / lag(ige), NA_real_),
    .by = c(subjid, allergen)
  ) |>
  mutate(
    any_obs_concentration = any(!is.na(ige)),
    all_obs_concentration = all(!is.na(ige)),
    .by = subjid
  ) |>
  mutate(
    allergen = fct_inorder(replace_values(
      as.character(allergen),
      "Dust" ~ "D. pteronyssinus",
      "Rye" ~ "Rye grass"
    ))
  )
```

```{r}
#| label: immunological-functions
multinorm <- cmdstan_model(here("stan", "multinorrm-cens.stan"))

make_stan_data <- function(
  dd,
  form = ~trt,
  var = log_ige,
  L = log10(c(0.005, 0.005, 0.0005))
) {
  tmp <- dd |>
    filter(row_number() == 1, .by = subjid)
  X <- model.matrix(form, data = tmp)
  X <- cbind(
    X[, 1:2, drop = FALSE],
    scale(X[, -(1:2), drop = FALSE], scale = F)
  )
  Y <- dd |>
    select(subjid, visit_age, {{ var }}) |>
    spread(visit_age, {{ var }}) |>
    select(-subjid) |>
    mutate(across(everything(), ~ replace_na(.x, -99))) |>
    as.matrix()
  out <- list(
    N = nrow(tmp),
    J = ncol(Y),
    X = X,
    P = ncol(X),
    Y = Y,
    R = dd |>
      mutate(
        R = case_when(
          is.na({{ var }}) ~ 1,
          {{ var }} == -Inf ~ 2,
          !is.na({{ var }}) ~ 0
        )
      ) |>
      select(subjid, visit_age, R) |>
      spread(visit_age, R) |>
      select(-subjid) |>
      as.matrix(),
    L = L
  )
  return(out)
}

gmr_table_dat <- function(gmr) {
  tdat <- gmr |>
    mutate(
      across(
        V1,
        list(
          mean = ~ mean(.x),
          sd = ~ sd(.x),
          med = ~ median(.x),
          lo = ~ quantile(.x, 0.025)[1, ],
          hi = ~ quantile(.x, 0.975)[1, ],
          pr = ~ Pr(.x < 1)
        )
      )
    ) |>
    select(-V1)
  return(tdat)
}

gmr_table <- function(tdat) {
  gt(tdat, rowname_col = "visit_age", groupname_col = "group") |>
    fmt_number(everything(), decimals = 2) |>
    cols_merge(
      c(V1_mean, V1_sd),
      pattern = "<<{1} $\\pm$ {2}>>"
    ) |>
    cols_merge(
      c(V1_lo, V1_hi),
      pattern = "<<({1}, {2})>>"
    ) |>
    cols_label(
      V1_mean = md("Mean &#177; std"),
      V1_med = "Median",
      V1_lo = md("95% CrI"),
      V1_pr = "Pr(< 1)",
    ) |>
    tab_stub_indent(
      rows = everything(),
      indent = 1
    ) |>
    tab_options(table.font.size = px(9))
}

make_ate <- function(dat, fit) {
  nd <- filter(dat, visit_age == "6-month")
  ndat <- expand_grid(
    visit_age = c("6-month", "7-month", "19-month"),
    trt = c("aP", "wP")
  )
  tmp <- brmsmargins(
    fit,
    newdata = nd,
    at = ndat,
    effects = "integrateoutRE",
    CI = 0.95,
    k = 100L,
    seed = 123,
    contrasts = cbind(
      "6-month_wP - aP" = c(-1, 1, 0, 0, 0, 0),
      "7-month_wP - aP" = c(0, 0, -1, 1, 0, 0),
      "19-month_wP - aP" = c(0, 0, 0, 0, -1, 1)
    )
  )
  r <- rvar(cbind(tmp$Posterior, tmp$Contrasts))
  names(r)[seq_len(nrow(ndat))] <- apply(ndat, 1, paste, collapse = "_")
  out <- as_tibble(t(r)) |>
    pivot_longer(
      everything(),
      names_to = c("visit_age", "trt"),
      names_sep = "_",
      values_to = "posterior"
    ) |>
    mutate(measure = "RD")
  return(out)
}


miss_pattern_trt_tab_ige <- function(dat, var = concentration) {
  pic <- c("X", "-")
  pal <- setNames(c("black", "red"), pic)
  tab <- dat |>
    to_factor() |>
    mutate(missing = pic[1 + as.integer(is.na({{ var }}))]) |>
    select(subjid, allergen, trt, visit_age, missing) |>
    pivot_wider(
      names_from = visit_age,
      values_from = missing,
      values_fill = "-"
    ) |>
    count(allergen, trt, `6-month`, `7-month`, `19-month`) |>
    arrange(trt, desc(n)) |>
    mutate(p = n / sum(n), .by = c(allergen, trt)) |>
    pivot_wider(
      names_from = "trt",
      values_from = c("n", "p"),
      values_fill = 0
    ) |>
    arrange(
      desc(`6-month`),
      desc(`7-month`),
      desc(`19-month`)
    )

  gt(tab, groupname_col = "allergen") |>
    fmt_percent(starts_with("p_"), decimals = 0) |>
    cols_merge(ends_with("_aP"), pattern = "<<{1} ({2})>>") |>
    cols_merge(ends_with("_wP"), pattern = "<<{1} ({2})>>") |>
    cols_label(
      `n_aP` = md("aP<br>N = 75"),
      `n_wP` = md("wP<br>N = 75")
    ) |>
    cols_align(align = "center", columns = 2:4) |>
    tab_spanner(
      label = "Count (%)",
      columns = 2:4
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(columns = everything())
    )
}
```

## Blood collection

For participants enrolled under stage 1, blood collection was to occur at visit 3 (6 to <7-months), visit 5 (21 to 35 days after visit 3, i.e. at 7-months), visit 7 (18 to <19-months), and visit 8 (21 to 35 days after visit 7, i.e. at 19-months). For those enrolled under stage 2, blood collection was at 18 to <19-months and 21 to 35 days after that collection.

The number and percentage of infants with blood collected at each visit is reported in @tbl-blood-collection. Most collections occurred at the scheduled ages (@fig-blood-collection-age) and within the protocolised window (@fig-blood-collection-gaps) but some collections did occur outside the windows.

```{r}
#| label: tbl-blood-collection
#| tbl-cap: |
#|   Number and proportion of infants who had blood collected at the specified ages.
#| tbl-pos: "H"
#| output: asis
dat_bc |>
  mutate(lbipvst = if_else(is.na(lbipvst) | !lbipvst, FALSE, TRUE)) |>
  left_join(select(dat_base, subjid, rand_stage, trt), join_by(subjid)) |>
  summarise(
    .by = c(rand_stage, visage, trt),
    y = sum(lbipvst),
    n = n(),
    p = y / n,
    lab = sprintf("%i/%i (%.0f%%)", y, n, 100 * p)
  ) |>
  arrange(rand_stage, visage) |>
  select(-y, -n, -p) |>
  pivot_wider(names_from = trt, values_from = lab) |>
  mutate(rand_stage = paste("Stage ", rand_stage)) |>
  gt(groupname_col = "rand_stage") |>
  cols_label(
    visage ~ "Visit age"
  ) |>
  cols_label(
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    wP ~ "\\makecell{\\textbf{wP} \\\\ N = 75}"
  ) |>
  tab_options(table.font.size = px(9), table.width = pct(35)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: fig-blood-collection-age
#| fig-cap: |
#|   Age (months) at blood collection by scheduled visit based on date of birth and documented visit dates.
#| fig-height: 2.5
dat_bc |>
  left_join(select(dat_base, subjid, birthdat)) |>
  mutate(bc_age = interval(birthdat, lbipdat) %/% months(1)) |>
  ggplot(aes(bc_age)) +
  facet_wrap(~visage, scales = "free_x") +
  geom_histogram(boundary = 0.5, binwidth = 1) +
  scale_x_continuous(breaks = seq(5, 20, 1)) +
  labs(y = "Count", x = "Age (months) at blood collection")
```

```{r}
#| label: fig-blood-collection-gaps
#| fig-cap: |
#|   Gap times (days) between blood collection and previous visit for 7-month (21-35 days after 6-month vaccination) and 19-month (21-35 days after 18-month vaccination) based on documented visit dates. The protocolised windows of 21 to 35 days are indicated by dotted lines.
#| fig-height: 2.5
dat_bc |>
  filter(visage %in% c("7-month", "19-month")) |>
  ggplot(aes(days_gap)) +
  facet_wrap(~visage) +
  geom_histogram(boundary = 0.5, binwidth = 1) +
  geom_vline(xintercept = c(20.5, 35.5), linetype = 2) +
  scale_x_continuous(breaks = c(10, 15, 21, 28, 35, 40, 45, 50, 55, 60)) +
  labs(y = "Count", x = "Gap (days) between vaccination and blood collection")
```

\clearpage

## IgE concentrations {#sec-ige}

### Missingness

Even if blood collection occurred, IgE concentrations may have been missing (e.g. insufficient sample collected). The majority of participants had at least one IgE measurement and a small majority had all planned measurements (@tbl-overall-missingness-ige). Missingness varied by the specific allergen being tested (@tbl-igg-missingness-ige).

```{r}
#| label: tbl-overall-missingness-ige
#| tbl-cap: |
#|    Number and proportion of participants with at least one observed IgE response (from any visit or allergen),
#|    and with all planned IgE responses observed (all visits and allergens).
#| tbl-pos: "H"
#| output: asis
dat_ige |>
  select(record_id, trt, any_obs_concentration, all_obs_concentration) |>
  distinct() |>
  tbl_summary(
    include = c(any_obs_concentration, all_obs_concentration),
    label = list(
      any_obs_concentration ~ "At least one IgE response available",
      all_obs_concentration ~ "All planned IgE responses available"
    ),
    by = trt
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = row_type == "label"
    )
  ) |>
  tab_options(
    table.font.size = px(8),
    table.width = pct(60)
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-igg-missingness-ige
#| tbl-cap:
#|    Missingness pattern of IgE concentrations by allergen, "X" indicates a measurement was available and "-" means it was missing.
#| tbl-pos: H
miss_pattern_trt_tab_ige(filter(dat_ige), var = ige) |>
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_row_groups(groups = "D. pteronyssinus")
  ) |>
  tab_options(table.font.size = px(8), table.width = pct(50))
```

\clearpage

### Descriptive

Descriptive summaries of IgE concentrations (kU/L) are reported in @tbl-sample-ige-gmc including the number of observations (n), the geometric mean concentration (GMC), the range of concentrations (min-max), and the proportion of concentrations exceeding either at least 0.01 or 0.1 kU/L.

```{r}
#| label: tbl-sample-ige-gmc
#| code-summary: Sample summaries
#| tbl-cap: |
#|    Descriptive summary of IgE concentrations (kU/L) by allergen type, visit age,
#|    and assigned vaccination schedule.
#| tbl-pos: "H"
#| output: asis
tab <- dat_ige |>
  mutate(
    ige = if_else(ige == 0, 0.0005, ige)
  ) |>
  summarise(
    obs = sum(!is.na(ige)),
    gmc = 10^mean(log10(ige), na.rm = TRUE),
    min = min(ige, na.rm = TRUE),
    max = max(ige, na.rm = TRUE),
    det = mean(ige >= 0.01, na.rm = TRUE),
    con = mean(ige >= 0.1, na.rm = TRUE),
    .by = c(allergen, visit_age, trt)
  ) |>
  pivot_wider(
    names_from = trt,
    values_from = c(obs, gmc, min, max, det, con),
    names_vary = "slowest"
  ) |>
  to_factor() |>
  mutate(
    allergen = fct_relevel(allergen, "Total IgE")
  ) |>
  arrange(allergen, visit_age)

gt(tab, groupname_col = "allergen", rowname_col = "visit_age") |>
  fmt_integer(starts_with("obs")) |>
  fmt_number(starts_with("gmc"), decimals = 2) |>
  fmt_number(starts_with(c("min", "max")), decimals = 2) |>
  fmt_percent(starts_with("con"), decimals = 0) |>
  fmt_percent(starts_with("det"), decimals = 0) |>
  cols_merge(
    columns = c(min_aP, max_aP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(min_wP, max_wP),
    pattern = "{1}–{2}"
  ) |>
  cols_label(
    obs_aP = "n",
    obs_wP = "n",
    gmc_aP = "GMC",
    gmc_wP = "GMC",
    min_aP = "Min–Max",
    min_wP = "Min–Max",
    det_aP = md("&ge;0.01"),
    det_wP = md("&ge;0.01"),
    con_aP = md("&ge;0.1"),
    con_wP = md("&ge;0.1")
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{aP} \\\\ N = 75}",
    columns = ends_with("_aP")
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{wP} \\\\ N = 75}",
    columns = ends_with("_wP")
  ) |>
  tab_footnote(
    footnote = "Values of 0.00 kU/L and 0.000 kU/L set to 0.005 and 0.0005 respectively for calculation of GMC.",
    locations = cells_column_labels(columns = starts_with("gmc"))
  ) |>
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_row_groups(groups = "D. pteronyssinus")
  ) |>
  tab_options(table.font.size = px(9)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-sample-ige-gmr
#| code-summary: Sample summaries
#| tbl-cap: |
#|    Descriptive summary of geometric mean fold-rise (GMFR), geometric mean ratio (GMR), and geometric mean fold-rise ratio (GMFRR) of wP compared to aP. Number in brackets (n) indicates the number of infants contributing to the calculation.
#| tbl-pos: "H"
#| output: asis
sample_gmr <- dat_ige |>
  # filter(allergen %in% c("Total IgE", "Tetanus-toxoid", "Egg white", "Whole egg")) |>
  mutate(
    ige = if_else(ige == 0, 0.0005, ige)
  ) |>
  mutate(
    allergen = fct_relevel(allergen, "Total IgE")
  ) |>
  arrange(subjid, allergen, visit_age) |>
  mutate(
    fold_rise = ige / lag(ige),
    .by = c(subjid, allergen)
  ) |>
  summarise(
    .by = c(allergen, trt, visit_age),
    obs_gmc = sum(!is.na(ige)),
    gmc = gmc(ige, na.rm = TRUE),
    obs_gmfr = sum(!is.na(fold_rise)),
    gmfr = gmc(fold_rise, na.rm = TRUE)
  ) |>
  pivot_wider(names_from = trt, values_from = c(obs_gmc, gmc, obs_gmfr, gmfr)) |>
  mutate(gmr = gmc_wP / gmc_aP, gmfrr = gmfr_wP / gmfr_aP) |>
  mutate(
    obs_gmfr_aP = replace_values(obs_gmfr_aP, 0 ~ NA),
    obs_gmfr_wP = replace_values(obs_gmfr_wP, 0 ~ NA)
  )

gt(sample_gmr, groupname_col = "allergen", rowname_col = "visit_age") |>
  fmt_integer(starts_with("obs")) |>
  fmt_number(starts_with(c("gmc", "gmr", "gmfr")), decimals = 2) |>
  cols_merge(
    columns = c(gmc_aP, obs_gmc_aP),
    pattern = "{1} ({2})"
  ) |>
  cols_merge(
    columns = c(gmc_wP, obs_gmc_wP),
    pattern = "{1} ({2})"
  ) |>
  cols_merge(
    columns = c(gmfr_aP, obs_gmfr_aP),
    pattern = "{1} ({2})"
  ) |>
  cols_merge(
    columns = c(gmfr_wP, obs_gmfr_wP),
    pattern = "{1} ({2})"
  ) |>
  cols_label(
    obs_gmc_aP = "n",
    obs_gmc_wP = "n",
    gmc_aP = "GMC (n)",
    gmc_wP = "GMC (n)",
    gmfr_aP = "GMFR (n)",
    gmfr_wP = "GMFR (n)",
    gmr ~ "GMR",
    gmfrr ~ "GMFRR"
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{aP} \\\\ N = 75}",
    columns = ends_with("_aP")
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{wP} \\\\ N = 75}",
    columns = ends_with("_wP")
  ) |>
  sub_missing() |>
  tab_footnote(
    footnote = "Values of 0.00 kU/L and 0.000 kU/L set to 0.005 and 0.0005 respectively for calculation of GMC.",
    locations = cells_column_labels(columns = starts_with("gmc"))
  ) |>
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_row_groups(groups = "D. pteronyssinus")
  ) |>
  tab_options(table.font.size = px(9)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

### Geometric mean concentration and fold-rise ratio

Model estimated geometric mean ratios (GMRs) and geometric mean fold-rise ratios (GMFRRs) are reported in @tbl-gmr-gmfir-ige-total. For completeness, all IgE types are reported, however, in some cases the GMR and GMFRR may not be meaningful due to the small number of participants with measurable quantities (e.g. tetanus toxoid at 6-months and 7-months) resulting in GMC's close to zero and GMR's which are ratios of two near zero values.

For total IgE, the GMFRR from 6 to 7 months was 0.94 (0.72, 1.22) and the probability of a reduction in the GMFR was 0.69.

```{r}
#| label: model-ige-total
#| cache: true
fit_ige_model <- function(ige_ref) {
  age_labs <- c("6-month", "7-month", "19-month")

  mdat <- dat_ige |>
    filter(allergen == ige_ref) |>
    filter(!all(is.na(ige)), .by = subjid) |>
    mutate(log_ige = log10(ige))

  sdat1 <- make_stan_data(mdat, var = log_ige)
  sdat2 <- make_stan_data(
    mdat,
    var = log_ige,
    ~ trt + gender + bfed + fborn + parinc_imp + ces + fha
  )
  snk <- capture.output(
    sfit1 <- multinorm$sample(
      data = sdat1,
      seed = 71356,
      refresh = 0,
      chains = 8,
      iter_sampling = 1250
    )
  )
  snk <- capture.output(
    sfit2 <- multinorm$sample(
      data = sdat2,
      seed = 71356,
      refresh = 0,
      chains = 8,
      iter_warmup = 500,
      iter_sampling = 1250
    )
  )
  sdrw1 <- as_draws_rvars(sfit1)
  sdrw2 <- as_draws_rvars(sfit2)

  rownames(sdrw1$b) <- age_labs
  rownames(sdrw2$b) <- age_labs

  gmr1 <- 10^sdrw1$b[, 2]
  gmr2 <- 10^sdrw2$b[, 2]
  gmfir1 <- 10^(sdrw1$b[-1, 2] - sdrw1$b[-3, 2])
  gmfir2 <- 10^(sdrw2$b[-1, 2] - sdrw2$b[-3, 2])
  rownames(gmfir1) <- rownames(gmfir2) <- c(
    "7/6-month (GMFR)",
    "19/7-month (GMFR)"
  )

  # Fill missing observations with conditional predictions
  # yfill <- t(as_rvar(sdat1$Y))
  # yfill[which(t(sdat1$R == 1))] <- sdrw1$y_missing
  # yfill[which(t(sdat1$R == 2))] <- sdrw1$y_censored
  # yfill <- t(yfill)
  return(as_tibble(rbind(gmr2, gmfir2), rownames = "visit_age"))
}

res_total_ige <- fit_ige_model("Total IgE")
res_tetanus_ige <- fit_ige_model("Tetanus-toxoid")
res_eggwhite_ige <- fit_ige_model("Egg white")
res_wholeegg_ige <- fit_ige_model("Whole egg")

gmr <- bind_rows(
  mutate(res_total_ige, group = "Total IgE"),
  mutate(res_tetanus_ige, group = "Tetanus toxoid"),
  mutate(res_eggwhite_ige, group = "Egg white"),
  mutate(res_wholeegg_ige, group = "Whole egg")
)
gmr_dat <- gmr_table_dat(gmr)
```

```{r}
#| label: tbl-gmr-gmfir-ige-total
#| tbl-cap: Geometric mean ratio (GMR) and geometric mean fold rise ratio (GMFRR) for wP compared to aP by age.
#| tbl-pos: "H"
#| output: asis
gmr_table(gmr_dat) |>
  tab_options(table.width = pct(65)) |>
  tab_footnote(
    "All models adjsuted for sex, birth order, breastfeeding status, delivery method, family history of atopic disease, and parental income at baseline."
  ) |>
  as_latex()
```

\clearpage

### Quantitation

Model estimate difference in probability of concentrations ≥0·01 kU/L, and conditional odds ratios, comparing wP with aP, are reported in @tbl-logistic-regression-ige-quant. Estimates had a large magnitude of uncertainty, but generally, there was only weak support for any direction of effect on probability of IgE concentration exceeding 0·01 kU/L.

```{r}
#| label: univariate-beta-binomial-ige-check
#| include: false
#| eval: false
# For checks sake, I first run univariate Binomial models for each visit by allergen combination
# similar to the primary outcome
dat_summ <- dat_ige |>
  filter(allergen != "Total IgE") |>
  summarise(
    .by = c(allergen, visit_age, trt),
    y = sum(ige >= 0.01, na.rm = T),
    n = sum(!is.na(ige)),
    p = y / n,
    a = 1 + y,
    b = 1 + n - y,
    post = dist_beta(a, b),
  ) |>
  arrange(allergen, visit_age, trt)
dat_summ_diff <- dat_summ |>
  mutate(post_drws = generate(post, 1e4)) |>
  select(allergen, visit_age, trt, post_drws) |>
  unnest(post_drws) |>
  group_by(allergen, visit_age, trt) |>
  mutate(.draw = 1:n()) |>
  ungroup() |>
  pivot_wider(names_from = "trt", values_from = "post_drws") |>
  mutate(delta = wP - aP)

diff_agg <- dat_summ_diff |>
  group_by(allergen, visit_age) |>
  nest_rvars() |>
  ungroup() |>
  mutate(
    across(
      aP:delta,
      list(
        med = median,
        lo = ~ quantile2(.x, 0.025),
        hi = ~ quantile2(.x, 0.975),
        pr = ~ Pr(.x < 0)
      ),
      .names = "{.fn}_{.col}"
    )
  ) |>
  select(-aP, -wP, -delta)

dat_summ |>
  select(allergen, visit_age, trt, n, y, p) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p)
  ) |>
  left_join(diff_agg, join_by(allergen, visit_age)) |>
  gt(rowname_col = "visit_age", groupname_col = "allergen") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_aP, lo_aP, hi_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_wP, lo_wP, hi_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "aP"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "wP"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_aP, med_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_row_groups(groups = "D. pteronyssinus")
  ) |>
  tab_options(table.font.size = px(9)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: univariate-logistic-regression-ige-check
#| include: false
#| eval: false
# Similar to above, run univariate analyses, but assuming logistic regression model, should be similar.
prior <- c(
  prior(student_t(4, 0, 1.75), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = b)
)

mdat <- dat_ige |>
  filter(allergen != "Total IgE", !is.na(ige)) |>
  mutate(y = as.numeric(ige >= 0.01)) |>
  nest(dat = -c(allergen, visit_age))

mdat <- mdat |>
  mutate(
    fit = map(
      dat,
      ~ brm(
        y ~ trt,
        data = .x,
        family = bernoulli,
        backend = "cmdstanr",
        prior = prior,
        seed = 71235,
        refresh = 0,
        chains = 8,
        iter = 625,
        warmup = 500
      )
    ),
    rd = map2(
      dat,
      fit,
      ~ .x |>
        select(trt) |>
        distinct() |>
        add_epred_rvars(.y)
    )
  )

d1 <- mdat |>
  select(visit_age, allergen, dat) |>
  unnest(dat) |>
  summarise(
    .by = c(visit_age, allergen, trt),
    y = sum(y),
    n = n(),
    p = y / n
  )
d2 <- mdat |>
  select(visit_age, allergen, rd) |>
  unnest(rd)

td <- left_join(d1, d2, join_by(visit_age, allergen, trt)) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p, .epred)
  ) |>
  mutate(delta = .epred_wP - .epred_aP) |>
  mutate(across(
    c(.epred_aP, .epred_wP, delta),
    list(
      med = median,
      lo = ~ quantile2(.x, 0.025),
      hi = ~ quantile2(.x, 0.975),
      pr = ~ Pr(.x < 0)
    ),
    .names = "{.fn}_{.col}"
  )) |>
  select(-pr_.epred_aP, -pr_.epred_wP, -.epred_aP, -.epred_wP, -delta)

td |>
  gt(rowname_col = "visit_age", groupname_col = "allergen") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_aP, lo_.epred_aP, hi_.epred_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_wP, lo_.epred_wP, hi_.epred_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "aP"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "wP"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_.epred_aP, med_.epred_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  tab_options(table.font.size = px(9))
```

```{r}
#| label: logistic-regression-ige-quant
#| include: false
#| cache: true
# Now run the actual model that was specified to correlate outcomes of participants
# with a random intercept term.
prior1 <- c(
  prior(student_t(4, 0, 1.75), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = b)
)

prior2 <- c(
  prior(student_t(4, 0, 1.75), class = "b"),
  prior(student_t(4, 0, 1.75), class = "Intercept"),
  prior(exponential(1), class = "sd")
)

mdat <- dat_ige |>
  filter(allergen != "Total IgE") |>
  mutate(y = as.numeric(ige >= 0.01)) |>
  nest(dat = -allergen)

mdat <- mdat |>
  mutate(
    fit = map(
      dat,
      ~ if (all(.x$visit_age == "19-month")) {
        # No subject specific random intercept
        brm(
          y ~ trt + gender + bfed + fborn + ces + fha + parinc_imp,
          data = filter(
            .x,
            !is.na(y)
          ),
          family = bernoulli,
          backend = "cmdstanr",
          prior = prior1,
          seed = 71235,
          refresh = 0,
          chains = 8,
          iter = 1750,
          warmup = 500
        )
      } else {
        # Subject specific random intercept to correlate outcomes, and visit specific effects
        brm(
          y ~ visit_age * (trt + gender + bfed + fborn + ces + fha + parinc_imp) + (1 | subjid),
          data = filter(.x, !is.na(y)),
          family = bernoulli,
          backend = "cmdstanr",
          prior = prior2,
          seed = 71235,
          refresh = 0,
          chains = 8,
          iter = 1750,
          warmup = 500
        )
      }
    )
  )

# mdat <- mdat |>
#   mutate(
#     rd = map2(
#       dat,
#       fit,
#       ~ .x |>
#         # Super-population, so sample random effects, need new IDs
#         mutate(subjid = gsub("01-", "", subjid)) |>
#         add_epred_rvars(
#           .y,
#           allow_new_levels = TRUE,
#           sample_new_levels = "gaussian"
#         ) |>
#         summarise(
#           .by = c(visit_age, trt),
#           .epred = rvar_mean(.epred)
#         )
#     )
#   )

mdat <- mdat |>
  mutate(
    rd2 = map2(
      dat,
      fit,
      ~ {
        if (all(.x$visit_age == "19-month")) {
          epred <- rvar(
            prediction(
              .y,
              .x,
              summarize = FALSE,
              posterior = TRUE,
              raw = TRUE,
              effects = "fixedonly",
              k = 100,
              seed = 123
            )$Posterior
          )
        } else {
          # Integrate out random effects for marginal predictions
          epred <- rvar(
            prediction(
              .y,
              .x,
              summarize = FALSE,
              posterior = TRUE,
              raw = TRUE,
              effects = "integrateoutRE",
              k = 100,
              seed = 123
            )$Posterior
          )
        }
        .x |>
          mutate(.epred = epred) |>
          summarise(
            .by = c(visit_age, trt),
            .epred = rvar_mean(.epred)
          )
      }
    )
  )
```

```{r}
#| label: tbl-logistic-regression-ige-quant
#| tbl-cap: |
#|   The model estimated standardised risk, risk difference, and conditional odds ratio for IgE ≥ 0.01 kU/L comparing wP with aP.
#| tbl-pos: "H"
#| output: asis
d1 <- mdat |>
  select(allergen, dat) |>
  unnest(dat) |>
  summarise(
    .by = c(visit_age, allergen, trt),
    e = sum(y, na.rm = TRUE),
    n = sum(!is.na(y)),
    p = e / n
  )
d2 <- mdat |>
  select(allergen, rd2) |>
  unnest(rd2)

d3 <- mdat |>
  select(allergen, fit) |>
  mutate(
    b = map(
      fit,
      ~ {
        rvs <- rvar(as_draws_matrix(.x))
        bs <- rvs[grepl("trt", names(rvs))]
        if (length(bs) == 1) {
          tibble(
            visit_age = "19-month",
            bs = exp(bs)
          )
        } else {
          tibble(
            visit_age = c("6-month", "7-month", "19-month"),
            bs = exp(c(bs[1], bs[1] + bs[2], bs[1] + bs[3]))
          )
        }
      }
    )
  )

td <- left_join(d1, d2, join_by(visit_age, allergen, trt)) |>
  rename(y = e) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p, .epred)
  ) |>
  mutate(delta = .epred_wP - .epred_aP) |>
  mutate(across(
    c(.epred_aP, .epred_wP, delta),
    list(
      med = median,
      lo = ~ quantile2(.x, 0.025),
      hi = ~ quantile2(.x, 0.975),
      pr = ~ Pr(.x < 0)
    ),
    .names = "{.fn}_{.col}"
  )) |>
  select(-pr_.epred_aP, -pr_.epred_wP, -.epred_aP, -.epred_wP, -delta) |>
  left_join(
    d3 |>
      select(allergen, b) |>
      unnest(b) |>
      mutate(across(
        bs,
        list(
          med = median,
          lo = ~ quantile2(.x, 0.025),
          hi = ~ quantile2(.x, 0.975)
        ),
        .names = "{.fn}_{.col}"
      )) |>
      select(-bs),
    join_by(allergen, visit_age)
  ) |>
  mutate(visit_age = factor(visit_age, levels = c("6-month", "7-month", "19-month"))) |>
  arrange(allergen, visit_age)

td |>
  gt(rowname_col = "visit_age", groupname_col = "allergen") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_aP, lo_.epred_aP, hi_.epred_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_wP, lo_.epred_wP, hi_.epred_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  cols_merge(
    c(med_bs, lo_bs, hi_bs),
    pattern = "{1} ({2}, {3})"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "\\makecell{\\textbf{aP} \\\\ N = 75}"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "\\makecell{\\textbf{wP} \\\\ N = 75}"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  tab_spanner(
    columns = ends_with("bs"),
    label = "Conditional odds ratio"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_.epred_aP, med_.epred_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  cols_label(
    c(med_bs) ~ "Median (95% CrI)"
  ) |>
  tab_footnote(
    "All models adjust for and standardise over sex, birth order, breastfeeding status, delivery method, family history of atopic disease, and parental income at baseline."
  ) |>
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_row_groups(groups = "D. pteronyssinus")
  ) |>
  tab_options(table.font.size = px(8)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: model-ige-quant
#| cache: true
fit_ige_sens_model <- function(ige_ref, ref_val) {
  mdat <- dat_ige |>
    filter(allergen == ige_ref) |>
    filter(!is.na(ige)) |>
    mutate(sens = as.numeric(ige >= ref_val))

  glmm_prior <- c(
    prior(student_t(4, 0, 1.75), class = "b"),
    prior(student_t(4, 0, 1.75), class = "Intercept"),
    prior(exponential(1), class = "sd")
  )

  mfit1 <- brm(
    sens ~ visit_age * trt + (1 | subjid),
    data = mdat,
    family = bernoulli(),
    backend = "cmdstanr",
    prior = glmm_prior,
    seed = 71235,
    refresh = 0,
    chains = 8,
    iter = 1750,
    warmup = 500
  )

  mfit2 <- update(
    mfit1,
    formula = ~ . +
      visit_age *
        (gender +
          bfed +
          fborn +
          ces +
          fha +
          parinc_imp),
    newdata = mdat
  )

  # Marginal risk difference estimate
  dat_ige |>
    mutate(trt = "aP") |>
    add_epred_rvars(mfit2, value = "aP", allow_new_levels = TRUE) |>
    mutate(trt = "wP") |>
    add_epred_rvars(mfit2, value = "wP", allow_new_levels = TRUE) |>
    mutate(dif = wP - aP) |>
    summarise(across(c(aP, wP, dif), rvar_mean), .by = visit_age)

  # mdrws1 <- rvar(as_draws_matrix(mfit1))
  # beta1 <- mdrws1[grepl("(^b_Inter|^b_age|^b_tr)", names(mdrws1))]
  # omega1 <- mdrws1["sd_subjectid__Intercept"]
  # beta1_marg <- do.call(c, as_draws_rvars(t(marginalcoef(mfit1, posterior = TRUE)$Posterior)))
  # names(beta1_marg) <- names(beta1)
  # cor1 <- exp(c(beta1[3], beta1[3] + beta1[4]))
  # mor1 <- exp(c(beta1_marg[3], beta1_marg[3] + beta1_marg[4]))

  mdrws2 <- rvar(as_draws_matrix(mfit2))
  beta2 <- mdrws2[grepl("(^b_Inter|^b_visit_age|^b_tr)", names(mdrws2))]
  omega2 <- mdrws2["sd_subjid__Intercept"]
  beta2_marg <- do.call(
    c,
    as_draws_rvars(t(marginalcoef(mfit2, posterior = TRUE)$Posterior))
  )
  names(beta2_marg) <- variables(mfit2)[1:30]
  beta2_marg <- beta2_marg[grepl(
    "(^b_Inter|^b_visit_age|^b_tr)",
    names(beta2_marg)
  )]
  cor2 <- exp(c(beta2[4], beta2[4] + beta2[5], beta2[4] + beta2[6]))
  mor2 <- exp(c(
    beta2_marg[4],
    beta2_marg[4] + beta2_marg[5],
    beta2_marg[4] + beta2_marg[6]
  ))
  p1 <- make_ate(dat_ige, mfit1) |>
    pivot_wider(names_from = trt, values_from = posterior)
  p2 <- make_ate(dat_ige, mfit2) |>
    pivot_wider(names_from = trt, values_from = posterior)
  tdat2 <- p2 |>
    select(-measure) |>
    bind_cols(cor = cor2, mor = mor2) |>
    mutate(group = ige_ref)
  return(tdat2)
}

snk <- capture.output({
  res_sens_tetanus <- fit_ige_sens_model("Tetanus-toxoid", 0.01)
  res_sens_eggwhite <- fit_ige_sens_model("Egg white", 0.01)
  res_sens_wholeegg <- fit_ige_sens_model("Whole egg", 0.01)
})

res_sens <- bind_rows(res_sens_tetanus, res_sens_eggwhite, res_sens_wholeegg) |>
  mutate(age = fct_rev(fct_inorder(visit_age)), group = fct_inorder(group)) |>
  relocate(group, .before = visit_age)
```

```{r}
#| label: tbl-quantitation-summary
#| tbl-cap: Summary of quantitation model estimates.
#| tbl-pos: "H"
#| include: false
out <- res_sens |>
  transmute(
    group,
    age,
    across(
      3:7,
      list(
        med = ~ median(.x),
        lo = ~ quantile(.x, 0.025)[1, ],
        hi = ~ quantile(.x, 0.975)[1, ]
      )
    ),
    across(5, list(pr = ~ Pr(.x < 0)))
  ) |>
  relocate(`wP - aP_pr`, .after = `wP - aP_hi`) |>
  gt(rowname_col = "age", groupname_col = "group") |>
  fmt_number(-(1:2), decimals = 2) |>
  cols_merge(
    starts_with("wP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("aP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    c("wP - aP_med", "wP - aP_lo", "wP - aP_hi"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("cor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("mor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_label(
    "aP_med" = "aP",
    "wP_med" = "wP",
    "wP - aP_med" = "wP \u2212 aP",
    "wP - aP_pr" = "Pr(wP \u2212 aP < 0)",
    "cor_med" = "Conditional",
    "mor_med" = "Marginal"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    columns = c("aP_med", "wP_med", "wP - aP_med"),
    label = md("Standardised Pr(IgE $\\geq$ 0.1), Median (95% CrI)"),
  ) |>
  tab_spanner(
    columns = c("cor_med", "mor_med"),
    label = "Odds ratios"
  ) |>
  tab_footnote(
    "All models adjsuted for sex, birth order, breastfeeding status, delivery method, family history of atopic disease, and parental income at baseline."
  ) |>
  tab_options(table.font.size = px(9))
out
```

\clearpage

### Sensitisation

Model estimate difference in probability of concentrations ≥0·1 kU/L, and conditional odds ratios, comparing wP with aP, are reported in @tbl-logistic-regression-ige-sens. Estimates had a large magnitude of uncertainty, but generally, there was only weak support for any direction of effect on probability of IgE concentration exceeding 0·1 kU/L.

```{r}
#| label: univariate-beta-binomial-ige-sens-check
#| include: false
#| eval: false
# For checks sake, I first run univariate Binomial models for each visit by allergen combination
# similar to the primary outcome
dat_summ <- dat_ige |>
  filter(allergen != "Total IgE") |>
  summarise(
    .by = c(allergen, visit_age, trt),
    y = sum(ige >= 0.1, na.rm = T),
    n = sum(!is.na(ige)),
    p = y / n,
    a = 1 + y,
    b = 1 + n - y,
    post = dist_beta(a, b),
  ) |>
  arrange(allergen, visit_age, trt)
dat_summ_diff <- dat_summ |>
  mutate(post_drws = generate(post, 1e4)) |>
  select(allergen, visit_age, trt, post_drws) |>
  unnest(post_drws) |>
  group_by(allergen, visit_age, trt) |>
  mutate(.draw = 1:n()) |>
  ungroup() |>
  pivot_wider(names_from = "trt", values_from = "post_drws") |>
  mutate(delta = wP - aP)

diff_agg <- dat_summ_diff |>
  group_by(allergen, visit_age) |>
  nest_rvars() |>
  ungroup() |>
  mutate(
    across(
      aP:delta,
      list(
        med = median,
        lo = ~ quantile2(.x, 0.025),
        hi = ~ quantile2(.x, 0.975),
        pr = ~ Pr(.x < 0)
      ),
      .names = "{.fn}_{.col}"
    )
  ) |>
  select(-aP, -wP, -delta)

dat_summ |>
  select(allergen, visit_age, trt, n, y, p) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p)
  ) |>
  left_join(diff_agg, join_by(allergen, visit_age)) |>
  gt(rowname_col = "visit_age", groupname_col = "allergen") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_aP, lo_aP, hi_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_wP, lo_wP, hi_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "aP"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "wP"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_aP, med_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  tab_options(table.font.size = px(9))
```

```{r}
#| label: univariate-logistic-regression-ige-sens-check
#| include: false
#| eval: false
# Similar to above, run univariate analyses, but assuming logistic regression model, should be similar.
prior <- c(
  prior(student_t(4, 0, 1.75), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = b)
)

mdat <- dat_ige |>
  filter(allergen != "Total IgE", !is.na(ige)) |>
  mutate(y = as.numeric(ige >= 0.1)) |>
  nest(dat = -c(allergen, visit_age))

mdat <- mdat |>
  mutate(
    fit = map(
      dat,
      ~ brm(
        y ~ trt,
        data = .x,
        family = bernoulli,
        backend = "cmdstanr",
        prior = prior,
        seed = 71235,
        refresh = 0,
        chains = 8,
        iter = 625,
        warmup = 500
      )
    ),
    rd = map2(
      dat,
      fit,
      ~ .x |>
        select(trt) |>
        distinct() |>
        add_epred_rvars(.y)
    )
  )

d1 <- mdat |>
  select(visit_age, allergen, dat) |>
  unnest(dat) |>
  summarise(
    .by = c(visit_age, allergen, trt),
    y = sum(y),
    n = n(),
    p = y / n
  )
d2 <- mdat |>
  select(visit_age, allergen, rd) |>
  unnest(rd)

td <- left_join(d1, d2, join_by(visit_age, allergen, trt)) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p, .epred)
  ) |>
  mutate(delta = .epred_wP - .epred_aP) |>
  mutate(across(
    c(.epred_aP, .epred_wP, delta),
    list(
      med = median,
      lo = ~ quantile2(.x, 0.025),
      hi = ~ quantile2(.x, 0.975),
      pr = ~ Pr(.x < 0)
    ),
    .names = "{.fn}_{.col}"
  )) |>
  select(-pr_.epred_aP, -pr_.epred_wP, -.epred_aP, -.epred_wP, -delta)

td |>
  gt(rowname_col = "visit_age", groupname_col = "allergen") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_aP, lo_.epred_aP, hi_.epred_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_wP, lo_.epred_wP, hi_.epred_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "aP"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "wP"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_.epred_aP, med_.epred_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  tab_options(table.font.size = px(9))
```


```{r}
#| label: logistic-regression-ige-sens
#| include: false
#| cache: true
# Now run the actual model that was specified to correlate outcomes of participants
# with a random intercept term.
prior1 <- c(
  prior(student_t(4, 0, 1.75), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = b)
)

prior2 <- c(
  prior(student_t(4, 0, 1.75), class = "b"),
  prior(student_t(4, 0, 1.75), class = "Intercept"),
  prior(exponential(1), class = "sd")
)

mdat <- dat_ige |>
  filter(allergen != "Total IgE") |>
  mutate(y = as.numeric(ige >= 0.1)) |>
  nest(dat = -allergen)

mdat <- mdat |>
  mutate(
    fit = map(
      dat,
      ~ if (all(.x$visit_age == "19-month")) {
        # No subject specific random intercept
        brm(
          y ~ trt + gender + bfed + fborn + ces + fha + parinc_imp,
          data = filter(
            .x,
            !is.na(y)
          ),
          family = bernoulli,
          backend = "cmdstanr",
          prior = prior1,
          seed = 71235,
          refresh = 0,
          chains = 8,
          iter = 1750,
          warmup = 500
        )
      } else {
        # Subject specific random intercept to correlate outcomes, and visit specific effects
        brm(
          y ~ visit_age * (trt + gender + bfed + fborn + ces + fha + parinc_imp) + (1 | subjid),
          data = filter(.x, !is.na(y)),
          family = bernoulli,
          backend = "cmdstanr",
          prior = prior2,
          seed = 71235,
          refresh = 0,
          chains = 8,
          iter = 1750,
          warmup = 500
        )
      }
    )
  )

# mdat <- mdat |>
#   mutate(
#     rd = map2(
#       dat,
#       fit,
#       ~ .x |>
#         # Super-population, so sample random effects, need new IDs
#         mutate(subjid = gsub("01-", "", subjid)) |>
#         add_epred_rvars(
#           .y,
#           allow_new_levels = TRUE,
#           sample_new_levels = "gaussian"
#         ) |>
#         summarise(
#           .by = c(visit_age, trt),
#           .epred = rvar_mean(.epred)
#         )
#     )
#   )

mdat <- mdat |>
  mutate(
    rd2 = map2(
      dat,
      fit,
      ~ {
        if (all(.x$visit_age == "19-month")) {
          epred <- rvar(
            prediction(
              .y,
              .x,
              summarize = FALSE,
              posterior = TRUE,
              raw = TRUE,
              effects = "fixedonly",
              k = 100,
              seed = 123
            )$Posterior
          )
        } else {
          # Integrate out random effects for marginal predictions
          epred <- rvar(
            prediction(
              .y,
              .x,
              summarize = FALSE,
              posterior = TRUE,
              raw = TRUE,
              effects = "integrateoutRE",
              k = 100,
              seed = 123
            )$Posterior
          )
        }
        .x |>
          mutate(.epred = epred) |>
          summarise(
            .by = c(visit_age, trt),
            .epred = rvar_mean(.epred)
          )
      }
    )
  )
```

```{r}
#| label: tbl-logistic-regression-ige-sens
#| tbl-cap: |
#|   The model estimated standardised risk, risk difference, and conditional odds ratio for IgE ≥ 0.1 kU/L comparing wP with aP.
#| tbl-pos: "H"
#| output: asis
d1 <- mdat |>
  select(allergen, dat) |>
  unnest(dat) |>
  summarise(
    .by = c(visit_age, allergen, trt),
    e = sum(y, na.rm = TRUE),
    n = sum(!is.na(y)),
    p = e / n
  )
d2 <- mdat |>
  select(allergen, rd2) |>
  unnest(rd2)

d3 <- mdat |>
  select(allergen, fit) |>
  mutate(
    b = map(
      fit,
      ~ {
        rvs <- rvar(as_draws_matrix(.x))
        bs <- rvs[grepl("trt", names(rvs))]
        if (length(bs) == 1) {
          tibble(
            visit_age = "19-month",
            bs = exp(bs)
          )
        } else {
          tibble(
            visit_age = c("6-month", "7-month", "19-month"),
            bs = exp(c(bs[1], bs[1] + bs[2], bs[1] + bs[3]))
          )
        }
      }
    )
  )

td <- left_join(d1, d2, join_by(visit_age, allergen, trt)) |>
  rename(y = e) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p, .epred)
  ) |>
  mutate(delta = .epred_wP - .epred_aP) |>
  mutate(across(
    c(.epred_aP, .epred_wP, delta),
    list(
      med = median,
      lo = ~ quantile2(.x, 0.025),
      hi = ~ quantile2(.x, 0.975),
      pr = ~ Pr(.x < 0)
    ),
    .names = "{.fn}_{.col}"
  )) |>
  select(-pr_.epred_aP, -pr_.epred_wP, -.epred_aP, -.epred_wP, -delta) |>
  left_join(
    d3 |>
      select(allergen, b) |>
      unnest(b) |>
      mutate(across(
        bs,
        list(
          med = median,
          lo = ~ quantile2(.x, 0.025),
          hi = ~ quantile2(.x, 0.975)
        ),
        .names = "{.fn}_{.col}"
      )) |>
      select(-bs),
    join_by(allergen, visit_age)
  ) |>
  mutate(visit_age = factor(visit_age, levels = c("6-month", "7-month", "19-month"))) |>
  arrange(allergen, visit_age)

td |>
  gt(rowname_col = "visit_age", groupname_col = "allergen") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_aP, lo_.epred_aP, hi_.epred_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_wP, lo_.epred_wP, hi_.epred_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  cols_merge(
    c(med_bs, lo_bs, hi_bs),
    pattern = "{1} ({2}, {3})"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "\\makecell{\\textbf{aP} \\\\ N = 75}"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "\\makecell{\\textbf{wP} \\\\ N = 75}"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  tab_spanner(
    columns = ends_with("bs"),
    label = "Conditional odds ratio"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_.epred_aP, med_.epred_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  cols_label(
    c(med_bs) ~ "Median (95% CrI)"
  ) |>
  tab_footnote(
    "All models adjust for and standardise over sex, birth order, breastfeeding status, delivery method, family history of atopic disease, and parental income at baseline."
  ) |>
  tab_style(
    style = cell_text(style = "italic"),
    locations = cells_row_groups(groups = "D. pteronyssinus")
  ) |>
  tab_options(table.font.size = px(8)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: model-ige-sens
#| cache: true
snk <- capture.output({
  res_sens_tetanus <- fit_ige_sens_model("Tetanus-toxoid", 0.1)
  res_sens_eggwhite <- fit_ige_sens_model("Egg white", 0.1)
  res_sens_wholeegg <- fit_ige_sens_model("Whole egg", 0.1)
})

res_sens <- bind_rows(res_sens_tetanus, res_sens_eggwhite, res_sens_wholeegg) |>
  mutate(age = fct_rev(fct_inorder(visit_age)), group = fct_inorder(group)) |>
  relocate(group, .before = visit_age)
```

```{r}
#| label: tbl-sensitisation-summary
#| tbl-cap: Summary of sensitisation model by response type.
#| tbl-pos: "H"
#| include: false
out <- res_sens |>
  transmute(
    group,
    age,
    across(
      3:7,
      list(
        med = ~ median(.x),
        lo = ~ quantile(.x, 0.025)[1, ],
        hi = ~ quantile(.x, 0.975)[1, ]
      )
    ),
    across(5, list(pr = ~ Pr(.x < 0)))
  ) |>
  relocate(`wP - aP_pr`, .after = `wP - aP_hi`) |>
  gt(rowname_col = "age", groupname_col = "group") |>
  fmt_number(-(1:2), decimals = 2) |>
  cols_merge(
    starts_with("wP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("aP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    c("wP - aP_med", "wP - aP_lo", "wP - aP_hi"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("cor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("mor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_label(
    "aP_med" = "aP",
    "wP_med" = "wP",
    "wP - aP_med" = "wP \u2212 aP",
    "wP - aP_pr" = "Pr(wP \u2212 aP < 0)",
    "cor_med" = "Conditional",
    "mor_med" = "Marginal"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    columns = c("aP_med", "wP_med", "wP - aP_med"),
    label = md("Standardised Pr(IgE $\\geq$ 0.1), Median (95% CrI)"),
  ) |>
  tab_spanner(
    columns = c("cor_med", "mor_med"),
    label = "Odds ratios"
  ) |>
  tab_footnote(
    "All models adjsuted for sex, birth order, breastfeeding status, delivery method, family history of atopic disease, and parental income at baseline."
  ) |>
  tab_options(table.font.size = px(9))
out
```

\clearpage

### 19-month only

In the following, analysis is restricted to only the 19-month IgE measurements.

```{r}
#| label: tbl-ige-gmr-19month
#| tbl-cap: |
#|   Estimated geometric mean ratio (GMR) at 19-months of age, wP versus aP, by allergen.
#| tbl-pos: "H"
pr1 <- c(
  prior(normal(0, 10), class = "Intercept"),
  prior(normal(0, 10), class = "b"),
  prior(cauchy(0, 10), class = "sigma")
)
pr2 <- c(
  prior(student_t(4, 0, 1.75), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = b)
)

dat_ige_19m <- dat_ige |>
  filter(visit_age == "19-month") |>
  mutate(
    ige_cens = if_else(ige == 0, 0.001, ige),
    cens = -1 * as.integer(ige == 0),
    ige_quant = as.numeric(ige >= 0.01),
    ige_sens = as.numeric(ige >= 0.1)
  ) |>
  nest(dat = -allergen) |>
  mutate(
    fit1 = map(
      dat,
      ~ brm(
        ige_cens | cens(cens) ~ trt,
        family = lognormal(),
        prior = pr1,
        data = .x,
        silent = 2,
        seed = 91835
      )
    ),
    gmr1 = map(fit1, ~ exp(as_draws_rvars(.x)$b_trtwP)),
    fit2 = map(
      dat,
      ~ brm(
        ige_quant ~ trt,
        family = bernoulli(),
        prior = pr2,
        data = .x,
        silent = 2,
        seed = 91835
      )
    ),
    rd2 = map(
      fit2,
      ~ tibble(trt = c("aP", "wP")) |>
        add_epred_rvars(.x) |>
        summarise(rd2 = 100 * (.epred[2] - .epred[1]))
    ),
    fit3 = map(
      dat,
      ~ brm(
        ige_sens ~ trt,
        family = bernoulli(),
        prior = pr2,
        data = .x,
        silent = 2,
        seed = 91835
      )
    ),
    rd3 = map(
      fit3,
      ~ tibble(trt = c("aP", "wP")) |>
        add_epred_rvars(.x) |>
        summarise(rd3 = 100 * (.epred[2] - .epred[1]))
    )
  )

tab_desc <- dat_ige_19m |>
  unnest(dat) |>
  mutate(ige = if_else(ige == 0, 0.0005, ige)) |>
  summarise(
    n = sum(!is.na(ige)),
    gmc = gmc(ige, na.rm = TRUE),
    quan = mean(ige_quant, na.rm = TRUE),
    sens = mean(ige_sens, na.rm = TRUE),
    .by = c(allergen, trt)
  ) |>
  mutate(allergen = fct_relevel(allergen, "Total IgE")) |>
  arrange(allergen, trt) |>
  pivot_wider(names_from = trt, values_from = c(n, gmc, quan, sens))
# mutate(gmr = c(NA, gmc[2] / gmc[1]), .by = allergen)

tab_mod <- dat_ige_19m |>
  unnest(gmr1) |>
  unnest(rd2) |>
  unnest(rd3) |>
  mutate(allergen = fct_relevel(allergen, "Total IgE")) |>
  arrange(allergen) |>
  mutate(
    across(
      c(gmr1, rd2, rd3),
      list(
        med = ~ median(.x),
        lo = ~ quantile2(.x, 0.025),
        hi = ~ quantile2(.x, 0.975)
      )
    )
  ) |>
  select(allergen, ends_with(c("med", "lo", "hi")))

# For memory
rm(dat_ige_19m)

left_join(tab_desc, tab_mod, join_by(allergen)) |>
  gt() |>
  cols_align("left", 1) |>
  fmt_number(gmc_aP:last_col(), decimals = 2) |>
  fmt_percent(starts_with(c("quan", "sens")), decimals = 0) |>
  fmt_number(starts_with("rd2"), decimals = 0) |>
  fmt_number(starts_with("rd3"), decimals = 0) |>
  cols_merge(starts_with("gmr1"), pattern = "<<{1} ({2}, {3})>>") |>
  cols_merge(starts_with("rd2"), pattern = "<<{1} ({2}, {3})>>") |>
  cols_merge(starts_with("rd3"), pattern = "<<{1} ({2}, {3})>>") |>
  tab_spanner(ends_with("_aP"), label = "aP") |>
  tab_spanner(ends_with("_wP"), label = "wP") |>
  tab_spanner(ends_with("med"), label = "Median (95% CrI)") |>
  cols_label(
    allergen ~ "IgE Type",
    starts_with("n_") ~ "N",
    starts_with("gmc_") ~ "GMC",
    starts_with("quan_") ~ md("&ge;0.01"),
    starts_with("sens_") ~ md("&ge;0.1"),
    gmr1_med ~ "GMR",
    rd2_med ~ md("Diff. &ge;0.01"),
    rd3_med ~ md("Diff. &ge;0.1")
  ) |>
  tab_options(table.font.size = px(9))
```

\clearpage

## IgG concentrations {#sec-igg}

### Missingness

```{r}
#| label: tbl-overall-missingness
#| tbl-cap: |
#|    Number and proportion of participants with at least one observed response,
#|    and with all planned responses observed.
#| tbl-pos: "H"
#| output: asis
source(here("reports", "R", "functions.R"))
multinorm <- cmdstan_model(file.path("stan", "multinorm.stan"))

dat_igg |>
  select(record_id, trt, any_obs_concentration, all_obs_concentration) |>
  distinct() |>
  tbl_summary(
    include = c(any_obs_concentration, all_obs_concentration),
    label = list(
      any_obs_concentration ~ "At least one IgG response available",
      all_obs_concentration ~ "All planned IgG responses available"
    ),
    by = trt
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = row_type == "label"
    )
  ) |>
  tab_options(
    table.font.size = px(8),
    table.width = pct(80)
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-igg-missingness-dtap
#| tbl-cap:
#|    Missingness pattern of IgG concentrations (anti-PT) by study stage.
#| tbl-pos: H
miss_pattern_trt_tab(filter(dat_igg, antigen == "PT")) |>
  tab_options(
    table.font.size = px(8),
    table.width = pct(60)
  )
```

\clearpage

### Descriptive summaries

Descriptive summaries of 13vPCV serotype-specific IgG concentrations are reported in @tbl-igg-summary-pnps. Summaries of fold-rises and ratios are reported in @tbl-sample-igg-gmr-dtap. The empirical distribution of IgG concentrations in each treatment group are in the appendix.

```{r}
#| label: tbl-igg-summary-dtap
#| tbl-cap: Descriptive summaries for IgG concentrations (IU/mL) by assigned treatment group.
#| output: asis
igg_gmc_summary_tab(
  filter(dat_igg, type != "PnPs")
) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(8)
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```


```{r}
#| label: tbl-sample-igg-gmr-dtap
#| code-summary: Sample summaries
#| tbl-cap: |
#|    Descriptive summary of geometric mean fold-rise (GMFR), percentage with 4-fold rise, geometric mean ratio (GMR), and geometric mean fold-rise ratio (GMFRR) of wP compared to aP. Number in brackets (n) indicates the number of infants contributing to the calculation of the quantity (GMC and GMFR).
#| tbl-pos: "H"
#| output: asis
sample_gmr <- dat_igg |>
  filter(type != "PnPs") |>
  mutate(
    fold_rise = if_else(!(visage %in% c("7-month", "19-month")), NA_real_, fold_rise),
    four_fold = as.numeric(fold_rise >= 4)
  ) |>
  arrange(subjid, antigen, visage) |>
  summarise(
    .by = c(antigen, trt, visage),
    obs_gmc = sum(!is.na(concentration)),
    gmc = gmc(concentration, na.rm = TRUE),
    obs_gmfr = sum(!is.na(fold_rise)),
    gmfr = gmc(fold_rise, na.rm = TRUE),
    fold4 = mean(four_fold, na.rm = TRUE)
  ) |>
  pivot_wider(names_from = trt, values_from = c(obs_gmc, gmc, obs_gmfr, gmfr, fold4)) |>
  mutate(gmr = gmc_wP / gmc_aP, gmfrr = gmfr_wP / gmfr_aP) |>
  mutate(
    obs_gmfr_aP = replace_values(obs_gmfr_aP, 0 ~ NA),
    obs_gmfr_wP = replace_values(obs_gmfr_wP, 0 ~ NA)
  )

gt(sample_gmr, groupname_col = "antigen", rowname_col = "visage") |>
  fmt_integer(starts_with("obs")) |>
  fmt_number(starts_with(c("gmc", "gmr", "gmfr")), decimals = 2) |>
  fmt_percent(starts_with("fold4"), decimals = 0) |>
  cols_merge(
    columns = c(gmc_aP, obs_gmc_aP),
    pattern = "<<{1} ({2})>>"
  ) |>
  cols_merge(
    columns = c(gmc_wP, obs_gmc_wP),
    pattern = "<<{1} ({2})>>"
  ) |>
  cols_merge(
    columns = c(gmfr_aP, obs_gmfr_aP),
    pattern = "<<{1} ({2})>>"
  ) |>
  cols_merge(
    columns = c(gmfr_wP, obs_gmfr_wP),
    pattern = "<<{1} ({2})>>"
  ) |>
  cols_label(
    obs_gmc_aP = "n",
    obs_gmc_wP = "n",
    gmc_aP = "GMC (n)",
    gmc_wP = "GMC (n)",
    gmfr_aP = "GMFR (n)",
    gmfr_wP = "GMFR (n)",
    fold4_aP = "4-fold rise",
    fold4_wP = "4-fold rise",
    gmr ~ "GMR",
    gmfrr ~ "GMFRR"
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{aP} \\\\ N = 150}",
    columns = ends_with("_aP")
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{wP} \\\\ N = 150}",
    columns = ends_with("_wP")
  ) |>
  sub_missing(c(fold4_aP, fold4_wP, gmfrr), missing_text = "") |>
  tab_options(table.font.size = px(9)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

```{r}
#| label: tbl-igg-summary-pnps
#| tbl-cap: Descriptive summaries for IgG concentrations (µg/mL) for 13-valent pneumococcal vaccine serotypes by assigned treatment group.
#| output: asis
igg_gmc_summary_tab(
  filter(dat_igg, type == "PnPs"),
  rn = "visage"
) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(8)
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: fig-igg-dtap
#| fig-cap: |
#|    IgG concentrations at 18 and 19-months of age for pertussis antigens by treatment group.
#| fig-width: 6
#| fig-height: 4
#| include: false
cols <- palette.colors(palette = "Okabe-Ito")[1:2]
dat_igg |>
  filter(type == "DTaP") |>
  ggplot(aes(interaction(trt, visage, sep = "_"), concentration)) +
  facet_wrap(~antigen) +
  geom_boxplot(aes(fill = trt)) +
  geom_hline(aes(yintercept = ref), linetype = 2) +
  scale_y_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  scale_x_discrete(
    "Age and treatment group",
    guide = legendry::guide_axis_nested(
      key = key_range_auto(sep = "_"),
      levels_text = list(element_text(), element_text(angle = 45, hjust = 0, vjust = 0.25))
    )
  ) +
  scale_fill_manual(values = c("aP" = cols[1], "wP" = cols[2])) +
  guides(fill = "none", colour = "none") +
  theme(axis.title = element_text(size = rel(0.9)))
```

```{r}
#| label: fig-igg-pnps
#| fig-cap: |
#|    IgG concentrations at 18-months of age for pneumococcal antigens by treatment group.
#| fig-width: 6
#| fig-height: 8
#| include: false
cols <- palette.colors(palette = "Okabe-Ito")[1:2]
dat_igg |>
  filter(type == "PnPs") |>
  ggplot(aes(interaction(trt, visage, sep = "_"), concentration)) +
  facet_wrap(~antigen, ncol = 3) +
  geom_boxplot(aes(fill = trt)) +
  geom_hline(aes(yintercept = ref), linetype = 2) +
  scale_y_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  scale_x_discrete(
    "Age and treatment group",
    guide = legendry::guide_axis_nested(
      key = key_range_auto(sep = "_"),
      levels_text = list(element_text(), element_text(angle = 45, hjust = 0, vjust = 0.25))
    )
  ) +
  scale_fill_manual(values = c("aP" = cols[1], "wP" = cols[2])) +
  guides(fill = "none", colour = "none") +
  theme(axis.title = element_text(size = rel(0.9)))
```

\clearpage

### Geometric mean ratio and fold-rise ratio

Model estimated geometric mean ratios and geometric mean fold-rise ratios are reported for all antigens in @tbl-igg-gmr and geometric mean ratios are displayed in @fig-gmr-igg-full-model. At 6-months, geometric mean concentrations were lower under wP compared to aP for FHA and PT. For other ages and antigens, concentrations were generally similar (GMR of at least 2/3) or higher under wP compared to aP.

```{r}
#| label: fit-igg-all
#| cache: true
#| cache-lazy: false
frm <- ~ trt + gender + ces + bfed + fborn + fha + parinc_imp
mdat <- dat_igg |>
  select(
    antigen,
    subjid,
    stage,
    visage,
    concentration,
    log_concentration,
    trt,
    gender,
    ces,
    bfed,
    fborn,
    fha,
    parinc_imp
  ) |>
  mutate(
    antigen = droplevels(antigen),
    visage = droplevels(visage)
  ) |>
  complete(
    antigen,
    nesting(subjid, trt, gender, ces, bfed, fborn, fha, parinc_imp),
    visage,
    fill = list(concentration = NA, log_concentration = NA)
  ) |>
  nest(dat = -antigen)
mdat <- mdat |>
  mutate(
    sdat = map(dat, ~ make_stan_data(.x, frm)),
    sfit = map(
      sdat,
      ~ {
        multinorm$sample(
          data = .x,
          seed = 71033,
          chains = 8,
          iter_sampling = 625,
          iter_warmup = 500,
          refresh = 0,
          show_messages = FALSE,
          show_exceptions = FALSE
        )
      }
    ),
    sdrw = map2(
      sfit,
      sdat,
      ~ {
        tmp <- as_draws_rvars(.x)
        rownames(tmp$beta) <- colnames(.y$Y)
        colnames(tmp$beta) <- colnames(.y$X)
        return(tmp)
      }
    ),
    gmr = map(
      sdrw,
      ~ {
        out <- 10^.x$beta[, 2]
        as_tibble(rbind(out), rownames = "age")
      }
    ),
    gmfrr = map(
      sdrw,
      ~ {
        out <- 10^rbind(.x$beta[2, 2] - .x$b[1, 2], .x$beta[4, 2] - .x$b[3, 2])
        rownames(out) <- c("7/6-month (GMFR)", "19/18-month (GMFR)")
        as_tibble(rbind(out), rownames = "age")
      }
    )
  )
gmr <- mdat |>
  select(antigen, gmr) |>
  unnest(gmr) |>
  mutate(group = "GMR") |>
  mutate(age = fct_inorder(age))
gmfrr <- mdat |>
  select(antigen, gmfrr) |>
  unnest(gmfrr) |>
  mutate(group = "GMFFR")
```

```{r}
#| label: tbl-igg-gmr
#| tbl-cap: |
#|    Adjusted IgG GMR (wP versus aP) posterior summaries by serotype/antigen and age.
posterior_gmr_table(filter(bind_rows(gmr, gmfrr))) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(9),
    table.width = pct(60)
  )
```

```{r}
#| label: tbl-igg-gmr-pnp
#| tbl-cap: |
#|    Adjusted IgG GMR (wP versus aP) posterior summaries by serotype/antigen and age.
#| include: false
posterior_gmr_table(filter(bind_rows(gmr, gmfrr), grepl("PnP", antigen))) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(9),
    table.width = pct(60)
  )
```

```{r}
#| label: fig-gmr-igg-full-model
#| fig-cap: Posterior GMR (median and 95% CrI) for wP versus aP by age and antigen.
#| fig-width: 8
#| fig-height: 10
des <- "
IA
JB
KC
LD
ME
NF
OG
PH
Q#
R#
S#
T#
U#
"
p <- gmr |>
  ggplot(aes(y = age, xdist = trtwP)) +
  ggh4x::facet_manual(vars(antigen), design = des, strip.position = "left") +
  stat_pointinterval(.width = 0.95) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  geom_vline(xintercept = 1, linetype = 2) +
  labs(x = "Geometric mean ratio", y = "Age") +
  scale_x_log10(breaks = c(0.4, 0.6, 0.8, 1, 1.2, 1.5, 2, 2.5, 3, 4)) +
  scale_y_discrete(limits = rev)
p
```

\clearpage

### Seroresponses and seroprotection to non-pertussis antigens

Model estimated differences in seroresponse and seroprotection (where applicable) are summarised in @tbl-logistic-regression-igg-seropos-dtap and @fig-gmr-igg-sero-dtap.

```{r}
#| label: model-igg-seropos-all
#| include: false
#| cache: true
#| cache-lazy: false
prior <- c(
  prior(student_t(4, 0, 1.75), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = b),
  prior(exponential(1), class = "sd")
)

mdat_igg_sero <- dat_igg |>
  # filter(type != "PnPs") |>
  mutate(y = positive) |>
  nest(dat = -antigen)

mdat_igg_sero <- mdat_igg_sero |>
  mutate(
    fit = map(
      dat,
      ~ brm(
        y ~ visage * (trt + gender + bfed + fborn + ces + fha + parinc_imp) + (1 | subjid),
        data = filter(.x, !is.na(y)),
        family = bernoulli,
        backend = "cmdstanr",
        prior = prior,
        seed = 71235,
        refresh = 0,
        chains = 8,
        iter = 625,
        warmup = 500
      )
    )
  )

mdat_igg_sero <- mdat_igg_sero |>
  mutate(
    rd2 = map2(
      dat,
      fit,
      ~ {
        epred <- rvar(
          prediction(
            .y,
            .x,
            summarize = FALSE,
            posterior = TRUE,
            raw = TRUE,
            effects = "integrateoutRE",
            k = 100,
            seed = 123
          )$Posterior
        )
        .x |>
          mutate(.epred = epred) |>
          summarise(
            .by = c(visage, trt),
            .epred = rvar_mean(.epred)
          )
      }
    )
  )
```

```{r}
#| label: tbl-logistic-regression-igg-seropos-dtap
#| tbl-cap: |
#|   The model estimated standardised risk, risk difference, and conditional odds ratio for seroresponses and seroprotection (where applicable) comparing recipients of wP as a first dose versus aP as a first dose.
#| tbl-pos: "H"
#| output: asis
d1_igg_sero <- mdat_igg_sero |>
  select(antigen, dat) |>
  unnest(dat) |>
  summarise(
    .by = c(visage, antigen, trt),
    e = sum(y, na.rm = TRUE),
    n = sum(!is.na(y)),
    p = e / n
  )
d2_igg_sero <- mdat_igg_sero |>
  select(antigen, rd2) |>
  unnest(rd2)

d3_igg_sero <- mdat_igg_sero |>
  select(antigen, fit) |>
  mutate(
    b = map(
      fit,
      ~ {
        rvs <- rvar(as_draws_matrix(.x))
        bs <- rvs[grepl("trt", names(rvs))]
        tibble(
          visage = c("6-month", "7-month", "18-month", "19-month"),
          bs = exp(c(bs[1], bs[1] + bs[2], bs[1] + bs[3], bs[1] + bs[4]))
        )
      }
    )
  )

td <- left_join(d1_igg_sero, d2_igg_sero, join_by(visage, antigen, trt)) |>
  rename(y = e) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p, .epred)
  ) |>
  mutate(delta = .epred_wP - .epred_aP) |>
  mutate(across(
    c(.epred_aP, .epred_wP, delta),
    list(
      med = median,
      lo = ~ quantile2(.x, 0.025),
      hi = ~ quantile2(.x, 0.975),
      pr = ~ Pr(.x < 0)
    ),
    .names = "{.fn}_{.col}"
  )) |>
  select(-pr_.epred_aP, -pr_.epred_wP, -.epred_aP, -.epred_wP, -delta) |>
  left_join(
    d3_igg_sero |>
      select(antigen, b) |>
      unnest(b) |>
      mutate(across(
        bs,
        list(
          med = median,
          lo = ~ quantile2(.x, 0.025),
          hi = ~ quantile2(.x, 0.975)
        ),
        .names = "{.fn}_{.col}"
      )) |>
      select(-bs),
    join_by(antigen, visage)
  ) |>
  mutate(visage = factor(visage, levels = c("6-month", "7-month", "18-month", "19-month"))) |>
  arrange(antigen, visage)

td |>
  gt(rowname_col = "visage", groupname_col = "antigen") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_aP, lo_.epred_aP, hi_.epred_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_wP, lo_.epred_wP, hi_.epred_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  cols_merge(
    c(med_bs, lo_bs, hi_bs),
    pattern = "{1} ({2}, {3})"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "\\makecell{\\textbf{aP} \\\\ N = 150}"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "\\makecell{\\textbf{wP} \\\\ N = 150}"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  tab_spanner(
    columns = ends_with("bs"),
    label = "Conditional odds ratio"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_.epred_aP, med_.epred_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  cols_label(
    c(med_bs) ~ "Median (95% CrI)"
  ) |>
  tab_footnote(
    "All models adjust for and standardise over sex, birth order, breastfeeding status, delivery method, family history of atopic disease, and parental income at baseline."
  ) |>
  tab_options(table.font.size = px(8), latex.use_longtable = TRUE) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: fig-gmr-igg-sero-dtap
#| fig-cap: Posterior difference in seroresponse and seroprotection (where applicable) probability (median and 95% CrI) for wP vs. aP by age and antigen.
#| fig-width: 8
#| fig-height: 10
des <- "
IA
JB
KC
LD
ME
NF
OG
PH
Q#
R#
S#
T#
U#
"
p <- d2_igg_sero |>
  pivot_wider(names_from = trt, values_from = .epred) |>
  mutate(delta = wP - aP) |>
  ggplot(aes(y = visage, xdist = delta)) +
  ggh4x::facet_manual(vars(antigen), design = des, strip.position = "left") +
  stat_pointinterval(.width = 0.95) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  geom_vline(xintercept = 0, linetype = 2) +
  labs(x = "Difference in seroresponse/seroprotection probability, wP versus aP", y = "Age") +
  scale_y_discrete(limits = rev)
p
```

\clearpage

### Four-fold rise

Model estimated differences in the probability of a 4-fold rise from 18-month vaccination to 21--35 days after vaccination are summarised in @tbl-igg-four-fold.

```{r}
#| label: model-igg-four-fold
#| include: false
#| cache: true
prior <- c(
  prior(student_t(4, 0, 1.75), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = b)
)

mdat_igg_fr <- dat_igg |>
  filter(visage == "19-month", type != "PnPs", antigen != "HBsAg") |>
  mutate(y = as.numeric(fold_rise >= 4)) |>
  nest(dat = -antigen) |>
  mutate(
    fit = map(
      dat,
      ~ brm(
        y ~ trt + gender + bfed + fborn + ces + fha + parinc_imp,
        data = filter(.x, !is.na(y)),
        family = bernoulli,
        backend = "cmdstanr",
        prior = prior,
        seed = 71235,
        refresh = 0,
        chains = 8,
        iter = 625,
        warmup = 500
      )
    )
  )

mdat_igg_fr <- mdat_igg_fr |>
  mutate(
    rd = map2(
      dat,
      fit,
      ~ .x |>
        add_epred_rvars(.y) |>
        summarise(
          .by = c(visage, trt),
          .epred = rvar_mean(.epred)
        )
    )
  )
```

```{r}
#| label: tbl-igg-four-fold
#| tbl-cap: Model estimated differences (wP compared to aP) in probability of 4-fold rise in IgG concentration from 18-months (pre-vaccination) to 21--35 days after vaccination by antigen.
#| tbl-pos: "H"
d1_igg_fr <- mdat_igg_fr |>
  select(antigen, dat) |>
  unnest(dat) |>
  summarise(
    .by = c(visage, antigen, trt),
    e = sum(y, na.rm = TRUE),
    n = sum(!is.na(y)),
    p = e / n
  )
d2_igg_fr <- mdat_igg_fr |>
  select(antigen, rd) |>
  unnest(rd)
d3_igg_fr <- mdat_igg_fr |>
  select(antigen, fit) |>
  mutate(
    b = map(
      fit,
      ~ {
        rvs <- rvar(as_draws_matrix(.x))
        bs <- rvs[grepl("trt", names(rvs))]
        tibble(
          visage = c("19-month"),
          bs = exp(bs[1])
        )
      }
    )
  )

td <- left_join(d1_igg_fr, d2_igg_fr, join_by(visage, antigen, trt)) |>
  rename(y = e) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p, .epred)
  ) |>
  mutate(delta = .epred_wP - .epred_aP) |>
  mutate(across(
    c(.epred_aP, .epred_wP, delta),
    list(
      med = median,
      lo = ~ quantile2(.x, 0.025),
      hi = ~ quantile2(.x, 0.975),
      pr = ~ Pr(.x < 0)
    ),
    .names = "{.fn}_{.col}"
  )) |>
  select(-pr_.epred_aP, -pr_.epred_wP, -.epred_aP, -.epred_wP, -delta) |>
  left_join(
    d3_igg_fr |>
      select(antigen, b) |>
      unnest(b) |>
      mutate(across(
        bs,
        list(
          med = median,
          lo = ~ quantile2(.x, 0.025),
          hi = ~ quantile2(.x, 0.975)
        ),
        .names = "{.fn}_{.col}"
      )) |>
      select(-bs),
    join_by(antigen, visage)
  ) |>
  mutate(visage = factor(visage, levels = c("6-month", "7-month", "18-month", "19-month"))) |>
  arrange(antigen, visage)

td |>
  select(-visage) |>
  gt(rowname_col = "antigen") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_aP, lo_.epred_aP, hi_.epred_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_.epred_wP, lo_.epred_wP, hi_.epred_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  cols_merge(
    c(med_bs, lo_bs, hi_bs),
    pattern = "{1} ({2}, {3})"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "aP"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "wP"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  tab_spanner(
    columns = ends_with("bs"),
    label = "Conditional odds ratio"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_.epred_aP, med_.epred_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  cols_label(
    c(med_bs) ~ "Median (95% CrI)"
  ) |>
  tab_footnote(
    "All models adjust for and standardise over sex, birth order, breastfeeding status, delivery method, family history of atopic disease, and parental income at baseline."
  ) |>
  tab_options(table.font.size = px(8), latex.use_longtable = TRUE)
```

\clearpage
