---
title: |
  OPTIMUM: OPTimising IMmunisation Using Mixed schedules
subtitle: |
  Clinical Study Report - Statistical Analyses
description: |
  A double-blind, randomised, controlled trial to compare allergic outcomes in children following vaccination with acellular pertussis antigen given at 2 months of age versus whole cell pertussis in the infant vaccine schedule
date: today
author: OPTIMUM study team
bibliography: refs.bib
csl: nature.csl
---

```{r}
#| label: packages-setup
library(here)

source(here("reports/R/_setup.R"))
```

```{r}
#| label: data-setup
source(here("reports/R/_data.R"))
```

\clearpage

```{=latex}
% \listoftoc[List of Abbreviations]{loa}
\phantomsection
```

# List of Abbreviations {.unnumbered}

**aP** acellular pertussis

**DT** diptheria toxoid

**FHA** filamentous haemagglutinin

**FIM2/3** fimbriae agglutinogens 2–3

**HBsAg** hepatitis B surface antigen

**Hib-PRP** capsular polysaccharide of Hib

**OFC** oral food challenge

**PRN** pertactin

**PT** anti-pertussis toxin

**SPT** skin prick test

**TT** anti-tetanus toxoid

**wP** whole-cell pertussis

\clearpage

{{< include _00-summary.qmd >}}

{{< include _01-introduction.qmd >}}

{{< include _02-study-participants.qmd >}}

{{< include _03-outcomes.qmd >}}

{{< include _04-safety.qmd >}}

{{< include _05-reactogenicity.qmd >}}

{{< include _06-immunogenicity.qmd >}}

{{< include _99-appendix.qmd >}}
