---
title: OPTIMUM
subtitle: Clinical Trial Report
date: today
author: James Totterdell
---

```{r}
#| label: packages-and-setup
suppressPackageStartupMessages({
  # Data
  library(here)
  library(qs)
  library(dplyr)
  library(forcats)
  library(lubridate)

  # Figures
  library(ggplot2)
  library(ggdist)
  library(showtext)
  library(patchwork)

  # Tables
  library(gt)
  library(gtsummary)

  # Models
  library(cmdstanr)
  library(posterior)
  library(brms)
  library(tidybayes)
  library(distributional)
})

options(brms.backend = "cmdstanr", mc.cores = 4)

readRenviron(here(".env"))
source(here("R", "util.R"))
source(here("R", "data", "process-combined-data.R"))

sysfonts::font_add(
  family = "TeX Gyre Pagella",
  regular = "texgyrepagella-regular.otf",
  bold = "texgyrepagella-bold.otf",
  italic = "texgyrepagella-italic.otf",
  bolditalic = "texgyrepagella-bolditalic.otf"
)

showtext::showtext_auto()
showtext::showtext_opts(dpi = 244)

theme_set(
  theme_bw(base_size = 10, base_family = "TeX Gyre Pagella") +
    theme(panel.grid.minor = element_blank())
)

theme_gtsummary_compact(font_size = 10)
```

```{r}
#| label: data-setup
com_file <- file.path(Sys.getenv("RDS_PATH"), config::get("combined_data_file"))
dat_raw <- qread(com_file)
```

```{r}
#| label: data-base
dat_base <- get_baseline_data(dat_raw, TRUE)
dat_spt <- get_skin_prick_long(dat_raw)
dat_ecz <- get_eczema_data(dat_raw)
dat_fa <- get_food_allergy(dat_raw)
dat_fc <- get_ofc_long(dat_raw)
```

```{r}
#| label: data-visits
dat_vs <- select_form(dat_raw, "participant_assessment") |>
  select(record_id, visdat, windyn, windreas, windothspec, visit, visage) |>
  mutate(
    visdat = if_else(
      record_id == "5785-83" & visit == 2,
      as_date("2023-03-30"),
      visdat
    )
  )

dat_grid <- dat_base |>
  crossing(
    visage = factor(
      unique(dat_vs$visage),
      levels = c(
        "4-month",
        "6-month",
        "6-month + 72 hrs",
        "7-month",
        "12-month",
        "18-month",
        "19-month"
      )
    )
  ) |>
  arrange(
    str_rank(record_id, numeric = TRUE),
    str_rank(visage, numeric = TRUE)
  ) |>
  mutate(discage = interval(birthdat, discdat) %/% months(1))

dat_grid_vs <- dat_grid |>
  left_join(dat_vs, join_by(record_id, visage)) |>
  mutate(record_id = fct_inorder(record_id), visage = fct_inorder(visage)) |>
  mutate(
    age_weeks = interval(birthdat, visdat) %/% weeks(1),
    age_months = interval(birthdat, visdat) %/% months(1)
  ) |>
  mutate(
    visit_gap = interval(lag(visdat), visdat) %/% days(1),
    .by = record_id
  ) |>
  mutate(visit_attended = !is.na(visdat))
```

```{r}
#| label: data-fhq
dat_fhq <- select_form(dat_raw, "food_and_household_questionnaire")

dat_grid <- dat_base |>
  crossing(
    visit_age = factor(
      unique(dat_fhq$visit_age),
      levels = c(
        "6-week",
        "4-month",
        "6-month",
        "7-month",
        "9-month",
        "12-month",
        "15-month",
        "18-month",
        "19-month"
      )
    )
  ) |>
  arrange(
    str_rank(record_id, numeric = TRUE),
    visit_age
  ) |>
  mutate(discage = interval(birthdat, discdat) %/% months(1)) |>
  select(-fecurr)

dat_grid_fhq <- dat_grid |>
  left_join(dat_fhq, join_by(record_id, visit_age)) |>
  mutate(
    record_id = fct_inorder(record_id),
    visit_age = fct_inorder(visit_age)
  ) |>
  mutate(
    age_weeks = interval(birthdat, fedat) %/% weeks(1),
    age_months = interval(birthdat, fedat) %/% months(1)
  ) |>
  mutate(
    visit_gap = interval(lag(fedat), fedat) %/% days(1),
    .by = record_id
  ) |>
  mutate(fhq_completed = !is.na(fedat)) |>
  filter(
    !(rand_stage == 2 & visit_age %in% c("4-month", "7-month", "19-month"))
  ) |>
  select(
    record_id,
    trt,
    rand_stage,
    discdat,
    disc_age_mth,
    visit_age,
    visit_gap,
    age_weeks,
    age_months,
    all_of(names(dat_fhq)),
    fhq_completed
  )
```

```{r}
#| label: data-additional-processing
dat_spt_pos <- summarise_spt_positive(dat_spt)
dat_spt_pos_any <- dat_spt_pos |>
  summarise(
    any_pos_spt = any(any_spt_pos_1mm & spt_age < 19),
    .by = record_id
  )
dat_spt_analysis <- dat_spt |>
  summarise(
    across(
      spt_0mm:spt_3mm,
      ~ if_else(
        all(!priyn) | all(spt_age > 18),
        NA,
        any(.x == 1 & spt_age < 19, na.rm = TRUE)
      )
    ),
    .by = c(record_id, trt)
  ) |>
  right_join(
    select(
      dat_base,
      record_id,
      rand_site,
      gender,
      fborn
    ),
    join_by(record_id)
  )

dat_ecz_spt <- dat_ecz |>
  left_join(dat_spt_pos_any, join_by(record_id))
```

\clearpage

# Introduction

## Trial outcomes

### Primary outcome (1)

IgE mediated food allergy with evidence of food sensitisation on skin prick test (SPT) by 12 months old and confirmed (where necessary) by medically supervised oral food challenge(s) (OFC). For the purpose of this trial, if either of the following conditions are met, the study participant will be considered to have reached the primary endpoint:

1.  Unequivocal evidence of IgE mediated food allergy, defined as:
    i.  A positive OFC with evidence of sensitisation on SPT to the food of interest *OR*
    ii. History of clinician-confirmed food anaphylaxis, with symptoms affecting at least two of the following: skin, gastrointestinal tract, respiratory tract, cardiovascular system; *AND* evidence of sensitisation on SPT to the food of interest.
2.  Highly probable IgE-mediated food allergy defined as a history of food allergic reaction (consistent with PRACTALL criteria), with evidence of sensitisation on SPT to the food of interest.

We will report on the number and proportion who satisfy this composite definition, and the number and proportion satisfying each of the three sub-components as per a hierarchy in order of 1.i, 1.ii, 2.

### Secondary outcomes (2)

a.  A history of parent-reported clinician-diagnosed new onset eczema by **6 months** *AND* a positive SPT to any allergen by approximately 12 months old.
b.  A history of parent-reported clinician-diagnosed new onset eczema by **12 months** *AND* a positive SPT to any allergen by approximately 12 months old.
c.  A SPT **\>1mm** greater than the negative control to at least one allergen by approximately 12 months of age
d.  A SPT **≥3mm** greater than the negative control to at least one allergen by approximately 12 months of age

### Details

For both food allergy and eczema, sensitisation on (positive) SPT was defined to be a result more than 1mm greater than negative control by 18 months of age.

For eczema, the age requirement (by 6 or 12 months of age) was determined by the "age at which allergy first occurred" (`outageval`). To be considered "new-onset" eczema, this age at which allergy first occurred needed to be after the infants enrolment date. Infants with clinician-diagnosed eczema first occurring prior to enrolment should be excluded as it's not possible for them to have "new-onset eczema".

## Analyses

### Primary outcome

The primary analysis was to assume independent Binomial models for occurence of food allergy under each treatment group with uniform priors on the outcome probabilities. A secondary analysis was to assume a logistic regression model for occurence of food allergy adjusting for site, sex, ...

### Secondary clinical outcomes

Analyses of secondary outcomes are analagous to those for the primary outcome.

\clearpage

# Enrolment, Characteristics and Completion

## Enrolment

```{r}
#| label: fig-enrolment-overall
#| fig-cap: Cumulative enrolment.
#| fig-height: 3
#| fig-width: 6
dat_enrol <- dat_base |>
  select(record_id, randdattim) |>
  mutate(randdate = as_date(randdattim)) |>
  arrange(randdate) |>
  summarise(n = n(), .by = randdate) |>
  mutate(cn = cumsum(n))

ggplot(
  dat_enrol,
  aes(randdate, cn)
) +
  geom_step() +
  scale_x_date("Calendar time", breaks = "3 months", date_labels = "%d %b %y") +
  scale_y_continuous("Number enrolled", breaks = seq(0, 1000, 100)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

\clearpage

## Characteristics

The baseline characteristics of participants are reported in @tbl-baseline-characteristics.

```{r}
#| label: tbl-baseline-characteristics
#| tbl-cap: Baseline characteristics of trial participants.
#| tbl-pos: "H"
#| output: "asis"
dat_base |>
  tbl_summary(
    include = c(gender, parinc, delt, wgt, len, inhc),
    by = trt,
    missing_text = "(Missing)"
  ) |>
  add_n() |>
  add_overall(last = TRUE) |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```

\clearpage

## Study completion

The study completion status as reported in the database is summarised in @tbl-discharge-outcome. Most of the *early termination - consent withdrawal* and *early termination - protocol violation/deviation* were due to the first 300 PCH participants who protocolised for the additional blood collections, a number of which withdrew consent for blood collection at the 18-month and 19-month visits (@tbl-discharge-outcome-site). Data was generally still collected for these participants with the exception of these blood collections.

```{r}
#| label: tbl-discharge-outcome
#| code-summary: Study completion
#| tbl-cap: Number of early study withdrawals and associated reasons.
#| tbl-pos: "H"
#| output: "asis"
tbl_study_termination <- function(dat) {
  st_by_trt <- dat |>
    count(streas, stetrreas, trt) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = trt) |>
    pivot_wider(names_from = trt, values_from = c(n, p))

  st_ovr <- dat |>
    count(streas, stetrreas) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n)) |>
    rename(n_ovr = n, p_ovr = p)

  left_join(st_by_trt, st_ovr, join_by(streas, stetrreas)) |>
    gt() |>
    cols_align(
      align = "left",
      columns = c(streas, stetrreas)
    ) |>
    fmt_percent(starts_with("p_"), decimals = 1) |>
    cols_merge(c(n_aP, p_aP), pattern = "{1} ({2})") |>
    cols_merge(c(n_wP, p_wP), pattern = "{1} ({2})") |>
    cols_merge(c(n_ovr, p_ovr), pattern = "{1} ({2})") |>
    cols_label(
      streas ~ md("**Participant status**"),
      stetrreas ~ md("**Reason**"),
      n_aP ~ "\\makecell{\\textbf{aP} \\\\ N = 486}",
      n_wP ~ "\\makecell{\\textbf{wP} \\\\ N = 486}",
      n_ovr ~ "\\makecell{\\textbf{Overall} \\\\ N = 972}"
    ) |>
    tab_options(table.font.size = "10px") |>
    sub_missing() |>
    tab_style(
      style = cell_text(align = "left"),
      locations = cells_footnotes()
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(n_aP:n_ovr)
    )
}

tbl_study_termination(dat_base) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-discharge-outcome-site
#| code-summary: Study completion by site
#| tbl-cap: Number of early study withdrawals and associated reasons by study site.
#| tbl-pos: "H"
#| output: "asis"
tbl_study_termination_by <- function(dat, by = rand_site) {
  st_by_trt <- dat |>
    count({{ by }}, streas, stetrreas, trt) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = c({{ by }}, trt)) |>
    pivot_wider(names_from = trt, values_from = c(n, p))

  st_ovr <- dat |>
    count({{ by }}, streas, stetrreas) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = {{ by }}) |>
    rename(n_ovr = n, p_ovr = p)

  left_join(st_by_trt, st_ovr, join_by({{ by }}, streas, stetrreas)) |>
    group_by({{ by }}) |>
    gt() |>
    cols_align(
      align = "left",
      columns = c(streas, stetrreas)
    ) |>
    fmt_percent(starts_with("p_"), decimals = 1) |>
    cols_merge(c(n_aP, p_aP), pattern = "{1} ({2})") |>
    cols_merge(c(n_wP, p_wP), pattern = "{1} ({2})") |>
    cols_merge(c(n_ovr, p_ovr), pattern = "{1} ({2})") |>
    cols_label(
      streas ~ md("**Participant status**"),
      stetrreas ~ md("**Reason**"),
      n_aP ~ "\\makecell{\\textbf{aP} \\\\ N = 486}",
      n_wP ~ "\\makecell{\\textbf{wP} \\\\ N = 486}",
      n_ovr ~ "\\makecell{\\textbf{Overall} \\\\ N = 972}"
    ) |>
    tab_options(table.font.size = "10px") |>
    sub_missing() |>
    tab_style(
      style = cell_text(align = "left"),
      locations = cells_footnotes()
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(n_aP:n_ovr)
    )
}

tbl_study_termination_by(dat_base) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

There were 17 participants where contact was lost at or before 15 months of age (based on `discdat`) (@tbl-listing-early-withdrawal).

```{r}
#| label: tbl-listing-early-withdrawal
#| tbl-cap: |
#|   Listing of infants with last contact occuring at or prior to 15 months of age.
dat_base |>
  select(record_id, stetrreas, birthdat, disc_age_mth) |>
  left_join(
    filter(dat_grid_vs, visage == "12-month") |>
      select(record_id, visit_attended, windreas, windothspec) |>
      rename(
        visit_attend_12m = visit_attended,
        reas_12m = windreas,
        oth_12m = windothspec
      ),
    join_by(record_id)
  ) |>
  left_join(
    filter(dat_grid_vs, visage == "18-month") |>
      select(record_id, visit_attended, windreas, windothspec) |>
      rename(
        visit_attend_18m = visit_attended,
        reas_18m = windreas,
        oth_18m = windothspec
      ),
    join_by(record_id)
  ) |>
  arrange(disc_age_mth) |>
  filter(disc_age_mth <= 15) |>
  select(-starts_with("reas"), -starts_with("oth")) |>
  gt() |>
  cols_label(
    stetrreas = "Withdrawal reason",
    disc_age_mth = "Age at last contact",
    visit_attend_12m = "Attend 12-month visit",
    visit_attend_18m = "Attend 18-month visit"
  ) |>
  fmt_date(
    c(birthdat)
  ) |>
  sub_missing() |>
  tab_options(table.font.size = px(10), table.width = pct(100))
```

\clearpage

## Missed visits

```{r}
#| label: tbl-stage1-visit-attendance
#| tbl-cap: |
#|   Number and proportion of infants who attended each visit in stage 1
#|   (based on presence of "visit date").
#|    6-month + 72 hour was optional visit.
#| tbl-pos: "H"
#| output: "asis"
vs_by_trt <- dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = c(trt, visage)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

vs_ovr <- dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = visage
  )

left_join(vs_by_trt, vs_ovr) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    wP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 150}"
  ) |>
  cols_align(align = "left", columns = visage) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-stage2-visit-attendance
#| tbl-cap: |
#|   Number and proportion of infants who attended each visit in stage 2
#|   (based on presence of "visit date").
#|   19-month visit was for first 150 at PCH only.
#| tbl-pos: "H"
#| output: "asis"
vs_by_trt <- dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = c(trt, visage)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

vs_ovr <- dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = visage
  )

left_join(vs_by_trt, vs_ovr) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    wP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 822}"
  ) |>
  cols_align(align = "left", columns = visage) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

## Missed food and household questionnaires

```{r}
#| label: tbl-stage1-fhq-completion
#| tbl-cap: |
#|   Number and proportion of infants who completed food and household questionnaire for each occasion in stage 1.
#| tbl-pos: "H"
#| output: "asis"
fhq_by_trt <- dat_grid_fhq |>
  filter(rand_stage == 1) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = c(trt, visit_age)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

fhq_ovr <- dat_grid_fhq |>
  filter(rand_stage == 1) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = visit_age
  )

left_join(fhq_by_trt, fhq_ovr) |>
  gt() |>
  cols_label(
    visit_age ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    wP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 150}"
  ) |>
  cols_align(align = "left", columns = visit_age) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-stage2-fhq-completion
#| tbl-cap: |
#|   Number and proportion of infants who completed food and household questionnaire for each occasion in stage 2.
#| tbl-pos: "H"
#| output: "asis"
fhq_by_trt <- dat_grid_fhq |>
  filter(rand_stage == 2) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = c(trt, visit_age)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

fhq_ovr <- dat_grid_fhq |>
  filter(rand_stage == 2) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = visit_age
  )

left_join(fhq_by_trt, fhq_ovr) |>
  gt() |>
  cols_label(
    visit_age ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    wP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 822}"
  ) |>
  cols_align(align = "left", columns = visit_age) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

# Outcomes

All of the trial outcomes depend upon a positive SPT to either a specific allergen or to any allergen. Therefore, this section first reports on the SPT data, and outcomes 2c and 2d, followed by the primary (food allergy) and other secondary (eczema) outcomes.

## Skin prick tests

```{r}
#| label: spt-no-records
no_spt_record <- dat_spt |>
  filter(row_number() == 1, .by = record_id) |>
  filter(!priyn)
```

There were `r nrow(no_spt_record)` participants with no documented SPTs (@tbl-spt-listing-no-records). In many cases this was due to early withdrawal from the study, with the remaining due to non-consent from the parent.

```{r}
#| label: tbl-spt-listing-no-records
#| tbl-cap: |
#|   Listing of infants with no SPT records.
#| tbl-pos: "H"
no_spt_record |>
  select(record_id, rand_site, streas, dis_age) |>
  gt() |>
  cols_label(
    dis_age = "Last contact age (months)"
  ) |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(100)
  )
```

All other participants had a record for at least the scheduled SPT, and in a few cases additional unscheduled SPT (@tbl-spt-number).

```{r}
#| label: tbl-spt-number
#| tbl-cap: |
#|   Number and percentage of infants with scheduled and unscheduled skin prick tests.
#| tbl-pos: "H"
#| output: "asis"
spt_nums <- dat_spt |>
  filter(row_number() == 1, .by = c(record_id, trt, spt_occasion, spt_num)) |>
  summarise(
    spt_num_sch = sum(spt_occasion == "scheduled" & priyn, na.rm = TRUE),
    spt_num_unsch = sum(
      spt_occasion %in% c("unscheduled", "other") & priyn,
      na.rm = TRUE
    ),
    .by = c(record_id, trt)
  )
tbl_summary(
  spt_nums,
  by = trt,
  include = c(spt_num_sch, spt_num_unsch),
  label = c(spt_num_sch ~ "Scheduled SPT", spt_num_unsch ~ "Unscheduled SPTs")
) |>
  add_overall(last = TRUE) |>
  remove_footnote_header(columns = everything()) |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

For the outcomes, a requirement was positive SPT by 18 months of age. Age of SPT in months was derived from `birthdat` (date of birth) and `pridat` (date of skin prick test) or in some cases `univdat` (date of unscheduled visit). Most SPTs were performed by 13 months of age, although some were delayed, with a few occurring after 18-months of age (@fig-spt-age). There were 5 infants who did have an SPT, but none by 18-months of age (@tbl-spt-listing-after-18months).

```{r}
#| label: fig-spt-age
#| fig-cap: |
#|   Age at skin prick test by type (scheduled or unscheduled).
#| fig-height: 2.5
dat_spt |>
  filter(row_number() == 1, .by = c(record_id, trt, spt_occasion, spt_num)) |>
  filter(!is.na(spt_age)) |>
  mutate(
    spt_occasion = str_to_sentence(if_else(
      spt_occasion == "other",
      "unscheduled",
      spt_occasion
    ))
  ) |>
  ggplot(aes(spt_age)) +
  facet_wrap(~spt_occasion, scales = "free_y") +
  geom_bar() +
  scale_x_continuous("Age (months)", breaks = seq(0, 30, 2)) +
  scale_y_continuous(
    "Number of skin prick tests",
    breaks = scales::pretty_breaks(10)
  )
```

```{r}
#| label: tbl-spt-listing-after-18months
#| tbl-cap: Listing of SPTs after 18-months of age.
#| tbl-pos: "H"
dat_spt |>
  filter(
    row_number() == 1,
    .by = c(record_id, trt, spt_occasion, spt_num, spt_age)
  ) |>
  filter(spt_age > 18) |>
  select(record_id, spt_occasion, spt_num, spt_age) |>
  mutate(
    spt_occasion = str_to_sentence(if_else(
      spt_occasion == "other",
      "unscheduled",
      spt_occasion
    ))
  ) |>
  distinct() |>
  gt() |>
  cols_label(
    spt_occasion ~ "SPT type",
    spt_num = "SPT occasion",
    spt_age = "Age (months)"
  ) |>
  tab_options(table.font.size = px(10), table.width = pct(50))
```

@tbl-spt-positive reports on the number and percentage of infants who had a SPT result of \>1mm and ≥3mm relative to negative control for at least one tested allergen (outcomes 2c and 2d). Any infants with no SPT records by 18 months of age are reported as missing.

```{r}
#| label: tbl-spt-positive
#| tbl-cap: |
#|   Number and percentage of infants with a positive SPT by 18 months of age.
#|   Infants with no SPT records counted as missing.
#| tbl-pos: "H"
#| results: "asis"
dat_spt |>
  summarise(
    across(
      spt_0mm:spt_3mm,
      ~ if_else(
        all(!priyn) | all(spt_age > 18),
        NA,
        any(.x == 1 & spt_age < 19, na.rm = TRUE)
      )
    ),
    .by = c(record_id, trt)
  ) |>
  tbl_summary(
    by = trt,
    include = c(spt_1mm, spt_3mm),
    label = c(
      # spt_0mm ~ "Any SPT >0mm",
      spt_1mm ~ "Any SPT >1mm vs. neg control",
      spt_3mm ~ "Any SPT ≥3mm vs neg. control"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(70)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

### Analyses

```{r}
#| label: spt-analysis
dat_outcome <- dat_spt_analysis |>
  pivot_longer(
    spt_0mm:spt_3mm,
    names_pattern = "spt_([0-3]mm)",
    names_to = "size",
    values_to = "outcome"
  ) |>
  filter(!is.na(outcome)) |>
  summarise(
    n = n(),
    y = sum(outcome),
    p = y / n,
    a = 1 + y,
    b = 1 + n - y,
    post = dist_beta(a, b),
    .by = c(trt, size)
  ) |>
  mutate(p_sup = c(NA, beta_ineq_sim(a, b)), .by = size)
diff_outcome <- dat_outcome |>
  mutate(post_drws = generate(post, 1e6)) |>
  select(size, trt, post_drws) |>
  unnest(post_drws) |>
  group_by(size, trt) |>
  mutate(.draw = 1:n()) |>
  ungroup() |>
  pivot_wider(names_from = "trt", values_from = "post_drws") |>
  mutate(delta = wP - aP)

prior <- c(
  prior(logistic(0, 1), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = "b")
)

fit_outcome <- dat_spt_analysis |>
  pivot_longer(
    spt_0mm:spt_3mm,
    names_pattern = "spt_([0-3]mm)",
    names_to = "size",
    values_to = "outcome"
  ) |>
  nest(-size) |>
  mutate(
    fit = map(
      data,
      \(x) {
        brm(
          outcome ~ trt + rand_site + gender + fborn,
          data = x,
          family = bernoulli(),
          prior = prior,
          warmup = 500,
          iter = 5500,
          refresh = 0,
          silent = 2
        )
      }
    ),
    rd = map2(data, fit, \(x, y) {
      wgt <- c(rdo(rdirichlet(1, rep(1, nrow(x))), ndraws = ndraws(y)))
      x |>
        mutate(trt = "aP") |>
        add_epred_rvars(y, value = ".epred_aP") |>
        mutate(trt = "wP") |>
        add_epred_rvars(y, value = ".epred_wP") |>
        mutate(wgt = wgt) |>
        summarise(
          .rd_empirical = rvar_mean(.epred_wP - .epred_aP),
          .rd_bootstrap = rvar_sum((.epred_wP - .epred_aP) * wgt)
        )
    })
  )
```

```{r}
#| label: fig-posterior-density-spt
#| fig-cap: |
#|    Posterior density for difference in probability of positive skin prick test by size,
#|    unadjusted primary model.
#| fig-height: 5
#| fig-width: 5
diff_outcome |>
  group_by(size) |>
  nest_rvars() |>
  ggplot(aes(xdist = delta)) +
  facet_wrap(~size, ncol = 1) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(
    fill = "Interval",
    x = "Difference in food allergy probability, wP - aP",
    y = "Density"
  )
```

```{r}
#| label: fig-posterior-difference-spt-adjusted
#| fig-cap: |
#|   Posterior difference in probability of food allergy for wP compared to aP, adjusted model.
#| fig-height: 5
#| fig-width: 5
fit_outcome |>
  unnest(rd) |>
  ggplot(aes(xdist = .rd_bootstrap)) +
  facet_wrap(~size, ncol = 1) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(
    fill = "Interval",
    x = "Difference in food allergy probability, wP - aP",
    y = "Density"
  )
```

\clearpage

## Food allergy

The primary source of information for food allergy was the `outcome_report` form. Reported outcomes were cross checked against the food and household questionnaire form (`food_and_household_questionnaire`) and the oral food challenge form (`food_challenge`).

Age at first reaction was derived from `outageval`. If this was missing then `birthdat` (date of birth) and `outdiagdat` (clinician-diagnosis date) was used to derive age in months.

In some cases participants may have withdrawn from the study without sufficient follow-up for food allergy to have been identified. The report makes no allowance for "missing" food allergy status, with absence of clinician diagnosed allergy (with or without adequate follow-up) counted as not meeting the primary outcome. Only 17 participants had an age of last contact 15 months or younger and may have been those most likely to have insufficient follow-up.

The following sections provide a summary of the number of infants who met each criteria. These infants are line listed in the appendix.

### Oral food challenge

There were `r nrow(filter(dat_fa, outalltp == "Oral food challenge"))` outcome records for food allergy diagnoses resulting from oral food challenge amongst `r length(unique(filter(dat_fa, outalltp == "Oral food challenge")$record_id))` infants. All reactions reportedly first occurred by 18 months of age and all diagnoses occurred by 18 months of age. All records had a matching positive skin prick test to the reported allergen and so all 22 infants met the primary outcome according to the definition (1.i).

```{r}
#| label: check-positive-spt-ofc
#| include: false
dat_po_ofc <- dat_fa |>
  filter(outalltp == "Oral food challenge") |>
  select(
    record_id,
    trt,
    outdiagdat,
    outageval_months2,
    out_age_months,
    outfrafood
  ) |>
  left_join(
    dat_spt_pos |>
      filter(any_spt_pos_1mm) |>
      select(record_id, spt_age, spt_pos_1mm),
    join_by(record_id),
    relationship = "many-to-many"
  ) |>
  rowwise() |>
  mutate(
    splt_foods = gsub("( |, )", "|", spt_pos_1mm),
    match_food = str_match(tolower(outfrafood), splt_foods)[1]
  ) |>
  ungroup() |>
  summarise(
    any_food_match = any(!is.na(match_food) & outageval_months2 <= 18),
    matched_foods = paste(
      unique(match_food[!is.na(match_food) & outageval_months2 <= 18]),
      collapse = ", "
    ),
    .by = record_id
  ) |>
  rename(po_ofc = any_food_match, po_ofc_match = matched_foods)
```

### Anaphylaxis

There were `r nrow(filter(dat_fa, outalltp == "Anaphylaxis"))` outcome records for food allergy diagnoses resulting from anaphylaxis amongst `r length(unique(filter(dat_fa, outalltp == "Anaphylaxis")$record_id))` infants. Two of these infants (`record_id: 68, 5785-283`) were diagnosed after 18 months of age with first reported reaction occuring after 18 months of age. One infant (`record_id = 8644-34`) had no skin prick test for the reported allergen. So of the infants with analphylaxis reported, 3 did not meet the criteria for the primary outcome according to the definition (1.ii).

```{r}
#| label: check-positive-spt-ana
#| include: false
dat_po_ana <- dat_fa |>
  filter(outalltp == "Anaphylaxis") |>
  select(
    record_id,
    trt,
    outdiagdat,
    outageval_months2,
    out_age_months,
    outfrafood
  ) |>
  left_join(
    dat_spt_pos |>
      filter(any_spt_pos_1mm) |>
      select(record_id, spt_age, spt_pos_1mm),
    join_by(record_id),
    relationship = "many-to-many"
  ) |>
  rowwise() |>
  mutate(
    splt_foods = gsub("( |, )", "|", spt_pos_1mm),
    match_food = str_match(tolower(outfrafood), splt_foods)[1]
  ) |>
  ungroup() |>
  summarise(
    any_food_match = any(!is.na(match_food) & outageval_months2 <= 18),
    matched_foods = paste(
      unique(match_food[!is.na(match_food) & outageval_months2 <= 18]),
      collapse = ", "
    ),
    .by = record_id
  ) |>
  rename(po_ana = any_food_match, po_ana_match = matched_foods)
```

### Highly probable IgE mediated food allergy

There were `r nrow(filter(dat_fa, outalltp == "Highly probably IgE mediated food allergy"))` outcome records for food allergy diagnoses resulting from highly probable IgE mediated allergy amongst `r length(unique(filter(dat_fa, outalltp == "Highly probably IgE mediated food allergy")$record_id))` infants. For one infant (`record_id: 68`), the age at first reaction and age at diagnosis were after 18 months of age. Four infants (`record_id: 75, 5785-239, 5785-243, 5785-331`) had no positive SPT to the reported allergen. Of the infants with IgE-mediated food allergy reported, 5 did not meet the primary outcome according to the definition (2.).

```{r}
#| label: check-positive-spt-ige
#| include: false
# Cross check food challenge data
summ_fc <- dat_fc |>
  filter(ofcresult != "Not challenged") |>
  summarise(
    ofcfood = paste(ofcfood, collapse = ", "),
    ofcresult = paste(ofcresult, collapse = ", "),
    ofcage = paste(ofcage, collapse = ", "),
    .by = record_id
  )

dat_po_ige <- dat_fa |>
  filter(outalltp == "Highly probably IgE mediated food allergy") |>
  select(
    record_id,
    trt,
    outdiagdat,
    outageval_months2,
    out_age_months,
    outfrafood
  ) |>
  left_join(
    dat_spt_pos |>
      filter(any_spt_pos_1mm) |>
      select(record_id, spt_age, spt_pos_1mm),
    join_by(record_id),
    relationship = "many-to-many"
  ) |>
  rowwise() |>
  mutate(
    splt_foods = gsub("( |, )", "|", spt_pos_1mm),
    outfrafood_sub = gsub("Dairy", "milk", outfrafood),
    match_food = str_match(tolower(outfrafood_sub), splt_foods)[1]
  ) |>
  ungroup() |>
  summarise(
    any_food_match = any(!is.na(match_food) & outageval_months2 <= 18),
    matched_foods = paste(
      unique(match_food[!is.na(match_food) & outageval_months2 <= 18]),
      collapse = ", "
    ),
    .by = record_id
  ) |>
  rename(po_ige = any_food_match, po_ige_match = matched_foods)

# dat_po_ige |>
#   left_join(summ_fc, join_by(record_id)) |>
#   print(n = Inf)
```

### Primary outcome

```{r}
#| label: primary-outcome-data
dat_po <- dat_base |>
  left_join(dat_po_ofc, join_by(record_id)) |>
  left_join(dat_po_ana, join_by(record_id)) |>
  left_join(dat_po_ige, join_by(record_id)) |>
  mutate(
    po_ofc = replace_na(po_ofc, 0),
    po_ana = replace_na(po_ana, 0),
    po_ige = replace_na(po_ige, 0),
    po = po_ofc | po_ana | po_ige,
    po_criteria = case_when(
      po_ofc == 1 ~ "Criteria 1.i",
      po_ana == 1 ~ "Criteria 1.ii",
      po_ige == 1 ~ "Criteria 2.",
      .default = "None"
    )
  ) |>
  select(
    record_id,
    rand_site,
    trt,
    gender,
    parinc,
    delt,
    fecurr,
    fborn,
    starts_with("po")
  )
```

@tbl-primary-outcome-criteria reports on the number of participants who met each of the primary outcome criteria (1.i, 1.ii, and 2) as outlined in the sections above.

@tbl-primary-outcome-overall reports the total number of infants who met the primary outcome, and the primary reason grouped hierarchically according to the ordering in the definition (i.e. OFC → Anaphylaxis → IgE mediated) (irrespective of their timing).

```{r}
#| label: tbl-primary-outcome-criteria
#| tbl-cap: |
#|   Number and percentage of infants who met each primary outcome condition and treatment group.
#| tbl-pos: "H"
#| results: "asis"
dat_po |>
  tbl_summary(
    by = trt,
    include = c(po_ofc, po_ana, po_ige),
    label = list(
      po_ofc ~ "Met condition 1.i.",
      po_ana ~ "Met condition 1.ii.",
      po_ige ~ "Met condition 2."
    ),
    missing = "no"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(70)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-primary-outcome-overall
#| tbl-cap: |
#|   Number and percentage of infants who met the primary outcome by primayr reason and treatment group.
#| tbl-pos: "H"
#| results: "asis"
dat_po |>
  tbl_summary(
    by = trt,
    include = c(po, po_criteria),
    label = list(
      po ~ "Met primary outcome",
      po_criteria ~ "Reason"
    ),
    missing = "no"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      filter(label != "None")
  ) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(70)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

### Analyses

```{r}
#| label: primary-outcome-analysis
po_sum <- dat_po |>
  summarise(
    n = n(),
    y = sum(po),
    p = y / n,
    a = 1 + y,
    b = 1 + n - y,
    post = dist_beta(a, b),
    .by = trt
  ) |>
  mutate(p_sup = c(NA, beta_ineq_sim(a, b)))
po_diff <- po_sum |>
  mutate(post_drws = generate(post, 1e6)) |>
  select(trt, post_drws) |>
  unnest(post_drws) |>
  group_by(trt) |>
  mutate(.draw = 1:n()) |>
  ungroup() |>
  pivot_wider(names_from = "trt", values_from = "post_drws") |>
  mutate(delta = wP - aP)

prior <- c(
  prior(logistic(0, 1), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = "b")
)
fit <- brm(
  po ~ trt + rand_site + gender + fborn,
  data = dat_po,
  family = bernoulli(),
  prior = prior,
  warmup = 500,
  iter = 5500,
  refresh = 0,
  silent = 2
)

wgt <- c(rdo(rdirichlet(1, rep(1, 972)), ndraws = ndraws(fit)))
dat_rd_po <- dat_po |>
  mutate(trt = "aP") |>
  add_epred_rvars(fit, value = ".epred_aP") |>
  mutate(trt = "wP") |>
  add_epred_rvars(fit, value = ".epred_wP") |>
  mutate(wgt = wgt) |>
  summarise(
    .rd_empirical = rvar_mean(.epred_wP - .epred_aP),
    .rd_bootstrap = rvar_sum((.epred_wP - .epred_aP) * wgt)
  )
```

```{r}
#| label: fig-posterior-density
#| fig-cap: Posterior density for difference in probability of primary outcome, unadjusted primary model.
#| fig-height: 2
#| fig-width: 5
po_diff |>
  nest_rvars() |>
  ggplot(aes(xdist = delta)) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(
    fill = "Interval",
    x = "Difference in food allergy probability, wP - aP",
    y = "Density"
  )
```

```{r}
#| label: fig-posterior-difference-adjusted
#| fig-cap: Posterior difference in probability of food allergy for wP compared to aP, adjusted model.
#| fig-height: 2
ggplot(dat_rd_po, aes(xdist = .rd_bootstrap)) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(
    fill = "Interval",
    x = "Difference in food allergy probability, wP - aP",
    y = "Density"
  )
```

\clearpage

## Eczema

The primary source of information for eczema was the `outcome_report` form. Reported outcomes were cross checked against the food and household questionnaire form and medical history forms. Infants whose first reaction date for clinician-diagnosed eczema was prior to enrolment were considered to have "pre-existing" eczema and were excluded from the analysis set. There were `r sum(dat_ecz$out_ecz_preexisting)` such infants.

The number and percentage of infants with any clinician-diagnosed eczema is reported in @tbl-eczema-trt. For the protocol defined outcomes \[2a. and 2b.\], the eczema diagnosis needed to be new-onset and accompanied by a positive SPT to at least one allergen by approximately 12 months of age. These are reported in @tbl-eczema-trt-newonset.

```{r}
#| label: tbl-eczema-trt
#| tbl-cap: |
#|   Number and percentage of infants with any clinician-diagnosed eczema.
#|   Those with missing under cases with positive SPT were due to no SPT records being available.
#| tbl-pos: "H"
#| results: "asis"

# dat_ecz |>
#   filter(out_ecz_preexisting) |>
#   summarise(n = sum(out_ecz_preexisting), .by = trt)

tbl0 <- dat_ecz_spt |>
  mutate(
    ecz = ecz,
    ecz_6m = ecz & outageval_months2 <= 6,
    ecz_12m = ecz & outageval_months2 <= 12,
    ecz_18m = ecz & outageval_months2 <= 18
  ) |>
  tbl_summary(
    by = trt,
    include = c(ecz, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      filter(label != "None")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(55)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-eczema-trt-newonset
#| tbl-cap: |
#|   Number and percentage of infants with new-onset clinician-diagnosed eczema.
#|   To be new-onset the first reaction date needed to be after the infants randomisation date.
#|   Those with missing under cases with positive SPT were due to no SPT records being available.
#| tbl-pos: "H"
#| results: "asis"
tbl1 <- dat_ecz_spt |>
  filter(!out_ecz_preexisting) |>
  mutate(
    ecz = ecz & !out_ecz_preexisting,
    ecz_6m = ecz & !out_ecz_preexisting & outageval_months2 <= 6,
    ecz_12m = ecz & !out_ecz_preexisting & outageval_months2 <= 12,
    ecz_18m = ecz & !out_ecz_preexisting & outageval_months2 <= 18
  ) |>
  tbl_summary(
    by = trt,
    include = c(ecz, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**")

# With positive SPT
tbl2 <- dat_ecz_spt |>
  filter(!out_ecz_preexisting) |>
  mutate(
    ecz = ecz & !out_ecz_preexisting & any_pos_spt,
    ecz_6m = ecz & !out_ecz_preexisting & outageval_months2 <= 6 & any_pos_spt,
    ecz_12m = ecz &
      !out_ecz_preexisting &
      outageval_months2 <= 12 &
      any_pos_spt,
    ecz_18m = ecz & !out_ecz_preexisting & outageval_months2 <= 18 & any_pos_spt
  ) |>
  tbl_summary(
    by = trt,
    include = c(ecz, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**")

tbl_stack(
  list(tbl1, tbl2),
  group_header = c(
    "New-onset cases",
    "New-onset cases with positive SPT"
  )
) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      filter(label != "None")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(55)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

# Safety

## Serious adverse events

A listing of serious adverse events are reported in @tbl-listing-sae. These are any events reported either in the SAE reporting log (`sae_reporting_log`, stage 2 only) or on the adverse events form (`adverse_events`) with `aetype = "SAE"`. In some cases, a record was on the SAE log but not the adverse event log.

```{r}
#| label: tbl-listing-sae
#| tbl-cap: |
#|   Listing of serious adverse events.
#|   Missing values were due to no adverse event record for corresponding SAE log record.
#| tbl-pos: "H"
dat_sae <- dat_raw |>
  select_form("serious_adverse_events") |>
  select(record_id, saestdat, saemeddra) |>
  distinct() |>
  arrange(str_rank(record_id, numeric = TRUE))

dat_ae_sae <- dat_raw |>
  select_form("adverse_events") |>
  filter(aetype == "SAE") |>
  select(record_id, aecode1, aesev, aetymead, aerel, aestdat, aeenddat) |>
  rename(saestdat = aestdat) |>
  arrange(str_rank(record_id, numeric = TRUE))

dat_saes <- full_join(dat_sae, dat_ae_sae, join_by(record_id, saestdat)) |>
  arrange(str_rank(record_id, numeric = TRUE), saestdat) |>
  mutate(saemeddra = if_else(is.na(aecode1), saemeddra, aecode1))

dat_saes |>
  left_join(select(dat_base, record_id, visdat1)) |>
  mutate(aeday = interval(visdat1, saestdat) %/% days(1)) |>
  select(-visdat1, -saestdat, -aeenddat, -aecode1) |>
  gt() |>
  cols_label(saemeddra ~ "MedDRA code", aeday ~ "Days after visit 1") |>
  sub_missing() |>
  tab_options(table.font.size = px(8), table.width = pct(100))
```

```{r}
#| label: tbl-listing-ae
#| tbl-cap: |
#|   Listing of adverse events.
#| tbl-pos: "H"
#| include: false
dat_ae <- dat_base |>
  left_join(
    dat_raw |>
      select_form("adverse_events"),
    join_by(record_id)
  )
```

\clearpage

# Appendix {.appendix}

## Primary outcome listings

```{r}
#| label: tbl-listing-ofc-matched-foods
#| tbl-cap: Listing of infants with oral food challenge outcome and SPT positive allergens.
#| tbl-pos: "H"
dat_po_ofc |>
  select(record_id, po_ofc_match) |>
  gt() |>
  cols_label(po_ofc_match ~ "SPT positive foods") |>
  tab_options(table.font.size = px(9), table.width = pct(40))
```

```{r}
#| label: tbl-listing-ana-matched-foods
#| tbl-cap: Listing of infants with anaphylaxis outcome and SPT positive allergens.
#| tbl-pos: "H"
dat_po_ana |>
  select(record_id, po_ana_match) |>
  gt() |>
  cols_label(po_ana_match ~ "SPT positive foods") |>
  tab_options(table.font.size = px(9), table.width = pct(40))
```

```{r}
#| label: tbl-listing-ige-matched-foods
#| tbl-cap: Listing of infants with IgE-mediated food allergy outcome and SPT positive allergens.
#| tbl-pos: "H"
dat_po_ige |>
  select(record_id, po_ige_match) |>
  gt() |>
  cols_label(po_ige_match ~ "SPT positive foods") |>
  tab_options(table.font.size = px(9), table.width = pct(40))
```

\clearpage

## Secondary outcome listings

```{r}
#| label: tbl-listing-eczema
#| tbl-cap: Listing of infants with eczema diagnosis and positive SPT to any allergen.
#| tbl-pos: "H"
dat_ecz_spt |>
  filter(!out_ecz_preexisting, ecz, any_pos_spt) |>
  select(record_id, outageval_months2) |>
  gt() |>
  fmt_number(decimals = 0) |>
  cols_label(
    outageval_months2 ~ "Age at first reaction (months)"
  ) |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(40),
    latex.use_longtable = TRUE
  )
```