---
title: OPTIMUM
subtitle: Clinical Trial Report
---

```{r}
#| label: packages-and-setup
suppressPackageStartupMessages({
  # Data
  library(here)
  library(qs)
  library(dplyr)
  library(forcats)
  library(lubridate)

  # Figures
  library(ggplot2)
  library(showtext)

  # Tables
  library(gt)
  library(gtsummary)
})

readRenviron(here(".env"))
source(here("R", "util.R"))
source(here("R", "data", "process-combined-data.R"))

sysfonts::font_add(
  family = "TeX Gyre Pagella",
  regular = "texgyrepagella-regular.otf",
  bold = "texgyrepagella-bold.otf",
  italic = "texgyrepagella-italic.otf",
  bolditalic = "texgyrepagella-bolditalic.otf"
)

showtext::showtext_auto()
showtext::showtext_opts(dpi = 244)

theme_set(
  theme_bw(base_size = 10, base_family = "TeX Gyre Pagella") +
    theme(panel.grid.minor = element_blank())
)

theme_gtsummary_compact(font_size = 10)
```

```{r}
#| label: data-setup
com_file <- file.path(Sys.getenv("RDS_PATH"), config::get("combined_data_file"))
dat_raw <- qread(com_file)
```

```{r}
#| label: data-base
dat_base <- get_baseline_data(dat_raw, TRUE)
```

```{r}
#| label: data-visits
dat_vs <- select_form(dat_raw, "participant_assessment") |>
  select(record_id, visdat, windyn, windreas, windothspec, visit, visage)
```

\clearpage

# Introduction

\clearpage

# Enrolment, Characteristics and Completion

## Enrolment

```{r}
#| label: fig-enrolment-overall
#| fig-cap: Cumulative enrolment.
#| fig-height: 3
#| fig-width: 6
dat_enrol <- dat_base |>
  select(record_id, randdattim) |>
  mutate(randdate = as_date(randdattim)) |>
  arrange(randdate) |>
  summarise(n = n(), .by = randdate) |>
  mutate(cn = cumsum(n))

ggplot(
  dat_enrol,
  aes(randdate, cn)
) +
  geom_step() +
  scale_x_date("Calendar time", breaks = "3 months", date_labels = "%d %b %y") +
  scale_y_continuous("Number enrolled", breaks = seq(0, 1000, 100)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

\clearpage

## Characteristics

The baseline characteristics of participants are reported in @tbl-baseline-characteristics.

```{r}
#| label: tbl-baseline-characteristics
#| tbl-cap: Baseline characteristics of trial participants.
#| tbl-pos: "H"
#| output: "asis"
dat_base |>
  tbl_summary(
    include = c(gender, parinc, delt, wgt, len, inhc),
    by = trt,
    missing_text = "(Missing)"
  ) |>
  add_n() |>
  add_overall(last = TRUE) |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```

\clearpage

## Study Completion

The study completion status as reported in the database is summarised in @tbl-discharge-outcome. Most of the *early termination - consent withdrawal* and *early termination - protocol violation/deviation* are due to the first 300 PCH participants with the additional blood collections, a number of which withdrew consent for blood collection at the 18-month visit (@tbl-discharge-outcome-site). Data was generally still collected for these participants with the exception of the blood collection.

```{r}
#| label: tbl-discharge-outcome
#| code-summary: Study completion
#| tbl-cap: Number of early study withdrawals and associated reasons.
#| tbl-pos: "H"
#| output: "asis"
tbl_study_termination <- function(dat) {
  st_by_trt <- dat |>
    count(streas, stetrreas, trt) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = trt) |>
    pivot_wider(names_from = trt, values_from = c(n, p))

  st_ovr <- dat |>
    count(streas, stetrreas) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n)) |>
    rename(n_ovr = n, p_ovr = p)

  left_join(st_by_trt, st_ovr, join_by(streas, stetrreas)) |>
    gt() |>
    cols_align(
      align = "left",
      columns = c(streas, stetrreas)
    ) |>
    fmt_percent(starts_with("p_"), decimals = 1) |>
    cols_merge(c(n_aP, p_aP), pattern = "{1} ({2})") |>
    cols_merge(c(n_wP, p_wP), pattern = "{1} ({2})") |>
    cols_merge(c(n_ovr, p_ovr), pattern = "{1} ({2})") |>
    cols_label(
      streas ~ md("**Participant status**"),
      stetrreas ~ md("**Reason**"),
      n_aP ~ "\\makecell{\\textbf{aP} \\\\ N = 486}",
      n_wP ~ "\\makecell{\\textbf{wP} \\\\ N = 486}",
      n_ovr ~ "\\makecell{\\textbf{Overall} \\\\ N = 972}"
    ) |>
    tab_options(table.font.size = "10px") |>
    sub_missing() |>
    tab_style(
      style = cell_text(align = "left"),
      locations = cells_footnotes()
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(n_aP:n_ovr)
    )
}

tbl_study_termination(dat_base) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-discharge-outcome-site
#| code-summary: Study completion by site
#| tbl-cap: Number of early study withdrawals and associated reasons by study site.
#| tbl-pos: "H"
#| output: "asis"
tbl_study_termination_by <- function(dat, by = rand_site) {
  st_by_trt <- dat |>
    count({{ by }}, streas, stetrreas, trt) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = c({{ by }}, trt)) |>
    pivot_wider(names_from = trt, values_from = c(n, p))

  st_ovr <- dat |>
    count({{ by }}, streas, stetrreas) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = {{ by }}) |>
    rename(n_ovr = n, p_ovr = p)

  left_join(st_by_trt, st_ovr, join_by({{ by }}, streas, stetrreas)) |>
    group_by({{ by }}) |>
    gt() |>
    cols_align(
      align = "left",
      columns = c(streas, stetrreas)
    ) |>
    fmt_percent(starts_with("p_"), decimals = 1) |>
    cols_merge(c(n_aP, p_aP), pattern = "{1} ({2})") |>
    cols_merge(c(n_wP, p_wP), pattern = "{1} ({2})") |>
    cols_merge(c(n_ovr, p_ovr), pattern = "{1} ({2})") |>
    cols_label(
      streas ~ md("**Participant status**"),
      stetrreas ~ md("**Reason**"),
      n_aP ~ "\\makecell{\\textbf{aP} \\\\ N = 486}",
      n_wP ~ "\\makecell{\\textbf{wP} \\\\ N = 486}",
      n_ovr ~ "\\makecell{\\textbf{Overall} \\\\ N = 972}"
    ) |>
    tab_options(table.font.size = "10px") |>
    sub_missing() |>
    tab_style(
      style = cell_text(align = "left"),
      locations = cells_footnotes()
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(n_aP:n_ovr)
    )
}

tbl_study_termination_by(dat_base) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

## Missed visits

```{r}
dat_grid <- dat_base |>
  crossing(visage = unique(dat_vs$visage)) |>
  arrange(
    str_rank(record_id, numeric = TRUE),
    str_rank(visage, numeric = TRUE)
  ) |>
  mutate(discage = interval(birthdat, discdat) %/% months(1))

dat_grid_vs <- dat_grid |>
  left_join(dat_vs, join_by(record_id, visage)) |>
  mutate(record_id = fct_inorder(record_id), visage = fct_inorder(visage)) |>
  mutate(
    age_weeks = interval(birthdat, visdat) %/% weeks(1),
    age_months = interval(birthdat, visdat) %/% months(1)
  ) |>
  mutate(
    visit_gap = interval(lag(visdat), visdat) %/% days(1),
    .by = record_id
  ) |>
  mutate(visit_attended = !is.na(visdat))
```

```{r}
#| label: tbl-stage1-visit-attendance
#| tbl-cap: |
#|   Number and proportion of infants who attended each visit in stage 1.
#|   (based on presence of "visit date").
#    6-month + 72 hour was optional visit.
#| tbl-pos: "H"
#| output: "asis"
vs_by_trt <- dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = c(trt, visage)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

vs_ovr <- dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = visage
  )

left_join(vs_by_trt, vs_ovr) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    wP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 150}"
  ) |>
  cols_align(align = "left", columns = visage) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-stage2-visit-attendance
#| tbl-cap: |
#|   Number and proportion of infants who attended each visit in stage 2.
#|   (based on presence of "visit date").
#|   19-month visit was for first 150 at PCH only.
#| tbl-pos: "H"
#| output: "asis"
vs_by_trt <- dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = c(trt, visage)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

vs_ovr <- dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = visage
  )

left_join(vs_by_trt, vs_ovr) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    wP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 822}"
  ) |>
  cols_align(align = "left", columns = visage) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```


\clearpage

# Primary Outcome

\clearpage

# Secondary Outcomes

\clearpage

# Safety
