---
title: OPTIMUM
subtitle: Clinical Trial Report
date: today
author: James Totterdell
---

```{r}
#| label: packages-and-setup
suppressPackageStartupMessages({
  # Data
  library(here)
  library(qs)
  library(dplyr)
  library(forcats)
  library(lubridate)
  library(labelled)
  library(arrow)

  # Figures
  library(ggplot2)
  library(ggdist)
  library(showtext)
  library(patchwork)
  library(legendry)
  library(ggh4x)
  library(scales)

  # Tables
  library(gt)
  library(gtsummary)

  # Models
  library(cmdstanr)
  library(posterior)
  library(brms)
  library(tidybayes)
  library(distributional)
  library(brmsmargins)
})

options(brms.backend = "cmdstanr", mc.cores = 8)

readRenviron(here(".env"))
source(here("R", "util.R"))
source(here("R", "data", "process-combined-data.R"))

sysfonts::font_add(
  family = "TeX Gyre Pagella",
  regular = "texgyrepagella-regular.otf",
  bold = "texgyrepagella-bold.otf",
  italic = "texgyrepagella-italic.otf",
  bolditalic = "texgyrepagella-bolditalic.otf"
)

showtext::showtext_auto()
showtext::showtext_opts(dpi = 244)

theme_set(
  theme_bw(base_size = 10, base_family = "TeX Gyre Pagella") +
    theme(panel.grid.minor = element_blank())
)

theme_gtsummary_compact(font_size = 10)

integer_breaks <- function(n = 5, ...) {
  fxn <- function(x) {
    breaks <- floor(pretty(x, n, ...))
    names(breaks) <- attr(breaks, "labels")
    breaks
  }
  return(fxn)
}
```

```{r}
#| label: data-setup
com_file <- file.path(Sys.getenv("RDS_PATH"), config::get("combined_data_file"))
dat_raw <- qread(com_file)
```

```{r}
#| label: data-base
dat_base <- get_baseline_data(dat_raw, TRUE)
dat_spt <- get_skin_prick_long(dat_raw)
dat_ecz <- get_eczema_data(dat_raw)
dat_fa <- get_food_allergy(dat_raw)
dat_fc <- get_ofc_long(dat_raw)

# For 4629-92 we do not consider their walnut allergy as valid
# therefore, this outcome is excluded here
dat_fa <- dat_fa |>
  filter(!(outfrafood == "walnut" & record_id == "4629-92"))

# Fix wrong coding of delt in stage 2
dat_base <- dat_base |>
  mutate(
    delt = case_match(delt, "Forceps/ vacuum assisted delivery" ~ "Forceps/vacuum assisted delivery", .default = delt)
  )
```

```{r}
#| label: data-visits
dat_vs <- select_form(dat_raw, "participant_assessment") |>
  select(record_id, visdat, windyn, windreas, windothspec, visit, visage) |>
  mutate(
    visdat = if_else(
      record_id == "5785-83" & visit == 2,
      as_date("2023-03-30"),
      visdat
    )
  )

dat_grid <- dat_base |>
  crossing(
    visage = factor(
      unique(dat_vs$visage),
      levels = c(
        "4-month",
        "6-month",
        "6-month + 72 hrs",
        "7-month",
        "12-month",
        "18-month",
        "19-month"
      )
    )
  ) |>
  arrange(
    str_rank(record_id, numeric = TRUE),
    str_rank(visage, numeric = TRUE)
  ) |>
  mutate(discage = interval(birthdat, discdat) %/% months(1))

dat_grid_vs <- dat_grid |>
  left_join(dat_vs, join_by(record_id, visage)) |>
  mutate(record_id = fct_inorder(record_id), visage = fct_inorder(visage)) |>
  mutate(
    age_weeks = interval(birthdat, visdat) %/% weeks(1),
    age_months = interval(birthdat, visdat) %/% months(1)
  ) |>
  mutate(
    visit_gap = interval(lag(visdat), visdat) %/% days(1),
    .by = record_id
  ) |>
  mutate(visit_attended = !is.na(visdat))
```

```{r}
#| label: data-fhq
dat_fhq <- select_form(dat_raw, "food_and_household_questionnaire")

dat_grid <- dat_base |>
  crossing(
    visit_age = factor(
      unique(dat_fhq$visit_age),
      levels = c(
        "6-week",
        "4-month",
        "6-month",
        "7-month",
        "9-month",
        "12-month",
        "15-month",
        "18-month",
        "19-month"
      )
    )
  ) |>
  arrange(
    str_rank(record_id, numeric = TRUE),
    visit_age
  ) |>
  mutate(discage = interval(birthdat, discdat) %/% months(1)) |>
  select(-fecurr, -fedog, -fecat, -fedcyn)

dat_grid_fhq <- dat_grid |>
  left_join(dat_fhq, join_by(record_id, visit_age)) |>
  mutate(
    record_id = fct_inorder(record_id),
    visit_age = fct_inorder(visit_age)
  ) |>
  mutate(
    age_weeks = interval(birthdat, fedat) %/% weeks(1),
    age_months = interval(birthdat, fedat) %/% months(1)
  ) |>
  mutate(
    visit_gap = interval(lag(fedat), fedat) %/% days(1),
    .by = record_id
  ) |>
  mutate(fhq_completed = !is.na(fedat)) |>
  filter(
    !(rand_stage == 2 & visit_age %in% c("4-month", "7-month", "19-month"))
  ) |>
  select(
    record_id,
    trt,
    rand_stage,
    discdat,
    disc_age_mth,
    visit_age,
    visit_gap,
    age_weeks,
    age_months,
    all_of(names(dat_fhq)),
    fhq_completed
  )
```

```{r}
#| label: data-additional-processing
dat_spt_pos <- summarise_spt_positive(dat_spt)
dat_spt_pos_any <- dat_spt_pos |>
  summarise(
    any_pos_spt = any(any_spt_pos_1mm & spt_age < 19),
    .by = record_id
  )

# Dataset used for SPT outcome analyses
dat_spt_analysis <- dat_spt |>
  summarise(
    across(
      spt_0mm:spt_3mm,
      ~ if_else(
        all(!priyn) | all(spt_age > 18),
        NA,
        any(.x == 1 & spt_age < 19, na.rm = TRUE)
      )
    ),
    .by = c(record_id, trt)
  ) |>
  right_join(
    select(
      dat_base,
      record_id,
      rand_site,
      gender,
      bfed,
      fborn,
      parinc_imp,
      ces,
      fha
    ),
    join_by(record_id)
  )

dat_ecz_spt <- dat_ecz |>
  left_join(dat_spt_pos_any, join_by(record_id))

# Dataset used for Eczema outcome analyses
dat_ecz_analysis <- dat_ecz_spt |>
  filter(!out_ecz_preexisting) |>
  mutate(
    ecz_6m = ecz &
      !out_ecz_preexisting &
      outageval_months2 <= 6 &
      any_pos_spt,
    ecz_12m = ecz &
      !out_ecz_preexisting &
      outageval_months2 <= 12 &
      any_pos_spt,
    ecz_18m = ecz &
      !out_ecz_preexisting &
      outageval_months2 <= 18 &
      any_pos_spt
  ) |>
  select(record_id, trt, ecz_6m, ecz_12m, ecz_18m) |>
  left_join(
    dat_base |>
      select(record_id, rand_site, gender, bfed, fborn, parinc_imp, ces, fha),
    join_by(record_id)
  )
```

\clearpage

# Introduction

## Purpose

This report summarises the data and results of the analyses for main outcomes of the OPTIMUM
trial. 
The report includes results for:

- baseline demographics
- skin prick tests
- food allergy
- eczema
- safety reporting

For full details of the trial processes, objectives, and outcomes refer to the protocol
and statistical analysis plan.

## Interventions

Participants were randomised 1:1 to receive either:

- Combined Diphtheria (D)-Tetanus (T)-acellular Pertussis (aP), Hepatitis B (HepB),
Inactivated Poliovirus (IPV) and Haemophilus influenzae type b (Hib) vaccine (DTaPHepB-
IPV-Hib) administered at approximately 2 months, together with the standard
schedule of vaccines (referred to as 3-aP + aP or aP for short), OR
- Combined Diphtheria (D)-Tetanus (T)-whole cell Pertussis (wP), Hepatitis B (HepB)
and Haemophilus influenzae type b (Hib) vaccine (DTwP-HepB- Hib) in place of
DTaP-HepB-IPV-Hib at approximately 2 months, together with the standard schedule
of vaccines (referred to as 1-wP/2-aP + aP or wP for short).

## Trial design

Originally, a sample size of 3,000 participants was planned. Under the adaptive design, as outlined in the protocol and statistical analysis plan, we estimated the trial had 85% power to detect a reduction in food allergy from 10% to 7% with type I error of 5%. If $\delta$ denotes the difference in probability of food allergy under each intervention, then trial success required that $\delta$ was very likely to be less than zero, specifically that $\Pr(\delta < 0) > 0.95$.

The maximum sample size was subsequently revised down to 2,000 participants due to feasability constraints. The trial was stopped early due to logistical issues with 972 participants randomised.

## Trial outcomes

### Primary outcome (1)

IgE mediated food allergy with evidence of food sensitisation on skin prick test (SPT) by 12 months old and confirmed (where necessary) by medically supervised oral food challenge(s) (OFC). For the purpose of this trial, if either of the following conditions are met, the study participant will be considered to have reached the primary endpoint:

1.  Unequivocal evidence of IgE mediated food allergy, defined as:
    i.  A positive OFC with evidence of sensitisation on SPT to the food of interest *OR*
    ii. History of clinician-confirmed food anaphylaxis, with symptoms affecting at least two of the following: skin, gastrointestinal tract, respiratory tract, cardiovascular system; *AND* evidence of sensitisation on SPT to the food of interest.
2.  Highly probable IgE-mediated food allergy defined as a history of food allergic reaction (consistent with PRACTALL criteria), with evidence of sensitisation on SPT to the food of interest.

We will report on the number and proportion who satisfy this composite definition, and the number and proportion satisfying each of the three sub-components as per a hierarchy in order of 1.i, 1.ii, 2.

### Secondary outcomes (2)

a.  A history of parent-reported clinician-diagnosed new onset eczema by **6 months** *AND* a positive SPT to any allergen by approximately 12 months old.
b.  A history of parent-reported clinician-diagnosed new onset eczema by **12 months** *AND* a positive SPT to any allergen by approximately 12 months old.
c.  A SPT **\>1mm** greater than the negative control to at least one allergen by approximately 12 months of age
d.  A SPT **≥3mm** greater than the negative control to at least one allergen by approximately 12 months of age

### Further Details

The protocol defined a panel of allergens to be tested. In addition to these, further allergens could also be tested at parental request or at discretion of the site PI. When the outcomes say "at least one allergen" [i.e. for eczema for positive SPT] we mean any allergen, including any additional allergens tested beyond the panel defined in the protocol.

For both food allergy and eczema, sensitisation on (positive) SPT was defined to be a result more than 1mm greater than negative control by 18 months of age. For food allergy, the positive SPT needed to match the specific food allergen. 

For eczema, the age requirement (by 6 or 12 months of age) was determined by the "age at which allergy first occurred" (`outageval`). To be considered "new-onset" eczema, this age at which allergy first occurred needed to be after the infants enrolment date. Infants with clinician-diagnosed eczema first occurring prior to enrolment were to be excluded as it's not possible for them to have "new-onset eczema".


## Analyses

### Primary outcome

The primary analysis was to assume independent Binomial models for occurence of food allergy under each treatment group with uniform priors on the outcome probabilities. 

::: {.callout-note}

## Primary model

Let $\theta_{\text{wP}}$ and $\theta_{\text{aP}}$ denote the probability of the outcome amongst infants who follow the respective strategy and $\delta = \theta_{\text{wP}} - \theta_{\text{aP}}$ the difference in probablity. If $n_{\text{wP}}$ and $n_{\text{aP}}$ infants were assigned to each respective strategy, with $y_{\text{wP}}$ and $y_{\text{aP}}$ events observed, than from our assumed model we estimate that $\theta_j \sim \text{Beta}(1 + y_j, 1 + n_j - y_j)$ represents our uncertainty and $\delta$ the difference between these.

:::

A secondary analysis was to assume a logistic regression model for occurence of food allergy adjusting for baseline factors: site, sex, breastfeeding status, birth order, parental income, mode of delivery, and family history of atopic disease.

### Secondary clinical outcomes

Analyses of secondary outcomes are analagous to those for the primary outcome, both the primary and secondary models.

\clearpage

# Enrolment, Characteristics and Completion

## Enrolment

There were `r nrow(dat_base)` participants randomised, `r nrow(dat_base)/2` per treatment group. The first participant was randomised on `r as_date(min(dat_base$randdattim))` and the final participant was randomised on `r as_date(max(dat_base$randdattim))`. A summary of enrolment over calendar time is presented in @fig-enrolment-overall

```{r}
#| label: fig-enrolment-overall
#| fig-cap: Cumulative enrolment.
#| fig-height: 3
#| fig-width: 6
dat_enrol <- dat_base |>
  select(record_id, randdattim) |>
  mutate(randdate = as_date(randdattim)) |>
  arrange(randdate) |>
  summarise(n = n(), .by = randdate) |>
  mutate(cn = cumsum(n))

ggplot(
  dat_enrol,
  aes(randdate, cn)
) +
  geom_step() +
  scale_x_date("Calendar time", breaks = "3 months", date_labels = "%d %b %y") +
  scale_y_continuous("Number enrolled", breaks = seq(0, 1000, 100)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

\clearpage

## Characteristics

The baseline characteristics of participants by their randomly assigned treatment group are reported in @tbl-baseline-characteristics. Characteristics were similar between the groups.

```{r}
#| label: tbl-baseline-characteristics
#| tbl-cap: Baseline characteristics of trial participants.
#| tbl-pos: "H"
#| output: "asis"
dat_base |>
  tbl_summary(
    include = c(
      # Demographics and infant history
      rand_stage,
      rand_site,
      gender,
      v1_age_days,
      parinc,
      fborn,
      fecurr,
      fecat,
      fedog,
      # Birth history
      mgrav_fac,
      mpar_fac,
      delt,
      ipab,
      neoab,
      gestadelw,
      wgt,
      len,
      inhc,
      aps1_fac,
      aps5_fac,
      # Maternal immunisation
      inmpertvp,
      vacad,
      prevpert,
      minfvp,
      # Family history
      fha
    ),
    label = list(
      rand_stage ~ "Protocol stage",
      rand_site ~ "Participating site",
      v1_age_days ~ "Age at enrolment (days)",
      fborn ~ "First born",
      mgrav_fac ~ "Maternal gravidity",
      mpar_fac ~ "Maternal parity",
      aps1_fac ~ "Apgar score (1 minute)",
      aps5_fac ~ "Apgar score (5 minutes)",
      prevpert ~ "Other maternal pertussis vaccination (last 5 years)",
      fha ~ "Family history of atopy",
      fecat ~ "Cat ownership",
      fedog ~ "Dog ownership"
    ),
    by = trt,
    missing_text = "(Missing)",
    digits = list(v1_age_days ~ 0, gestadelw ~ 0)
  ) |>
  add_n() |>
  add_overall(last = TRUE) |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  cols_width(label ~ pct(30)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = row_type == "label"
    )
  ) |>
  tab_options(
    table.width = pct(100),
    table.font.size = px(9),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```

\clearpage

## Study completion

The study completion status as reported in the database is summarised in @tbl-discharge-outcome and by site in @tbl-discharge-outcome-site. Most of the *early termination - consent withdrawal* and *early termination - protocol violation/deviation* were due to the first 300 PCH participants who were scheduled for the additional blood collections, a number of which withdrew consent for blood collection at the 18-month and 19-month visits (@tbl-discharge-outcome-site). Data was generally still collected for these participants with the exception of these blood collections.

```{r}
#| label: tbl-discharge-outcome
#| code-summary: Study completion
#| tbl-cap: Number of early study withdrawals and associated reasons.
#| tbl-pos: "H"
#| output: "asis"
tbl_study_termination <- function(dat) {
  st_by_trt <- dat |>
    count(streas, stetrreas, trt) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = trt) |>
    pivot_wider(names_from = trt, values_from = c(n, p))

  st_ovr <- dat |>
    count(streas, stetrreas) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n)) |>
    rename(n_ovr = n, p_ovr = p)

  left_join(st_by_trt, st_ovr, join_by(streas, stetrreas)) |>
    gt() |>
    cols_align(
      align = "left",
      columns = c(streas, stetrreas)
    ) |>
    fmt_percent(starts_with("p_"), decimals = 1) |>
    cols_merge(c(n_aP, p_aP), pattern = "{1} ({2})") |>
    cols_merge(c(n_wP, p_wP), pattern = "{1} ({2})") |>
    cols_merge(c(n_ovr, p_ovr), pattern = "{1} ({2})") |>
    cols_label(
      streas ~ md("**Participant status**"),
      stetrreas ~ md("**Reason**"),
      n_aP ~ "\\makecell{\\textbf{aP} \\\\ N = 486}",
      n_wP ~ "\\makecell{\\textbf{wP} \\\\ N = 486}",
      n_ovr ~ "\\makecell{\\textbf{Overall} \\\\ N = 972}"
    ) |>
    tab_options(table.font.size = "10px") |>
    sub_missing() |>
    tab_style(
      style = cell_text(align = "left"),
      locations = cells_footnotes()
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(n_aP:n_ovr)
    )
}

tbl_study_termination(dat_base) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-discharge-outcome-site
#| code-summary: Study completion by site
#| tbl-cap: Number of early study withdrawals and associated reasons by study site.
#| tbl-pos: "H"
#| output: "asis"
tbl_study_termination_by <- function(dat, by = rand_site) {
  st_by_trt <- dat |>
    count({{ by }}, streas, stetrreas, trt) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = c({{ by }}, trt)) |>
    pivot_wider(names_from = trt, values_from = c(n, p))

  st_ovr <- dat |>
    count({{ by }}, streas, stetrreas) |>
    mutate(across(c(streas), ~ fct_na_value_to_level(., level = "Missing"))) |>
    mutate(p = n / sum(n), .by = {{ by }}) |>
    rename(n_ovr = n, p_ovr = p)

  left_join(st_by_trt, st_ovr, join_by({{ by }}, streas, stetrreas)) |>
    group_by({{ by }}) |>
    gt() |>
    cols_align(
      align = "left",
      columns = c(streas, stetrreas)
    ) |>
    fmt_percent(starts_with("p_"), decimals = 1) |>
    cols_merge(c(n_aP, p_aP), pattern = "{1} ({2})") |>
    cols_merge(c(n_wP, p_wP), pattern = "{1} ({2})") |>
    cols_merge(c(n_ovr, p_ovr), pattern = "{1} ({2})") |>
    cols_label(
      streas ~ md("**Participant status**"),
      stetrreas ~ md("**Reason**"),
      n_aP ~ "\\makecell{\\textbf{aP} \\\\ N = 486}",
      n_wP ~ "\\makecell{\\textbf{wP} \\\\ N = 486}",
      n_ovr ~ "\\makecell{\\textbf{Overall} \\\\ N = 972}"
    ) |>
    tab_options(table.font.size = "10px") |>
    sub_missing() |>
    tab_style(
      style = cell_text(align = "left"),
      locations = cells_footnotes()
    ) |>
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(n_aP:n_ovr)
    )
}

tbl_study_termination_by(dat_base) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

There were 17 participants where contact was lost at or before 15 months of age (based on `discdat`) (@tbl-listing-early-withdrawal).

```{r}
#| label: tbl-listing-early-withdrawal
#| tbl-cap: |
#|   Listing of infants with last contact occuring at or prior to 15 months of age.
dat_base |>
  select(record_id, stetrreas, birthdat, disc_age_mth) |>
  left_join(
    filter(dat_grid_vs, visage == "12-month") |>
      select(record_id, visit_attended, windreas, windothspec) |>
      rename(
        visit_attend_12m = visit_attended,
        reas_12m = windreas,
        oth_12m = windothspec
      ),
    join_by(record_id)
  ) |>
  left_join(
    filter(dat_grid_vs, visage == "18-month") |>
      select(record_id, visit_attended, windreas, windothspec) |>
      rename(
        visit_attend_18m = visit_attended,
        reas_18m = windreas,
        oth_18m = windothspec
      ),
    join_by(record_id)
  ) |>
  arrange(disc_age_mth) |>
  filter(disc_age_mth <= 15) |>
  select(-starts_with("reas"), -starts_with("oth")) |>
  gt() |>
  cols_label(
    stetrreas = "Withdrawal reason",
    disc_age_mth = "Age at last contact",
    visit_attend_12m = "Attend 12-month visit",
    visit_attend_18m = "Attend 18-month visit"
  ) |>
  fmt_date(
    c(birthdat)
  ) |>
  sub_missing() |>
  tab_options(table.font.size = px(10), table.width = pct(100))
```

\clearpage

## Missed visits

```{r}
#| label: tbl-stage1-visit-attendance
#| tbl-cap: |
#|   Number and proportion of infants who attended each visit in stage 1
#|   (based on presence of "visit date").
#|    6-month + 72 hour was optional visit.
#| tbl-pos: "H"
#| output: "asis"
vs_by_trt <- dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = c(trt, visage)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

vs_ovr <- dat_grid_vs |>
  filter(rand_stage == 1) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = visage
  )

left_join(vs_by_trt, vs_ovr) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    wP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 150}"
  ) |>
  cols_align(align = "left", columns = visage) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-stage2-visit-attendance
#| tbl-cap: |
#|   Number and proportion of infants who attended each visit in stage 2
#|   (based on presence of "visit date").
#|   19-month visit was for first 150 at PCH only.
#| tbl-pos: "H"
#| output: "asis"
vs_by_trt <- dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = c(trt, visage)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

vs_ovr <- dat_grid_vs |>
  filter(rand_stage == 2, visage %in% c("12-month", "18-month", "19-month")) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(visit_attended),
      100 * mean(visit_attended)
    ),
    .by = visage
  )

left_join(vs_by_trt, vs_ovr) |>
  gt() |>
  cols_label(
    visage ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    wP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 822}"
  ) |>
  cols_align(align = "left", columns = visage) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

## Missed food and household questionnaires

```{r}
#| label: tbl-stage1-fhq-completion
#| tbl-cap: |
#|   Number and proportion of infants who completed food and household questionnaire for each occasion in stage 1.
#| tbl-pos: "H"
#| output: "asis"
fhq_by_trt <- dat_grid_fhq |>
  filter(rand_stage == 1) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = c(trt, visit_age)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

fhq_ovr <- dat_grid_fhq |>
  filter(rand_stage == 1) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = visit_age
  )

left_join(fhq_by_trt, fhq_ovr) |>
  gt() |>
  cols_label(
    visit_age ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    wP ~ "\\makecell{\\textbf{aP} \\\\ N = 75}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 150}"
  ) |>
  cols_align(align = "left", columns = visit_age) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-stage2-fhq-completion
#| tbl-cap: |
#|   Number and proportion of infants who completed food and household questionnaire for each occasion in stage 2.
#| tbl-pos: "H"
#| output: "asis"
fhq_by_trt <- dat_grid_fhq |>
  filter(rand_stage == 2) |>
  summarise(
    lab = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = c(trt, visit_age)
  ) |>
  pivot_wider(names_from = trt, values_from = lab)

fhq_ovr <- dat_grid_fhq |>
  filter(rand_stage == 2) |>
  summarise(
    Overall = sprintf(
      "%i (%.0f%%)",
      sum(fhq_completed),
      100 * mean(fhq_completed)
    ),
    .by = visit_age
  )

left_join(fhq_by_trt, fhq_ovr) |>
  gt() |>
  cols_label(
    visit_age ~ md("**Visit**"),
    aP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    wP ~ "\\makecell{\\textbf{aP} \\\\ N = 411}",
    Overall ~ "\\makecell{\\textbf{Overall} \\\\ N = 822}"
  ) |>
  cols_align(align = "left", columns = visit_age) |>
  cols_align(align = "right", columns = aP:Overall) |>
  tab_style(
    style = cell_text(align = "center"),
    location = cells_column_labels()
  ) |>
  tab_options(table.font.size = "10px", table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

# Outcomes

All of the trial outcomes depend upon a positive SPT to either a specific allergen or to any allergen. Therefore, this section first reports on the SPT data, and outcomes 2c and 2d, followed by the primary (food allergy) and other secondary (eczema) outcomes.

## Skin prick tests

```{r}
#| label: spt-no-records
no_spt_record <- dat_spt |>
  filter(row_number() == 1, .by = record_id) |>
  filter(!priyn)
```

There were `r nrow(no_spt_record)` participants with no documented SPTs (@tbl-spt-listing-no-records). In many cases this was due to early withdrawal from the study, with the remaining due to non-consent from the parent. All other participants had a record for at least the scheduled SPT, and in a few cases additional unscheduled SPT (@tbl-spt-number).

```{r}
#| label: tbl-spt-listing-no-records
#| tbl-cap: |
#|   Listing of infants with no SPT records.
#| tbl-pos: "H"
no_spt_record |>
  select(record_id, rand_site, streas, dis_age) |>
  gt() |>
  cols_label(
    dis_age = "Last contact age (months)"
  ) |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(100)
  )
```

```{r}
#| label: tbl-spt-number
#| tbl-cap: |
#|   Number and percentage of infants with scheduled and unscheduled skin prick tests.
#| tbl-pos: "H"
#| output: "asis"
spt_nums <- dat_spt |>
  filter(row_number() == 1, .by = c(record_id, trt, spt_occasion, spt_num)) |>
  summarise(
    spt_num_sch = sum(spt_occasion == "scheduled" & priyn, na.rm = TRUE),
    spt_num_unsch = sum(
      spt_occasion %in% c("unscheduled", "other") & priyn,
      na.rm = TRUE
    ),
    .by = c(record_id, trt)
  )
tbl_summary(
  spt_nums,
  by = trt,
  include = c(spt_num_sch, spt_num_unsch),
  label = c(spt_num_sch ~ "Scheduled SPT", spt_num_unsch ~ "Unscheduled SPTs")
) |>
  add_overall(last = TRUE) |>
  remove_footnote_header(columns = everything()) |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(50)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

For the outcomes, a requirement was positive SPT by 18 months of age. Age of SPT in months was derived from `birthdat` (date of birth) and `pridat` (date of skin prick test) or in some cases `univdat` (date of unscheduled visit) when `pridat` was missing. 

Most SPTs were performed by 13 months of age, although some were delayed, with a few occurring after 18-months of age (@fig-spt-age). There were 5 infants who did have an SPT, but none by 18-months of age (@tbl-spt-listing-after-18months). Participants who were enrolled under earlier stage 1 protocol versions had their SPT scheduled for 18 months of age rather than 12 months of age. Delays meant some of these SPT's did not occur until 19 or 20 months of age. For the outcomes, "approximately 12 months of age" was taken to mean "by 18 months of age" which then currently excludes some of these later SPTs.

```{r}
#| label: fig-spt-age
#| fig-cap: |
#|   Age at skin prick test by protocol stage and occasion type (scheduled or unscheduled).
#|   Y-axis scale varies by panel.
#| fig-height: 4
dat_spt |>
  filter(row_number() == 1, .by = c(record_id, trt, spt_occasion, spt_num)) |>
  filter(!is.na(spt_age)) |>
  mutate(
    rand_stage = paste0("Stage ", rand_stage),
    spt_occasion = str_to_sentence(if_else(
      spt_occasion == "other",
      "unscheduled",
      spt_occasion
    ))
  ) |>
  ggplot(aes(spt_age)) +
  ggh4x::facet_grid2(spt_occasion ~ rand_stage, scales = "free_y", independent = "y") +
  geom_bar() +
  scale_x_continuous("Age (months)", breaks = seq(0, 30, 2)) +
  scale_y_continuous(
    "Number of skin prick tests",
    breaks = integer_breaks(10)
  )
```

```{r}
#| label: tbl-spt-listing-after-18months
#| tbl-cap: Listing of SPTs after 18-months of age.
#| tbl-pos: "H"
dat_spt |>
  filter(
    row_number() == 1,
    .by = c(record_id, trt, spt_occasion, spt_num, spt_age)
  ) |>
  filter(spt_age > 18) |>
  select(record_id, spt_occasion, spt_num, spt_age) |>
  mutate(
    spt_occasion = str_to_sentence(if_else(
      spt_occasion == "other",
      "unscheduled",
      spt_occasion
    ))
  ) |>
  distinct() |>
  gt() |>
  cols_label(
    spt_occasion ~ "SPT type",
    spt_num = "SPT occasion",
    spt_age = "Age (months)"
  ) |>
  tab_options(table.font.size = px(10), table.width = pct(50))
```

@tbl-spt-positive reports on the number and percentage of infants who had a SPT result of \>1mm and ≥3mm relative to negative control for at least one tested allergen (outcomes 2c and 2d). Any infants with no SPT records by 18 months of age are reported as missing.

@tbl-spt-positive-tested provides a breakdown of the specific allergens from the standard panel which satisfied the sensitisation criteria. @tbl-spt-positive-tested-other reports on any other tested allergens beyond the standard panel.

Whole egg had a higher proportion of missingness than other allergens from the standard panel. This was usually because either egg white or egg yolk specifically was assessed and reported in the additional allergens instead of whole egg.

```{r}
#| label: tbl-spt-positive
#| tbl-cap: |
#|   Number and percentage of infants with a positive SPT by 18 months of age.
#|   Infants with no SPT records counted as missing.
#| tbl-pos: "H"
#| results: "asis"
dat_spt |>
  summarise(
    across(
      spt_0mm:spt_3mm,
      ~ if_else(
        all(!priyn) | all(spt_age > 18),
        NA,
        any(.x == 1 & spt_age < 19, na.rm = TRUE)
      )
    ),
    .by = c(record_id, trt)
  ) |>
  tbl_summary(
    by = trt,
    include = c(spt_1mm, spt_3mm),
    label = c(
      # spt_0mm ~ "Any SPT >0mm",
      spt_1mm ~ "Any SPT >1mm vs. neg control",
      spt_3mm ~ "Any SPT ≥3mm vs neg. control"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(70)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| include: false
std_allergens <- c(
  "D.pteronyssinus",
  "cat dander",
  "perennial ryegrass",
  "whole egg",
  "cashew",
  "cow's milk",
  "peanut",
  "seasme"
)
dat_spt |>
  select(
    record_id,
    trt,
    spt_occasion,
    spt_age,
    spt_tested,
    spt_result,
    spt_reaction,
    spt_0mm,
    spt_1mm,
    spt_3mm
  ) |>
  filter(spt_tested %in% std_allergens) |>
  filter(spt_occasion == "scheduled") |>
  summarise(
    N = n(),
    n = sum(!is.na(spt_1mm)),
    y = sum(spt_1mm, na.rm = TRUE),
    p = y / n,
    m = sum(is.na(spt_1mm)),
    .by = c(trt, spt_tested)
  ) |>
  pivot_wider(
    names_from = trt,
    values_from = c(N, n, y, p, m),
    names_vary = "slowest"
  ) |>
  gt(rowname_col = "spt_tested") |>
  tab_spanner(ends_with("aP"), label = "aP") |>
  tab_spanner(ends_with("wP"), label = "wP") |>
  cols_label(
    starts_with("n_", ignore.case = FALSE) ~ "Tests",
    starts_with("y_") ~ ">1mm",
    starts_with("m_") ~ "Missing"
  )
```

```{r}
#| label: tbl-spt-positive-tested
#| tbl-cap: |
#|   Number and percentage of infants with a positive SPT (scheduled only)
#|   by 18 months of age by tested allergen (for those included in the standard panel).
#|   Each cell gives positive/tested (%) [missing result%].
#| tbl-pos: "H"
#| results: "asis"
std_allergens <- c(
  "D.pteronyssinus",
  "cat dander",
  "perennial ryegrass",
  "whole egg",
  "cashew",
  "cow's milk",
  "peanut"
)
dat_spt |>
  filter(spt_tested %in% std_allergens) |>
  filter(spt_occasion == "scheduled") |>
  select(record_id, trt, spt_age, spt_tested, spt_0mm, spt_1mm, spt_3mm) |>
  droplevels() |>
  complete(nesting(select(dat_base, record_id, trt)), spt_tested) |>
  mutate(
    across(spt_0mm:spt_3mm, ~ .x * (spt_age <= 18))
  ) |>
  pivot_longer(
    spt_0mm:spt_3mm,
    names_pattern = "spt_([0-3]+mm)",
    names_to = "size",
    values_to = "diameter"
  ) |>
  pivot_wider(names_from = spt_tested, values_from = diameter) |>
  mutate(
    size = factor(
      size,
      levels = c("0mm", "1mm", "3mm"),
      labels = c(
        "Wheal size >0mm",
        "Wheal size >1mm vs. neg control",
        "Wheal size ≥3mm vs neg. control"
      )
    )
  ) |>
  tbl_strata(
    strata = size,
    .tbl_fun = ~ .x |>
      tbl_summary(
        include = `D.pteronyssinus`:peanut,
        by = trt,
        statistic = all_dichotomous() ~ "{n}/{N_nonmiss} ({p}%) [{p_miss}%]",
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Allergen**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_column_alignment(starts_with("stat"), align = "right"),
    .combine_with = "tbl_stack"
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(100)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-spt-positive-tested-other
#| tbl-cap: |
#|   Number and percentage of infants with a positive SPT (scheduled only)
#|   by 18 months of age by tested allergen for additional allergens which were tested.
#|   Each cell gives positive/tested (%). Totals restricted to only those infants with at least one
#|   additional allergen tested
#| tbl-pos: "H"
#| results: "asis"
std_allergens <- c(
  "D.pteronyssinus",
  "cat dander",
  "perennial ryegrass",
  "whole egg",
  "cashew",
  "cow's milk",
  "peanut"
)
dat_spt |>
  unique() |>
  filter(
    !(spt_tested %in% std_allergens),
    !is.na(spt_tested),
    spt_reaction != "Not Done"
  ) |>
  filter(spt_occasion == "scheduled") |>
  select(record_id, trt, spt_age, spt_tested, spt_0mm, spt_1mm, spt_3mm) |>
  droplevels() |>
  # complete(nesting(select(dat_base, record_id, trt)), spt_tested) |>
  mutate(
    across(spt_0mm:spt_3mm, ~ .x * (spt_age <= 18))
  ) |>
  pivot_longer(
    spt_0mm:spt_3mm,
    names_pattern = "spt_([0-3]+mm)",
    names_to = "size",
    values_to = "diameter"
  ) |>
  mutate(
    diameter = as.logical(diameter),
    spt_tested = fct_infreq(spt_tested)
  ) |>
  arrange(spt_tested) |>
  pivot_wider(names_from = spt_tested, values_from = diameter) |>
  mutate(
    size = factor(
      size,
      levels = c("0mm", "1mm", "3mm"),
      labels = c(
        "Wheal size >0mm",
        "Wheal size >1mm vs. neg control",
        "Wheal size ≥3mm vs neg. control"
      )
    )
  ) |>
  tbl_strata(
    strata = size,
    .tbl_fun = ~ .x |>
      tbl_summary(
        include = `sesame`:`d. farinae`,
        by = trt,
        type = list(everything() ~ "dichotomous"),
        value = list(everything() ~ TRUE),
        statistic = all_dichotomous() ~ "{n}/{N_nonmiss} ({p}%)",
        missing = "no"
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header() |>
      modify_header(label = "**Allergen**") |>
      modify_header(
        all_stat_cols() ~
          paste0(
            "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
            "N = {formatC(n, format='d', big.mark=',')}}}"
          )
      ) |>
      modify_column_alignment(starts_with("stat"), align = "right"),
    .combine_with = "tbl_stack"
  ) |>
  as_gt() |>
  tab_options(
    table.font.size = px(10),
    table.width = pct(70),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```


\clearpage

### Skin prick test model

@tbl-posterior-spt summarises the estimated model parameters for the probability of skin prick test sensitsation under each treatment group (see secondary outcomes in section 1.4.2 numbers 2c and 2d), and their differences (the posterior densities are shown in @fig-posterior-density-spt and @fig-posterior-difference-spt-adjusted).

The results between the primary analysis (unadjusted) and secondary (adjusted) analyses were near identical. Given the data, under the assumed model the most credible values (based on 95% equal-tailed credible interval) for the difference in skin prick test sensitisation at 1mm [2.c] between wP and aP were (-0.04, 0.06) and at 3mm [2.d] were (-0.04, 0.05). The probability of a reduction in skin prick test sensitisation under wP compared to aP was 0.33 and 0.41 respectively.

```{r}
#| label: spt-analysis
dat_outcome <- dat_spt_analysis |>
  pivot_longer(
    spt_0mm:spt_3mm,
    names_pattern = "spt_([0-3]mm)",
    names_to = "size",
    values_to = "outcome"
  ) |>
  filter(!is.na(outcome)) |>
  summarise(
    n = n(),
    y = sum(outcome),
    p = y / n,
    a = 1 + y,
    b = 1 + n - y,
    post = dist_beta(a, b),
    .by = c(trt, size)
  ) |>
  mutate(p_sup = c(NA, beta_ineq_sim(a, b)), .by = size)
diff_outcome <- dat_outcome |>
  mutate(post_drws = generate(post, 1e6)) |>
  select(size, trt, post_drws) |>
  unnest(post_drws) |>
  group_by(size, trt) |>
  mutate(.draw = 1:n()) |>
  ungroup() |>
  pivot_wider(names_from = "trt", values_from = "post_drws") |>
  mutate(delta = wP - aP)

prior <- c(
  prior(logistic(0, 1), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = "b")
)

fit_outcome <- dat_spt_analysis |>
  pivot_longer(
    spt_0mm:spt_3mm,
    names_pattern = "spt_([0-3]mm)",
    names_to = "size",
    values_to = "outcome"
  ) |>
  nest(-size) |>
  mutate(
    fit = map(
      data,
      \(x) {
        brm(
          outcome ~
            trt + rand_site + gender + bfed + fborn + parinc_imp + ces + fha,
          data = x,
          family = bernoulli(),
          prior = prior,
          warmup = 500,
          iter = 5500,
          refresh = 0,
          silent = 2,
          seed = 54123
        )
      }
    ),
    rd = map2(data, fit, \(x, y) {
      wgt <- c(rdo(rdirichlet(1, rep(1, nrow(x))), ndraws = ndraws(y)))
      x |>
        mutate(trt = "aP") |>
        add_epred_rvars(y, value = ".epred_aP") |>
        mutate(trt = "wP") |>
        add_epred_rvars(y, value = ".epred_wP") |>
        mutate(wgt = wgt) |>
        summarise(
          aP = rvar_mean(.epred_aP),
          wP = rvar_mean(.epred_wP),
          .rd_empirical = rvar_mean(.epred_wP - .epred_aP),
          .rd_bootstrap = rvar_sum((.epred_wP - .epred_aP) * wgt)
        )
    })
  )
```

```{r}
#| label: tbl-posterior-spt
#| tbl-cap: |
#|   Posterior summaries for probability and difference in probability of positive skin prick test by size,
#|    unadjusted primary model. Point indicates median.
#| tbl-pos: "H"
diff_agg <- diff_outcome |>
  group_by(size) |>
  nest_rvars() |>
  ungroup() |>
  mutate(
    model = "Unadjusted",
    across(
      aP:delta,
      list(
        med = median,
        lo = ~ quantile2(.x, 0.025),
        hi = ~ quantile2(.x, 0.975),
        pr = ~ Pr(.x < 0)
      ),
      .names = "{.fn}_{.col}"
    )
  ) |>
  select(-aP, -wP, -delta) |>
  bind_rows(
    fit_outcome |>
      unnest(rd) |>
      rename(delta = .rd_empirical) |>
      select(size, aP, wP, delta) |>
      mutate(model = "Adjusted") |>
      mutate(across(
        aP:delta,
        list(
          med = median,
          lo = ~ quantile2(.x, 0.025),
          hi = ~ quantile2(.x, 0.975),
          pr = ~ Pr(.x < 0)
        ),
        .names = "{.fn}_{.col}"
      )) |>
      select(-aP, -wP, -delta)
  ) |>
  select(-pr_aP, -pr_wP)
dat_outcome |>
  select(size, trt, n, y, p) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p)
  ) |>
  left_join(diff_agg, join_by(size)) |>
  mutate(
    size = factor(
      size,
      levels = c("0mm", "1mm", "3mm"),
      labels = c(">0mm", ">1mm vs ctrl. [2.c]", "≥3mm vs ctrl. [2.d]")
    )
  ) |>
  gt(groupname_col = "model", rowname_col = "size") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_align(align = "left", columns = 1) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_aP, lo_aP, hi_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_wP, lo_wP, hi_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "aP"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "wP"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_aP, med_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  tab_options(table.font.size = px(9))
```

```{r}
#| label: fig-posterior-density-spt
#| fig-cap: |
#|    Posterior density for difference in probability of positive skin prick test by size,
#|    unadjusted primary model.
#| fig-height: 3
#| fig-width: 5
diff_outcome |>
  group_by(size) |>
  nest_rvars() |>
  ggplot(aes(xdist = delta)) +
  facet_wrap(~size, ncol = 1) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(
    fill = "Interval",
    x = "Difference in skin prick test sensitisation probability, wP - aP",
    y = "Density"
  )
```

```{r}
#| label: fig-posterior-difference-spt-adjusted
#| fig-cap: |
#|   Posterior difference in probability of skin prick test sensitisation for wP compared to aP, adjusted model.
#| fig-height: 3
#| fig-width: 5
fit_outcome |>
  unnest(rd) |>
  ggplot(aes(xdist = .rd_bootstrap)) +
  facet_wrap(~size, ncol = 1) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(
    fill = "Interval",
    x = "Difference in skin prick test sensitisation probability, wP - aP",
    y = "Density"
  )
```

\clearpage

## Food allergy

The primary source of information for food allergy was the `outcome_report` form. Reported outcomes were cross checked against the food and household questionnaire form (`food_and_household_questionnaire`) and the oral food challenge form (`food_challenge`).

Age at first reaction was derived from `outageval`. If this was missing then `birthdat` (date of birth) and `outdiagdat` (clinician-diagnosis date) was used to derive age in months.

In some cases participants may have withdrawn from the study without sufficient follow-up for food allergy to have been identified. The report makes no allowance for "missing" food allergy status, with absence of clinician diagnosed allergy (with or without adequate follow-up) counted as not meeting the primary outcome. Only 17 participants had an age of last contact 15 months or younger and may have been those most likely to have insufficient follow-up.

The following sections provide a summary of the number of infants who met each criteria. These infants are line listed in the appendix.

### Oral food challenge

There were `r nrow(filter(dat_fa, outalltp == "Oral food challenge"))` outcome records for food allergy diagnoses resulting from oral food challenge amongst `r length(unique(filter(dat_fa, outalltp == "Oral food challenge")$record_id))` infants. One reaction (record ID `8644-34`) reportedly first occurred after 18 months of age and all others were by 18 months of age. All records had a matching positive skin prick test to the reported allergen and so all 22 infants met the primary outcome according to the definition (1.i).

```{r}
#| label: check-positive-spt-ofc
#| include: false
dat_po_ofc <- dat_fa |>
  filter(outalltp == "Oral food challenge", outageval_months2 <= 18) |>
  select(
    record_id,
    trt,
    outdiagdat,
    outageval_months2,
    out_age_months,
    outfrafood
  ) |>
  left_join(
    dat_spt_pos |>
      filter(any_spt_pos_1mm) |>
      select(record_id, spt_age, spt_pos_1mm),
    join_by(record_id),
    relationship = "many-to-many"
  ) |>
  rowwise() |>
  mutate(
    splt_foods = gsub("( |, )", "|", spt_pos_1mm),
    match_food = str_match(tolower(outfrafood), splt_foods)[1]
  ) |>
  ungroup() |>
  summarise(
    outfrafood = paste(unique(tolower(outfrafood)), collapse = ", "),
    any_food_match = any(!is.na(match_food) & outageval_months2 <= 18),
    matched_foods = paste(
      unique(match_food[!is.na(match_food) & outageval_months2 <= 18]),
      collapse = ", "
    ),
    .by = c(record_id, trt)
  ) |>
  rename(
    po_ofc = any_food_match,
    po_ofc_foods = outfrafood,
    po_ofc_match = matched_foods
  )
```

### Anaphylaxis

There were `r nrow(filter(dat_fa, outalltp == "Anaphylaxis"))` outcome records for food allergy diagnoses resulting from anaphylaxis amongst `r length(unique(filter(dat_fa, outalltp == "Anaphylaxis")$record_id))` infants. Two of these infants (`record_id: 68, 5785-283`) were diagnosed after 18 months of age with first reported reaction occuring after 18 months of age. One infant (`record_id = 8644-34`) had no skin prick test for the reported allergens. So of the infants with analphylaxis reported, 3 did not meet the criteria for the primary outcome according to the definition (1.ii).

```{r}
#| label: check-positive-spt-ana
#| include: false
dat_po_ana <- dat_fa |>
  filter(outalltp == "Anaphylaxis", outageval_months2 <= 18) |>
  select(
    record_id,
    trt,
    outdiagdat,
    outageval_months2,
    out_age_months,
    outfrafood
  ) |>
  left_join(
    dat_spt_pos |>
      filter(any_spt_pos_1mm) |>
      select(record_id, spt_age, spt_pos_1mm),
    join_by(record_id),
    relationship = "many-to-many"
  ) |>
  rowwise() |>
  mutate(
    splt_foods = gsub("( |, )", "|", spt_pos_1mm),
    match_food = str_match(tolower(outfrafood), splt_foods)[1]
  ) |>
  ungroup() |>
  summarise(
    outfrafood = paste(unique(tolower(outfrafood)), collapse = ", "),
    any_food_match = any(!is.na(match_food) & outageval_months2 <= 18),
    matched_foods = paste(
      unique(match_food[!is.na(match_food) & outageval_months2 <= 18]),
      collapse = ", "
    ),
    .by = c(record_id, trt)
  ) |>
  rename(
    po_ana = any_food_match,
    po_ana_foods = outfrafood,
    po_ana_match = matched_foods
  )
```

### Highly probable IgE mediated food allergy

There were `r nrow(filter(dat_fa, outalltp == "Highly probably IgE mediated food allergy"))` outcome records for food allergy diagnoses resulting from highly probable IgE mediated allergy amongst `r length(unique(filter(dat_fa, outalltp == "Highly probably IgE mediated food allergy")$record_id))` infants. For one infant (`record_id: 68`), the age at first reaction and age at diagnosis were after 18 months of age. Four infants (`record_id: 75, 5785-239, 5785-243, 5785-331`) had no positive SPT to the reported allergen. Of the infants with IgE-mediated food allergy reported, 5 did not meet the primary outcome according to the definition (2.).

```{r}
#| label: check-positive-spt-ige
#| include: false
# Cross check food challenge data
summ_fc <- dat_fc |>
  filter(ofcresult != "Not challenged") |>
  summarise(
    ofcfood = paste(ofcfood, collapse = ", "),
    ofcresult = paste(ofcresult, collapse = ", "),
    ofcage = paste(ofcage, collapse = ", "),
    .by = record_id
  )

dat_po_ige_1 <- dat_fa |>
  filter(
    outalltp == "Highly probably IgE mediated food allergy",
    outageval_months2 <= 18
  ) |>
  select(
    record_id,
    trt,
    outdiagdat,
    outageval_months2,
    out_age_months,
    outfrafood
  ) |>
  left_join(
    dat_spt_pos |>
      filter(any_spt_pos_1mm) |>
      select(record_id, spt_age, spt_pos_1mm),
    join_by(record_id),
    relationship = "many-to-many"
  ) |>
  rowwise() |>
  mutate(
    splt_foods = gsub("( |, )", "|", spt_pos_1mm),
    splt_foods = replace_na(splt_foods, "^$"),
    outfrafood_sub = gsub(
      "Whole Egg",
      "egg",
      gsub("cows ", "", gsub("Dairy", "milk", outfrafood))
    ),
    match_food = gsub(
      "NA",
      "",
      paste(
        unique(str_match_all(tolower(outfrafood_sub), splt_foods)[[1]]),
        collapse = ", "
      )
    )
  ) |>
  ungroup()
dat_po_ige <- dat_po_ige_1 |>
  summarise(
    outfrafood = paste(unique(tolower(outfrafood)), collapse = ", "),
    any_food_match = any(!is.na(match_food) & outageval_months2 <= 18),
    matched_foods = gsub(
      "^, ",
      "",
      gsub(
        "(, )+",
        ", ",
        paste(
          unique(match_food[!is.na(match_food) & outageval_months2 <= 18]),
          collapse = ", "
        )
      )
    ),
    .by = c(record_id, trt)
  ) |>
  rename(
    po_ige = any_food_match,
    po_ige_foods = outfrafood,
    po_ige_match = matched_foods
  )
```

### Primary outcome

```{r}
#| label: primary-outcome-data
dat_po <- dat_base |>
  left_join(select(dat_po_ofc, -trt), join_by(record_id)) |>
  left_join(select(dat_po_ana, -trt), join_by(record_id)) |>
  left_join(select(dat_po_ige, -trt), join_by(record_id)) |>
  mutate(
    po_ofc = replace_na(po_ofc, 0),
    po_ana = replace_na(po_ana, 0),
    po_ige = replace_na(po_ige, 0),
    po = po_ofc | po_ana | po_ige,
    po_criteria = case_when(
      po_ofc == 1 ~ "Criteria 1.i",
      po_ana == 1 ~ "Criteria 1.ii",
      po_ige == 1 ~ "Criteria 2.",
      .default = "None"
    )
  ) |>
  select(
    record_id,
    rand_site,
    trt,
    gender,
    parinc_imp,
    delt,
    fecurr,
    fborn,
    bfed,
    ces,
    fha,
    starts_with("po")
  )
```

@tbl-primary-outcome-criteria reports on the number of participants who met each of the primary outcome criteria (1.i, 1.ii, and 2) as outlined in the sections above.

@tbl-primary-outcome-overall reports the total number of infants who met the primary outcome, and the primary reason grouped hierarchically according to the ordering in the definition (i.e. OFC → Anaphylaxis → IgE mediated) (irrespective of their timing).

```{r}
#| label: tbl-primary-outcome-criteria
#| tbl-cap: |
#|   Number and percentage of infants who met each primary outcome condition and treatment group.
#| tbl-pos: "H"
#| results: "asis"
dat_po |>
  tbl_summary(
    by = trt,
    include = c(po_ofc, po_ana, po_ige),
    label = list(
      po_ofc ~ "Met condition 1.i.",
      po_ana ~ "Met condition 1.ii.",
      po_ige ~ "Met condition 2."
    ),
    missing = "no"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(70)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-primary-outcome-overall
#| tbl-cap: |
#|   Number and percentage of infants who met the primary outcome by primary reason and treatment group.
#| tbl-pos: "H"
#| results: "asis"
dat_po |>
  tbl_summary(
    by = trt,
    include = c(po, po_criteria),
    label = list(
      po ~ "Met primary outcome",
      po_criteria ~ "Reason"
    ),
    missing = "no"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      filter(label != "None")
  ) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(70)) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

### Primary outcome model

@tbl-posterior-primary summarises the estimated model parameters for the probability of food allergy under each treatment group, and their difference (the posterior densities are shown in @fig-posterior-density and @fig-posterior-difference-adjusted).

The results between the primary analysis (unadjusted) and secondary (adjusted) analyses were near identical. Given the data, under the assumed model the most credible values (based on 95% equal-tailed credible interval) for the difference in food allergy between wP and aP were (-0.02, 0.04). The probability of a reduction in food allergy under wP compared to aP was 0.35.

```{r}
#| label: primary-outcome-analysis
po_sum <- dat_po |>
  summarise(
    n = n(),
    y = sum(po),
    p = y / n,
    a = 1 + y,
    b = 1 + n - y,
    post = dist_beta(a, b),
    .by = trt
  ) |>
  mutate(p_sup = c(NA, beta_ineq_sim(a, b)))
po_diff <- po_sum |>
  mutate(post_drws = generate(post, 1e6)) |>
  select(trt, post_drws) |>
  unnest(post_drws) |>
  group_by(trt) |>
  mutate(.draw = 1:n()) |>
  ungroup() |>
  pivot_wider(names_from = "trt", values_from = "post_drws") |>
  mutate(delta = wP - aP)

prior <- c(
  prior(logistic(0, 1), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = "b")
)
fit <- brm(
  po ~ trt + rand_site + gender + bfed + fborn + parinc_imp + ces + fha,
  data = dat_po,
  family = bernoulli(),
  prior = prior,
  warmup = 500,
  iter = 5500,
  refresh = 0,
  silent = 2,
  seed = 54123
)

wgt <- c(rdo(rdirichlet(1, rep(1, 972)), ndraws = ndraws(fit)))
dat_rd_po <- dat_po |>
  mutate(trt = "aP") |>
  add_epred_rvars(fit, value = ".epred_aP") |>
  mutate(trt = "wP") |>
  add_epred_rvars(fit, value = ".epred_wP") |>
  mutate(wgt = wgt) |>
  summarise(
    aP = rvar_mean(.epred_aP),
    wP = rvar_mean(.epred_wP),
    .rd_empirical = rvar_mean(.epred_wP - .epred_aP),
    .rd_bootstrap = rvar_sum((.epred_wP - .epred_aP) * wgt)
  )
```

```{r}
#| label: tbl-posterior-primary
#| tbl-cap: |
#|   Posterior summaries for probability and difference in probability of food allergy,
#|    unadjusted primary model.
#| tbl-pos: "H"

diff_agg <- po_diff |>
  nest_rvars() |>
  ungroup() |>
  mutate(
    model = "Unadjusted",
    across(
      aP:delta,
      list(
        med = median,
        lo = ~ quantile2(.x, 0.025),
        hi = ~ quantile2(.x, 0.975),
        pr = ~ Pr(.x < 0)
      ),
      .names = "{.fn}_{.col}"
    )
  ) |>
  select(-aP, -wP, -delta) |>
  bind_rows(
    dat_rd_po |>
      rename(delta = .rd_empirical) |>
      select(aP, wP, delta) |>
      mutate(model = "Adjusted") |>
      mutate(across(
        aP:delta,
        list(
          med = median,
          lo = ~ quantile2(.x, 0.025),
          hi = ~ quantile2(.x, 0.975),
          pr = ~ Pr(.x < 0)
        ),
        .names = "{.fn}_{.col}"
      )) |>
      select(-aP, -wP, -delta)
  ) |>
  select(-pr_aP, -pr_wP)

po_sum |>
  select(trt, n, y, p) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p)
  ) |>
  bind_cols(diff_agg) |>
  gt(rowname_col = "model") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_aP, lo_aP, hi_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_wP, lo_wP, hi_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "aP"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "wP"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_aP, med_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  tab_options(table.font.size = px(9))
```

```{r}
#| label: fig-posterior-density
#| fig-cap: Posterior density for difference in probability of primary outcome, unadjusted primary model.
#| fig-height: 2
#| fig-width: 5
po_diff |>
  nest_rvars() |>
  ggplot(aes(xdist = delta)) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(-0.1, 0.1)) +
  labs(
    fill = "Interval",
    x = "Difference in food allergy probability, wP - aP",
    y = "Density"
  )
```


```{r}
#| label: fig-posterior-difference-adjusted
#| fig-cap: Posterior difference in probability of food allergy for wP compared to aP, adjusted model.
#| fig-height: 2
#| fig-width: 5
ggplot(dat_rd_po, aes(xdist = .rd_bootstrap)) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10), limits = c(-0.1, 0.1)) +
  labs(
    fill = "Interval",
    x = "Difference in food allergy probability, wP - aP",
    y = "Density"
  )
```

\clearpage

## Eczema

The primary source of information for eczema was the `outcome_report` form. Reported outcomes were cross checked against the food and household questionnaire form and medical history forms. Infants whose first reaction date for clinician-diagnosed eczema was prior to enrolment were considered to have "pre-existing" eczema and were excluded from the analysis set. There were `r sum(dat_ecz$out_ecz_preexisting)` such infants.

The number and percentage of infants with any clinician-diagnosed eczema is reported in @tbl-eczema-trt. For the protocol defined outcomes \[2a. and 2b.\], the eczema diagnosis needed to be new-onset and accompanied by a positive SPT to at least one allergen by approximately 12 months of age. These are reported in @tbl-eczema-trt-newonset.

```{r}
#| label: tbl-eczema-trt
#| tbl-cap: |
#|   Number and percentage of infants with any clinician-diagnosed eczema.
#| tbl-pos: "H"
#| results: "asis"

# dat_ecz |>
#   filter(out_ecz_preexisting) |>
#   summarise(n = sum(out_ecz_preexisting), .by = trt)

tbl0 <- dat_ecz_spt |>
  mutate(
    ecz = ecz,
    ecz_6m = ecz & outageval_months2 <= 6,
    ecz_12m = ecz & outageval_months2 <= 12,
    ecz_18m = ecz & outageval_months2 <= 18
  ) |>
  tbl_summary(
    by = trt,
    include = c(ecz, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      filter(label != "None")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(55)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-eczema-trt-newonset
#| tbl-cap: |
#|   Number and percentage of infants with new-onset clinician-diagnosed eczema.
#|   To be new-onset the first reaction date needed to be after the infants randomisation date.
#|   Those with missing were due to eczema diagnosis being present but no SPT records being available.
#| tbl-pos: "H"
#| results: "asis"
tbl1 <- dat_ecz_spt |>
  filter(!out_ecz_preexisting) |>
  mutate(
    ecz = ecz & !out_ecz_preexisting,
    ecz_6m = ecz & !out_ecz_preexisting & outageval_months2 <= 6,
    ecz_12m = ecz & !out_ecz_preexisting & outageval_months2 <= 12,
    ecz_18m = ecz & !out_ecz_preexisting & outageval_months2 <= 18
  ) |>
  tbl_summary(
    by = trt,
    include = c(ecz, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**")

# With positive SPT
tbl2 <- dat_ecz_spt |>
  filter(!out_ecz_preexisting) |>
  mutate(
    ecz = ecz & !out_ecz_preexisting & any_pos_spt,
    ecz_6m = ecz & !out_ecz_preexisting & outageval_months2 <= 6 & any_pos_spt,
    ecz_12m = ecz &
      !out_ecz_preexisting &
      outageval_months2 <= 12 &
      any_pos_spt,
    ecz_18m = ecz & !out_ecz_preexisting & outageval_months2 <= 18 & any_pos_spt
  ) |>
  tbl_summary(
    by = trt,
    include = c(ecz, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**")

tbl_stack(
  list(tbl1, tbl2),
  group_header = c(
    "New-onset cases",
    "New-onset cases with positive SPT"
  )
) |>
  remove_footnote_header() |>
  modify_header(label = "**Outcome**") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      filter(label != "None")
  ) |>
  as_gt() |>
  tab_options(table.font.size = px(10), table.width = pct(55)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

### Eczema model

@tbl-posterior-eczema summarises the estimated model parameters for the probability of eczema by age under each treatment group, and their difference (the posterior densities are shown in @fig-posterior-density-eczema and @fig-posterior-difference-eczema-adjusted).

The results between the primary analysis (unadjusted) and secondary (adjusted) analyses were near identical. Given the data, under the assumed primary model the most credible values (based on 95% equal-tailed credible interval) for the difference in eczema at 6 months [2.a] between wP and aP were (-0.04, 0.02) and at 12 months [2.b] (-0.04, 0.03). The probability of a reduction in eczema under wP compared to aP was 0.77 and 0.66 respectively.

```{r}
#| label: eczema-analysis
dat_outcome <- dat_ecz_analysis |>
  pivot_longer(
    ecz_6m:ecz_18m,
    names_pattern = "ecz_([0-9]+m)",
    names_to = "age",
    values_to = "outcome"
  ) |>
  mutate(age = factor(age, levels = c("6m", "12m", "18m"), labels = c("6m [2.a]", "12m [2.b]", "18m"))) |>
  filter(!is.na(outcome)) |>
  summarise(
    n = n(),
    y = sum(outcome),
    p = y / n,
    a = 1 + y,
    b = 1 + n - y,
    post = dist_beta(a, b),
    .by = c(trt, age)
  ) |>
  mutate(p_sup = c(NA, beta_ineq_sim(a, b)), .by = age)
diff_outcome <- dat_outcome |>
  mutate(post_drws = generate(post, 1e6)) |>
  select(age, trt, post_drws) |>
  unnest(post_drws) |>
  group_by(age, trt) |>
  mutate(.draw = 1:n()) |>
  ungroup() |>
  pivot_wider(names_from = "trt", values_from = "post_drws") |>
  mutate(delta = wP - aP)

prior <- c(
  prior(logistic(0, 1), class = "Intercept"),
  prior(student_t(4, 0, 1.75), class = "b")
)

fit_outcome <- dat_ecz_analysis |>
  pivot_longer(
    ecz_6m:ecz_18m,
    names_pattern = "ecz_([0-9]+m)",
    names_to = "age",
    values_to = "outcome"
  ) |>
  mutate(age = factor(age, levels = c("6m", "12m", "18m"), labels = c("6m [2.a]", "12m [2.b]", "18m"))) |>
  nest(-age) |>
  mutate(
    fit = map(
      data,
      \(x) {
        brm(
          outcome ~
            trt + rand_site + gender + bfed + fborn + parinc_imp + ces + fha,
          data = x,
          family = bernoulli(),
          prior = prior,
          warmup = 500,
          iter = 5500,
          refresh = 0,
          silent = 2,
          seed = 54123
        )
      }
    ),
    rd = map2(data, fit, \(x, y) {
      wgt <- c(rdo(rdirichlet(1, rep(1, nrow(x))), ndraws = ndraws(y)))
      x |>
        mutate(trt = "aP") |>
        add_epred_rvars(y, value = ".epred_aP") |>
        mutate(trt = "wP") |>
        add_epred_rvars(y, value = ".epred_wP") |>
        mutate(wgt = wgt) |>
        summarise(
          aP = rvar_mean(.epred_aP),
          wP = rvar_mean(.epred_wP),
          .rd_empirical = rvar_mean(.epred_wP - .epred_aP),
          .rd_bootstrap = rvar_sum((.epred_wP - .epred_aP) * wgt)
        )
    })
  )
```

```{r}
#| label: tbl-posterior-eczema
#| tbl-cap: |
#|   Posterior summaries for probability and difference in probability of eczema by age of onset,
#|    unadjusted primary model.
#| tbl-pos: "H"
diff_agg <- diff_outcome |>
  group_by(age) |>
  nest_rvars() |>
  ungroup() |>
  mutate(
    model = "Unadjusted",
    across(
      aP:delta,
      list(
        med = median,
        lo = ~ quantile2(.x, 0.025),
        hi = ~ quantile2(.x, 0.975),
        pr = ~ Pr(.x < 0)
      ),
      .names = "{.fn}_{.col}"
    )
  ) |>
  select(-aP, -wP, -delta) |>
  bind_rows(
    fit_outcome |>
      unnest(rd) |>
      rename(delta = .rd_empirical) |>
      select(age, aP, wP, delta) |>
      mutate(model = "Adjusted") |>
      mutate(across(
        aP:delta,
        list(
          med = median,
          lo = ~ quantile2(.x, 0.025),
          hi = ~ quantile2(.x, 0.975),
          pr = ~ Pr(.x < 0)
        ),
        .names = "{.fn}_{.col}"
      )) |>
      select(-aP, -wP, -delta)
  ) |>
  select(-pr_aP, -pr_wP)

dat_outcome |>
  select(age, trt, n, y, p) |>
  pivot_wider(
    names_from = trt,
    values_from = c(n, y, p)
  ) |>
  left_join(diff_agg, join_by(age)) |>
  gt(groupname_col = "model", rowname_col = "age") |>
  fmt_percent(starts_with("p_"), decimals = 0) |>
  fmt_number(starts_with(c("med", "lo", "hi", "pr")), decimals = 2) |>
  cols_merge(
    c(y_aP, n_aP, p_aP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_aP, lo_aP, hi_aP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(y_wP, n_wP, p_wP),
    pattern = "{1}/{2} ({3})"
  ) |>
  cols_merge(
    c(med_wP, lo_wP, hi_wP),
    pattern = "{1} ({2}, {3})"
  ) |>
  cols_merge(
    c(med_delta, lo_delta, hi_delta, pr_delta),
    pattern = "{1} ({2}, {3}), {4}"
  ) |>
  tab_spanner(
    columns = ends_with("aP"),
    label = "aP"
  ) |>
  tab_spanner(
    columns = ends_with("wP"),
    label = "wP"
  ) |>
  tab_spanner(
    columns = ends_with("delta"),
    label = "Difference"
  ) |>
  cols_label(
    c(y_aP, y_wP) ~ "y/n (%)"
  ) |>
  cols_label(
    c(med_aP, med_wP) ~ "Median (95% CrI)"
  ) |>
  cols_label(
    c(med_delta) ~ "Median (95% CrI), Pr(<0)"
  ) |>
  tab_options(table.font.size = px(9))
```

```{r}
#| label: fig-posterior-density-eczema
#| fig-cap: |
#|    Posterior density for difference in probability of eczema diagnosis by age at reaction,
#|    unadjusted primary model.
#| fig-height: 3
#| fig-width: 5
diff_outcome |>
  group_by(age) |>
  nest_rvars() |>
  ggplot(aes(xdist = delta)) +
  facet_wrap(~age, ncol = 1) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(
    fill = "Interval",
    x = "Difference in eczema probability, wP - aP",
    y = "Density"
  )
```


```{r}
#| label: fig-posterior-difference-eczema-adjusted
#| fig-cap: |
#|   Posterior difference in probability of eczema diagnosis for wP compared to aP, adjusted model.
#| fig-height: 3
#| fig-width: 5
fit_outcome |>
  unnest(rd) |>
  ggplot(aes(xdist = .rd_bootstrap)) +
  facet_wrap(~age, ncol = 1) +
  stat_histinterval(
    aes(fill = after_stat(level)),
    breaks = 100,
    .width = c(0.5, 0.8, 0.95),
    slab_linewidth = 0.5,
    slab_colour = "black"
  ) +
  geom_vline(xintercept = 0, linetype = 2) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  labs(
    fill = "Interval",
    x = "Difference in eczema probability, wP - aP",
    y = "Density"
  )
```

\clearpage

# Safety

## Serious adverse events

A listing of serious adverse events are reported in @tbl-listing-sae. These are any events reported either in the SAE reporting log (`sae_reporting_log`, stage 2 only) or on the adverse events form (`adverse_events`) with `aetype = "SAE"`. In some cases, a record was on the SAE log but not the adverse event log.

```{r}
#| label: tbl-listing-sae
#| tbl-cap: |
#|   Listing of serious adverse events.
#|   Missing values were due to no adverse event record for corresponding SAE log record.
#| tbl-pos: "H"
dat_sae <- dat_raw |>
  select_form("serious_adverse_events") |>
  select(record_id, saestdat, saemeddra) |>
  distinct() |>
  arrange(str_rank(record_id, numeric = TRUE))

dat_ae_sae <- dat_raw |>
  select_form("adverse_events") |>
  filter(aetype == "SAE") |>
  select(record_id, aecode1, aesev, aetymead, aerel, aestdat, aeenddat) |>
  rename(saestdat = aestdat) |>
  arrange(str_rank(record_id, numeric = TRUE))

dat_saes <- full_join(dat_sae, dat_ae_sae, join_by(record_id, saestdat)) |>
  arrange(str_rank(record_id, numeric = TRUE), saestdat) |>
  mutate(saemeddra = if_else(is.na(aecode1), saemeddra, aecode1)) |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  relocate(trt, .after = record_id)

dat_saes |>
  left_join(select(dat_base, record_id, visdat1)) |>
  mutate(aeday = interval(visdat1, saestdat) %/% days(1)) |>
  select(-visdat1, -saestdat, -aeenddat, -aecode1) |>
  gt() |>
  cols_label(
    trt ~ "Treatment",
    saemeddra ~ "MedDRA code",
    aeday ~ "Days after visit 1"
  ) |>
  sub_missing() |>
  tab_options(table.font.size = px(8), table.width = pct(100)) |>
  cols_width(c(aerel, aetymead) ~ px(100))
```

\clearpage

## Adverse events

```{r}
dat_ae <- dat_raw |>
  select_form("adverse_events") |>
  left_join(select(dat_base, record_id, rand_site), join_by(record_id))
dat_base_ae <- dat_base |>
  select(
    record_id,
    rand_site,
    rand_stage,
    randdattim,
    trt,
    birthdat,
    visdat1
  ) |>
  left_join(
    select_form(dat_raw, "vaccine_administration_v3") |>
      select(record_id, vac7admdat),
    join_by(record_id)
  ) |>
  left_join(dat_ae, join_by(record_id, rand_site)) |>
  mutate(
    ae_age = interval(birthdat, aestdat) %/% days(1),
    aeday_v1 = interval(visdat1, aestdat) %/% days(1),
    aeday_v3 = interval(vac7admdat, aestdat) %/% days(1)
  )
```

There were `r nrow(dat_ae)` adverse events documented during the trial.
Most adverse events were for infants randomised at PCH (`r sum(dat_ae$rand_site == "PCH")`) with only `r sum(dat_ae$rand_site != "PCH")` adverse events for infants randomised at a site other than PCH (@tbl-events-by-stage-and-site and listing in @tbl-listing-ae-non-pch).

There was a higher number of adverse events in the wP group compared to the aP group, but for both groups the majority of events were documented as unrelated to the blinded study vaccine (@tbl-summary-ae-pch). A listing of the related AEs is given in @tbl-listing-ae-related-pch.

A summary of the number and proportion of infants randomised at PCH with any reported adverse event by the MeDRA preferred term is given in @tbl-any-event-ae.  A summary of the total number of events is given in @tbl-events-ae.

```{r}
#| label: tbl-events-by-stage-and-site
#| tbl-cap: |
#|   Number of adverse events by study stage and site.
#| tbl-pos: "H"
dat_base_ae |>
  summarise(
    participants = length(unique(record_id)),
    aes = sum(!is.na(aecode1)),
    .by = c(trt, rand_site, rand_stage)
  ) |>
  pivot_wider(names_from = trt, values_from = c(participants, aes)) |>
  mutate(rand_stage = paste0("Stage ", rand_stage)) |>
  gt(groupname_col = "rand_stage", rowname_col = "rand_site") |>
  cols_merge(
    c(aes_aP, participants_aP),
    pattern = "{1}/{2}"
  ) |>
  cols_merge(
    c(aes_wP, participants_wP),
    pattern = "{1}/{2}"
  ) |>
  cols_label(
    aes_aP ~ "aP (events/participants)",
    aes_wP ~ "wP (events/participants)"
  ) |>
  tab_options(table.font.size = px(9), table.width = pct(50))
```

```{r}
#| label: tbl-listing-ae-non-pch
#| tbl-cap: |
#|   Adverse event listing of infants randomised at a site other than PCH.
#| tbl-pos: "H"
dat_base_ae |>
  filter(rand_site != "PCH", !is.na(aecode1)) |>
  select(record_id, rand_site, trt, aecode1, aesev, aerel, aetymead) |>
  gt() |>
  cols_label(
    rand_site ~ "Site",
    trt ~ "Treatment"
  ) |>
  tab_options(table.font.size = px(9), table.width = pct(100)) |>
  cols_width(aecode1 ~ px(100), aetymead ~ px(100), aerel ~ px(100)) |>
  cols_align(align = "left")
```

```{r}
#| label: tbl-summary-ae-pch
#| tbl-cap: |
#|   Number and proportion of infants with any adverse events and the number of adverse events per infant, at PCH.
#| tbl-pos: "H"
#| results: "asis"
dat_base_ae |>
  filter(rand_site == "PCH") |>
  summarise(
    `
Any
AE
` = any(!is.na(aecode1)),
    `
Any
SAE
` = any(aetype == "SAE", na.rm = TRUE),
    `
Any
related
AE
` = any(aerel == "Related", na.rm = TRUE),
    `
Any
severe
AE
` = any(aesev == "Severe", na.rm = TRUE),
    `
Number
of
AEs
` = sum(!is.na(aecode1)),
    `
Number
of
related
AEs
` = sum(aerel == "Related", na.rm = TRUE),
    .by = c(record_id, trt)
  ) |>
  tbl_summary(
    by = "trt",
    include = `
Any
AE
`:`
Number
of
related
AEs
`,
    digits = ~0
  ) |>
  add_overall(last = TRUE) |>
  modify_column_alignment(columns = starts_with("stat"), align = "right") |>
  remove_footnote_header() |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(100),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-listing-ae-related-pch
#| tbl-cap: |
#|   Listing of adverse events designated as "Related".
#| tbl-pos: "H"
dat_base_ae |>
  filter(aerel == "Related") |>
  select(record_id, trt, aecode1, aeday_v1, aetype, aesev, aerel, aetymead) |>
  gt() |>
  cols_label(
    trt ~ "Treatment",
    aeday_v1 ~ "Days after visit 1"
  ) |>
  tab_options(table.font.size = px(9), table.width = pct(100)) |>
  cols_width(
    record_id ~ px(80),
    trt ~ px(80),
    aecode1 ~ px(100),
    aetymead ~ px(100),
    aerel ~ px(80),
    aetype ~ px(80),
    aesev ~ px(80),
    aerel ~ px(80),
    aeday_v1 ~ px(60)
  ) |>
  cols_align(align = "left") |>
  cols_align(aeday_v1, align = "right") |>
  sub_missing()
```

```{r}
#| label: tbl-any-event-ae
#| tbl-cap: |
#|   Proportion of infants with any event by MeDRA preferred terms, PCH only.
#| tbl-pos: "H"
#| results: "asis"
dat_base_ae |>
  filter(rand_site == "PCH") |>
  tbl_hierarchical(
    variables = aecode1,
    by = trt,
    id = record_id,
    overall_row = TRUE,
    denominator = filter(dat_base, rand_site == "PCH")
  ) |>
  add_overall(last = TRUE) |>
  modify_column_alignment(columns = starts_with("stat"), "right") |>
  remove_footnote_header() |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(100),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-events-ae
#| tbl-cap: |
#|   Number of events by MeDRA preferred terms, PCH only.
#| tbl-pos: "H"
#| results: "asis"
dat_base_ae |>
  filter(rand_site == "PCH") |>
  tbl_hierarchical_count(
    variables = aecode1,
    by = trt,
    overall_row = TRUE,
    denominator = filter(dat_base, rand_site == "PCH")
  ) |>
  add_overall(last = TRUE) |>
  modify_column_alignment(columns = starts_with("stat"), "right") |>
  remove_footnote_header() |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(100),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

## Diary Cards

Combined stage 1 and stage 2 diary card responses are reported in the following sections.

```{r}
#| label: diary-data
st1_pth <- file.path(
  Sys.getenv("RDS_PATH"),
  "data",
  "derived",
  "tidy",
  "stage1"
)

plot_react <- function(
  dat,
  trt = treatment,
  var = intensity,
  slab = "Intensity",
  ref = 0,
  ymax = 1
) {
  pdat <- dat |>
    group_by({{ trt }}, vaxage, day) |>
    count({{ var }}) |>
    mutate(
      p = n / if_else(is.na({{ var }}), sum(n), sum(n[!is.na({{ var }})]))
    ) |>
    ungroup() |>
    complete({{ trt }}, vaxage, day, {{ var }}, fill = list(n = 0, p = 0))
  plab <- dat |>
    summarise(
      n = sum({{ var }} != ref, na.rm = TRUE),
      p = mean({{ var }} != ref, na.rm = TRUE),
      lab = sprintf(
        "%i/%i",
        sum({{ var }} != ref, na.rm = TRUE),
        sum(!is.na({{ var }}))
      ),
      .by = c({{ trt }}, vaxage, day)
    )
  p <- ggplot(
    pdat |> filter({{ var }} != ref),
    aes(interaction(day, to_factor({{ trt }}), sep = "!"), p)
  ) +
    facet_wrap(~ fct_na_value_to_level(to_factor(vaxage)), ncol = 2) +
    geom_col(aes(fill = to_factor({{ var }})), width = 0.8) +
    geom_text(
      data = plab,
      aes(label = lab),
      vjust = -1,
      size = 1.5,
      family = "TeX Gyre Pagella"
    ) +
    scale_fill_viridis_d(slab) +
    scale_x_discrete(
      guide = legendry::guide_axis_nested(key = key_range_auto(sep = "!")),
      name = "Day following vaccination"
    ) +
    scale_y_continuous("Proportion", limit = c(0, ymax), expand = c(0, 0)) +
    labs(y = "Proportion") +
    theme(legend.position = "bottom", panel.grid.major.x = element_blank())
  return(p)
}

plot_sys_reaction <- function(diary_sol, x, ymax = 1) {
  p <- diary_sol |>
    mutate(intensity = ordered(as.numeric(intensity) - 1)) |>
    filter(term == x) |>
    plot_react(trt, intensity, x, "0", ymax = ymax) +
    geom_vline(xintercept = 7.5, linetype = 1, colour = "grey80") +
    labs(
      caption = paste0(
        "Grade 1: easily tolerated\n",
        "Grade 2: interferes with normal everyday activities\n",
        "Grade 3: prevents normal everyday activities or requires significant medical intervention\n"
      )
    ) +
    theme(
      plot.caption = element_text(hjust = 0, size = rel(0.5), lineheight = 1)
    )
  return(p)
}

plot_isr <- function(
  dat,
  trt = treatment,
  var = size_cat,
  ref = 0,
  ylim = 1,
  var_lab = "Size (mm)"
) {
  make_plot_data <- function(dat, trt = trt, var = var, ref = ref) {
    pdat <- dat |>
      group_by({{ trt }}, vaxage, vaccine, day) |>
      count({{ var }}) |>
      mutate(
        p = n / if_else(is.na({{ var }}), sum(n), sum(n[!is.na({{ var }})]))
      ) |>
      ungroup() |>
      complete(
        {{ trt }},
        nesting(vaxage, vaccine),
        day,
        {{ var }},
        fill = list(n = 0, p = 0)
      ) |>
      to_factor()
    plab <- dat |>
      summarise(
        n = sum({{ var }} != ref, na.rm = TRUE),
        p = mean({{ var }} != ref, na.rm = TRUE),
        lab = sprintf(
          "%i/%i",
          sum({{ var }} != ref, na.rm = TRUE),
          sum(!is.na({{ var }}))
        ),
        .by = c({{ trt }}, vaxage, vaccine, day)
      )
    return(list(pdat, plab))
  }
  out <- make_plot_data(dat, {{ trt }}, {{ var }}, ref)
  pdat <- out[[1]]
  plab <- out[[2]]

  base_plot <- function(pdat, plab) {
    ggplot(
      pdat |> filter({{ var }} != ref),
      aes(interaction(day, {{ trt }}, sep = "!"), p)
    ) +
      facet_wrap2(
        vaxage ~ vaccine,
        strip = strip_nested(
          background_x = elem_list_rect(
            fill = c("grey90", "grey80"),
            colour = c("white", "grey20")
          ),
          by_layer_x = TRUE
        ),
        ncol = 4
      ) +
      geom_col(aes(fill = {{ var }})) +
      geom_vline(xintercept = 7.5, colour = "grey80") +
      scale_fill_viridis_d(var_lab) +
      scale_x_discrete(
        guide = legendry::guide_axis_nested(key_range_auto(sep = "!")),
        name = "Day following vaccination"
      ) +
      scale_y_continuous("Proportion", limit = c(0, ylim), expand = c(0, 0)) +
      theme(panel.grid.major.x = element_blank(), legend.position = "top")
  }
  g <- base_plot(pdat, plab)
  return(g)
}

st1_diary <- read_parquet(file.path(st1_pth, "diary-temperature.parquet"))
st2_diary <- select_form(dat_raw, "diary")
```

### Returned

```{r}
tab_dat <- st1_diary |>
  to_factor() |>
  select(subjectid, vaxage, diaretyn) |>
  filter(!is.na(vaxage)) |>
  distinct() |>
  mutate(stage = "Stage 1") |>
  mutate(subjectid = gsub("^01", "01-", subjectid)) |>
  left_join(select(dat_base, subjid, trt), join_by(subjectid == subjid)) |>
  rename(record_id = subjectid) |>
  bind_rows(
    st2_diary |>
      filter(reaction == "shared") |>
      pull(data) |>
      pluck(1) |>
      select(record_id, vaxage, diaretyn) |>
      distinct() |>
      mutate(
        diaretyn = if_else(!diaretyn | is.na(diaretyn), "No", "Yes"),
        stage = "Stage 2"
      ) |>
      left_join(select(dat_base, record_id, trt), join_by(record_id))
  ) |>
  mutate(
    vaxage = factor(
      vaxage,
      levels = c("6-week", "4-month", "6-month", "18-month")
    )
  )
```

```{r}
#| label: tbl-diary-returned-stage1
#| tbl-cap: |
#|   Returned diary cards from stage 1.
#| tbl-pos: "H"
#| results: "asis"
tab_dat |>
  filter(stage == "Stage 1") |>
  tbl_strata2(
    strata = vaxage,
    .tbl_fun = ~ .x |>
      tbl_summary(
        include = diaretyn,
        by = trt,
        label = list(diaretyn = .y)
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header(contains("stat_")),
    .combine_with = "tbl_stack",
    .combine_args = list(group_header = NULL)
  ) |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(75),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: tbl-diary-returned-stage2
#| tbl-cap: |
#|   Returned diary cards from stage 2.
#| tbl-pos: "H"
#| results: "asis"
tab_dat |>
  filter(stage == "Stage 2") |>
  tbl_strata2(
    strata = vaxage,
    .tbl_fun = ~ .x |>
      tbl_summary(
        include = diaretyn,
        by = trt,
        label = list(diaretyn = .y)
      ) |>
      add_overall(last = TRUE) |>
      remove_footnote_header(contains("stat_")),
    .combine_with = "tbl_stack",
    .combine_args = list(group_header = NULL)
  ) |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  as_gt() |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(75),
    latex.use_longtable = TRUE
  ) |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

\clearpage

### Solicited systemic reactions

```{r}
#| label: diary-sol-data
st1_sol <- read_parquet(file.path(
  st1_pth,
  "diary-solicited-reactions.parquet"
)) |>
  to_factor() |>
  filter(!is.na(vaxage)) |>
  mutate(subjid = gsub("^01", "01-", subjectid)) |>
  left_join(select(dat_base, subjid, trt), join_by(subjid)) |>
  rename(record_id = subjid) |>
  select(record_id, trt, vaxage, term, day, intensity) |>
  mutate(term = str_to_sentence(term), stage = 1)

st2_sol <- filter(st2_diary, reaction == "sol") |>
  pull(data) |>
  pluck(1) |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  select(record_id, trt, vaxage, term, day, intensity) |>
  mutate(day = as.integer(day), stage = 2)

diary_sol <- bind_rows(st1_sol, st2_sol) |>
  mutate(
    vaxage = factor(
      vaxage,
      levels = c("6-week", "4-month", "6-month", "18-month")
    )
  )
```

#### Irritability

```{r}
#| label: fig-irritability-daily
#| fig-cap: Irritability day 0 to 6 by vaccination age and assigned treatment.
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Irritability")
```

\clearpage

#### Vomiting

@fig-vomiting-daily presents the distribution of daily grades for vomiting following each vaccination occassion by assigned treatment.

```{r}
#| label: fig-vomiting-daily
#| fig-cap: Vomiting day 0 to 6 by vaccination age and assigned treatment.
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Vomiting")
```

\clearpage

#### Diarrhoea

@fig-diarrhoea-daily presents the distribution of daily grades for diarrhoea following each vaccination occassion by assigned treatment.

```{r}
#| label: fig-diarrhoea-daily
#| fig-cap: Diarrhoea day 0 to 6 by vaccination age and assigned treatment.
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Diarrhoea")
```

\clearpage

#### Decreased Feeding

@fig-decreasedfeeding-daily presents the distribution of daily intensity grades for decreased feeding following each vaccination occassion by assigned treatment.

```{r}
#| label: fig-decreasedfeeding-daily
#| fig-cap: Decreased feeding day 0 to 6 by vaccination age and assigned treatment.
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Decreased feeding")
```

\clearpage

#### Drowsiness

@fig-drowsiness-daily presents the distribution of daily intensity grades for drowsiness following each vaccination occassion by assigned treatment.

```{r}
#| label: fig-drowsiness-daily
#| fig-cap: Drowsiness day 0 to 6 by vaccination age and assigned treatment.
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Drowsiness")
```

\clearpage

#### Restlessness

@fig-restlessness-daily presents the distribution of daily intensity grades for restlessness following each vaccination occassion by assigned treatment.

```{r}
#| label: fig-restlessness-daily
#| fig-cap: Restlessness day 0 to 6 by vaccination age and assigned treatment.
#| fig-height: 4
#| fig-width: 6
plot_sys_reaction(diary_sol, "Restlessness")
```

\clearpage

#### Fever

```{r}
#| label: diary-temp-data
st1_temp <- read_parquet(file.path(
  st1_pth,
  "diary-temperature.parquet"
)) |>
  to_factor() |>
  filter(!is.na(vaxage)) |>
  rename(temp_fac = temp_cat) |>
  mutate(subjid = gsub("^01", "01-", subjectid)) |>
  left_join(select(dat_base, subjid, trt), join_by(subjid)) |>
  rename(record_id = subjid) |>
  select(record_id, trt, vaxage, day, temp_fac)

st2_temp <- filter(st2_diary, reaction == "temp") |>
  pull(data) |>
  pluck(1) |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  select(record_id, trt, vaxage, day, temp_fac) |>
  mutate(day = as.integer(day))

diary_temp <- bind_rows(st1_temp, st2_temp) |>
  mutate(
    vaxage = factor(
      vaxage,
      levels = c("6-week", "4-month", "6-month", "18-month")
    )
  )
```

```{r}
#| label: fig-fever
#| fig-cap: Fever by vaccination age and assigned treatment, stage 1 and 2 combined.
plot_fever <- function() {
  p <- diary_temp |>
    plot_react(trt, temp_fac, "Temperature", "None (<38)") +
    ylim(0, 0.1)
  return(p)
}
plot_fever()
```

\clearpage

### Solicited Injection Site Reactions

```{r}
#| label: data-isr
st1_isr <- read_parquet(file.path(st1_pth, "diary-site-reactions.parquet")) |>
  to_factor() |>
  mutate(
    vaccine = fct_recode(
      vaccine,
      "aP/wP" = "aP or wP",
      "MMR" = "MMRV",
      "DTPa-hepB-IPV-Hib" = "DTaP-Hib-HepB-IPV",
      "PCV-13" = "13vPCV"
    )
  ) |>
  mutate(subjid = gsub("^01", "01-", subjectid)) |>
  left_join(select(dat_base, subjid, trt), join_by(subjid)) |>
  rename(record_id = subjid, size_fac = size_cat) |>
  select(record_id, trt, term, vaxage, vaccine, day, size, size_fac) |>
  mutate(stage = 1)

st2_isr <- st2_diary |>
  filter(reaction == "isr") |>
  pull(data) |>
  pluck(1) |>
  left_join(
    st2_diary |>
      filter(reaction == "shared") |>
      pull(data) |>
      pluck(1),
    join_by(record_id, visit, vaxage, vaxloc_id)
  ) |>
  mutate(
    day = as.integer(day),
    vaccine = fct_recode(vaccine, "PCV-13" = "PCv13")
  ) |>
  select(record_id, vaxage, term, vaccine, day, size, size_fac) |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  mutate(stage = 2) |>
  filter(!is.na(vaccine))


diary_isr <- bind_rows(st1_isr, st2_isr) |>
  mutate(
    vaxage = factor(
      vaxage,
      levels = c("6-week", "4-month", "6-month", "18-month")
    )
  )
```

```{r}
#| label: fig-redness-daily
#| fig-cap: |
#|    Injection site erythema by vaccination age, vaccine, and assigned treatment.
#|    Stage 1: 6-week, 4-month, 6-month, 18-month.
#|    Stage 2: 6-week, 18-month.
#| fig-width: 7
#| fig-height: 4
plot_isr(
  diary_isr |> filter(term == "Erythema"),
  trt = trt,
  var = size_fac,
  ref = "None",
  var_lab = "Redness size (mm)"
)
```

```{r}
#| label: fig-swelling-daily
#| fig-cap: |
#|   Injection site swelling by vaccination age, vaccine, and assigned treatment.
#|   Stage 1: 6-week, 4-month, 6-month, 18-month.
#|   Stage 2: 6-week, 18-month.
#| fig-width: 7
#| fig-height: 4
plot_isr(
  diary_isr |> filter(term == "Swelling"),
  trt = trt,
  var = size_fac,
  ref = "None",
  var_lab = "Swelling size (mm)"
)
```

```{r}
#| label: fig-induration-daily
#| fig-cap:  |
#|   Injection site induration by vaccination age, vaccine, and assigned treatment.
#|   Stage 1: 6-week, 4-month, 6-month, 18-month.
#|   Stage 2: 6-week, 18-month.
#| fig-width: 7
#| fig-height: 4
plot_isr(
  diary_isr |> filter(term == "Induration"),
  trt = trt,
  var = size_fac,
  ref = "None",
  var_lab = "Induration size (mm)"
)
```

```{r}
#| label: diary-pain-data
st1_pain <- read_parquet(file.path(st1_pth, "diary-pain.parquet")) |>
  to_factor() |>
  mutate(
    vaccine = fct_recode(
      vaccine,
      "aP/wP" = "aP or wP",
      "MMR" = "MMRV",
      "DTPa-hepB-IPV-Hib" = "DTaP-Hib-HepB-IPV",
      "PCV-13" = "13vPCV"
    )
  ) |>
  mutate(subjid = gsub("^01", "01-", subjectid)) |>
  left_join(select(dat_base, subjid, trt), join_by(subjid)) |>
  rename(record_id = subjid) |>
  select(record_id, trt, vaxage, vaccine, day, pain) |>
  mutate(stage = 1)
st2_pain <- st2_diary |>
  filter(reaction == "pain") |>
  pull(data) |>
  pluck(1) |>
  left_join(
    st2_diary |>
      filter(reaction == "shared") |>
      pull(data) |>
      pluck(1),
    join_by(record_id, visit, vaxage, vaxloc_id)
  ) |>
  mutate(
    day = as.integer(day),
    vaccine = fct_recode(vaccine, "PCV-13" = "PCv13")
  ) |>
  select(record_id, vaxage, vaccine, day, pain) |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  mutate(stage = 2) |>
  filter(!is.na(vaccine))

# Fill in missing records
# ids <- st2_diary |>
#   filter(reaction == "pain") |>
#   pull(data) |>
#   pluck(1) |> pull(record_id) |> unique()
# st2_pain <- st2_pain |>
#   complete(record_id = ids, nesting(vaxage, vaccine), day)

diary_pain <- bind_rows(st1_pain, st2_pain) |>
  mutate(
    vaxage = factor(
      vaxage,
      levels = c("6-week", "4-month", "6-month", "18-month")
    )
  )
```

```{r}
#| label: fig-pain-daily
#| fig-cap: Injection site pain by vaccination age, vaccine, and assigned treatment.
#| fig-width: 7
#| fig-height: 4
plot_isr(diary_pain, trt, pain, ref = "None", var_lab = "Pain intensity")
```
  
\clearpage

# Appendix {.appendix}

## Primary outcome listings
  
```{r}
#| label: tbl-listing-ofc-matched-foods
#| tbl-cap: |
#|    Listing of infants with oral food challenge outcome by 18 months of age and corresponding SPT positive allergens.
#| tbl-pos: "H"
dat_po_ofc |>
  select(record_id, trt, po_ofc_foods, po_ofc_match) |>
  gt() |>
  cols_label(
    trt ~ "Treatment",
    po_ofc_foods ~ "OFC positive foods",
    po_ofc_match ~ "Foods with positive SPT"
  ) |>
  tab_options(table.font.size = px(9), table.width = pct(75))
```

```{r}
#| label: tbl-listing-ana-matched-foods
#| tbl-cap: Listing of infants with anaphylaxis outcome by 18 months of age and corresponding SPT positive allergens.
#| tbl-pos: "H"
dat_po_ana |>
  select(record_id, trt, po_ana_foods, po_ana_match) |>
  gt() |>
  cols_label(
    trt ~ "Treatment",
    po_ana_foods ~ "Anaphylaxis foods",
    po_ana_match ~ "Foods with positive SPT"
  ) |>
  tab_options(table.font.size = px(9), table.width = pct(75))
```

```{r}
#| label: tbl-listing-ige-matched-foods
#| tbl-cap: |
#|    Listing of infants with IgE-mediated food allergy outcome by 18 months of age and corresponding
#|    SPT positive allergens.
#| tbl-pos: "H"
dat_po_ige |>
  select(record_id, trt, po_ige_foods, po_ige_match) |>
  gt() |>
  cols_label(
    trt ~ "Treatment",
    po_ige_foods ~ "IgE mediated foods",
    po_ige_match ~ "Foods with positive SPT"
  ) |>
  tab_options(table.font.size = px(9), table.width = pct(75))
```

\clearpage

## Secondary outcome listings

```{r}
#| label: tbl-listing-eczema
#| tbl-cap: Listing of infants with eczema diagnosis and positive SPT to any allergen.
#| tbl-pos: "H"
dat_ecz_spt |>
  filter(!out_ecz_preexisting, ecz, any_pos_spt) |>
  select(record_id, trt, outageval_months2) |>
  gt() |>
  fmt_number(decimals = 0) |>
  cols_label(
    trt ~ "Treatment",
    outageval_months2 ~ "Age at first reaction (months)"
  ) |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(60),
    latex.use_longtable = TRUE
  )
```

\clearpage

## IgE responses

```{r}
dat_ige <- select_form(dat_raw, "ige")
dat_ige <- dat_base |>
  right_join(dat_ige, join_by(record_id, subjid)) |>
  select(
    record_id,
    subjid,
    trt,
    gender,
    bfed,
    fborn,
    parinc_imp,
    ces,
    fha,
    visit_age,
    allergen,
    ige,
    comment
  )

multinorm <- cmdstan_model(here("stan", "multinorrm-cens.stan"))

make_stan_data <- function(
  dd,
  form = ~trt,
  var = log_ige,
  L = log10(c(0.005, 0.005, 0.0005))
) {
  tmp <- dd |>
    filter(row_number() == 1, .by = subjid)
  X <- model.matrix(form, data = tmp)
  X <- cbind(
    X[, 1:2, drop = FALSE],
    scale(X[, -(1:2), drop = FALSE], scale = F)
  )
  out <- list(
    N = nrow(tmp),
    J = 3,
    X = X,
    P = ncol(X),
    Y = dd |>
      select(subjid, visit_age, {{ var }}) |>
      spread(visit_age, {{ var }}) |>
      select(-subjid) |>
      mutate(across(everything(), ~ replace_na(.x, -99))) |>
      as.matrix(),
    R = dd |>
      mutate(
        R = case_when(
          is.na({{ var }}) ~ 1,
          {{ var }} == -Inf ~ 2,
          !is.na({{ var }}) ~ 0
        )
      ) |>
      select(subjid, visit_age, R) |>
      spread(visit_age, R) |>
      select(-subjid) |>
      as.matrix(),
    L = L
  )
  return(out)
}

gmr_table_dat <- function(gmr) {
  tdat <- gmr |>
    mutate(
      across(V1, list(
        mean = ~ mean(.x),
        sd = ~ sd(.x),
        med = ~ median(.x),
        lo = ~ quantile(.x, 0.025)[1, ],
        hi = ~ quantile(.x, 0.975)[1, ],
        pr = ~ Pr(.x < 1)
      ))
    ) |>
    select(-V1)
  return(tdat)
}

gmr_table <- function(tdat) {
  gt(tdat, rowname_col = "visit_age", groupname_col = "group") |>
    fmt_number(everything(), decimals = 2) |>
    cols_merge(
      c(V1_mean, V1_sd),
      pattern = "<<{1} $\\pm$ {2}>>"
    ) |>
    cols_merge(
      c(V1_lo, V1_hi),
      pattern = "<<({1}, {2})>>"
    ) |>
    cols_label(
      V1_mean = md("Mean &#177; std"),
      V1_med = "Median",
      V1_lo = md("95% CrI"),
      V1_pr = "Pr(< 1)",
    ) |>
    tab_stub_indent(
      rows = everything(),
      indent = 1
    ) |>
    tab_options(table.font.size = px(9))
}

make_ate <- function(dat, fit) {
  nd <- filter(dat, visit_age == "6-month")
  ndat <- expand_grid(
    visit_age = c("6-month", "7-month", "19-month"),
    trt = c("aP", "wP")
  )
  tmp <- brmsmargins(
    fit,
    newdata = nd,
    at = ndat,
    effects = "integrateoutRE",
    CI = 0.95,
    k = 100L,
    seed = 123,
    contrasts = cbind(
      "6-month_wP - aP" = c(-1, 1, 0, 0, 0, 0),
      "7-month_wP - aP" = c(0, 0, -1, 1, 0, 0),
      "19-month_wP - aP" = c(0, 0, 0, 0, -1, 1)
    )
  )
  r <- rvar(cbind(tmp$Posterior, tmp$Contrasts))
  names(r)[seq_len(nrow(ndat))] <- apply(ndat, 1, paste, collapse = "_")
  out <- as_tibble(t(r)) |>
    pivot_longer(
      everything(),
      names_to = c("visit_age", "trt"),
      names_sep = "_",
      values_to = "posterior"
    ) |>
    mutate(measure = "RD")
  return(out)
}
```

The 6 and 7 month IgE responses were previously reported. 
What follows additionally reports on 19-month IgE responses.
Descriptive statistics are reported in @tbl-sample-ige-gmc.

```{r}
#| label: tbl-sample-ige-gmc
#| code-summary: Sample summaries
#| tbl-cap: |
#|    Sample IgE concentration summaries (kU/L) by allergen type, visit age,
#|    and assigned vaccination schedule.
#| tbl-pos: "H"
#| results: "asis"
tab <- dat_ige |>
  mutate(
    ige = if_else(ige == 0, 0.0005, ige)
  ) |>
  summarise(
    obs = sum(!is.na(ige)),
    gmc = 10^mean(log10(ige), na.rm = TRUE),
    min = min(ige, na.rm = TRUE),
    max = max(ige, na.rm = TRUE),
    det = mean(ige >= 0.01, na.rm = TRUE),
    con = mean(ige >= 0.1, na.rm = TRUE),
    .by = c(allergen, visit_age, trt)
  ) |>
  pivot_wider(
    names_from = trt,
    values_from = c(obs, gmc, min, max, det, con),
    names_vary = "slowest"
  ) |>
  to_factor() |>
  arrange(allergen, visit_age)

gt(tab, groupname_col = "allergen", rowname_col = "visit_age") |>
  fmt_integer(starts_with("obs")) |>
  fmt_number(starts_with("gmc"), decimals = 2) |>
  fmt_number(starts_with(c("min", "max")), decimals = 2) |>
  fmt_percent(starts_with("con"), decimals = 0) |>
  fmt_percent(starts_with("det"), decimals = 0) |>
  cols_merge(
    columns = c(min_aP, max_aP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(min_wP, max_wP),
    pattern = "{1}–{2}"
  ) |>
  cols_label(
    obs_aP = "n",
    obs_wP = "n",
    gmc_aP = "GMC",
    gmc_wP = "GMC",
    min_aP = "Min–Max",
    min_wP = "Min–Max",
    det_aP = md("&ge;0.01"),
    det_wP = md("&ge;0.01"),
    con_aP = md("&ge;0.1"),
    con_wP = md("&ge;0.1")
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    label = "aP",
    columns = ends_with("_aP")
  ) |>
  tab_spanner(
    label = "wP",
    columns = ends_with("_wP")
  ) |>
  tab_footnote(
    footnote = "Values of 0.00 kU/L and 0.000 kU/L set to 0.005 and 0.0005 respectively for calculation of sample GMC.",
    locations = cells_column_labels(columns = starts_with("gmc"))
  ) |>
  tab_options(table.font.size = px(9)) |>
  as_latex() |>
  # Fix slash bug
  gsub(pattern = "\\\\textendash", replacement = "\\\\textendash ") |>
  cat()
```

```{r}
#| include: false
#| eval: false
dat_po |>
  right_join(dat_ige, join_by(record_id)) |>
  filter(grepl("egg|total", tolower(allergen))) |>
  mutate(
    egg = factor(
      grepl("egg", po_ofc_foods) |
        grepl("egg", po_ana_foods) |
        grepl("egg", po_ige_foods)
    )
  ) |>
  mutate(ige_imp = if_else(ige == 0, 0.0005, ige)) |>
  ggplot(aes(egg, ige_imp)) +
  facet_grid(allergen ~ visit_age) +
  geom_boxplot(aes(colour = egg)) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  labs(colour = "Egg allergy")
```

### Geometric mean concentration

```{r}
#| label: model-ige-total
#| cache: true
fit_ige_model <- function(ige_ref) {
  age_labs <- c("6-month", "7-month", "19-month")

  mdat <- dat_ige |>
    filter(allergen == ige_ref) |>
    filter(!all(is.na(ige)), .by = subjid) |>
    mutate(log_ige = log10(ige))

  sdat1 <- make_stan_data(mdat, var = log_ige)
  sdat2 <- make_stan_data(
    mdat,
    var = log_ige,
    ~ trt + gender + bfed + fborn + parinc_imp + ces + fha
  )
  snk <- capture.output(
    sfit1 <- multinorm$sample(
      data = sdat1,
      seed = 71356,
      refresh = 0,
      chains = 8,
      iter_sampling = 1250
    )
  )
  snk <- capture.output(
    sfit2 <- multinorm$sample(
      data = sdat2,
      seed = 71356,
      refresh = 0,
      chains = 8,
      iter_warmup = 500,
      iter_sampling = 1250
    )
  )
  sdrw1 <- as_draws_rvars(sfit1)
  sdrw2 <- as_draws_rvars(sfit2)

  rownames(sdrw1$b) <- age_labs
  rownames(sdrw2$b) <- age_labs

  gmr1 <- 10^sdrw1$b[, 2]
  gmr2 <- 10^sdrw2$b[, 2]
  gmfir1 <- 10^(sdrw1$b[-1, 2] - sdrw1$b[-3, 2])
  gmfir2 <- 10^(sdrw2$b[-1, 2] - sdrw2$b[-3, 2])
  rownames(gmfir1) <- rownames(gmfir2) <- c("7/6-month (GMFR)", "19/7-month (GMFR)")

  # Fill missing observations with conditional predictions
  # yfill <- t(as_rvar(sdat1$Y))
  # yfill[which(t(sdat1$R == 1))] <- sdrw1$y_missing
  # yfill[which(t(sdat1$R == 2))] <- sdrw1$y_censored
  # yfill <- t(yfill)
  return(as_tibble(rbind(gmr2, gmfir2), rownames = "visit_age"))
}

res_total_ige <- fit_ige_model("Total IgE")
res_tetanus_ige <- fit_ige_model("Tetanus-toxoid")
res_eggwhite_ige <- fit_ige_model("Egg white")
res_wholeegg_ige <- fit_ige_model("Whole egg")

gmr <- bind_rows(
  mutate(res_total_ige, group = "Total IgE"),
  mutate(res_tetanus_ige, group = "Tetanus toxoid"),
  mutate(res_eggwhite_ige, group = "Egg white"),
  mutate(res_wholeegg_ige, group = "Whole egg")
)
gmr_dat <- gmr_table_dat(gmr)
```

```{r}
#| label: tbl-gmr-gmfir-ige-total
#| tbl-cap: Geometric mean ratio (GMR) and geometric mean fold rise ratio (GMFRR) for total IgE.
#| tbl-pos: "H"
#| results: "asis"
gmr_table(gmr_dat) |>
  as_latex()
```

### Sensitisation

```{r}
#| label: model-ige-sens
#| cache: true
fit_ige_sens_model <- function(ige_ref, ref_val) {
  mdat <- dat_ige |>
    filter(allergen == ige_ref) |>
    filter(!is.na(ige)) |>
    mutate(sens = as.numeric(ige >= ref_val))

  glmm_prior <- c(
    prior(student_t(4, 0, 1.75), class = "b"),
    prior(student_t(4, 0, 1.75), class = "Intercept"),
    prior(exponential(1), class = "sd")
  )

  mfit1 <- brm(
    sens ~ visit_age * trt + (1 | subjid),
    data = mdat,
    family = bernoulli(),
    backend = "cmdstanr",
    prior = glmm_prior,
    seed = 71235,
    refresh = 0,
    chains = 8,
    iter = 1750,
    warmup = 500
  )

  mfit2 <- update(
    mfit1,
    formula = ~ . +
      visit_age *
        (gender +
          bfed +
          fborn +
          ces +
          fha +
          parinc_imp),
    newdata = mdat
  )

  # Marginal risk difference estimate
  dat_ige |>
    mutate(trt = "aP") |>
    add_epred_rvars(mfit2, value = "aP", allow_new_levels = TRUE) |>
    mutate(trt = "wP") |>
    add_epred_rvars(mfit2, value = "wP", allow_new_levels = TRUE) |>
    mutate(dif = wP - aP) |>
    summarise(across(c(aP, wP, dif), rvar_mean), .by = visit_age)

  # mdrws1 <- rvar(as_draws_matrix(mfit1))
  # beta1 <- mdrws1[grepl("(^b_Inter|^b_age|^b_tr)", names(mdrws1))]
  # omega1 <- mdrws1["sd_subjectid__Intercept"]
  # beta1_marg <- do.call(c, as_draws_rvars(t(marginalcoef(mfit1, posterior = TRUE)$Posterior)))
  # names(beta1_marg) <- names(beta1)
  # cor1 <- exp(c(beta1[3], beta1[3] + beta1[4]))
  # mor1 <- exp(c(beta1_marg[3], beta1_marg[3] + beta1_marg[4]))

  mdrws2 <- rvar(as_draws_matrix(mfit2))
  beta2 <- mdrws2[grepl("(^b_Inter|^b_visit_age|^b_tr)", names(mdrws2))]
  omega2 <- mdrws2["sd_subjid__Intercept"]
  beta2_marg <- do.call(
    c,
    as_draws_rvars(t(marginalcoef(mfit2, posterior = TRUE)$Posterior))
  )
  names(beta2_marg) <- variables(mfit2)[1:30]
  beta2_marg <- beta2_marg[grepl(
    "(^b_Inter|^b_visit_age|^b_tr)",
    names(beta2_marg)
  )]
  cor2 <- exp(c(beta2[4], beta2[4] + beta2[5], beta2[4] + beta2[6]))
  mor2 <- exp(c(
    beta2_marg[4],
    beta2_marg[4] + beta2_marg[5],
    beta2_marg[4] + beta2_marg[6]
  ))
  p1 <- make_ate(dat_ige, mfit1) |>
    pivot_wider(names_from = trt, values_from = posterior)
  p2 <- make_ate(dat_ige, mfit2) |>
    pivot_wider(names_from = trt, values_from = posterior)
  tdat2 <- p2 |>
    select(-measure) |>
    bind_cols(cor = cor2, mor = mor2) |>
    mutate(group = ige_ref)
  return(tdat2)
}

snk <- capture.output({
  res_sens_tetanus <- fit_ige_sens_model("Tetanus-toxoid", 0.1)
  res_sens_eggwhite <- fit_ige_sens_model("Egg white", 0.1)
  res_sens_wholeegg <- fit_ige_sens_model("Whole egg", 0.1)
})

res_sens <- bind_rows(res_sens_tetanus, res_sens_eggwhite, res_sens_wholeegg) |>
  mutate(age = fct_rev(fct_inorder(visit_age)), group = fct_inorder(group)) |>
  relocate(group, .before = visit_age)
```

```{r}
#| label: tbl-sensitisation-summary
#| tbl-cap: Summary of sensitisation model by response type.
#| tbl-pos: "H"
out <- res_sens |>
  transmute(
    group,
    age,
    across(
      3:7,
      list(
        med = ~ median(.x),
        lo = ~ quantile(.x, 0.025)[1, ],
        hi = ~ quantile(.x, 0.975)[1, ]
      )
    ),
    across(5, list(pr = ~ Pr(.x < 0)))
  ) |>
  relocate(`wP - aP_pr`, .after = `wP - aP_hi`) |>
  gt(rowname_col = "age", groupname_col = "group") |>
  fmt_number(-(1:2), decimals = 2) |>
  cols_merge(
    starts_with("wP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("aP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    c("wP - aP_med", "wP - aP_lo", "wP - aP_hi"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("cor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("mor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_label(
    "aP_med" = "aP",
    "wP_med" = "wP",
    "wP - aP_med" = "wP \u2212 aP",
    "wP - aP_pr" = "Pr(wP \u2212 aP < 0)",
    "cor_med" = "Conditional",
    "mor_med" = "Marginal"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    columns = c("aP_med", "wP_med", "wP - aP_med"),
    label = md("Standardised Pr(IgE $\\geq$ 0.1), Median (95% CrI)"),
  ) |>
  tab_spanner(
    columns = c("cor_med", "mor_med"),
    label = "Odds ratios"
  ) |>
  tab_options(table.font.size = px(9))
out
```

\clearpage

### Quantitation

```{r}
#| label: model-ige-quant
#| cache: true
snk <- capture.output({
  res_sens_tetanus <- fit_ige_sens_model("Tetanus-toxoid", 0.01)
  res_sens_eggwhite <- fit_ige_sens_model("Egg white", 0.01)
  res_sens_wholeegg <- fit_ige_sens_model("Whole egg", 0.01)
})

res_sens <- bind_rows(res_sens_tetanus, res_sens_eggwhite, res_sens_wholeegg) |>
  mutate(age = fct_rev(fct_inorder(visit_age)), group = fct_inorder(group)) |>
  relocate(group, .before = visit_age)
```

```{r}
#| label: tbl-sensitisation-tetanus-summary
#| tbl-cap: Summary of sensitisation model for tetanus toxoid IgE.
#| tbl-pos: "H"
out <- res_sens |>
  transmute(
    group,
    age,
    across(
      3:7,
      list(
        med = ~ median(.x),
        lo = ~ quantile(.x, 0.025)[1, ],
        hi = ~ quantile(.x, 0.975)[1, ]
      )
    ),
    across(5, list(pr = ~ Pr(.x < 0)))
  ) |>
  relocate(`wP - aP_pr`, .after = `wP - aP_hi`) |>
  gt(rowname_col = "age", groupname_col = "group") |>
  fmt_number(-(1:2), decimals = 2) |>
  cols_merge(
    starts_with("wP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("aP_"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    c("wP - aP_med", "wP - aP_lo", "wP - aP_hi"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("cor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_merge(
    starts_with("mor"),
    pattern = "<<{1} ({2}, {3})>>"
  ) |>
  cols_label(
    "aP_med" = "aP",
    "wP_med" = "wP",
    "wP - aP_med" = "wP \u2212 aP",
    "wP - aP_pr" = "Pr(wP \u2212 aP < 0)",
    "cor_med" = "Conditional",
    "mor_med" = "Marginal"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    columns = c("aP_med", "wP_med", "wP - aP_med"),
    label = md("Standardised Pr(IgE $\\geq$ 0.1), Median (95% CrI)"),
  ) |>
  tab_spanner(
    columns = c("cor_med", "mor_med"),
    label = "Odds ratios"
  ) |>
  tab_options(table.font.size = px(9))
out
```
