---
title: OPTIMUM Stage 2 IgG Concentrations
subtitle: Analysis Report
---

```{r}
#| label: packages-and-setup
suppressPackageStartupMessages({
  # Data
  library(qs)
  library(dplyr)
  library(tidyr)
  library(forcats)
  library(lubridate)
  library(labelled)

  # Figures
  library(ggplot2)
  library(scales)
  library(legendry)

  # Tables
  library(gt)
  library(gtsummary)

  # Models
  library(mmrm)
  library(emmeans)
  library(brms)
  library(cmdstanr)
  library(posterior)
  library(tidybayes)
  library(ggdist)
  library(bayesplot)
  library(loo)
})

readRenviron(".env")

source("R/util.R")
source("reports/functions.R")

options(
  brms.backend = "cmdstanr",
  mc.cores = 8
)

sysfonts::font_add(
  family = "TeX Gyre Pagella",
  regular = "texgyrepagella-regular.otf",
  bold = "texgyrepagella-bold.otf",
  italic = "texgyrepagella-italic.otf",
  bolditalic = "texgyrepagella-bolditalic.otf"
)

showtext::showtext_auto()
showtext::showtext_opts(dpi = 244)

bs <- 9

theme_set(
  theme_bw(base_size = bs, base_family = "TeX Gyre Pagella") +
    theme(
      legend.position = "bottom",
      panel.grid.minor = element_blank(),
      axis.text = element_text(colour = "black"),
      legend.key.size = unit(0.75, "lines")
    )
)

it_warmup <- 1250
it_sample <- 500
```

```{r}
#| label: model
multinorm <- cmdstan_model(file.path("stan", "multinorm.stan"))
multinorm_het <- cmdstan_model(file.path("stan", "multinorm-het.stan"))
multit <- cmdstan_model(file.path("stan", "multit.stan"))
```

```{r}
#| label: data
dat_rand_list <- read_randomisation_lists()
dat_all <- qread(file.path(
  Sys.getenv("RDS_PATH"),
  config::get("combined_data_file")
))

dat_base <- left_join(
  select_form(dat_all, "randomisation"),
  dat_rand_list,
  join_by(rand)
) |>
  left_join(
    select_form(dat_all, "demographics"),
    join_by(record_id)
  ) |>
  left_join(
    select_form(dat_all, "birth_history"),
    join_by(record_id)
  ) |>
  left_join(
    select_form(dat_all, "food_and_household_questionnaire") |>
      filter(visit_age == "6-week") |>
      select(record_id, fecurr),
    join_by(record_id)
  ) |>
  left_join(
    select_form(dat_all, "family_history_of_atopy") |>
      select(record_id, fha),
    join_by(record_id)
  ) |>
  mutate(
    dage = interval(birthdat, visdat1) %/% days(1),
    mpar = factor(if_else(mpar > 4, 4, mpar), labels = c("1", "2", "3", "4+")),
    sex = factor(if_else(gender == "Male", 1, 0), labels = c("Female", "Male")),
    caes = factor(
      case_when(
        is.na(delt) ~ NA_real_,
        grepl("caesarean", delt) ~ 1,
        TRUE ~ 0
      ),
      labels = c("Vaginal", "Caesarean")
    ),
    bfed = factor(
      case_when(
        is.na(fecurr) ~ NA_real_,
        fecurr == "Exclusively Breastfed" ~ 0,
        fecurr == "Exclusively formula-fed" ~ 2,
        TRUE ~ 1
      ),
      labels = c("Complete", "Partial", "None")
    ),
    inc = factor(
      case_when(
        is.na(parinc) ~ NA_real_,
        parinc == ">$180,000" ~ 1,
        parinc == "$87,001 to $180,000" ~ 2,
        TRUE ~ 3
      ),
      labels = c(">$180,000", "$87,001-$180,000", "≤$87,000")
    ),
    sibnum = if_else(
      sibnum == "Not Applicable",
      "0",
      if_else(sibnum == "Unknown", NA_character_, sibnum)
    ),
    fborn = factor(
      if_else(sibnum == "0" | is.na(sibnum), 1, 0),
      labels = c("No", "Yes")
    ),
    fha = factor(fha, labels = c("No", "Yes"))
  )

dat_bc <- select_form(dat_all, "blood_collection") |>
  left_join(select(dat_base, record_id, subjid), join_by(record_id)) |>
  select(-record_id, -starts_with("lbsite___"))

dat_igg <- select_form(dat_all, "igg") |>
  left_join(dat_bc, join_by(subjid, visage)) |>
  left_join(dat_base, join_by(subjid)) |>
  mutate(stage = if_else(substr(subjid, 1, 2) == "01", "1", "2")) |>
  filter(!(antigen %in% c("PnPs 11A"))) |>
  mutate(antigen = fct_inorder(antigen))

dat_base <- dat_igg |>
  select(all_of(colnames(dat_base)), concentration) |>
  mutate(
    any_obs_concentration = any(!is.na(concentration)),
    all_obs_concentration = all(!is.na(concentration)),
    .by = subjid
  ) |>
  select(-concentration) |>
  distinct()
```

\clearpage

# Background

In previous papers, we reported on the immunogenicity of a mixed wP/aP primary schedule compared to the standard aP-only primary schedule in terms of:

- IgG responses to PT, FHA, PRN, TT, and DT
- IgE-mediated immune responses to total IgE, hens' egg IgE, and TT

immediately before the aP dose at 6-months of age and 1-month after (at 7-months of age).

Subsequently, we reported on IgG-antibody responses to concurrently administered 13-valent pneumococcal conjugated vaccine (13vPCV), Hib capsular polysaccharide (Hib-PRP) and HBsAg at 6, 7, 18, and 19-months of age (i.e. immediately before and 1-month after the 6 and 18-month doses).

In both cases, these outcomes were restricted to the first 150 participants at PCH recruited as part of OPTIMUM stage 1.

As part of OPTIMUM stage 2, the next 150 participants enrolled at PCH were also invited to have bloods collected at 18 and 19-months of age. The previously analyses of IgG concentrations are updated to include these additional 150 participants, and to report on the 18 and 19-month IgG responses for PT, FHA, PRN, TT, and DT.

\clearpage

# Baseline

```{r}
#| label: tbl-baseline-characteristics
#| tbl-cap: Baseline characteristics of the combined 300 trial participants.
dat_base |>
  tbl_summary(
    include = c(
      gender,
      mpar,
      inmpertvp,
      vacad,
      delt,
      gestadelw,
      dage
    ),
    by = trt,
    missing_text = "(Missing)",
    label = list(
      gender = "Sex of child",
      mpar = "Maternal parity",
      dage = "Age (days) at enrolment"
    ),
    type = list(
      gestadelw ~ "continuous"
    ),
    digits = list(
      gestadelw = 0,
      dage = 0
    )
  ) |>
  add_n() |>
  add_overall(last = TRUE) |>
  modify_header(stat_by = md("**{level}**<br>N = {N}")) |>
  as_gt() |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = row_type == "label"
    )
  ) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(8)
  )
```

```{r}
#| label: tbl-baseline-characteristics-stage
#| tbl-cap: Baseline characteristics of the combined 300 trial participants, by stage.
dat_base |>
  tbl_summary(
    include = c(
      gender,
      mpar,
      inmpertvp,
      vacad,
      delt,
      gestadelw,
      dage
    ),
    by = rand_stage,
    missing_text = "(Missing)",
    label = list(
      gender = "Sex of child",
      mpar = "Maternal parity",
      dage = "Age (days) at enrolment"
    ),
    type = list(
      gestadelw ~ "continuous"
    ),
    digits = list(
      gestadelw = 0,
      dage = 0
    )
  ) |>
  add_n() |>
  add_overall(last = TRUE) |>
  modify_header(stat_by = md("**{level}**<br>N = {N}")) |>
  as_gt() |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = row_type == "label"
    )
  ) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(8)
  )
```

\clearpage

# Descriptive

In stage 1, bloods were collected at 6-months, 7-months, 18-months, and 19-months.
In stage 2, bloods were only collected at 18-months and 19-months and not all antigens were assessed for both ages.

- at 18-months, all responses except FIM2/3 were assessed (DT, FHA, PRN, PT, TT, HBsAg, Hib-PRP, and 13vPCV).
- at 19-months, only the DTaP responses (DT, FHA, PRN, PT, TT) and Hib-PRP were assessed.

## Missingness

```{r}
#| label: tbl-overall-missingness
#| tbl-cap: |
#|    Number and proportion of participants with at least one observed response,
#|    and with all planned responses observed.
#| tbl-pos: "H"
dat_base |>
  tbl_summary(
    include = c(any_obs_concentration, all_obs_concentration),
    label = list(
      any_obs_concentration ~ "At least one IgG response available",
      all_obs_concentration ~ "All planned IgG responses available"
    ),
    by = trt
  ) |>
  add_overall(last = TRUE) |>
  modify_header(stat_by = md("**{level}**<br>N = {N}")) |>
  as_gt() |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = row_type == "label"
    )
  ) |>
  tab_options(
    table.font.size = px(8)
  )
```

```{r}
#| label: tbl-igg-missingness-dtap
#| tbl-cap:
#|    Missingness pattern of IgG concentrations for DTaP and Hib-PRP by study stage.
#| tbl-pos: H
miss_pattern_trt_tab(filter(dat_igg, antigen == "PT")) |>
  tab_options(table.font.size = px(8))
```

```{r}
#| label: tbl-igg-missingness-pnp
#| tbl-cap:
#|    Missingness pattern of IgG concentrations for 13vPCV by study stage.
miss_pattern_trt_tab(filter(dat_igg, antigen == "PnPs 1")) |>
  tab_options(table.font.size = px(8))
```

\clearpage

## Concentrations

```{r}
#| label: tbl-igg-summary-dtap
#| tbl-cap: Descriptive statistics for DTaP IgG concentrations.
igg_gmc_summary_tab(
  filter(dat_igg, type != "PnPs")
) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(8)
  )
```

```{r}
#| label: tbl-igg-summary-pnps
#| tbl-cap: Descriptive statistics for PnP IgG concentrations.
igg_gmc_summary_tab(
  filter(dat_igg, type == "PnPs"),
  rn = "visage"
) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(8)
  )
```

```{r}
#| label: fig-igg-dtap
#| fig-cap: |
#|    IgG concentrations at 18 and 19-months of age for pertussis antigens by treatment group, stage 1 and 2 combined.
#| fig-width: 6
#| fig-height: 4
cols <- palette.colors(palette = "Okabe-Ito")[1:2]
dat_igg |>
  filter(type == "DTaP") |>
  ggplot(aes(interaction(trt, visage, sep = "_"), concentration)) +
  facet_wrap(~antigen) +
  geom_boxplot(aes(fill = trt)) +
  geom_hline(aes(yintercept = ref), linetype = 2) +
  scale_y_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  scale_x_discrete(
    "Age and treatment group",
    guide = guide_axis_nested(key = key_range_auto(sep = "_"))
  ) +
  scale_fill_manual(values = c("aP" = cols[1], "wP" = cols[2])) +
  guides(fill = "none", colour = "none") +
  theme(axis.title = element_text(size = rel(0.9)))
```

```{r}
#| label: fig-igg-pnps
#| fig-cap: |
#|    IgG concentrations at 18-months of age for pneumococcal antigens by treatment group, stage 1 and 2 combined.
#| fig-width: 6
#| fig-height: 8
cols <- palette.colors(palette = "Okabe-Ito")[1:2]
dat_igg |>
  filter(type == "PnPs") |>
  ggplot(aes(interaction(trt, visage, sep = "_"), concentration)) +
  facet_wrap(~antigen, ncol = 3) +
  geom_boxplot(aes(fill = trt)) +
  geom_hline(aes(yintercept = ref), linetype = 2) +
  scale_y_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  scale_x_discrete(
    "Age and treatment group",
    guide = guide_axis_nested(key = key_range_auto(sep = "_"))
  ) +
  scale_fill_manual(values = c("aP" = cols[1], "wP" = cols[2])) +
  guides(fill = "none", colour = "none") +
  theme(axis.title = element_text(size = rel(0.9)))
```

\clearpage

# Modelling

To estimate the geometric mean ratio (GMR) (wP versus aP) of each antibody type, a multivariate normal model with unstructured covariance was assumed for the $\log$-transformed IgG concentrations (at 6-month, 7-month, 18-month, and 19-months of age) with the age-specific mean response adjusting for the baseline covariates: sex, delivery type, breastfeeding status, first born, and family history of atopy. For infants $i=1,\ldots,n$ measured at age $j=1,2,3,4$ 

$$
\begin{aligned}
Y_{i} &\overset{\text{ind.}}{\sim} \text{Normal}(\mu_i, \Sigma) \\
\mu_{ij} &= \alpha_j + \sum_{p=1}^P \beta_{jp}x_{ip} \\
\Sigma &= \Lambda\Omega\Lambda \\
\Lambda &= \text{diag}(\sigma_1, \sigma_2, \sigma_3, \sigma_4) \\ 
\Omega &= \begin{pmatrix}
1 & \rho_{12} & \rho_{13} & \rho_{14} \\
\rho_{12} & 1 & \rho_{23} & \rho_{24} \\
\rho_{13} & \rho_{23} & 1 & \rho_{34} \\
\rho_{14} & \rho_{24} & \rho_{34} & 1
\end{pmatrix} \\
\alpha_j, \beta_{jp} &\overset{\text{ind.}}{\sim} \text{Normal}(0, 10) \\
\sigma_j &\overset{\text{ind.}}{\sim} \text{Cauchy}^+(0, 10) \\
\Omega &\sim \text{LKJ}(1).
\end{aligned}
$$

The analyses use all available responses.

## Stage 1 and 2 combined

```{r}
#| label: fit-igg-all
frm <- ~ trt + sex + caes + bfed + fborn + fha
mdat <- dat_igg |>
  select(
    antigen,
    subjid,
    stage,
    visage,
    concentration,
    log_concentration,
    trt,
    sex,
    caes,
    bfed,
    fborn,
    fha,
    inc
  ) |>
  mutate(antigen = droplevels(antigen)) |>
  complete(
    antigen,
    nesting(subjid, trt, sex, caes, bfed, fborn, fha, inc),
    visage,
    fill = list(concentration = NA, log_concentration = NA)
  ) |>
  nest(dat = -antigen)
mdat <- mdat |>
  mutate(
    sdat = map(dat, ~ make_stan_data(.x, frm)),
    sfit = map(
      sdat,
      ~ {
        multinorm$sample(
          data = .x,
          seed = 71033,
          chains = 8,
          iter_sampling = it_sample,
          iter_warmup = it_warmup,
          refresh = 0,
          show_messages = FALSE,
          show_exceptions = FALSE
        )
      }
    ),
    sdrw = map2(
      sfit,
      sdat,
      ~ {
        tmp <- as_draws_rvars(.x)
        rownames(tmp$beta) <- colnames(.y$Y)
        colnames(tmp$beta) <- colnames(.y$X)
        return(tmp)
      }
    ),
    gmr = map(
      sdrw,
      ~ {
        out <- 10^.x$beta[, 2]
        as_tibble(rbind(out), rownames = "age")
      }
    )
  )
gmr <- mdat |>
  select(antigen, gmr) |>
  unnest(gmr) |>
  mutate(group = "GMR") |>
  mutate(age = fct_inorder(age))
```

```{r}
#| label: fit-igg-all-conditional-pred
#| eval: false
ex_dat <- mdat$sdat[[1]]
ex_mod <- mdat$sdrw[[1]]

sub_matrix <- function(M, idx) {
  N <- length(idx)
  out <- matrix(0, N, N)
  for (i in 1:N) {
    for (j in 1:N) {
      out[i, j] <- M[idx[i], idx[j]]
    }
  }
  out
}

sub_offdiagonal_matrix <- function(M, idx1, idx2) {
  Nr <- length(idx1)
  Nc <- length(idx2)
  out <- matrix(0, Nr, Nc)
  for (i in 1:Nr) {
    for (j in 1:Nc) {
      out[i, j] <- M[idx1[i], idx2[j]]
    }
  }
  out
}

partition_matrix <- function(M, idx) {
  N <- nrow(M)
  J <- length(idx)
  idx_left <- setdiff(1:N, idx)
  S1 <- sub_matrix(M, idx)
  S2 <- sub_matrix(M, idx_left)
  S3 <- sub_diagonal_matrix(M, idx, idx_left)
  return(list(S1, S2, S3))
}

rvar_sub_matrix <- rfun(sub_matrix)
rvar_sub_offdiagonal_matrix <- rfun(sub_offdiagonal_matrix)

id <- 3
idx_obs <- which(ex_dat$R[id, ] == 0)
idx_mis <- which(ex_dat$R[id, ] == 1)

sigma2 <- rvar_sub_matrix(ex_mod$Sigma, idx_obs)
sigma1 <- rvar_sub_matrix(ex_mod$Sigma, idx_mis)
sigma12 <- rvar_sub_offdiagonal_matrix(ex_mod$Sigma, idx_mis, idx_obs)

s <- sigma1 - sigma3 %*% rfun(solve)(sigma2) %*% t(sigma3)
m <- ex_mod$eta[id, ][idx_mis] +
  sigma12 %*%
    rfun(solve)(sigma2) %*%
    (ex_dat$Y[id, ][idx_obs] - ex_mod$eta[id, ][idx_obs])
```

```{r}
#| label: tbl-igg-gmr
#| tbl-cap: |
#|    Adjusted IgG GMR posterior summaries by serotype/antigen and age.
posterior_gmr_table(gmr) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(9)
  )
```

```{r}
#| label: fig-gmr-igg-full-model
#| fig-cap: Posterior GMR (median and 95% CrI) for wP vs. aP by age and antigen, all available data.
#| fig-width: 8
#| fig-height: 10
des <- "
IA
JB
KC
LD
ME
NF
OG
PH
Q#
R#
S#
T#
U#
"
p <- gmr |>
  ggplot(aes(y = age, xdist = trtwP)) +
  ggh4x::facet_manual(vars(antigen), design = des, strip.position = "left") +
  stat_pointinterval(.width = 0.95) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  geom_vline(xintercept = 1, linetype = 2) +
  labs(fill = "Interval", x = "Geometric mean ratio", y = "Age") +
  scale_x_log10(breaks = c(0.4, 0.6, 0.8, 1, 1.2, 1.5, 2, 2.5, 3, 4)) +
  scale_y_discrete(limits = rev)
p
```

\clearpage

# Appendix {.unnumbered}

## Descriptive

```{r}
#| label: fig-igg-dtap-ecdf
#| fig-cap: |
#|    IgG concentrations at 18-months of age for pneumococcal antigens by treatment group, stage 1 and 2 combined.
#| fig-width: 6
#| fig-height: 6
cols <- c("black", "orange")
dat_igg |>
  filter(type != "PnPs") |>
  ggplot(aes(concentration)) +
  facet_grid(antigen ~ visage) +
  # stat_ecdf(aes(colour = trt)) +
  geom_step(aes(y = 1 - after_stat(y), colour = trt), stat = "ecdf") +
  scale_x_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  labs(y = "complementary eCDF") +
  scale_colour_manual("Treatment", values = c("aP" = cols[1], "wP" = cols[2])) +
  theme(
    axis.title = element_text(size = rel(0.9)),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r}
#| label: fig-igg-pnps-ecdf
#| fig-cap: |
#|    IgG concentrations at 18-months of age for pneumococcal antigens by treatment group, stage 1 and 2 combined.
#| fig-width: 6
#| fig-height: 8
cols <- c("black", "orange")
dat_igg |>
  filter(type == "PnPs") |>
  ggplot(aes(concentration)) +
  facet_grid(antigen ~ visage) +
  # stat_ecdf(aes(colour = trt)) +
  geom_step(aes(y = 1 - after_stat(y), colour = trt), stat = "ecdf") +
  scale_x_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  labs(y = "complementary eCDF") +
  scale_colour_manual("Treatment", values = c("aP" = cols[1], "wP" = cols[2])) +
  theme(
    axis.title = element_text(size = rel(0.9)),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

\clearpage

## Modelling

In the following, each stage is modelled independently.

### Stage 1 only

As a reference, the analyses were re-run restricting to only stage 1 participants.

```{r}
#| label: fit-igg-stage-1
frm <- ~ trt + sex + caes + bfed + fborn + fha
mdat <- dat_igg |>
  select(
    antigen,
    type,
    subjid,
    stage,
    visage,
    concentration,
    log_concentration,
    trt,
    sex,
    caes,
    bfed,
    fborn,
    fha,
    inc
  ) |>
  filter(stage == "1") |>
  mutate(antigen = droplevels(antigen)) |>
  complete(
    antigen,
    nesting(subjid, trt, sex, caes, bfed, fborn, fha, inc),
    visage,
    fill = list(concentration = NA, log_concentration = NA)
  ) |>
  nest(dat = -c(type, antigen)) |>
  mutate(
    sdat = map(dat, ~ make_stan_data(.x, frm)),
    sfit = map(
      sdat,
      ~ {
        multinorm$sample(
          data = .x,
          seed = 71033,
          chains = 8,
          iter_sampling = it_sample,
          iter_warmup = it_warmup,
          refresh = 0,
          show_messages = FALSE,
          show_exceptions = FALSE
        )
      }
    ),
    sdrw = map2(
      sfit,
      sdat,
      ~ {
        tmp <- as_draws_rvars(.x)
        rownames(tmp$beta) <- colnames(.y$Y)
        colnames(tmp$beta) <- colnames(.y$X)
        return(tmp)
      }
    ),
    gmr = map(
      sdrw,
      ~ {
        out <- 10^.x$beta[, 2]
        as_tibble(rbind(out), rownames = "age")
      }
    )
  )
gmr1 <- mdat |>
  select(antigen, gmr) |>
  unnest(gmr) |>
  mutate(group = "GMR") |>
  mutate(age = fct_inorder(age))
```

```{r}
#| label: tbl-igg-gmr-stage-1
#| tbl-cap: |
#|    Adjusted IgG GMR posterior summaries by serotype/antigen and age.
posterior_gmr_table(gmr1) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(9)
  )
```

```{r}
#| label: fig-gmr-igg-full-model-stage-1
#| fig-cap: Posterior GMR (median, 95% CrI) for wP vs. aP by age and antigen, all available data.
#| fig-width: 8
#| fig-height: 10
des <- "
IA
JB
KC
LD
ME
NF
OG
PH
Q#
R#
S#
T#
U#
"
p <- gmr1 |>
  ggplot(aes(y = age, xdist = trtwP)) +
  ggh4x::facet_manual(vars(antigen), design = des, strip.position = "left") +
  stat_pointinterval(.width = 0.95) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  geom_vline(xintercept = 1, linetype = 2) +
  labs(fill = "Interval", x = "Geometric mean ratio", y = "Age") +
  scale_x_log10(breaks = c(0.4, 0.6, 0.8, 1, 1.2, 1.5, 2, 2.5, 3, 4)) +
  scale_y_discrete(limits = rev)
p
```

\clearpage

### Stage 2 only

```{r}
#| label: fit-igg-stage-2
frm <- ~ trt + sex + caes + bfed + fborn + fha
mdat <- dat_igg |>
  select(
    antigen,
    type,
    subjid,
    stage,
    visage,
    concentration,
    log_concentration,
    trt,
    sex,
    caes,
    bfed,
    fborn,
    fha,
    inc
  ) |>
  filter(stage == "2") |>
  mutate(antigen = droplevels(antigen)) |>
  nest(dat = -c(type, antigen)) |>
  mutate(
    sdat = map(dat, ~ make_stan_data(.x, frm)),
    sfit = map(
      sdat,
      ~ {
        multinorm$sample(
          data = .x,
          seed = 71033,
          chains = 8,
          iter_sampling = it_sample,
          iter_warmup = it_warmup,
          refresh = 0,
          show_messages = FALSE,
          show_exceptions = FALSE
        )
      }
    ),
    sdrw = map2(
      sfit,
      sdat,
      ~ {
        tmp <- as_draws_rvars(.x)
        rownames(tmp$beta) <- colnames(.y$Y)
        colnames(tmp$beta) <- colnames(.y$X)
        return(tmp)
      }
    ),
    gmr = map(
      sdrw,
      ~ {
        out <- 10^.x$beta[, 2]
        as_tibble(rbind(out), rownames = "age")
      }
    )
  )
gmr2 <- mdat |>
  select(antigen, gmr) |>
  unnest(gmr) |>
  mutate(group = "GMR") |>
  mutate(age = fct_inorder(age))
```

```{r}
#| label: tbl-igg-gmr-stage-2
#| tbl-cap: |
#|    Adjusted IgG GMR posterior summaries by serotype/antigen and age.
posterior_gmr_table(gmr2) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(9)
  )
```

```{r}
#| label: fig-gmr-igg-full-model-stage-2
#| fig-cap: Posterior GMR (median, 95% CrI) for wP vs. aP by age and antigen, all available data.
#| fig-width: 8
#| fig-height: 7
des <- "
HA
IB
JC
KD
LE
MF
NG
O#
P#
Q#
R#
S#
T#
"
p <- gmr2 |>
  ggplot(aes(y = age, xdist = trtwP)) +
  ggh4x::facet_manual(vars(antigen), design = des, strip.position = "left") +
  stat_pointinterval(.width = 0.95) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  geom_vline(xintercept = 1, linetype = 2) +
  labs(fill = "Interval", x = "Geometric mean ratio", y = "Age") +
  scale_x_log10(breaks = c(0.4, 0.6, 0.8, 1, 1.2, 1.5, 2, 2.5, 3, 4)) +
  scale_y_discrete(limits = rev)
p
```

\clearpage

### Comparison

```{r}
#| label: fig-gmr-igg-full-model-stage-1vs2
#| fig-cap: Posterior GMR (median, 95% CrI) for wP vs. aP by age and antigen, all available data.
#| fig-width: 8
#| fig-height: 10
des <- "
IA
JB
KC
LD
ME
NF
OG
PH
Q#
R#
S#
T#
U#
"
gmr <- bind_rows(gmr1, gmr2, .id = "stage")
cols <- palette.colors(palette = "Okabe-Ito")[1:2]
p <- gmr |>
  ggplot(aes(y = age, xdist = trtwP)) +
  ggh4x::facet_manual(vars(antigen), design = des, strip.position = "left") +
  stat_pointinterval(
    aes(colour = stage),
    position = position_dodge(width = 0.6),
    .width = 0.95
  ) +
  scale_colour_manual(values = cols) +
  geom_vline(xintercept = 1, linetype = 2) +
  labs(fill = "Interval", x = "Geometric mean ratio", y = "Age") +
  scale_x_log10(breaks = c(0.4, 0.6, 0.8, 1, 1.2, 1.5, 2, 2.5, 3, 4)) +
  scale_y_discrete(limits = rev)
p
```
