---
title: OPTIMUM Stage 1 OPA Results
subtitle: Report
---

```{r}
#| label: packages-and-setup
suppressPackageStartupMessages({
  # Data
  library(qs)
  library(dplyr)
  library(tidyr)
  library(forcats)
  library(lubridate)
  library(labelled)

  # Figures
  library(ragg)
  library(ggplot2)
  library(scales)
  library(legendry)
  library(ggbeeswarm)

  # Tables
  library(gt)
  library(gtsummary)

  # Models
  library(mmrm)
  library(emmeans)
  library(brms)
  library(cmdstanr)
  library(posterior)
  library(tidybayes)
  library(ggdist)
  library(bayesplot)
  library(loo)
})

readRenviron(".env")

source("R/util.R")
source("reports/R/functions.R")

options(
  brms.backend = "cmdstanr",
  mc.cores = 8
)

sysfonts::font_add(
  family = "TeX Gyre Pagella",
  regular = "texgyrepagella-regular.otf",
  bold = "texgyrepagella-bold.otf",
  italic = "texgyrepagella-italic.otf",
  bolditalic = "texgyrepagella-bolditalic.otf"
)

theme_set(
  theme_bw(base_size = 11, base_family = "TeX Gyre Pagella") +
    theme(
      legend.position = "bottom",
      panel.grid.minor = element_blank(),
      axis.text = element_text(colour = "black"),
      legend.key.size = unit(0.75, "lines")
    )
)
```

```{r}
#| label: data-raw
com_file <- file.path(Sys.getenv("RDS_PATH"), config::get("combined_data_file"))
dat_raw <- qread(com_file)
dat_opa <- readxl::read_excel(
  file.path(
    Sys.getenv("RDS_PATH"),
    "data",
    "raw",
    "stage1",
    "240326 Optimum MOPA Results.xlsx"
  ),
  skip = 2
)
```

```{r}
dat_rand <- dat_raw |>
  select_form("randomisation") |>
  left_join(
    dat_raw |> select_form("allocations"),
    join_by(rand)
  )

dat_bc <- dat_raw |>
  select_form("blood_collection") |>
  select(record_id, visage, lbipdat) |>
  left_join(dat_rand, join_by(record_id))

dat_igg <- dat_raw |>
  select_form("igg") |>
  filter(
    grepl("01-", subjid),
    antigen %in% c("PnPs 3", "PnPs 19F", "PnPs 23F")
  ) |>
  left_join(dat_bc, join_by(subjid, visage))

dat_opa <- dat_opa |>
  rename(
    subjid = ID,
    visit = Visit,
    date = `Sample Date`,
    `PnPs 3` = `Serotype 3`,
    `PnPs 19F` = `Serotype 19F`,
    `PnPs 23F` = `Serotype 23F`
  ) |>
  mutate(subjid = gsub(" V3| V7", "", subjid)) |>
  mutate(
    date = as_date(date),
    visage = factor(visit, levels = c("V3", "V7"), labels = c("6-month", "18-month"))
  ) |>
  select(-visit) |>
  pivot_longer(`PnPs 3`:`PnPs 23F`, names_to = "antigen", values_to = "opa")

dat_com <- dat_igg |>
  filter(subjid %in% dat_opa$subjid, visage %in% c("6-month", "18-month")) |>
  left_join(dat_opa, join_by(subjid, visage, antigen)) |>
  mutate(antigen = fct_inorder(antigen))

n_subj <- length(unique(dat_com$subjid))
```

# Background


- multiplexed opsonophagocytic killing assay (MOPA)
- opsonophagocytic assay (OPA)
- Opsonisation titers/index (OI) (?serum dilution required to kill x% of bacteria)

Functional opsonophagocytosis antibody (OPA) levels may be a better predictor of protection than serum IgG.

Multiplexed opsonophagocytic killing assay (MOPA) measurements were provided for `{r} n_subj` stage 1 OPTIMUM participants. Measurements were taken on blood collections from visit 3 (6-month) and visit 7 (18-month) where available for each of these participants. Measurements were available for serotypes: PnPs 3, PnPs 19F, and PnPs 23F.

Aim is to:

- compare OPA between assigned treatment groups
- compare OPA to IgG

\clearpage

# OPA, IgG and Treatment Group

```{r}
#| label: fig-opa-antigen-visit
#| fig-cap: OPA by visit age and serotype.
#| fig-height: 3
#| fig-width: 5
dat_com |>
  ggplot(aes(visage, opa)) +
  facet_wrap(~antigen) +
  geom_boxplot(width = 0.25) +
  geom_beeswarm(shape = 21) +
  scale_y_continuous(
    "OPA",
    transform = "log2",
    breaks = scales::breaks_log(n = 10, base = 2),
    labels = label_log(base = 2)
  ) +
  labs(x = "Visit age")
```

```{r}
#| label: fig-opa-antigen-visit-trt
#| fig-cap: OPA by visit age, serotype, and assigned treatment group.
#| fig-height: 3
#| fig-width: 5
dat_com |>
  ggplot(aes(interaction(trt, visage, sep = "!"), opa)) +
  facet_wrap(~antigen) +
  geom_boxplot() +
  geom_beeswarm(shape = 21) +
  scale_y_continuous(
    "OPA",
    transform = "log2",
    breaks = scales::breaks_log(n = 10, base = 2),
    labels = label_log(base = 2)
  ) +
  guides(x = legendry::guide_axis_nested(key = key_range_auto(sep = "!"))) +
  labs(x = "Visit age | treatment")
```

```{r}
#| label: fig-igg-antigen-visit
#| fig-cap: IgG by visit age and serotype.
#| fig-height: 3
#| fig-width: 5
dat_com |>
  ggplot(aes(visage, concentration)) +
  facet_wrap(~antigen) +
  geom_boxplot(width = 0.25) +
  geom_beeswarm(shape = 21) +
  scale_y_continuous(
    "IgG",
    transform = "log2",
    breaks = scales::breaks_log(n = 10, base = 2),
    labels = label_log(base = 2)
  ) +
  labs(x = "Visit age")
```

```{r}
#| label: fig-igg-antigen-visit-trt
#| fig-cap: IgG by visit age, serotype, and assigned treatment group.
#| fig-height: 3
#| fig-width: 5
dat_com |>
  ggplot(aes(interaction(trt, visage, sep = "!"), concentration)) +
  facet_wrap(~antigen) +
  geom_boxplot() +
  geom_beeswarm(shape = 21) +
  scale_y_continuous(
    "IgG",
    transform = "log2",
    breaks = scales::breaks_log(n = 10, base = 2),
    labels = label_log(base = 2)
  ) +
  guides(x = legendry::guide_axis_nested(key = key_range_auto(sep = "!"))) +
  labs(x = "Visit age | treatment")
```

\clearpage

# OPA and IgG

```{r}
#| label: fig-igg-vs-opa-antigen
#| fig-cap: Correlation between IgG and OPA by antigen.
#| fig-height: 3
#| fig-width: 6
dat_cor <- dat_com |>
  select(subjid, visage, antigen, concentration, opa) |>
  nest(dat = -antigen) |>
  mutate(
    fit = map(dat, ~ lm(log(opa) ~ log(concentration), data = .x)),
    cor = map_dbl(dat, ~ cor(log(.x$concentration), log(.x$opa), use = "complete.obs")),
    rho = sprintf("Correlation = %.2f", cor)
  )

dat_com |>
  ggplot() +
  facet_wrap(~antigen) +
  geom_point(aes(concentration, opa)) +
  geom_smooth(aes(concentration, opa), method = "lm", se = FALSE) +
  geom_label(data = dat_cor, aes(label = rho, x = 1, y = 1), size = 3) +
  scale_x_continuous(
    "IgG",
    transform = "log2",
    breaks = 2^seq(-6, 6, by = 2),
    labels = label_log(base = 2)
  ) +
  scale_y_continuous(
    "OPA",
    transform = "log2",
    breaks = 2^seq(0, 12, by = 2),
    labels = label_log(base = 2)
  )
```

```{r}
#| label: fig-igg-vs-opa-antigen-visit-age
#| fig-cap: Correlation between IgG and OPA by antigen and visit age.
#| fig-height: 5
#| fig-width: 6
dat_cor <- dat_com |>
  select(subjid, visage, antigen, concentration, opa) |>
  nest(dat = -c(antigen, visage)) |>
  mutate(
    fit = map(dat, ~ lm(log(opa) ~ log(concentration), data = .x)),
    cor = map_dbl(dat, ~ cor(log(.x$concentration), log(.x$opa), use = "complete.obs")),
    rho = sprintf("Correlation = %.2f", cor)
  )

dat_com |>
  ggplot() +
  facet_grid(visage ~ antigen) +
  geom_point(aes(concentration, opa)) +
  geom_smooth(aes(concentration, opa), method = "lm", se = FALSE) +
  geom_label(data = dat_cor, aes(label = rho, x = 1, y = 1), size = 3) +
  scale_x_continuous(
    "IgG",
    transform = "log2",
    breaks = 2^seq(-6, 6, by = 2),
    labels = label_log(base = 2)
  ) +
  scale_y_continuous(
    "OPA",
    transform = "log2",
    breaks = 2^seq(0, 12, by = 2),
    labels = label_log(base = 2)
  )
```
