---
title: OPTIMUM Stage 1 OPA Results
subtitle: Report
---

```{r}
#| label: packages-and-setup
suppressPackageStartupMessages({
  # Data
  library(qs2)
  library(dplyr)
  library(tidyr)
  library(forcats)
  library(lubridate)
  library(labelled)

  # Figures
  library(ragg)
  library(ggplot2)
  library(scales)
  library(legendry)
  library(ggbeeswarm)
  library(patchwork)

  # Tables
  library(gt)
  library(gtsummary)

  # Models
  library(mmrm)
  library(emmeans)
  library(brms)
  library(cmdstanr)
  library(posterior)
  library(tidybayes)
  library(ggdist)
  library(bayesplot)
  library(loo)
})

readRenviron(".env")

source("R/util.R")
source("reports/R/functions.R")
source(file.path("R", "data", "process-combined-data.R"))

options(
  brms.backend = "cmdstanr",
  mc.cores = 8
)

sysfonts::font_add(
  family = "TeX Gyre Pagella",
  regular = "texgyrepagella-regular.otf",
  bold = "texgyrepagella-bold.otf",
  italic = "texgyrepagella-italic.otf",
  bolditalic = "texgyrepagella-bolditalic.otf"
)

theme_set(
  theme_bw(base_size = 11, base_family = "TeX Gyre Pagella") +
    theme(
      legend.position = "bottom",
      panel.grid.minor = element_blank(),
      axis.text = element_text(colour = "black"),
      legend.key.size = unit(0.75, "lines")
    )
)

multinorm <- cmdstan_model(file.path("stan", "multinorm.stan"))
```

```{r}
#| label: data-raw
com_file <- file.path(Sys.getenv("RDS_PATH"), config::get("combined_data_file"))
dat_raw <- qd_read(com_file)
dat_opa <- readxl::read_excel(
  file.path(
    Sys.getenv("RDS_PATH"),
    "data",
    "raw",
    "stage1",
    "240326 Optimum MOPA Results.xlsx"
  ),
  skip = 2
)

dat_opa <- dat_opa |>
  rename(
    subjid = ID,
    visit = Visit,
    date = `Sample Date`,
    `PnPs 3` = `Serotype 3`,
    `PnPs 19F` = `Serotype 19F`,
    `PnPs 23F` = `Serotype 23F`
  ) |>
  mutate(subjid = gsub(" V3| V7", "", subjid)) |>
  mutate(
    date = as_date(date),
    visage = factor(visit, levels = c("V3", "V7"), labels = c("6-month", "18-month"))
  ) |>
  select(-visit) |>
  pivot_longer(`PnPs 3`:`PnPs 23F`, names_to = "antigen", values_to = "opa")

opa_subj <- unique(dat_opa$subjid)
n_subj <- length(opa_subj)

dat_base <- get_baseline_data(dat_raw, TRUE) |>
  filter(subjid %in% opa_subj)

dat_bc <- dat_raw |>
  select_form("blood_collection") |>
  select(record_id, visage, lbipdat) |>
  left_join(select(dat_base, record_id, subjid), join_by(record_id)) |>
  select(-record_id)

dat_igg <- dat_raw |>
  select_form("igg") |>
  filter(
    grepl("01-", subjid),
    antigen %in% c("PnPs 3", "PnPs 19F", "PnPs 23F")
  ) |>
  left_join(dat_bc, join_by(subjid, visage)) |>
  filter(subjid %in% opa_subj, visage %in% c("6-month", "18-month"))

dat_com <- dat_base |>
  left_join(dat_igg, join_by(subjid)) |>
  left_join(dat_opa, join_by(subjid, visage, antigen)) |>
  mutate(antigen = fct_inorder(antigen))
```

# Background

Multiplexed opsonophagocytic killing assay (MOPA) measurements were provided for `{r} n_subj` stage 1 OPTIMUM participants. Measurements were taken on blood collections from visit 3 (6-month) and visit 7 (18-month) where available for each of these participants. Measurements were available for serotypes: PnPs 3, PnPs 19F, and PnPs 23F.

Aim is to:

- compare OPA between assigned treatment groups
- compare OPA to IgG

@sec-characteristics summarises the baseline characteristics of the infants with available OPA. @sec-opa-trt looks at OPA measurements by visit age and treatment group, in particular, the GMR of wP relative to aP at each age. @sec-igg-trt looks at the same, but for IgG restricted to those infants with OPA available. @sec-opa-igg-trt compares these. @sec-opa-igg-cor looks at correlations between OPA and IgG values.

```{=latex}
\clearpage
```

# Characteristics {#sec-characteristics}

@tbl-baseline-characteristics summarisese the characteristics of those infants who had OPA processed.

```{r}
#| label: tbl-baseline-characteristics
#| tbl-cap: Baseline characteristics of trial participants.
#| tbl-pos: "H"
#| output: "asis"
tbl <- dat_base |>
  set_variable_labels(
    etomthr___1 = "Indigenous Australia",
    etomthr___2 = "European Caucasian",
    etomthr___3 = "Asian",
    etomthr___4 = "Indian subcontinent",
    etomthr___5 = "Middle Eastern",
    etomthr___6 = "Pacific Islander",
    etomthr___7 = "South American",
    etomthr___8 = "Black African",
    etofthr___1 = "Indigenous Australia",
    etofthr___2 = "European Caucasian",
    etofthr___3 = "Asian",
    etofthr___4 = "Indian subcontinent",
    etofthr___5 = "Middle Eastern",
    etofthr___6 = "Pacific Islander",
    etofthr___7 = "South American",
    etofthr___8 = "Black African",
    etoinf___1 = "Indigenous Australia",
    etoinf___2 = "European Caucasian",
    etoinf___3 = "Asian",
    etoinf___4 = "Indian subcontinent",
    etoinf___5 = "Middle Eastern",
    etoinf___6 = "Pacific Islander",
    etoinf___7 = "South American",
    etoinf___8 = "Black African",
  ) |>
  mutate(
    bcmoth = fct_lump_min(bcmoth, min = 3) |>
      # fct_recode(Other = "OTHER") |>
      fct_infreq() |>
      fct_relevel("Other", after = Inf),
    bcfath = fct_lump_min(bcfath, min = 3) |>
      fct_recode(Other = "OTHER") |>
      fct_infreq() |>
      fct_relevel("Other", after = Inf)
  ) |>
  tbl_summary(
    include = c(
      # Demographics and infant history
      # rand_stage,
      # rand_site,
      gender,
      v1_age_days,
      parinc,
      bcmoth,
      bcfath,
      etomthr___2,
      etomthr___3,
      etomthr___4,
      etomthr___7,
      etomthr___1,
      etomthr___6,
      etomthr___8,
      etofthr___2,
      etofthr___3,
      etofthr___4,
      etofthr___6,
      etofthr___7,
      etofthr___1,
      etofthr___8,
      etoinf___2,
      etoinf___3,
      etoinf___4,
      etoinf___7,
      etoinf___6,
      etoinf___1,
      etoinf___8,
      fborn,
      fecurr,
      fecat,
      fedog,
      # Birth history
      mgrav_fac,
      mpar_fac,
      delt,
      ipab,
      neoab,
      gestadelw,
      wgt,
      len,
      inhc,
      aps1_fac,
      aps5_fac,
      # Maternal immunisation
      inmpertvp,
      vacad,
      prevpert,
      minfvp,
      # Family history
      fha
    ),
    label = list(
      # rand_stage ~ "Protocol stage",
      # rand_site ~ "Participating site",
      bcmoth ~ "Birth country of mother",
      bcfath ~ "Birth country of father",
      v1_age_days ~ "Age at enrolment (days)",
      fborn ~ "First born",
      mgrav_fac ~ "Maternal gravidity",
      mpar_fac ~ "Maternal parity",
      aps1_fac ~ "Apgar score (1 minute)",
      aps5_fac ~ "Apgar score (5 minutes)",
      prevpert ~ "Other maternal pertussis vaccination (last 5 years)",
      fha ~ "Family history of atopy",
      fecat ~ "Cat ownership",
      fedog ~ "Dog ownership",
      fecurr ~ "Feeding status"
    ),
    by = trt,
    missing_text = "(Missing)",
    digits = ~0
  ) |>
  add_n() |>
  add_overall(last = TRUE) |>
  add_variable_group_header(header = "Mother's ethnicity", variables = starts_with("etomthr")) |>
  add_variable_group_header(header = "Father's ethnicity", variables = starts_with("etofthr")) |>
  add_variable_group_header(header = "Infant's ethnicity", variables = starts_with("etoinf")) |>
  modify_column_alignment(all_stat_cols(), "right") |>
  modify_header(
    all_stat_cols() ~
      paste0(
        "\\makecell{{\\textbf{{{level}}} \\\\\\\\ ",
        "N = {formatC(n, format='d', big.mark=',')}}}"
      )
  ) |>
  modify_table_body(
    ~ .x |>
      mutate(
        n = if_else(grepl("___", variable), NA_character_, n),
        n = if_else(row_type == "variable_group", '53', n)
      )
  )
tbl |>
  as_gt() |>
  cols_width(label ~ pct(30)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(
      columns = label,
      rows = (row_type == "label" | row_type == "variable_group") & !grepl("___", variable)
    )
  ) |>
  tab_options(
    table.width = pct(100),
    table.font.size = px(9),
    latex.use_longtable = TRUE
  ) |>
  tab_footnote(
    footnote = "Multiple may apply to each participant",
    locations = cells_body(rows = row_type == "variable_group", columns = label)
  ) |>
  as_latex() |>
  latex_unescape() |>
  fix_footnote() |>
  cat()
```

```{=latex}
\clearpage
```

# OPA by age and treatment group {#sec-opa-trt}

A summary of the available OPA measurements visit age is shown in @fig-opa-antigen-visit.
A summary by visit age and treatment group is shown in @fig-opa-antigen-visit-trt.

```{r}
#| label: fig-opa-antigen-visit
#| fig-cap: OPA by visit age and serotype. Y-axis is $\log_2$ transformed.
#| fig-height: 3
#| fig-width: 5
dat_com |>
  ggplot(aes(visage, opa)) +
  facet_wrap(~antigen) +
  geom_boxplot(width = 0.25) +
  geom_beeswarm(shape = 21, alpha = 0.75) +
  scale_y_continuous(
    "OPA",
    transform = "log2",
    breaks = scales::breaks_log(n = 10, base = 2),
    labels = label_log(base = 2)
  ) +
  labs(x = "Visit age")
```

```{r}
#| label: fig-opa-antigen-visit-trt
#| fig-cap: OPA by visit age, serotype, and assigned treatment group. Y-axis is $\log_2$ transformed.
#| fig-height: 3
#| fig-width: 5
dat_com |>
  ggplot(aes(interaction(trt, visage, sep = "!"), opa)) +
  facet_wrap(~antigen) +
  geom_boxplot() +
  geom_beeswarm(shape = 21) +
  scale_y_continuous(
    "OPA",
    transform = "log2",
    breaks = scales::breaks_log(n = 10, base = 2),
    labels = label_log(base = 2)
  ) +
  guides(x = legendry::guide_axis_nested(key = key_range_auto(sep = "!"))) +
  labs(x = "Visit age | treatment")
```

```{r}
#| label: fig-opa-antigen-visit-trt-ecdf
#| fig-cap: OPA ceCDF by visit age, serotype, and assigned treatment group. X-axis is $\log_2$ transformed.
#| fig-height: 4
#| fig-width: 6
dat_com |>
  ggplot(aes(opa)) +
  facet_grid(visage ~ antigen) +
  stat_ecdf(aes(colour = trt, y = after_stat(1 - ecdf))) +
  scale_x_continuous(
    "OPA",
    transform = "log2",
    breaks = scales::breaks_log(n = 10, base = 2),
    labels = label_log(base = 2)
  ) +
  labs(y = "eCDF", colour = "Treatment") +
  scale_colour_manual(values = c("black", "orange"))
```

```{r}
#| label: tbl-opa-summaries
#| tbl-cap: |
#|   Descriptive summaries of OPA measurements by age and treatment group.
#| tbl-pos: "H"
#| results: "asis"
var <- sym("opa")
tab_dat <- dat_com |>
  summarise(
    obs = sum(!is.na({{ var }})),
    gmc = gmc({{ var }}, na.rm = TRUE),
    gsd = gsd({{ var }}, na.rm = TRUE),
    q1 = quantile({{ var }}, probs = 0.25, na.rm = TRUE),
    q3 = quantile({{ var }}, probs = 0.75, na.rm = TRUE),
    min = min({{ var }}, na.rm = TRUE),
    max = max({{ var }}, na.rm = TRUE),
    .by = c(antigen, visage, trt)
  ) |>
  arrange(trt) |>
  pivot_wider(
    names_from = trt,
    values_from = c(obs, gmc, gsd, q1, q3, min, max),
    names_vary = "slowest"
  ) |>
  to_factor() |>
  arrange(antigen, visage)
tab <- gt(tab_dat, groupname_col = "antigen", rowname_col = "visage") |>
  fmt_integer(starts_with("obs")) |>
  fmt_number(starts_with(c("gmc", "gsd")), decimals = 2) |>
  fmt(starts_with(c("q1", "q3")), fns = fmt_num) |>
  fmt(starts_with(c("min", "max")), fns = fmt_num) |>
  cols_merge(
    columns = c(min_aP, max_aP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(min_wP, max_wP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(q1_aP, q3_aP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(q1_wP, q3_wP),
    pattern = "{1}–{2}"
  ) |>
  cols_label(
    obs_aP = "n",
    obs_wP = "n",
    gmc_aP = "GMC",
    gmc_wP = "GMC",
    gsd_aP = "GSD",
    gsd_wP = "GSD",
    q1_aP = "Q1–Q3",
    q1_wP = "Q1–Q3",
    min_aP = "Min–Max",
    min_wP = "Min–Max"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{aP} \\\\ N = 29}",
    columns = ends_with("_aP")
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{wP} \\\\ N = 24}",
    columns = ends_with("_wP")
  ) |>
  sub_missing(
    columns = everything(),
    rows = everything(),
    missing_text = "---"
  ) |>
  tab_footnote(
    footnote = "Geometric mean concentration",
    locations = cells_column_labels(columns = starts_with("gmc"))
  ) |>
  tab_footnote(
    footnote = "Geometric standard deviation",
    locations = cells_column_labels(columns = starts_with("gsd"))
  ) |>
  tab_footnote(
    footnote = "Q1 - 25th sample percentile, Q3 - 75th sample percentile",
    locations = cells_column_labels(columns = starts_with("q1"))
  ) |>
  tab_footnote(
    footnote = "Seroprotection at specified levels",
    locations = cells_column_labels(columns = starts_with("pos"))
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_column_labels(columns = everything())
  ) |>
  tab_options(table.font.size = px(9))
tab |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: fit-opa-all
make_stan_data <- function(dd, var, form = ~trt) {
  tmp <- dd |>
    filter(row_number() == 1, .by = subjid) |>
    select(-visage) |>
    distinct() |>
    to_factor(drop_unused_labels = TRUE)
  X <- model.matrix(form, data = tmp)
  Y <- dd |>
    select(subjid, visage, {{ var }}) |>
    spread(visage, {{ var }}) |>
    select(-subjid) |>
    mutate(across(everything(), ~ replace_na(.x, -99))) |>
    as.matrix()
  R <- dd |>
    mutate(
      R = case_when(
        is.na({{ var }}) ~ 1,
        !is.na({{ var }}) ~ 0
      )
    ) |>
    select(subjid, visage, R) |>
    spread(visage, R) |>
    select(-subjid) |>
    as.matrix()
  out <- list(
    N = nrow(tmp),
    J = ncol(Y),
    X = X,
    P = ncol(X),
    Y = Y,
    R = R
  )
  return(out)
}


mdat <- dat_com |>
  select(
    antigen,
    subjid,
    visage,
    trt,
    opa
  ) |>
  mutate(
    log_opa = log10(opa),
    antigen = droplevels(antigen),
    visage = droplevels(visage)
  ) |>
  nest(dat = -antigen)

mdat_opa <- mdat |>
  mutate(
    sdat = map(dat, ~ make_stan_data(.x, log_opa, ~trt)),
    sfit = map(
      sdat,
      ~ {
        multinorm$sample(
          data = .x,
          seed = 71033,
          chains = 8,
          iter_sampling = 625,
          iter_warmup = 500,
          refresh = 0,
          show_messages = FALSE,
          show_exceptions = FALSE
        )
      }
    ),
    sdrw = map2(
      sfit,
      sdat,
      ~ {
        tmp <- as_draws_rvars(.x)
        rownames(tmp$beta) <- colnames(.y$Y)
        colnames(tmp$beta) <- colnames(.y$X)
        return(tmp)
      }
    ),
    gmr = map(
      sdrw,
      ~ {
        out <- 10^.x$beta[, 2]
        as_tibble(rbind(out), rownames = "visage")
      }
    )
  )

gmr_opa <- mdat_opa |>
  select(antigen, gmr) |>
  unnest(gmr) |>
  mutate(group = "GMR") |>
  mutate(visage = fct_inorder(visage))
```

```{r}
#| label: opa-tab-data
tab_dat_opa <- gmr_opa |>
  arrange(antigen) |>
  mutate(
    antigen = fct_inorder(antigen),
    visage = fct_inorder(visage),
    group = fct_inorder(group)
  ) |>
  arrange(antigen, group, visage) |>
  mutate(
    mean = mean(trtwP),
    sd = sd(trtwP),
    med = median(trtwP),
    lo = quantile(trtwP, 0.025)[1, ],
    hi = quantile(trtwP, 0.975)[1, ],
    pr = Pr(trtwP > 1),
    pr_ninf = if_else(group == "GMR", Pr(trtwP > 2 / 3), NA_real_)
  ) |>
  select(-trtwP, -group)

tab_opa <- gt(tab_dat_opa, groupname_col = "antigen", rowname_col = "visage") |>
  fmt_number(2:9, decimals = 2) |>
  cols_merge(
    columns = c(mean, sd),
    pattern = "<<{1} ± {2}>>"
  ) |>
  cols_merge(
    columns = c(lo, hi),
    pattern = "<<({1}, {2})>>"
  ) |>
  cols_label(
    visage = "Age",
    mean = md("Mean ± std"),
    med = "Median",
    lo = "95% CrI",
    pr = "Pr(> 1)",
    pr_ninf = "Pr(> 2/3)"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  cols_align(
    columns = 2,
    align = "left"
  ) |>
  sub_missing(
    columns = everything(),
    rows = everything(),
    missing_text = "---"
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_column_labels(columns = everything())
  )
```

```{r}
#| label: fig-gmr-opa-model
#| fig-cap: Posterior GMR (median and 95% CrI) for wP versus aP by age and antigen for OPA.
#| fig-width: 5
#| fig-height: 3
p <- gmr_opa |>
  ggplot(aes(y = visage, xdist = trtwP)) +
  facet_wrap(vars(antigen), ncol = 1, strip.position = "left") +
  stat_pointinterval(.width = c(0.8, 0.95)) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  geom_vline(xintercept = 1, linetype = 2) +
  labs(x = "Geometric mean ratio", y = "Age") +
  scale_x_log10(breaks = c(0.4, 0.6, 0.8, 1, 1.2, 1.5, 2, 2.5, 3, 4)) +
  scale_y_discrete(limits = rev)
p
```

```{r}
#| label: fig-gmr-opa-model-ppc
#| fig-cap: Posterior predictive eCDF by antigen and treatment group.
#| fig-width: 5
#| fig-height: 5
plot_pp_ecdf <- function(sfit, sdat) {
  Ypred <- c(as_draws_rvars(sfit)[["Ypred"]])
  Y <- c(sdat$Y)
  G <- factor(sdat$X[, 2], levels = 0:1, labels = c("aP", "wP"))
  Ypred <- as_draws_matrix(Ypred[Y != -99])[1:100, ]
  G <- c(G, G)[Y != -99]
  Y <- Y[Y != -99]
  ppc_ecdf_overlay_grouped(
    Y,
    Ypred,
    G
  )
}
pp <- map2(mdat_opa$sfit, mdat_opa$sdat, plot_pp_ecdf)
(pp[[1]] + labs(title = mdat_opa$antigen[[1]])) /
  (pp[[2]] + labs(title = mdat_opa$antigen[[2]])) /
  (pp[[3]] + labs(title = mdat_opa$antigen[[3]])) +
  plot_layout(guides = "collect") &
  theme(title = element_text(size = 8))
```

```{=latex}
\clearpage
```

# IgG by age and treatment group {#sec-igg-trt}

```{r}
#| label: fig-igg-antigen-visit
#| fig-cap: IgG by visit age and serotype.
#| fig-height: 3
#| fig-width: 5
dat_com |>
  ggplot(aes(visage, concentration)) +
  facet_wrap(~antigen) +
  geom_boxplot(width = 0.25) +
  geom_beeswarm(shape = 21) +
  scale_y_continuous(
    "IgG",
    transform = "log2",
    breaks = scales::breaks_log(n = 10, base = 2),
    labels = label_log(base = 2)
  ) +
  labs(x = "Visit age")
```

```{r}
#| label: fig-igg-antigen-visit-trt
#| fig-cap: IgG by visit age, serotype, and assigned treatment group.
#| fig-height: 3
#| fig-width: 5
dat_com |>
  ggplot(aes(interaction(trt, visage, sep = "!"), concentration)) +
  facet_wrap(~antigen) +
  geom_boxplot() +
  geom_beeswarm(shape = 21) +
  scale_y_continuous(
    "IgG",
    transform = "log2",
    breaks = scales::breaks_log(n = 10, base = 2),
    labels = label_log(base = 2)
  ) +
  guides(x = legendry::guide_axis_nested(key = key_range_auto(sep = "!"))) +
  labs(x = "Visit age | treatment")
```

```{r}
#| label: fig-igg-antigen-visit-trt-ecdf
#| fig-cap: IgG ceCDF by visit age, serotype, and assigned treatment group. X-axis is $\log_2$ transformed.
#| fig-height: 4
#| fig-width: 6
dat_com |>
  ggplot(aes(concentration)) +
  facet_grid(visage ~ antigen) +
  stat_ecdf(aes(colour = trt, y = after_stat(1 - ecdf))) +
  scale_x_continuous(
    "IgG",
    transform = "log2",
    breaks = scales::breaks_log(n = 10, base = 2),
    labels = label_log(base = 2)
  ) +
  labs(y = "eCDF", colour = "Treatment") +
  scale_colour_manual(values = c("black", "orange"))
```

```{r}
#| label: tbl-igg-summaries
#| tbl-cap: |
#|   Descriptve summaries of IgG concentrations by age and treatment group.
#| tbl-pos: "H"
#| results: "asis"
var <- sym("concentration")
tab_dat <- dat_com |>
  summarise(
    obs = sum(!is.na({{ var }})),
    gmc = gmc({{ var }}, na.rm = TRUE),
    gsd = gsd({{ var }}, na.rm = TRUE),
    q1 = quantile({{ var }}, probs = 0.25, na.rm = TRUE),
    q3 = quantile({{ var }}, probs = 0.75, na.rm = TRUE),
    min = min({{ var }}, na.rm = TRUE),
    max = max({{ var }}, na.rm = TRUE),
    .by = c(antigen, visage, trt)
  ) |>
  arrange(trt) |>
  pivot_wider(
    names_from = trt,
    values_from = c(obs, gmc, gsd, q1, q3, min, max),
    names_vary = "slowest"
  ) |>
  to_factor() |>
  arrange(antigen, visage)
tab <- gt(tab_dat, groupname_col = "antigen", rowname_col = "visage") |>
  fmt_integer(starts_with("obs")) |>
  fmt_number(starts_with(c("gmc", "gsd")), decimals = 2) |>
  fmt(starts_with(c("q1", "q3")), fns = fmt_num) |>
  fmt(starts_with(c("min", "max")), fns = fmt_num) |>
  cols_merge(
    columns = c(min_aP, max_aP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(min_wP, max_wP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(q1_aP, q3_aP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(q1_wP, q3_wP),
    pattern = "{1}–{2}"
  ) |>
  cols_label(
    obs_aP = "n",
    obs_wP = "n",
    gmc_aP = "GMC",
    gmc_wP = "GMC",
    gsd_aP = "GSD",
    gsd_wP = "GSD",
    q1_aP = "Q1–Q3",
    q1_wP = "Q1–Q3",
    min_aP = "Min–Max",
    min_wP = "Min–Max"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{aP} \\\\ N = 29}",
    columns = ends_with("_aP")
  ) |>
  tab_spanner(
    label = "\\makecell{\\textbf{wP} \\\\ N = 24}",
    columns = ends_with("_wP")
  ) |>
  sub_missing(
    columns = everything(),
    rows = everything(),
    missing_text = "---"
  ) |>
  tab_footnote(
    footnote = "Geometric mean concentration",
    locations = cells_column_labels(columns = starts_with("gmc"))
  ) |>
  tab_footnote(
    footnote = "Geometric standard deviation",
    locations = cells_column_labels(columns = starts_with("gsd"))
  ) |>
  tab_footnote(
    footnote = "Q1 - 25th sample percentile, Q3 - 75th sample percentile",
    locations = cells_column_labels(columns = starts_with("q1"))
  ) |>
  tab_footnote(
    footnote = "Seroprotection at specified levels",
    locations = cells_column_labels(columns = starts_with("pos"))
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_column_labels(columns = everything())
  ) |>
  tab_options(table.font.size = px(9))
tab |>
  as_latex() |>
  latex_unescape() |>
  cat()
```

```{r}
#| label: fit-igg-all
mdat <- dat_com |>
  select(
    antigen,
    subjid,
    visage,
    trt,
    log_concentration
  ) |>
  mutate(
    antigen = droplevels(antigen),
    visage = droplevels(visage)
  ) |>
  nest(dat = -antigen)

mdat_igg <- mdat |>
  mutate(
    sdat = map(dat, ~ make_stan_data(.x, log_concentration, ~trt)),
    sfit = map(
      sdat,
      ~ {
        multinorm$sample(
          data = .x,
          seed = 71033,
          chains = 8,
          iter_sampling = 625,
          iter_warmup = 500,
          refresh = 0,
          show_messages = FALSE,
          show_exceptions = FALSE
        )
      }
    ),
    sdrw = map2(
      sfit,
      sdat,
      ~ {
        tmp <- as_draws_rvars(.x)
        rownames(tmp$beta) <- colnames(.y$Y)
        colnames(tmp$beta) <- colnames(.y$X)
        return(tmp)
      }
    ),
    gmr = map(
      sdrw,
      ~ {
        out <- 10^.x$beta[, 2]
        as_tibble(rbind(out), rownames = "visage")
      }
    )
  )

gmr_igg <- mdat_igg |>
  select(antigen, gmr) |>
  unnest(gmr) |>
  mutate(group = "GMR") |>
  mutate(visage = fct_inorder(visage))
```

```{r}
#| label: igg-tab-data
tab_dat_igg <- gmr_igg |>
  arrange(antigen) |>
  mutate(
    antigen = fct_inorder(antigen),
    visage = fct_inorder(visage),
    group = fct_inorder(group)
  ) |>
  arrange(antigen, group, visage) |>
  mutate(
    mean = mean(trtwP),
    sd = sd(trtwP),
    med = median(trtwP),
    lo = quantile(trtwP, 0.025)[1, ],
    hi = quantile(trtwP, 0.975)[1, ],
    pr = Pr(trtwP > 1),
    pr_ninf = if_else(group == "GMR", Pr(trtwP > 2 / 3), NA_real_)
  ) |>
  select(-trtwP, -group)

tab_igg <- gt(tab_dat_igg, groupname_col = "antigen", rowname_col = "visage") |>
  fmt_number(2:9, decimals = 2) |>
  cols_merge(
    columns = c(mean, sd),
    pattern = "<<{1} ± {2}>>"
  ) |>
  cols_merge(
    columns = c(lo, hi),
    pattern = "<<({1}, {2})>>"
  ) |>
  cols_label(
    visage = "Age",
    mean = md("Mean ± std"),
    med = "Median",
    lo = "95% CrI",
    pr = "Pr(> 1)",
    pr_ninf = "Pr(> 2/3)"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  cols_align(
    columns = 2,
    align = "left"
  ) |>
  sub_missing(
    columns = everything(),
    rows = everything(),
    missing_text = "---"
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_column_labels(columns = everything())
  )
```

```{r}
#| label: fig-gmr-igg-model
#| fig-cap: Posterior GMR (median and 95% CrI) for wP versus aP by age and antigen for IgG.
#| fig-width: 5
#| fig-height: 3
p <- gmr_igg |>
  ggplot(aes(y = visage, xdist = trtwP)) +
  facet_wrap(vars(antigen), ncol = 1, strip.position = "left") +
  stat_pointinterval(.width = c(0.8, 0.95)) +
  scale_fill_brewer(palette = "Reds", na.translate = FALSE) +
  geom_vline(xintercept = 1, linetype = 2) +
  labs(x = "Geometric mean ratio", y = "Age") +
  scale_x_log10(breaks = c(0.4, 0.6, 0.8, 1, 1.2, 1.5, 2, 2.5, 3, 4)) +
  scale_y_discrete(limits = rev)
p
```

```{r}
#| label: fig-gmr-igg-model-ppc
#| fig-cap: Posterior predictive eCDF by antigen and treatment group.
#| fig-width: 5
#| fig-height: 5
pp <- map2(mdat_igg$sfit, mdat_igg$sdat, plot_pp_ecdf)
(pp[[1]] + labs(title = mdat_igg$antigen[[1]])) /
  (pp[[2]] + labs(title = mdat_igg$antigen[[2]])) /
  (pp[[3]] + labs(title = mdat_igg$antigen[[3]])) +
  plot_layout(guides = "collect") &
  theme(title = element_text(size = 8))
```

```{=latex}
\clearpage
```

# OPA and IgG GMR {#sec-opa-igg-trt}

The estimated GMR for wP versus aP by visit age for OPA and IgG are shown in @tbl-gmr-igg-opa.
The GMR for OPA were generally lower and closer to 1 compared to IgG.
For PnPs 3, GMR at 18-months was higher than 6-months for IgG but was lower for OPA.

```{r}
#| label: tbl-gmr-igg-opa
#| tbl-cap: |
#|   Summary of estimated posterior geometric mean ratios (wP versus aP) for IgG and OPA by age.
#| tbl-pos: "H"
tab_dat <- left_join(
  tab_dat_igg |>
    rename_with(.fn = ~ paste0(.x, "_igg"), mean:pr_ninf),
  tab_dat_opa |>
    rename_with(.fn = ~ paste0(.x, "_opa"), mean:pr_ninf),
  join_by(antigen, visage)
)

gt(tab_dat, groupname_col = "antigen", rowname_col = "visage") |>
  fmt_number(2:16, decimals = 2) |>
  cols_merge(
    columns = c(mean_igg, sd_igg),
    pattern = "<<{1} ± {2}>>"
  ) |>
  cols_merge(
    columns = c(lo_igg, hi_igg),
    pattern = "<<({1}, {2})>>"
  ) |>
  cols_merge(
    columns = c(mean_opa, sd_opa),
    pattern = "<<{1} ± {2}>>"
  ) |>
  cols_merge(
    columns = c(lo_opa, hi_opa),
    pattern = "<<({1}, {2})>>"
  ) |>
  cols_label(
    visage ~ "Age",
    starts_with("mean") ~ md("Mean ± std"),
    starts_with("med") ~ "Median",
    starts_with("lo") ~ "95% CrI",
    starts_with("pr") ~ "Pr(> 1)",
    starts_with("pr_ninf") ~ "Pr(> 2/3)"
  ) |>
  tab_stub_indent(
    rows = everything(),
    indent = 1
  ) |>
  cols_align(
    columns = 2,
    align = "left"
  ) |>
  sub_missing(
    columns = everything(),
    rows = everything(),
    missing_text = "---"
  ) |>
  tab_spanner(ends_with("_igg"), label = "IgG") |>
  tab_spanner(ends_with("_opa"), label = "OPA") |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_column_labels(columns = everything())
  ) |>
  tab_options(table.font.size = px(8))
```

```{=latex}
\clearpage
```

# OPA and IgG correlations {#sec-opa-igg-cor}

```{r}
#| label: fig-igg-vs-opa-antigen
#| fig-cap: Correlation between IgG and OPA by antigen.
#| fig-height: 2.5
#| fig-width: 6
dat_cor <- dat_com |>
  select(subjid, visage, antigen, concentration, opa) |>
  nest(dat = -antigen) |>
  mutate(
    fit = map(dat, ~ lm(log(opa) ~ log(concentration), data = .x)),
    cor = map_dbl(dat, ~ cor(log(.x$concentration), log(.x$opa), use = "complete.obs")),
    rho = sprintf("Correlation = %.2f", cor)
  )

dat_com |>
  ggplot() +
  facet_wrap(~antigen) +
  geom_point(aes(concentration, opa)) +
  geom_smooth(aes(concentration, opa), method = "lm", se = FALSE) +
  geom_label(data = dat_cor, aes(label = rho, x = 1, y = 1), size = 3) +
  scale_x_continuous(
    "IgG",
    transform = "log2",
    breaks = 2^seq(-6, 6, by = 2),
    labels = label_log(base = 2)
  ) +
  scale_y_continuous(
    "OPA",
    transform = "log2",
    breaks = 2^seq(0, 12, by = 2),
    labels = label_log(base = 2)
  )
```

```{r}
#| label: fig-igg-vs-opa-antigen-visit-age
#| fig-cap: Correlation between IgG and OPA by antigen and visit age.
#| fig-height: 4
#| fig-width: 6
dat_cor <- dat_com |>
  select(subjid, visage, antigen, concentration, opa) |>
  nest(dat = -c(antigen, visage)) |>
  mutate(
    fit = map(dat, ~ lm(log(opa) ~ log(concentration), data = .x)),
    cor = map_dbl(dat, ~ cor(log(.x$concentration), log(.x$opa), use = "complete.obs")),
    rho = sprintf("Correlation = %.2f", cor)
  )

dat_com |>
  ggplot() +
  facet_grid(visage ~ antigen) +
  geom_point(aes(concentration, opa)) +
  geom_smooth(aes(concentration, opa), method = "lm", se = FALSE) +
  geom_label(data = dat_cor, aes(label = rho, x = 1, y = 1), size = 3) +
  scale_x_continuous(
    "IgG",
    transform = "log2",
    breaks = 2^seq(-6, 6, by = 2),
    labels = label_log(base = 2)
  ) +
  scale_y_continuous(
    "OPA",
    transform = "log2",
    breaks = 2^seq(0, 12, by = 2),
    labels = label_log(base = 2)
  )
```

```{r}
#| label: fig-igg-vs-opa-antigen-visit-age-trt
#| fig-cap: Correlation between IgG and OPA by antigen and visit age.
#| fig-height: 4
#| fig-width: 6
dat_cor <- dat_com |>
  select(subjid, visage, trt, antigen, concentration, opa) |>
  nest(dat = -c(antigen, visage, trt)) |>
  mutate(
    fit = map(dat, ~ lm(log(opa) ~ log(concentration), data = .x)),
    cor = map_dbl(dat, ~ cor(log(.x$concentration), log(.x$opa), use = "complete.obs")),
    rho = sprintf("Corr. = %.2f", cor),
    x = if_else(trt == "aP", 2^(-4), 2^3)
  )

dat_com |>
  ggplot(aes(colour = trt)) +
  facet_grid(visage ~ antigen) +
  geom_point(aes(concentration, opa)) +
  geom_smooth(aes(concentration, opa), method = "lm", se = FALSE) +
  geom_label(data = dat_cor, aes(label = rho, x = x, y = 1), size = 2.5) +
  scale_x_continuous(
    "IgG",
    transform = "log2",
    breaks = 2^seq(-6, 6, by = 2),
    labels = label_log(base = 2)
  ) +
  scale_y_continuous(
    "OPA",
    transform = "log2",
    breaks = 2^seq(0, 12, by = 2),
    labels = label_log(base = 2)
  ) +
  scale_colour_manual(values = c("black", "orange"))
```
