---
title: OPTIMUM Stage 1 and 2 IgG Concentrations
subtitle: Investigation Report
---

```{r}
#| label: packages-and-setup
suppressPackageStartupMessages({
  # Data
  library(qs)
  library(dplyr)
  library(tidyr)
  library(forcats)
  library(lubridate)
  library(labelled)

  # Figures
  library(ggplot2)
  library(scales)
  library(legendry)

  # Tables
  library(gt)
  library(gtsummary)

  # Models
  library(mmrm)
  library(emmeans)
  library(brms)
  library(cmdstanr)
  library(posterior)
  library(tidybayes)
  library(ggdist)
  library(bayesplot)
  library(loo)
})

readRenviron(".env")

source("R/util.R")
source("reports/functions.R")

options(
  brms.backend = "cmdstanr",
  mc.cores = 8
)

sysfonts::font_add(
  family = "TeX Gyre Pagella",
  regular = "texgyrepagella-regular.otf",
  bold = "texgyrepagella-bold.otf",
  italic = "texgyrepagella-italic.otf",
  bolditalic = "texgyrepagella-bolditalic.otf"
)

bs <- 9

theme_set(
  theme_bw(base_size = bs, base_family = "TeX Gyre Pagella") +
    theme(
      legend.position = "bottom",
      panel.grid.minor = element_blank(),
      axis.text = element_text(colour = "black"),
      legend.key.size = unit(0.75, "lines")
    )
)

it_warmup <- 1250
it_sample <- 500
```

```{r}
#| label: model
multinorm <- cmdstan_model(file.path("stan", "multinorm.stan"))
multinorm_het <- cmdstan_model(file.path("stan", "multinorm-het.stan"))
multit <- cmdstan_model(file.path("stan", "multit.stan"))
```

```{r}
#| label: data
dat_rand_list <- read_randomisation_lists()
dat_all <- qread(file.path(
  Sys.getenv("RDS_PATH"),
  config::get("combined_data_file")
))

dat_base <- left_join(
  select_form(dat_all, "randomisation") |>
    mutate(randdate = date(randdattim)),
  dat_rand_list,
  join_by(rand)
) |>
  left_join(
    select_form(dat_all, "demographics"),
    join_by(record_id)
  ) |>
  left_join(
    select_form(dat_all, "birth_history"),
    join_by(record_id)
  ) |>
  left_join(
    select_form(dat_all, "food_and_household_questionnaire") |>
      filter(visit_age == "6-week") |>
      select(record_id, fecurr),
    join_by(record_id)
  ) |>
  left_join(
    select_form(dat_all, "family_history_of_atopy") |>
      select(record_id, fha),
    join_by(record_id)
  ) |>
  mutate(
    dage = interval(birthdat, visdat1) %/% days(1),
    mpar = factor(if_else(mpar > 4, 4, mpar), labels = c("1", "2", "3", "4+")),
    sex = factor(if_else(gender == "Male", 1, 0), labels = c("Female", "Male")),
    caes = factor(
      case_when(
        is.na(delt) ~ NA_real_,
        grepl("caesarean", delt) ~ 1,
        TRUE ~ 0
      ),
      labels = c("Vaginal", "Caesarean")
    ),
    bfed = factor(
      case_when(
        is.na(fecurr) ~ NA_real_,
        fecurr == "Exclusively Breastfed" ~ 0,
        fecurr == "Exclusively formula-fed" ~ 2,
        TRUE ~ 1
      ),
      labels = c("Complete", "Partial", "None")
    ),
    inc = factor(
      case_when(
        is.na(parinc) ~ NA_real_,
        parinc == ">$180,000" ~ 1,
        parinc == "$87,001 to $180,000" ~ 2,
        TRUE ~ 3
      ),
      labels = c(">$180,000", "$87,001-$180,000", "≤$87,000")
    ),
    sibnum = if_else(
      sibnum == "Not Applicable",
      "0",
      if_else(sibnum == "Unknown", NA_character_, sibnum)
    ),
    fborn = factor(
      if_else(sibnum == "0" | is.na(sibnum), 1, 0),
      labels = c("No", "Yes")
    ),
    fha = factor(fha, labels = c("No", "Yes"))
  )

dat_bc <- select_form(dat_all, "blood_collection") |>
  left_join(select(dat_base, record_id, subjid), join_by(record_id)) |>
  select(-record_id, -starts_with("lbsite___"))

dat_igg <- select_form(dat_all, "igg") |>
  left_join(dat_bc, join_by(subjid, visage)) |>
  left_join(dat_base, join_by(subjid)) |>
  mutate(stage = if_else(substr(subjid, 1, 2) == "01", "1", "2")) |>
  filter(!(antigen %in% c("PnPs 11A"))) |>
  mutate(antigen = fct_inorder(antigen))

dat_base <- dat_igg |>
  select(all_of(colnames(dat_base)), concentration) |>
  mutate(
    any_obs_concentration = any(!is.na(concentration)),
    all_obs_concentration = all(!is.na(concentration)),
    .by = subjid
  ) |>
  select(-concentration) |>
  distinct()
```

\clearpage

# Background

When summarising the stage 1 and stage 2 IgG concentration data we noted that the GMR for PRN changed direction. The aim here is to explore this further for any possible explanation.

# PRN

There is a discrepancy between stage 1 and stage 2 concentrations for PRN by treatment groups. This may just be due to noise, but the following explores this a bit further.

```{r}
#| label: tbl-prn-comparison
#| tbl-cap: |
#|    IgG concentration for PRN and PT by age, study stage, and treatment group.
var <- sym("concentration")
var2 <- sym("positive")
tab <- dat_igg |>
  filter(antigen %in% c("PRN")) |>
  summarise(
    obs = sum(!is.na({{ var }})),
    gmc = gmc({{ var }}, na.rm = TRUE),
    gsd = gsd({{ var }}, na.rm = TRUE),
    m = median({{ var }}, na.rm = TRUE),
    q1 = quantile({{ var }}, probs = 0.25, na.rm = TRUE),
    q3 = quantile({{ var }}, probs = 0.75, na.rm = TRUE),
    min = min({{ var }}, na.rm = TRUE),
    max = max({{ var }}, na.rm = TRUE),
    pos = mean({{ var2 }}, na.rm = TRUE),
    .by = c(antigen, stage, visage, trt)
  ) |>
  pivot_wider(
    names_from = trt,
    values_from = c(obs, gmc, gsd, m, q1, q3, min, max, pos),
    names_vary = "slowest"
  ) |>
  to_factor() |>
  arrange(antigen, visage, stage) |>
  mutate(
    grp = paste0("Stage - ", stage)
  ) |>
  select(-c(antigen, stage))

gt(tab, groupname_col = "grp", rowname_col = "visage") |>
  fmt_integer(starts_with("obs")) |>
  fmt_number(starts_with(c("gmc", "gsd")), decimals = 2) |>
  fmt(starts_with(c("m", "q1", "q3")), fns = fmt_num) |>
  fmt(starts_with(c("min", "max")), fns = fmt_num) |>
  fmt_percent(starts_with(c("pos")), decimals = 0) |>
  cols_merge(
    columns = c(min_aP, max_aP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(min_wP, max_wP),
    pattern = "{1}–{2}"
  ) |>
  cols_merge(
    columns = c(m_aP, q1_aP, q3_aP),
    pattern = "{1} ({2}–{3})"
  ) |>
  cols_merge(
    columns = c(m_wP, q1_wP, q3_wP),
    pattern = "{1} ({2}–{3})"
  ) |>
  cols_label(
    obs_aP = "n",
    obs_wP = "n",
    gmc_aP = "GMC",
    gmc_wP = "GMC",
    gsd_aP = "GSD",
    gsd_wP = "GSD",
    m_aP = "Q2 (Q1–Q3)",
    m_wP = "Q2 (Q1–Q3)",
    min_aP = "Min–Max",
    min_wP = "Min–Max",
    pos_aP = md("S+"),
    pos_wP = md("S+")
  ) |>
  tab_spanner(
    label = md("**aP**"),
    columns = ends_with("_aP")
  ) |>
  tab_spanner(
    label = md("**wP**"),
    columns = ends_with("_wP")
  ) |>
  sub_missing(
    columns = everything(),
    rows = everything(),
    missing_text = "---"
  ) |>
  tab_footnote(
    footnote = "Geometric mean concentration",
    locations = cells_column_labels(columns = starts_with("gmc"))
  ) |>
  tab_footnote(
    footnote = "Geometric standard deviation",
    locations = cells_column_labels(columns = starts_with("gsd"))
  ) |>
  tab_footnote(
    footnote = "Q1 - 25th sample percentile, Q3 - 75th sample percentile",
    locations = cells_column_labels(columns = starts_with("q1"))
  ) |>
  tab_footnote(
    footnote = "Seroprotection at specified levels",
    locations = cells_column_labels(columns = starts_with("pos"))
  ) |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_column_labels(columns = everything())
  ) |>
  tab_options(
    latex.use_longtable = TRUE,
    table.font.size = px(8)
  )
```

```{r}
#| label: fig-igg-prn
#| fig-cap: |
#|    IgG concentrations for PRN by age, treatment, and stage.
#| fig-width: 6
#| fig-height: 4
cols <- palette.colors(palette = "Okabe-Ito")[1:2]
dat_igg |>
  filter(antigen == "PRN") |>
  mutate(
    side = if_else(trt == "aP", "left", "right"),
    stage = paste("Stage", stage)
  ) |>
  ggplot(aes(interaction(trt, visage, sep = "_"), concentration)) +
  facet_wrap(~stage, ncol = 3) +
  geom_boxplot(aes(fill = trt)) +
  geom_hline(aes(yintercept = ref), linetype = 2) +
  scale_y_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  scale_x_discrete(
    "Age and treatment group",
    guide = guide_axis_nested(key = key_range_auto(sep = "_"))
  ) +
  scale_fill_manual(values = c("aP" = cols[1], "wP" = cols[2])) +
  guides(fill = "none", colour = "none") +
  theme(axis.title = element_text(size = rel(0.9)))
```

```{r}
#| label: fig-igg-prn-ecdf
#| fig-cap: |
#|    IgG concentrations for PRN by age, treatment, and stage.
#| fig-width: 6
#| fig-height: 4
cols <- palette.colors(palette = "Okabe-Ito")[1:2]
p <- dat_igg |>
  filter(antigen == "PRN") |>
  mutate(stage = paste("Stage", stage)) |>
  ggplot(aes(concentration)) +
  facet_grid(stage ~ visage) +
  # stat_ecdf(aes(colour = trt)) +
  geom_step(aes(y = 1 - after_stat(y), colour = trt), stat = "ecdf") +
  scale_x_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  labs(y = "complementary eCDF") +
  scale_colour_manual("Treatment", values = c("aP" = cols[1], "wP" = cols[2])) +
  theme(
    axis.title = element_text(size = rel(0.9)),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
p
```

\clearpage

# Time-varying GMR

Here GMR is considered as a function of enrolment time.

## PRN

I expect the difference between stages is by chance and attributable to noise. To explore this, consider the GMR of wP vs aP as a time-varying function of randomisation date for all antigens. PRN does appear to stand out with a slightly stronger signal, but there are also varying trends in other antigens.

```{r}
#| label: data-calendar-time
dat_sub <- dat_igg |>
  filter(
    antigen %in% c("DT", "TT", "PT", "FHA", "PRN", "Hib-PRP"),
    visage %in% c("18-month", "19-month")
  ) |>
  mutate(caltime = interval(min(randdate), randdate) %/% days(1)) |>
  mutate(
    caltime = caltime / max(caltime),
    trtord = ordered(trt),
    sexord = ordered(sex)
  ) |>
  mutate(caltime_st2 = min(caltime[stage == "2"]))

st1_first <- dat_sub |>
  pull(randdate) |>
  min()
st2_first <- dat_sub |>
  filter(stage == "2") |>
  pull(randdate) |>
  min()
```

::: {.landscape}

```{r}
#| label: fig-explore-time-varying-gmr
#| fig-cap: |
#|   IgG concentrations by randomisation date, antigen, and treatment group, with smooth.
#| fig-width: 8
#| fig-height: 5
cols <- palette.colors(palette = "Okabe-Ito")[1:2]
ggplot(
  dat_sub,
  aes(randdate, concentration, colour = trt)
) +
  ggh4x::facet_grid2(visage ~ antigen, scales = "free_y", independent = "y") +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE) +
  geom_vline(xintercept = st2_first, linetype = 2, colour = "grey50") +
  scale_y_continuous(
    "IgG Concentration",
    trans = "log10",
    breaks = c(1, 10, 100, 1000)
  ) +
  scale_colour_manual("Treatment", values = cols) +
  labs(x = "Randomisation date")
```

:::

```{r}
#| label: fit-time-varying-gmr
frm <- log_concentration ~
  s(caltime) +
    trtord +
    s(caltime, by = trtord)

dat_sub_fits <- dat_sub |>
  nest(dat = -c(antigen, visage)) |>
  mutate(
    fit = map(
      dat,
      ~ brm(
        frm,
        data = .,
        backend = "cmdstanr",
        control = list(adapt_delta = 0.95),
        refresh = 0,
        silent = 2
      )
    ),
    dat = map2(
      dat,
      fit,
      ~ .x |>
        expand(nesting(caltime, randdate), trtord) |>
        add_epred_rvars(.y) |>
        pivot_wider(names_from = trtord, values_from = .epred) |>
        mutate(gmr = 10^(wP - aP))
    )
  )
```

::: {.landscape}

```{r}
#| label: fig-model-gmr-time-varying
#| fig-cap: |
#|   Varying estimates of GMR (wP vs aP) by randomisation date (smoothed), antigen, and visit.
#|   Red line indicates average GMR.
#|   Vertical dotted grey line indicates first stage 2 participant.
#| fig-width: 8
#| fig-height: 5
dat_sub_avg <- dat_sub_fits |>
  unnest(dat) |>
  summarise(
    gmr = exp(rvar_mean(log(gmr))),
    .by = c(antigen, visage)
  )

dat_sub_fits |>
  unnest(dat) |>
  ggplot(aes(randdate, ydist = gmr)) +
  ggh4x::facet_grid2(visage ~ antigen) +
  stat_lineribbon() +
  geom_hline(data = dat_sub_avg, aes(yintercept = mean(gmr)), colour = "red") +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey50") +
  geom_vline(xintercept = st2_first, linetype = 2, colour = "grey50") +
  annotate(
    x = date("2019-01-01"),
    y = 0.25,
    label = "Stage 1",
    size = 2,
    geom = "text",
    family = "TeX Gyre Pagella"
  ) +
  annotate(
    x = date("2021-06-01"),
    y = 0.25,
    label = "Stage 2",
    size = 2,
    geom = "text",
    family = "TeX Gyre Pagella"
  ) +
  scale_y_continuous(
    "Geometric mean ratio",
    trans = "log10",
    breaks = c(0.2, 0.5, 1, 2, 5, 10)
  ) +
  labs(x = "Randomisation date") +
  scale_fill_brewer()
```

:::

## Sex

For examples sake, consider the time-varying GMR of sex (Male vs Female) instead of treatment (wP vs aP). We see similar time-varying trends with calendar time in some antigens (e.g. DT). Presumably also just due to chance. Although the PRN reversal is slightly stronger in signal, I wouldn't consider the reversal of PRN to be surprising given the inherent noise in IgG concentrations.

::: {.landscape}

```{r}
#| label: fig-explore-time-varying-gmr-sex
#| fig-cap: |
#|   IgG concentrations by randomisation date, antigen, and treatment group, with smooth.
#| fig-width: 8
#| fig-height: 5
cols <- palette.colors(palette = "Okabe-Ito")[1:2]
ggplot(
  dat_sub,
  aes(randdate, concentration, colour = sex)
) +
  ggh4x::facet_grid2(visage ~ antigen, scales = "free_y", independent = "y") +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE) +
  geom_vline(xintercept = st2_first, linetype = 2, colour = "grey50") +
  scale_y_continuous(
    "IgG Concentration",
    trans = "log10",
    breaks = c(1, 10, 100, 1000)
  ) +
  scale_colour_manual("Treatment", values = cols) +
  labs(x = "Randomisation date")
```

:::

```{r}
#| label: fit-time-varying-gmr-sex
frm <- log_concentration ~
  s(caltime) +
    sexord +
    s(caltime, by = sexord)

dat_sub_fits <- dat_sub |>
  nest(dat = -c(antigen, visage)) |>
  mutate(
    fit = map(
      dat,
      ~ brm(
        frm,
        data = .,
        backend = "cmdstanr",
        control = list(adapt_delta = 0.95),
        refresh = 0,
        silent = 2
      )
    ),
    dat = map2(
      dat,
      fit,
      ~ .x |>
        expand(nesting(caltime, randdate), sexord) |>
        add_epred_rvars(.y) |>
        pivot_wider(names_from = sexord, values_from = .epred) |>
        mutate(gmr = 10^(Male - Female))
    )
  )
```

::: {.landscape}

```{r}
#| label: fig-model-gmr-time-varying-sex
#| fig-cap: |
#|   Varying estimates of GMR (sex) by randomisation date (smoothed), antigen, and visit.
#|   Red line indicates average GMR.
#|   Vertical dotted grey line indicates first stage 2 participant.
#| fig-width: 8
#| fig-height: 5
dat_sub_avg <- dat_sub_fits |>
  unnest(dat) |>
  summarise(
    gmr = exp(rvar_mean(log(gmr))),
    .by = c(antigen, visage)
  )

dat_sub_fits |>
  unnest(dat) |>
  ggplot(aes(randdate, ydist = gmr)) +
  ggh4x::facet_grid2(visage ~ antigen) +
  stat_lineribbon() +
  geom_hline(data = dat_sub_avg, aes(yintercept = mean(gmr)), colour = "red") +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey50") +
  geom_vline(xintercept = st2_first, linetype = 2, colour = "grey50") +
  annotate(
    x = date("2019-01-01"),
    y = 0.25,
    label = "Stage 1",
    size = 2,
    geom = "text",
    family = "TeX Gyre Pagella"
  ) +
  annotate(
    x = date("2021-06-01"),
    y = 0.25,
    label = "Stage 2",
    size = 2,
    geom = "text",
    family = "TeX Gyre Pagella"
  ) +
  scale_y_continuous(
    "Geometric mean ratio",
    trans = "log10",
    breaks = c(0.2, 0.5, 1, 2, 5, 10)
  ) +
  labs(x = "Randomisation date") +
  scale_fill_brewer()
```

:::

```{r}
#| include: false
#| eval: false

# dat_sub_fits |>
#   unnest(dat) |>
#   ggplot(aes(caltime, ydist = 10^aP)) +
#   ggh4x::facet_grid2(visage ~ antigen, scales = "free_y") +
#   stat_lineribbon() +
#   stat_lineribbon(aes(ydist = 10^wP), fill = "red") +
#   scale_y_continuous("GMR", trans = "log10", breaks = c(0.1, 0.5, 1, 5, 10)) +
#   labs(x = "Calendar time (scaled by min and max)") +
#   scale_fill_brewer()

cols <- palette.colors(palette = "Okabe-Ito")[1:2]
dat_sub <- dat_igg |>
  filter(antigen == "PRN", visage %in% c("18-month", "19-month")) |>
  arrange(visage, subjid)

x <- dat_sub |>
  filter(visage == "18-month", trt == "aP", stage == "2") |>
  pull(log_concentration)
y <- dat_sub |>
  filter(visage == "18-month", trt == "wP", stage == "2") |>
  pull(log_concentration)
wilcox.test(x, y)

sdat_0 <- make_stan_data(dat_sub, form = ~trt)
sdat_1 <- make_stan_data(dat_sub, form = ~ trt + stage)
sdat_2 <- make_stan_data(dat_sub, form = ~ trt * stage)
sdat_2$x <- sdat_2$X[, 2] + 1

mfit_0 <- multinorm$sample(
  data = sdat_0,
  show_messages = FALSE,
  show_exceptions = FALSE
)
mfit_1 <- multinorm$sample(
  data = sdat_1,
  show_messages = FALSE,
  show_exceptions = FALSE
)
mfit_2 <- multinorm$sample(
  data = sdat_2,
  show_messages = FALSE,
  show_exceptions = FALSE
)
mfit_3 <- multinorm_het$sample(
  data = sdat_2,
  show_messages = FALSE,
  show_exceptions = FALSE
)
mfit_4 <- multit$sample(
  data = sdat_0,
  show_messages = FALSE,
  show_exceptions = FALSE
)

sdrw_0 <- as_draws_rvars(mfit_0)
sdrw_1 <- as_draws_rvars(mfit_1)
sdrw_2 <- as_draws_rvars(mfit_2)
sdrw_3 <- as_draws_rvars(mfit_3)
sdrw_4 <- as_draws_rvars(mfit_4)

colnames(sdrw_1$beta) <- colnames(sdat_1$X)
rownames(sdrw_1$beta) <- colnames(sdat_1$Y)
colnames(sdrw_2$beta) <- colnames(sdat_2$X)
rownames(sdrw_2$beta) <- colnames(sdat_2$Y)

dat_pred <- dat_sub |>
  arrange(visage, subjid) |>
  mutate(stage = paste("Stage", stage)) |>
  mutate(
    pred_0 = c(10^sdrw_0$Ypred),
    pred_1 = c(10^sdrw_1$Ypred),
    pred_2 = c(10^sdrw_2$Ypred),
    pred_3 = c(10^sdrw_3$Ypred),
    pred_4 = c(10^sdrw_4$Ypred)
  )

dat_pred |>
  unnest_rvars() |>
  filter(.draw < 251) |>
  ggplot(aes(pred_0)) +
  ggh4x::facet_grid2(stage + trt ~ visage, scales = "free_x") +
  stat_ecdf(aes(group = .draw, y = 1 - after_stat(y)), alpha = 0.1) +
  stat_ecdf(aes(concentration, y = 1 - after_stat(y)), colour = "red") +
  scale_x_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  labs(y = "complementary eCDF") +
  # coord_cartesian(xlim = c(0.01, 1e3)) +
  theme(
    axis.title = element_text(size = rel(0.9)),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

dat_pred |>
  unnest_rvars() |>
  filter(.draw < 251) |>
  ggplot(aes(pred_1)) +
  ggh4x::facet_grid2(stage + trt ~ visage, scales = "free_x") +
  stat_ecdf(aes(group = .draw, y = 1 - after_stat(y)), alpha = 0.1) +
  stat_ecdf(aes(concentration, y = 1 - after_stat(y)), colour = "red") +
  scale_x_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  labs(y = "complementary eCDF") +
  coord_cartesian(xlim = c(0.01, 1e3)) +
  theme(
    axis.title = element_text(size = rel(0.9)),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

dat_pred |>
  unnest_rvars() |>
  filter(.draw < 251) |>
  ggplot(aes(pred_2)) +
  ggh4x::facet_grid2(stage + trt ~ visage, scales = "free_x") +
  stat_ecdf(aes(group = .draw, y = 1 - after_stat(y)), alpha = 0.1) +
  stat_ecdf(aes(concentration, y = 1 - after_stat(y)), colour = "red") +
  scale_x_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  labs(y = "complementary eCDF") +
  coord_cartesian(xlim = c(0.01, 1e3)) +
  theme(
    axis.title = element_text(size = rel(0.9)),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

dat_pred |>
  unnest_rvars() |>
  filter(.draw < 251) |>
  ggplot(aes(pred_3)) +
  ggh4x::facet_grid2(stage + trt ~ visage, scales = "free_x") +
  stat_ecdf(aes(group = .draw, y = 1 - after_stat(y)), alpha = 0.1) +
  stat_ecdf(aes(concentration, y = 1 - after_stat(y)), colour = "red") +
  scale_x_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  labs(y = "complementary eCDF") +
  coord_cartesian(xlim = c(0.01, 1e3)) +
  theme(
    axis.title = element_text(size = rel(0.9)),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

dat_pred |>
  unnest_rvars() |>
  filter(.draw < 251) |>
  ggplot(aes(pred_4)) +
  ggh4x::facet_grid2(stage + trt ~ visage, scales = "free_x") +
  stat_ecdf(aes(group = .draw, y = 1 - after_stat(y)), alpha = 0.1) +
  stat_ecdf(aes(concentration, y = 1 - after_stat(y)), colour = "red") +
  scale_x_continuous(
    "IgG concentration, log scale",
    breaks = c(0.01, 0.1, 1, 10, 100, 1000),
    labels = comma,
    trans = "log10"
  ) +
  labs(y = "complementary eCDF") +
  coord_cartesian(xlim = c(0.01, 1e3)) +
  theme(
    axis.title = element_text(size = rel(0.9)),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

ids <- which(!is.na(dat_sub$concentration))
quantile_fun <- \(y) quantile(y, 0.25)
gmc <- \(x) exp(mean(log(x)))
gsd <- \(x) exp(sd(log(x)))

y <- dat_sub$concentration[ids]
yrep <- as_draws_matrix(c(10^sdrw_0$Ypred)[ids])
grp <- dat_sub$trt[ids]

ppc_stat_grouped(
  y,
  yrep,
  stat = "min",
  bins = 100,
  group = grp
)

ppc_stat_grouped(
  y,
  yrep,
  stat = "quantile_fun",
  bins = 100,
  group = grp
)

ppc_stat_grouped(
  y,
  yrep,
  stat = "gsd",
  bins = 100,
  group = grp
)

ppc_stat_2d(y, yrep, stat = c("median", "mad"))
ppc_stat_2d(y, yrep, stat = c("mean", "sd"))
ppc_stat_2d(y, yrep, stat = c("min", "max")) +
  scale_y_log10()
```
