---
title: Baseline Characteristics
description: Summary of enrolment and baseline characteristics.
code-fold: true
---

{{< include _setup.qmd >}}

## Enrolment

Trial enrolment is summarised in @fig-enrolment-overall.

```{r}
#| label: fig-enrolment-overall
#| code-summary: Enrolment figure
#| fig-cap: Cumulative enrolment.
#| fig-height: 3
#| fig-width: 6
dat_enrol <- dat_base |>
  select(record_id, randdattim) |>
  mutate(randdate = as_date(randdattim)) |>
  arrange(randdate) |>
  summarise(n = n(), .by = randdate) |>
  mutate(cn = cumsum(n))

ggplot(
  dat_enrol,
  aes(randdate, cn)
) +
  geom_step() +
  scale_x_date("Calendar time", breaks = "3 months", date_labels = "%d %b %y") +
  scale_y_continuous("Number enrolled", breaks = seq(0, 1000, 100)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Vaccination Administration

For one participant the pharmacy reportedly dispensed the alternative study vaccine to the participants randomly allocated type. The IMP is monitored for all participants, so we are confident this was the only time this occurred. This error was documented in the protocol deviations log [subject 02-333, record id 4629-184]. 

To strictly adhere to ITT, this participant would be analysed according to their randomly allocated group to avoid any potential accidental bias. However, due to the strict IMP monitoring, the one-off occurrence, the fact that the intervention is a single point-treatment, and the objective nature of determining whether the correct vaccine was given at this single point treatment, I'm inclined to consider a modified-ITT where participants are grouped according to what was actually recieved at the first visit (which in this case is equivalent to what was randomly allocated for all but this one participant).

That being said, givein one participant is unlikely to make much difference, what follows currently sticks with strict ITT, grouping participants as assigned rather than as received. If there's general agreement with switching to the modified-ITT, then these will be updated.

## Baseline Characteristics

Baseline characteristics are summarised in @tbl-baseline-characteristics.

```{r}
#| label: tbl-baseline-characteristics
#| code-summary: Baseline table
#| tbl-cap: Baseline characteristics of trial participants.
dat_base |>
  tbl_summary(
    include = c(gender, parinc, inmpertvp, minfvp, ipab, delt, gestadelw, wgt, len, inhc),
    by = trt,
    missing_text = "(Missing)"
  ) |>
  add_n() |>
  add_overall(last = TRUE) |>
  as_gt() |>
  tab_options(table.font.size = "10px") |>
  tab_style(style = cell_text(align = "left"), locations = cells_footnotes())
```
