---
title: Trial Outcomes
description: Document derivation, summaries, and analyses related to primary and secondary clinical outcomes.
---

{{< include _setup.qmd >}}

# Trial Outcome Definitions

## Primary outcome (1)

**IgE mediated food allergy**

IgE mediated food allergy with evidence of food sensitisation on skin prick test (SPT) by 12 months old and confirmed (where necessary) by medically supervised oral food challenge(s) (OFC). For the purpose of this trial, if either of the following conditions are met, the study participant will be considered to have reached the primary endpoint:

1. Unequivocal evidence of IgE mediated food allergy, defined as:
     i. A positive OFC with evidence of sensitisation on SPT to the food of interest *OR*
     ii. History of clinician-confirmed food anaphylaxis, with symptoms affecting at least two of the following: skin, gastrointestinal tract, respiratory tract, cardiovascular system; *AND* evidence of sensitisation on SPT to the food of interest.
2. Highly probable IgE-mediated food allergy defined as a history of food allergic reaction (consistent with PRACTALL criteria), with evidence of sensitisation on SPT to the food of interest.

We will report on the number and proportion who satisfy this composite definition, and the number and proportion satisfying each of the three sub-components as per a hierarchy in order of 1.i, 1.ii, 2.

## Secondary outcomes (2)

a. A history of parent-reported clinician-diagnosed new onset eczema by **6 months** *AND* a positive SPT to any allergen by approximately 12 months old.
b. A history of parent-reported clinician-diagnosed new onset eczema by **12 months** *AND* a positive SPT to any allergen by approximately 12 months old.
c. A SPT **>1mm** greater than the negative control to at least one allergen by approximately 12 months of age
d. A SPT **≥3mm** greater than the negative control to at least one allergen by approximately 12 months of age

## Details

Some of the above is open to interpretation, but generally:

- we consider any SPT before 18-months of age to be valid (i.e. approximately 12 months of age)
- sensitisation on SPT requires a reaction greater than 1mm compared to negative control solution (i.e. >1mm)
- new-onset eczema means we exclude any infants with pre-existing eczema at 6-weeks of age (enrolment)
- by at least one allergen, we mean any allergen tested, either from the base set or any additionally tested allergens

```{r}
#| label: data-derive
# Randomisation
dat_ran <- select_form(dat_raw, "randomisation") |>
  mutate(randdat = date(randdattim)) |>
  left_join(select_form(dat_raw, "allocations"), join_by(rand))

# Add study termination info to baseline data
dat_st <- select_form(dat_raw, "study_termination")
dat_dem <- select_form(dat_raw, "demographics") |>
  select(record_id, visdat1, birthdat, gender) |>
  left_join(dat_st, join_by(record_id)) |>
  mutate(discage = interval(birthdat, discdat) %/% months(1)) |>
  relocate(discage, .after = discdat)

dat_mh <- select_form(dat_raw, "medical_history")

# Main baseline information
dat_base <- dat_ran |>
  left_join(dat_dem, join_by(record_id))

# Outcomes
dat_out <- select_form(dat_raw, "outcome_report")
dat_fhq <- select_form(dat_raw, "food_and_household_questionnaire")
dat_spt <- select_form(dat_raw, "skin_prick_test")
dat_fch <- select_form(dat_raw, "food_challenge")
```

# Skin Prick Test

I start with the skin prick tests (SPTs) [outcomes 2c. and 2d.], as all other outcomes rely on positive SPTs in addition to further conditions. For the food allergy and eczema outcomes, sensitisation to food allergens or a positive SPT for a specific food was defined as a SPT wheal size >1 mm than that produced by a negative control solution.

```{r}
#| label: data-spt
dat_spt <- get_skin_prick_long(dat_raw)

dat_spt_pos <- dat_spt |>
  summarise(
    any_spt_pos_0mm = if_else(all(!priyn), NA, any(spt_0mm == 1, na.rm = TRUE)),
    spt_pos_0mm = if_else(
      all(!priyn),
      NA_character_,
      paste(
        spt_tested[spt_0mm == 1 & !is.na(spt_0mm)],
        collapse = ", "
      )
    ),
    any_spt_pos_1mm = if_else(all(!priyn), NA, any(spt_1mm == 1, na.rm = TRUE)),
    spt_pos_1mm = if_else(
      all(!priyn),
      NA_character_,
      paste(
        spt_tested[spt_1mm == 1 & !is.na(spt_1mm)],
        collapse = ", "
      )
    ),
    any_spt_pos_3mm = if_else(all(!priyn), NA, any(spt_3mm == 1, na.rm = TRUE)),
    spt_pos_3mm = if_else(
      all(!priyn),
      NA_character_,
      paste(
        spt_tested[spt_3mm == 1 & !is.na(spt_3mm)],
        collapse = ", "
      )
    ),
    .by = c(record_id, pridat, spt_occasion, spt_age)
  ) |>
  arrange(str_rank(record_id, numeric = TRUE), pridat)
```

## Derivation

For reference, the following REDCap form fields correspond to these foods:

- `prires4` - whole egg
- `prires5` - cashew
- `prires6` - cow's milk
- `prires7` - peanut
- `prires8` - sesame

The protocol defined a panel of allergens to be tested: cow's milk, hen egg, peanut, cashew, house dust mite, cat fur, rye grass. In addition to these, further allergens could also be tested at parental request or at discretion of the site PI. When the outcomes say "at least one allergen" we mean any allergen, including any additional allergens tested beyond the panel defined in the protocol.

In the REDCap form, there is a field `prireact[1-9]` which indicates if a reaction occurred. However, this only indicates that the wheal diameter was greater than zero, not that it was >1mm more than negative control. Therefore, positive SPTs need to be determined. For example, in the following some participants have `prireact1` "Yes", because the reaction was >0mm, but the reaction may not have been >1mm more than the negative control (`pripos1`).

There are two types of skin prick tests: those conducted as part of a scheduled visit as soon as possible after 12-months of age and no later than 18-months of age, and unscheduled skin prick tests which may be performed at any time. When we say "by approximately 12 months of age", my understanding is we mean "no more than 19-months of age" given the window allowed for performing SPTs, but need to confirm.

The following participants had *no* SPT records. In some cases this was likely due to early termination, however, some participants completed the protocol.

```{r}
#| label: tbl-infants-no-spt-record
dat_spt |>
  filter(!priyn) |>
  filter(row_number() == 1, .by = c(record_id, spt_occasion)) |>
  left_join(dat_st, join_by(record_id)) |>
  select(
    record_id,
    birthdat,
    visdat1,
    discdat,
    streas,
    stetrreas,
    spt_occasion
  ) |>
  mutate(discage = interval(birthdat, discdat) %/% months(1)) |>
  gt() |>
  tab_options(table.font.size = "10px")
```

All other participants had at least one scheduled SPT.

```{r}
#| label: tbl-infants-count-spts
#| tbl-cap: |
#|   Number and percentage of infants with scheduled and unscheduled skin prick tests.
spt_nums <- dat_spt |>
  left_join(dat_ran, join_by(record_id)) |>
  filter(row_number() == 1, .by = c(record_id, trt, spt_occasion, spt_num)) |>
  summarise(
    spt_num_sch = sum(spt_occasion == "scheduled" & priyn, na.rm = TRUE),
    spt_num_unsch = sum(spt_occasion == "unscheduled" & priyn, na.rm = TRUE),
    .by = c(record_id, trt)
  )
tbl_summary(
  spt_nums,
  by = trt,
  include = c(spt_num_sch, spt_num_unsch),
  label = c(spt_num_sch ~ "Scheduled SPT", spt_num_unsch ~ "Unscheduled SPTs")
) |>
  add_overall(last = TRUE) |>
  as_gt() |>
  tab_options(table.font.size = "10px") |>
  tab_style(style = cell_text(align = "left"), locations = cells_footnotes())
```

```{r}
#| label: fig-age-at-spt
#| fig-cap: |
#|   Age at skin prick test by type (scheduled or unscheduled).
#| fig-height: 3
dat_spt |>
  left_join(dat_ran, join_by(record_id)) |>
  filter(row_number() == 1, .by = c(record_id, trt, spt_occasion, spt_num)) |>
  filter(!is.na(spt_age)) |>
  ggplot(aes(spt_age)) +
  facet_wrap( ~ spt_occasion, scales = "free_y") +
  geom_bar() +
  scale_x_continuous("Age at SPT (months)", breaks = seq(0, 30, 2)) +
  labs(y = "Skin prick tests")
```

If a participant has no (scheduled) SPT record, then we should treat their SPT outcomes as missing, and therefore as unknown whether they had a positive SPT or not.

## Summary

There was no difference between groups with respect to SPT sensitisation.

```{r}
#| label: tbl-infants-spt-positive
#| tbl-cap: |
#|   Number and percentage of infants with a positive SPT by 18-months of age.
#|   Infants with no SPT records counted as missing.
dat_spt |>
  left_join(dat_ran, join_by(record_id)) |>
  summarise(
    across(
      spt_0mm:spt_3mm,
      ~ if_else(
        all(!priyn),
        NA,
        any(.x == 1 & spt_age < 19, na.rm = TRUE)
      )
    ),
    .by = c(record_id, trt)
  ) |>
  tbl_summary(
    by = trt,
    include = c(spt_0mm, spt_1mm, spt_3mm),
    label = c(
      spt_0mm ~ "Any SPT >0mm",
      spt_1mm ~ "Any SPT >1mm vs. neg control",
      spt_3mm ~ "Any SPT ≥3mm vs neg. control"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**") |>
  as_gt() |>
  tab_options(table.font.size = "10px") |>
  tab_style(style = cell_text(align = "left"), locations = cells_footnotes())
```

# Eczema

```{r}
#| label: data-eczema
dat_ecz <- get_eczema_data(dat_raw)

dat_ecz |>
  filter(!is.na(outageval)) |>
  select(record_id, birthdat, outdiagdat, outageval, outageunit) |>
  mutate(
    outageval = as.numeric(outageval),
    outagedat = case_when(
      outageunit == "Weeks" ~ birthdat + weeks(round(outageval)),
      outageunit == "Months" ~ birthdat + months(round(outageval)),
      outageunit == "Years" ~ birthdat + years(round(outageval))
    ),
    outageval2 = interval(birthdat, outagedat) %/% months(1),
    outagediag = interval(birthdat, outdiagdat) %/% months(1),
    diff = outdiagdat - outagedat
  ) |>
    arrange(desc(diff)) |>
    print(n = Inf)
```

## Derivation

There are multiple sources of information for new eczema diagnoses: the outcome report, the food and household questionnaire (FHQ), and for stage 1 the adverse event of special interest form (AESI). The primary source of truth for clinician-diagnosed eczema should be the outcome form, but we would like to check the sources for consistency.

The protocol definitions specify "new onset" eczema, which means we do not wish to count infants who already had an eczema diagnosis at enrolment (6-weeks). Again, this may be on the FHQ or could also be reported in the medical history (MH) form. It may also be reported on the outcome report in some cases.

It's also not entirely clear as to what timing should be used. The outcome requires "clinicain-diagnosed", so I presume we should only consider eczema present from the data of clinician diagnosis, however, the "age at which allergy was first present" may be much earlier than the diagnosis date. Need clarification on this point.

On the FHQ, the response to `feecz` may be either `Yes` or `N/A (child has already been diagnosed with eczema before the last visit)`. In both cases, we should check that there is also an eczema record on the outcome report. If not, we should query.

Check for differences between FHQ and outcome report. For those with eczema diagnosis on FHQ:

```{r}
# FHA eczema but no outcome report
filter(dat_ecz, fhq_ecz == "Yes" & fhq_ecz_date > visdat1 & ecz == 0) |>
  select(
    record_id,
    birthdat,
    visdat1,
    fhq_ecz,
    fhq_ecz_date,
    fhq_ecz_age_months
  ) |>
  gt() |>
  tab_options(table.font.size = "10px")
```

For this with "previous eczema" on FHQ:

```{r}
# FHA "prior eczema" but no outcome report
filter(dat_ecz, fhq_ecz == "Previous diagnosis" & ecz == 0 & fhq_vis_age != "6-week") |>
  select(
    record_id,
    birthdat,
    visdat1,
    fhq_ecz,
    fhq_ecz_date,
    fhq_ecz_age_months
  ) |>
  gt() |>
  tab_options(table.font.size = "10px")
```

I identified 16 infants with past diagnosis of eczema at baseline: 6 due to medical history, 11 due to food and household questionnaire at 6 weeks of age, and 8 due to outcome report with diagnosis prior to enrolment. However, two appear to be a coding errors (4629-198 and 5785-111) given the reported dates are prior to date of birth.

```{r}
filter(
  dat_ecz,
  mh_ecz_preexisting | fhq_ecz_preexisting | out_ecz_preexisting
) |>
  select(
    record_id,
    birthdat,
    visdat1,
    mh_ecz_age_weeks,
    mh_ecz_preexisting,
    fhq_ecz_preexisting,
    fhq_ecz_date,
    fhq_ecz_age_weeks,
    out_ecz_preexisting,
    outdiagdat,
    ecz_age_weeks
  )
```

For the protocol defined outcomes [2a. and 2b.], the eczema diagnosis needs to be accompanied by an SPT to at least one allergen by approximately 12 months of age.

## Summary

```{r}
#| label: tbl-eczema-trt
#| tbl-cap: |
#|   Number and percentage of infants with an clinician-diagnosed eczema.
#|   Those with "Missing" were because the date of clinician diagnosis was missing.
dat_ecz |>
  left_join(select(dat_ran, record_id, trt)) |>
  filter(ecz_preexisting) |>
  summarise(n = sum(ecz_preexisting), .by = trt)
dat_ecz |>
  left_join(select(dat_ran, record_id, trt)) |>
  filter(!ecz_preexisting) |>
  mutate(
    ecz = ecz & !ecz_preexisting,
    ecz_6m = ecz & !ecz_preexisting & ecz_age_months <= 6,
    ecz_12m = ecz & !ecz_preexisting & ecz_age_months <= 12,
    ecz_18m = ecz & !ecz_preexisting & ecz_age_months <= 18
  ) |>
    tbl_summary(
      by = trt,
      include = c(ecz, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
    add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**") |>
  as_gt() |>
  tab_options(table.font.size = "10px") |>
  tab_style(style = cell_text(align = "left"), locations = cells_footnotes())
```

# Food allergy

The primary source of information is the outcome form, in particular, records where `outalltp` is one of *Anaphylaxis*, *Highly probably IgE mediated food allergy*, or *Oral food challenge*. There is also a form `food_challenge` for the oral food challenges themselves.

```{r}
fa_recs <- dat_out |>
  filter(!is.na(outalltp), outalltp != "Eczema")
fa_recs |> count(outalltp)
```

Generally, there is one record per allergen and reaction type. For example,

```{r}
fa_recs |>
  filter(record_id == "4629-92") |>
  select(record_id, outalltp, outrepdat, outdiagdat, outfrasrcespec , outfrafood)
```

How many individuals had some kind of food allergy outcome recorded, and how many records did they?

```{r}
fa_recs |>
  count(record_id, name = "records") |>
  count(records, name = "infants")
```
As mentioned above, to meet the primary outcome we require a positive skin prick test to the relevant allergen in addition to an outcome report of anaphylaxis, positive oral food challenge, or highly probable IgE mediated food allergy. Rather than checking for *any* positive SPT, need to check specifically for one which is positive to the relevant food.

## Anaphylaxis

There were 6 participants with a record of anaphylaxis. One participant did not have a positive skin prick test to the reported food (5785-283), however, this appears to have been a data entry issue. Apparently there was a positive SPT to sesame, it's just not recorded in REDCap, awaiting an update to this record.

```{r}
ana_recs <- dat_out |>
  filter(outalltp == "Anaphylaxis") |>
  select(
    record_id,
    outalltp,
    outsev,
    outdiagdat,
    outfrasource,
    outfrafood,
    outfraexptp
  ) |>
  arrange(record_id) |>
  summarise(
    foods = paste(outfrafood, collapse = ", "),
    .by = record_id
  )
```

## Highly Probably IgE Mediated Food Allergy

There were 46 individuals with a record of highly probable IgE mediated food allergy.
From manual review, a few were missing positive SPT:

- 5785-239 (eggplant)
- 5785-243 (cows milk, rice)
- 5785-331 (egg)
- 75

```{r}
ige_recs <- dat_out |>
  filter(outalltp == "Highly probably IgE mediated food allergy") |>
  select(
    record_id,
    outalltp,
    outsev,
    outdiagdat,
    outfrasource,
    outfrafood,
    outfraexptp
  ) |>
  arrange(record_id) |>
  summarise(
    foods = paste(outfrafood, collapse = ", "),
    .by = record_id
  )
```

## Food Challenge

From the outcome report, there were 19 infants with a positive oral food challenge. 
Again, some did not have a positive SPT:

- 4630-45 (peanut)
- 5785-39 (sesame)

```{r}
ofc_recs <- dat_out |>
  filter(outalltp == "Oral food challenge") |>
  select(
    record_id,
    outalltp,
    outsev,
    outdiagdat,
    outfrasource,
    outfrafood,
    outfraexptp
  ) |>
  arrange(record_id) |>
  summarise(
    foods = paste(outfrafood, collapse = ", "),
    .by = record_id
  )
```

For reference, in the oral food challenge records:

- `ofcfoodtp1` - peanut
- `ofcfoodtp2` - egg
- `ofcfoodtp3` - milk
- `ofcfoodtp4` - cashew


Some infants may have multiple food challenges, others may have only one.

```{r}
dat_fch |>
  count(record_id, name = "records") |>
  count(records, name = "infants")
```

For example, the following infant had three separate food challenges, each for a different allergen: peanut, milk, and egg, all being positive. However, in the outcomes form, they only have two oral food challenge records (no record for the egg). To be safe, need to check both sources.

```{r}
fa_recs |>
  filter(record_id == "5785-333") |>
  select(record_id, outalltp, outsev, outrepdat, outfrafood, outfraexptp)
dat_fch |>
  filter(record_id == "5785-333") |>
  select(record_id, ofc_num, ofcdat, ofcwindyn, starts_with("ofcfoodtp"))
```