---
title: Trial Outcomes
description: Document derivation, summaries, and analyses related to primary and secondary clinical outcomes.
code-fold: true
---

{{< include _setup.qmd >}}

# Trial Outcome Definitions

## Primary outcome (1)

**IgE mediated food allergy**

IgE mediated food allergy with evidence of food sensitisation on skin prick test (SPT) by 12 months old and confirmed (where necessary) by medically supervised oral food challenge(s) (OFC). For the purpose of this trial, if either of the following conditions are met, the study participant will be considered to have reached the primary endpoint:

1. Unequivocal evidence of IgE mediated food allergy, defined as:
     i. A positive OFC with evidence of sensitisation on SPT to the food of interest *OR*
     ii. History of clinician-confirmed food anaphylaxis, with symptoms affecting at least two of the following: skin, gastrointestinal tract, respiratory tract, cardiovascular system; *AND* evidence of sensitisation on SPT to the food of interest.
2. Highly probable IgE-mediated food allergy defined as a history of food allergic reaction (consistent with PRACTALL criteria), with evidence of sensitisation on SPT to the food of interest.

We will report on the number and proportion who satisfy this composite definition, and the number and proportion satisfying each of the three sub-components as per a hierarchy in order of 1.i, 1.ii, 2.

## Secondary outcomes (2)

a. A history of parent-reported clinician-diagnosed new onset eczema by **6 months** *AND* a positive SPT to any allergen by approximately 12 months old.
b. A history of parent-reported clinician-diagnosed new onset eczema by **12 months** *AND* a positive SPT to any allergen by approximately 12 months old.
c. A SPT **>1mm** greater than the negative control to at least one allergen by approximately 12 months of age
d. A SPT **≥3mm** greater than the negative control to at least one allergen by approximately 12 months of age

## Details

Some of the above is open to interpretation, but generally:

- we consider any SPT before 18-months of age to be valid (i.e. approximately 12 months of age)
- sensitisation on SPT requires a reaction greater than 1mm compared to negative control solution (i.e. >1mm)
- new-onset eczema means we exclude any infants with pre-existing eczema at 6-weeks of age (enrolment)
- for timing (both food allergy and eczema), we consider the age at which a reaction was first observed, although it's unclear how accurate this is, and may be subject to recall bias. If a clinician diagnosis date is missing ("outdiagdat"), then we consider that it was not clinician-diagnosed.
- by at least one allergen, we mean any allergen tested, either from the base set or any additionally tested allergens

```{r}
#| label: data-derive
#| code-summary: Raw datasets

# Medical history
dat_mh <- select_form(dat_raw, "medical_history")

# Outcomes
dat_out_raw <- select_form(dat_raw, "outcome_report")
dat_fhq_raw <- select_form(dat_raw, "food_and_household_questionnaire")
dat_spt_raw <- select_form(dat_raw, "skin_prick_test")
dat_fch_raw <- select_form(dat_raw, "food_challenge")
```

# Skin Prick Test

I start with the skin prick tests (SPTs) [outcomes 2c. and 2d.], as all other outcomes rely on positive SPTs in addition to further conditions. For the food allergy and eczema outcomes, sensitisation to food allergens or a positive SPT for a specific food was defined as a SPT wheal size >1 mm than that produced by a negative control solution.

```{r}
#| label: data-spt
#| code-summary: Summarise positve SPTs
dat_spt <- get_skin_prick_long(dat_raw)
dat_spt_pos <- dat_spt |>
  summarise(
    any_spt_pos_0mm = if_else(all(!priyn), NA, any(spt_0mm == 1, na.rm = TRUE)),
    spt_pos_0mm = if_else(
      all(!priyn),
      NA_character_,
      paste(
        spt_tested[spt_0mm == 1 & !is.na(spt_0mm)],
        collapse = ", "
      )
    ),
    any_spt_pos_1mm = if_else(all(!priyn), NA, any(spt_1mm == 1, na.rm = TRUE)),
    spt_pos_1mm = if_else(
      all(!priyn),
      NA_character_,
      paste(
        spt_tested[spt_1mm == 1 & !is.na(spt_1mm)],
        collapse = ", "
      )
    ),
    any_spt_pos_3mm = if_else(all(!priyn), NA, any(spt_3mm == 1, na.rm = TRUE)),
    spt_pos_3mm = if_else(
      all(!priyn),
      NA_character_,
      paste(
        spt_tested[spt_3mm == 1 & !is.na(spt_3mm)],
        collapse = ", "
      )
    ),
    .by = c(record_id, pridat, spt_occasion, spt_age)
  ) |>
  arrange(str_rank(record_id, numeric = TRUE), pridat)
dat_spt_pos_any <- dat_spt_pos |>
  summarise(
    any_pos_spt = any(any_spt_pos_1mm & spt_age < 19),
    .by = record_id
  )
```

## Derivation

For reference, the following REDCap form fields correspond to these foods:

- `prires1` - D.pteronyssinus (dust mite)
- `prires2` - cat dander
- `prires3` - perennial ryegrass
- `prires4` - whole egg
- `prires5` - cashew
- `prires6` - cow's milk
- `prires7` - peanut
- `prires8` - sesame

The protocol defined a panel of allergens to be tested (those listed above). In addition to these, further allergens could also be tested at parental request or at discretion of the site PI. When the outcomes say "at least one allergen" [i.e. for eczema] we mean any allergen, including any additional allergens tested beyond the panel defined in the protocol.

In the REDCap form, there is a field `prireact[1-9]` which indicated if a reaction occurred. However, this appears to only indicate that the wheal diameter was greater than zero, not that it was >1mm more than negative control. Therefore, positive SPTs need to be determined. For example, in the following some participants have `prireact1` "Yes", because the reaction was >0mm, but the reaction may not have been >1mm more than the negative control (`pripos1`).

There are two types of skin prick tests: those conducted as part of a scheduled visit as soon as possible after 12-months of age and no later than 18-months of age, and unscheduled skin prick tests which may be performed at any time. When we say "by approximately 12 months of age", my understanding is we mean "no more than 19-months of age" given the window allowed for performing SPTs, but need to confirm.

The following participants had *no* SPT records. In some cases this was likely due to early termination, however, some participants completed the protocol.

```{r}
#| label: tbl-infants-no-spt-record
#| code-summary: Listing - no SPT
dat_spt |>
  filter(!priyn) |>
  filter(row_number() == 1, .by = c(record_id, spt_occasion)) |>
  select(
    record_id,
    birthdat,
    visdat1,
    discdat,
    streas,
    stetrreas,
    spt_occasion
  ) |>
  mutate(discage = interval(birthdat, discdat) %/% months(1)) |>
  gt() |>
  tab_options(table.font.size = "10px")
```

All other participants had at least one scheduled SPT.

```{r}
#| label: tbl-infants-count-spts
#| code-summary: Number and proportion with SPTs
#| tbl-cap: |
#|   Number and percentage of infants with scheduled and unscheduled skin prick tests.
spt_nums <- dat_spt |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  filter(row_number() == 1, .by = c(record_id, trt, spt_occasion, spt_num)) |>
  summarise(
    spt_num_sch = sum(spt_occasion == "scheduled" & priyn, na.rm = TRUE),
    spt_num_unsch = sum(spt_occasion == "unscheduled" & priyn, na.rm = TRUE),
    .by = c(record_id, trt)
  )
tbl_summary(
  spt_nums,
  by = trt,
  include = c(spt_num_sch, spt_num_unsch),
  label = c(spt_num_sch ~ "Scheduled SPT", spt_num_unsch ~ "Unscheduled SPTs")
) |>
  add_overall(last = TRUE) |>
  as_gt() |>
  tab_options(table.font.size = "10px") |>
  tab_style(style = cell_text(align = "left"), locations = cells_footnotes())
```

The were a few SPTs which were performed after 18 months of age. If adhereing to the "approximately 12 months of age" as any SPT by 18 months of age, these SPTs should be excluded.

```{r}
#| label: fig-age-at-spt
#| code-summary: Age at SPT
#| fig-cap: |
#|   Age at skin prick test by type (scheduled or unscheduled).
#| fig-height: 3
dat_spt |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  filter(row_number() == 1, .by = c(record_id, trt, spt_occasion, spt_num)) |>
  filter(!is.na(spt_age)) |>
  ggplot(aes(spt_age)) +
  facet_wrap(~spt_occasion, scales = "free_y") +
  geom_bar() +
  scale_x_continuous("Age at SPT (months)", breaks = seq(0, 30, 2)) +
  labs(y = "Skin prick tests")
```

```{r}
#| label: tbl-spt-after-18-months
#| code-summary: SPTs after 18-months of age
#| tbl-cap: Listing of SPTs after 18-months of age.
dat_spt |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  filter(
    row_number() == 1,
    .by = c(record_id, trt, spt_occasion, spt_num, spt_age)
  ) |>
  filter(spt_age > 18) |>
  select(record_id, birthdat, spt_occasion, pridat, spt_num, spt_age) |>
  gt() |>
  tab_options(table.font.size = "10px")
```

If a participant has no (scheduled) SPT record, then we should treat their SPT outcomes as missing, and therefore as unknown whether they had a positive SPT or not.

## Summary

The number of proportion of infants with a positive SPT by 18 months of age is given in @tbl-infants-spt-positive.

```{r}
#| label: tbl-infants-spt-positive
#| tbl-cap: |
#|   Number and percentage of infants with a positive SPT by 18-months of age.
#|   Infants with no SPT records counted as missing.
tab <- dat_spt |>
  left_join(select(dat_base, record_id, trt), join_by(record_id)) |>
  summarise(
    across(
      spt_0mm:spt_3mm,
      ~ if_else(
        all(!priyn),
        NA,
        any(.x == 1 & spt_age < 19, na.rm = TRUE)
      )
    ),
    .by = c(record_id, trt)
  ) |>
  tbl_summary(
    by = trt,
    include = c(spt_0mm, spt_1mm, spt_3mm),
    label = c(
      spt_0mm ~ "Any SPT >0mm",
      spt_1mm ~ "Any SPT >1mm vs. neg control",
      spt_3mm ~ "Any SPT ≥3mm vs neg. control"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**") |>
  as_gt() |>
  tab_options(table.font.size = "10px") |>
  tab_style(style = cell_text(align = "left"), locations = cells_footnotes())
tab
```

---

# Eczema

```{r}
#| label: data-eczema
dat_ecz <- get_eczema_data(dat_raw) |>
  left_join(dat_spt_pos_any, join_by(record_id))
```

## Derivation

There are multiple sources of information for new eczema diagnoses: the outcome report, the food and household questionnaire (FHQ), and for stage 1 the adverse event of special interest form (AESI). The primary source of truth for clinician-diagnosed eczema should be the outcome form, but we would like to check the sources for consistency.

The protocol definitions specify "new onset" eczema, which means we do not wish to count infants who already had an eczema diagnosis at enrolment (6-weeks). Again, this may be on the FHQ or could also be reported in the medical history (MH) form. It may also be reported on the outcome report in some cases.

It's not transparent as to what timing should be used. The outcome requires "clinicain-diagnosed", however, the "age at which allergy was first present" may be much earlier than the diagnosis date. From discussions, it sounds as though age at first reaction was what was of interest. Therefore, this will be used to decide whether eczema was present at 6 or 12 months of age.

On the FHQ, the response to `feecz` may be either `Yes` or `N/A (child has already been diagnosed with eczema before the last visit)`. In both cases, we should check that there is also an eczema record on the outcome report. If not, we should query.

Check for differences between FHQ and outcome report. For those with eczema diagnosis on FHQ:

```{r}
#| label: tbl-fha-eczema-no-outcome
#| code-summary: FHA eczema but no outcome report
#| tbl-cap: |
#|   Eczema diagnosis reported on FHQ, but no outcome report available.
filter(dat_ecz, fhq_ecz == "Yes" & fhq_ecz_date > visdat1 & ecz == 0) |>
  select(
    record_id,
    birthdat,
    visdat1,
    fhq_ecz,
    fhq_ecz_date,
    fhq_ecz_age_months
  ) |>
  gt() |>
  tab_options(table.font.size = "10px")
```

For this with "previous eczema" on FHQ:

```{r}
# FHA "prior eczema" but no outcome report
filter(
  dat_ecz,
  fhq_ecz == "Previous diagnosis" & ecz == 0 & fhq_vis_age != "6-week"
) |>
  select(
    record_id,
    birthdat,
    visdat1,
    fhq_ecz,
    fhq_ecz_date,
    fhq_ecz_age_months
  ) |>
  gt() |>
  tab_options(table.font.size = "10px")
```

I identified 15 infants with past diagnosis of eczema at baseline: 6 due to medical history, 11 due to food and household questionnaire at 6 weeks of age, and 8 due to outcome report with diagnosis prior to enrolment. However, two appear to be a coding errors (4629-198 and 5785-111) given the reported dates are prior to date of birth.

```{r}
filter(
  dat_ecz,
  mh_ecz_preexisting | fhq_ecz_preexisting | out_ecz_preexisting
) |>
  select(
    record_id,
    birthdat,
    visdat1,
    mh_ecz_age_weeks,
    mh_ecz_preexisting,
    fhq_ecz_preexisting,
    fhq_ecz_date,
    fhq_ecz_age_weeks,
    out_ecz_preexisting,
    outdiagdat,
    ecz_age_weeks
  ) |>
  gt() |>
  tab_options(table.font.size = px(9))
```

For some infants who have an outcome report, the date of first reaction is missing (`outageval` and `outageunit`).
In these cases we assume that the diagnosis date was the date of first reaction (`outdiagdat`).

```{r}
dat_out_raw |>
  filter(outalltp == "Eczema", is.na(outageval), nchar(record_id) > 3) |>
  print(n = Inf)
```

For the protocol defined outcomes [2a. and 2b.], the eczema diagnosis needs to be accompanied by an SPT to at least one allergen by approximately 12 months of age.

## Summary

The number and percentage of infants with eczema by a given age is reported in @tbl-eczema-trt.

```{r}
#| label: tbl-eczema-trt
#| tbl-cap: |
#|   Number and percentage of infants with an clinician-diagnosed eczema.
#|   Those with "Missing" under total cases were because the date of reaction was missing.
#|   Additional missing under cases with positive SPT were due to missing SPT status.
dat_ecz |>
  left_join(select(dat_ran, record_id, trt)) |>
  filter(ecz_preexisting) |>
  summarise(n = sum(ecz_preexisting), .by = trt)

# With eczema
tbl1 <- dat_ecz |>
  left_join(select(dat_ran, record_id, trt), join_by(record_id)) |>
  filter(!ecz_preexisting) |>
  mutate(
    ecz = ecz & !ecz_preexisting,
    ecz_6m = ecz & !ecz_preexisting & outageval_months <= 6,
    ecz_12m = ecz & !ecz_preexisting & outageval_months <= 12,
    ecz_18m = ecz & !ecz_preexisting & outageval_months <= 18
  ) |>
  tbl_summary(
    by = trt,
    include = c(ecz, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**")

# With positive SPT
tbl2 <- dat_ecz |>
  left_join(select(dat_ran, record_id, trt), join_by(record_id)) |>
  filter(!ecz_preexisting) |>
  mutate(
    ecz = ecz & !ecz_preexisting & any_pos_spt,
    ecz_6m = ecz & !ecz_preexisting & outageval_months <= 6 & any_pos_spt,
    ecz_12m = ecz & !ecz_preexisting & outageval_months <= 12 & any_pos_spt,
    ecz_18m = ecz & !ecz_preexisting & outageval_months <= 18 & any_pos_spt
  ) |>
  tbl_summary(
    by = trt,
    include = c(ecz, ecz_6m, ecz_12m, ecz_18m),
    label = c(
      ecz ~ "Eczema",
      ecz_6m ~ "Eczema by 6-months",
      ecz_12m ~ "Eczema by 12-months",
      ecz_18m ~ "Eczema by 18-months"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**")

tbl_stack(
  list(tbl1, tbl2),
  group_header = c("Total cases", "Cases with positive SPT")
) |>
  as_gt() |>
  tab_options(table.font.size = px(10)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
  )
```

A summary of the characteristics of these eczema diagnoses are given in @tbl-eczema-characteristics.

```{r}
#| label: tbl-eczema-characteristics
#| tbl-cap: |
#|   Characteristics of eczema cases.
dat_ecz |>
  filter(!is.na(outalltp), !ecz_preexisting) |>
  left_join(select(dat_ran, record_id, trt), join_by(record_id)) |>
  mutate(outeczsrce = fct_infreq(outeczsrce)) |>
  tbl_summary(
    by = trt,
    include = c(outsev, outeczsrce, outageval_months, ecz_age_months),
    label = list(
      outeczsrce ~ "Source of information",
      outageval_months ~ "Age at first reaction",
      ecz_age_months ~ "Age at diagnosis"
    ),
    missing_text = "Missing"
  ) |>
  add_overall(last = TRUE) |>
  modify_header(label = "**Outcome**") |>
  as_gt() |>
  tab_options(table.font.size = "10px") |>
  tab_style(style = cell_text(align = "left"), locations = cells_footnotes())
```

```{r}
dat_ecz |>
  filter(outageval_months < 1.5) |>
  select(
    record_id,
    ecz_age_months,
    outageval_months,
    outageval,
    outageunit,
    ecz_preexisting,
    fhq_ecz
  )
dat_ecz |>
  filter(!is.na(outalltp), outageval_months > 1.5) |>
  ggplot(aes(ecz_age_months, floor(outageval_months))) +
  geom_point() +
  geom_abline() +
  coord_fixed() +
  labs(
    x = "Age allergy confirmed by clinician (months)",
    y = "Age allergy first occurred (months)"
  ) +
  scale_x_continuous(breaks = 1:20) +
  scale_y_continuous(breaks = 1:20)
```

---

# Food allergy

The primary source of information is the outcome form, in particular, records where `outalltp` is one of *Anaphylaxis*, *Highly probably IgE mediated food allergy*, or *Oral food challenge*. There is also a form `food_challenge` for the oral food challenges themselves.

```{r}
#| code-summary: Food allergy records
dat_fa <- dat_out_raw |>
  filter(!is.na(outalltp), outalltp != "Eczema")
dat_fa |> count(outalltp)
```

Generally, there is one record per allergen and reaction type. For example,

```{r}
#| code-summary: Example infant
dat_fa |>
  filter(record_id == "4629-92") |>
  select(
    record_id,
    outalltp,
    outrepdat,
    outdiagdat,
    outfrasrcespec,
    outfrafood,
    outageval,
    outageunit
  )
```

How many individuals had some kind of food allergy outcome recorded, and how many records did they?

```{r}
#| code-summary: Food allergy records per infant
dat_fa |>
  count(record_id, name = "records") |>
  count(records, name = "infants")
```

To meet the primary outcome we require a positive skin prick test to the relevant allergen in addition to an outcome report of anaphylaxis, positive oral food challenge, or highly probable IgE mediated food allergy. Rather than checking for *any* positive SPT, need to check specifically for one which is positive to the relevant food.

## Food Challenge

For reference, in the oral food challenge records:

- `ofcfoodtp1` - peanut
- `ofcfoodtp2` - egg
- `ofcfoodtp3` - milk
- `ofcfoodtp4` - cashew

From the outcome report, there were 21 infants with a positive oral food challenge outcome. 

```{r}
ofc_recs <- dat_out_raw |>
  filter(outalltp == "Oral food challenge") |>
  arrange(record_id) |>
  select(
    record_id,
    outsev,
    outrepdat,
    outawardat,
    outdiagdat,
    outageval,
    outageunit,
    outageval_months,
    outfrafood,
    outfraeldose
  )
ofc_recs |>
  arrange(outageval_months) |>
  gt() |>
  tab_options(table.font.size = "10px")
```

All had a positive SPT to the relevant food by 18 months of age.

```{r}
out_ofc_foods <- dat_out_raw |>
  filter(outalltp == "Oral food challenge") |>
  arrange(record_id) |>
  select(
    record_id,
    outfrafood
  ) |>
  summarise(
    outfrafood = paste(tolower(outfrafood), collapse = ", "),
    .by = record_id
  ) |>
  left_join(dat_spt_pos, join_by(record_id)) |>
  select(record_id, outfrafood, spt_age, spt_pos_1mm) |>
  arrange(str_rank(record_id, numeric = TRUE))
out_ofc_foods |>
  gt() |>
  tab_options(table.font.size = "10px")
```

Some infants had oral food challenges but no outcome report for oral food challenge, despite a positive result. 

```{r}
ofc_positive <- dat_fch_raw |>
  filter(
    ofcfoodtp1 == "Positive" |
      ofcfoodtp2 == "Positive" |
      ofcfoodtp3 == "Positive" |
      ofcfoodtp4 == "Positive" |
      ofcoutoth1 == "Positive" |
      ofcoutoth2 == "Positive"
  ) |>
  arrange(str_rank(record_id, numeric = TRUE))
right_join(out_ofc_foods, ofc_positive, relationship = "many-to-many") |>
  filter(is.na(outfrafood))
```

Usually, either anaphylaxis or IgE mediated food allergy was reported. An exception was 5785-123 who have only eczema reported.

```{r}
ids <- right_join(out_ofc_foods, ofc_positive, relationship = "many-to-many") |>
  filter(is.na(outfrafood)) |>
  pull(record_id)
dat_out_raw |>
  filter(record_id %in% ids) |>
  print(n = Inf)
```

## Anaphylaxis

There were 7 participants with a record of anaphylaxis. 5785-283 did not have a positive skin prick test to the reported food (sesame) however, this appears to have been a data entry issue. Apparently there was a positive SPT to sesame, it's just not recorded in REDCap, awaiting an update to this record. Record ID 68 had a positive SPT, however it was at 19 months of age which does not meet the 18 month cut-off. Participant 4629-92 did not have a positive SPT for the two anaphylaxis foods, cow's milk and walnut, however, for cow's milk this was due to a missing negative control. The response to cow's milk was 9mm, which may have been positive.

```{r}
ana_recs <- dat_out_raw |>
  filter(outalltp == "Anaphylaxis") |>
  arrange(record_id) |>
  select(
    record_id,
    outsev,
    outrepdat,
    outawardat,
    outdiagdat,
    outageval,
    outageunit,
    outageval_months,
    outfrafood,
    outfraeldose
  )
ana_recs |>
  arrange(outageval_months) |>
  gt() |>
  tab_options(table.font.size = "10px")
```

```{r}
out_ana_foods <- dat_out_raw |>
  filter(outalltp == "Anaphylaxis") |>
  arrange(record_id) |>
  select(
    record_id,
    outfrafood
  ) |>
  summarise(
    outfrafood = paste(tolower(outfrafood), collapse = ", "),
    .by = record_id
  ) |>
  left_join(dat_spt_pos, join_by(record_id)) |>
  select(record_id, outfrafood, spt_age, spt_pos_1mm) |>
  arrange(str_rank(record_id, numeric = TRUE))
out_ana_foods |>
  gt() |>
  tab_options(table.font.size = "10px")
```

```{r}
dat_spt |>
  filter(record_id == "4629-92") |>
  filter(spt_tested == "cow's milk", !is.na(spt_result))
```

## Highly Probably IgE Mediated Food Allergy

There were 46 individuals with a record of highly probable IgE mediated food allergy.
From manual review, a few were missing positive SPT:

- 5785-239 (eggplant)
- 5785-243 (cows milk, rice)
- 5785-331 (egg)
- 75

```{r}
ige_recs <- dat_out_raw |>
  filter(outalltp == "Highly probably IgE mediated food allergy") |>
  select(
    record_id,
    outalltp,
    outsev,
    outdiagdat,
    outfrasource,
    outfrafood,
    outfraexptp
  ) |>
  arrange(record_id) |>
  summarise(
    foods = paste(outfrafood, collapse = ", "),
    .by = record_id
  )
```
