---
title: Infant vaccinations
subtitle: Check both study and non-study vaccination receipts.
---

{{< include _setup.qmd >}}

As part of our due diligence for OPTIMUM we need to confirm that all the infants have received 3 polio doses by the time they are 18 months old. This should happen as part of them attending V3, but we have had some infants not attend so we have had sites confirming that they have received the vaccine outside of the trial using the AIR, follow up and the non-study vaccine log.

All participants should have received the following:

- Infanrix Hexa at 4 months and 6 months (recorded in the non-study vaccine log), and
- Infanrix-IPV at the 18-month visit (V3) (recorded in Vaccine Administration V3).

In some cases, the V3 polio dose was given outside the study, in which case it would also be documented in the non-study vaccine log.

- Some participants have been LTFP or withdrawn
- Some have received three doses including the blinded study dose, which is not documented in REDCap because of the blinding (aP contains polio, wP does not).
- Occasionally, participants were given an alternative vaccine instead of Infanrix Hexa or Infanrix-IPV. It could be recorded as the following in the non-study vaccine log:
  - IPOL
  - Polio
  - Poliomyelitis
  - Quadracel
  - Vaxelis

We do not necessarily care about the timing, just that 3 polio vaccines were received by 18 months of age.

```{r}
dat_st2 <- qread(st2_file) |>
  filter(
    redcap_form_name %in%
      c(
        "randomisation",
        "demographics",
        "study_termination",
        "participant_assessment",
        "vaccine_administration_v1",
        "vaccine_administration_v3",
        "nonstudy_vaccination_log"
      )
  )
dat_allocations <- select_form(dat_raw, "allocations")
```

```{r}
dat_rand <- extract_tibble(dat_st2, "randomisation") |>
  select(record_id, subjid, rand, randdattim) |>
  mutate(rand = as.character(rand)) |>
  left_join(dat_allocations, join_by(rand))

dat_demo <- extract_tibble(dat_st2, "demographics")
dat_st <- extract_tibble(dat_st2, "study_termination")
dat_pa <- extract_tibble(dat_st2, "participant_assessment")
dat_vax1 <- extract_tibble(dat_st2, "vaccine_administration_v1")
dat_vax3 <- extract_tibble(dat_st2, "vaccine_administration_v3")
dat_nstv <- extract_tibble(dat_st2, "nonstudy_vaccination_log")

dat_nstv_long <- dat_nstv |>
  select(-redcap_event, -redcap_data_access_group, -form_status_complete) |>
  pivot_longer(
    ipvac1:ipnstvacdat6,
    names_pattern = "(ipvac|vacothspec|ipnstvacdat|vacaddyn)([1-8])",
    names_to = c(".value", "vaxseq")
  ) |>
  filter(!is.na(ipvac)) |>
  select(-instvacyn, -vacaddyn)
```

```{r}
dat_ref <- dat_demo |>
  select(record_id, birthdat, visdat1) |>
  left_join(dat_rand, join_by(record_id)) |>
  left_join(
    dat_st |>
      select(record_id, discdat, streas, stetrreas, ippvspec),
    join_by(record_id)
  ) |>
  mutate(
    discage_wks = interval(birthdat, discdat) %/% weeks(1),
    discage_mth = interval(birthdat, discdat) %/% months(1)
  )

# Visit 1 and 3 vaccinations which include polio
dat_ref_vax3 <- dat_ref |>
  left_join(
    dat_pa |>
      filter(redcap_event == "visit_3") |>
      select(record_id, visdat, windyn, windreas, windothspec, ipnsvac),
    join_by(record_id)
  ) |>
  left_join(
    dat_vax3 |>
      select(record_id, vac7admdat, vac7iploc, vac7adyn),
    join_by(record_id)
  ) |>
  mutate(
    polio_v1 = as.numeric(trt == "aP"),
    polio_v3_age = interval(birthdat, visdat) %/% months(1),
    polio_v3 = as.numeric(!is.na(vac7iploc) & (vac7iploc != "Not given"))
  )

# Non-study vaccinations which include polio
dat_ref_nstv <- dat_ref |>
  left_join(dat_nstv_long, join_by(record_id)) |>
  mutate(nstv_age = interval(birthdat, ipnstvacdat) %/% months(1))

polio_str_1 <- "infan|ifan|ipv|ipol|polio|quad|vaxe"
polio_str_2 <- "hexa|ipv|ipol|polio|quad|vaxe"
dat_ref_nstv_polio <- dat_ref_nstv |>
  summarise(
    polio_nstv_count = sum(
      (grepl(polio_str_2, tolower(ipvac)) |
        grepl(polio_str_2, tolower(vacothspec)))
    ),
    .by = record_id
  )

dat_ref_vax <- left_join(dat_ref_vax3, dat_ref_nstv_polio, join_by(record_id))
dat_ref_vax_total <- dat_ref_vax |>
  select(record_id, discage_mth, polio_v1, polio_v3, polio_nstv_count) |>
  mutate(polio_total = polio_v1 + polio_v3 + polio_nstv_count)
```

Records to check as having fewer than 3 polio vaccinations... I just ran through all these manually to confirm.

```{r}
dat_ref_vax_total |>
  filter(polio_total < 3) |>
  print(n = Inf)

# dat_ref_vax3 |>
#   filter(record_id == "8644-61") |>
#   select(record_id, streas, stetrreas, ippvspec, discage_mth, trt, birthdat, visdat1, visdat, vac7iploc)
# dat_ref_nstv |>
#   filter(record_id == "8644-61") |>
#   select(record_id, discage_mth, trt, birthdat, visdat1, ipnstvacdat, ipvac, vacothspec, nstv_age) |>
#   arrange(nstv_age)
```
